---
title: "What does Crossref add to our journal data?"
subtitle: "Meeting update · 15 September 2026"
lang: en
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    code-fold: show
    code-summary: "R code"
    code-copy: true
    embed-resources: true
    css: crossref-meeting-report.css
execute:
  echo: true
  message: false
  warning: false
  error: false
  cache: false
params:
  data_file: "../artifacts/ojs_journal_enrichment_simple_ver/pkp-ojs-multisource-enriched-simple_ver.parquet"
  crossref_cache: "../artifacts/ojs_journal_enrichment/full-v7/crossref-journals.rds"
  bq_cli: "/Users/yann.jy/.local/google-cloud-sdk/bin/bq"
---

```{r}
#| label: read-saved-data
library(arrow)
library(dplyr)
library(jsonlite)
fmt <- function(x) format(x, big.mark = ",", scientific = FALSE, trim = TRUE)

journals <- read_parquet(params$data_file)
crossref <- readRDS(params$crossref_cache)
```

```{r}
#| label: display-settings
#| include: false
n_valid <- sum(journals$identifier_status == "valid")
snapshot_date <- substr(crossref$retrieved_at, 1, 10)
output_dir <- "../artifacts/crossref-meeting-report"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
stopifnot(nrow(journals) == 98273, !"admin_email" %in% names(journals))
```

We use the existing **PKP V7 list enriched with OpenAlex and Crossref**:
`r fmt(nrow(journals))` records, including `r fmt(n_valid)` with a usable ISSN.
The saved Crossref data were collected on `r snapshot_date`. The results below
are calculated from that merged file when this report runs.

## 1. Crossref fields

**Agreed task:** Compare every Crossref source field with related OpenAlex
fields, using actual values from the same journal.

```{r}
#| label: select-one-journal
example <- journals |>
  filter(openalex_id == "https://openalex.org/S4210240683",
         openalex_match_status == "unique",
         crossref_match_status == "unique")
stopifnot(nrow(example) == 1,
          example$crossref_title == example$openalex_display_name,
          setequal(fromJSON(example$crossref_ISSN),
                   fromJSON(example$openalex_issn)))
```

The example throughout both tables is **`r example$crossref_title`**
(ISSNs `r paste(fromJSON(example$crossref_ISSN), collapse = " and ")`).
All 11 Crossref source columns in the saved data are shown below.
The examples are read from this one row of the merged file. Long nested fields
show selected entries; **Full saved value** opens the complete content.
`NA` means the corresponding field exists but has no value for this journal.

```{r}
#| label: field-mapping
#| code-fold: true
#| code-summary: "R code: field names and explanations"
field_map <- tibble::tribble(
  ~cr, ~cr_meaning, ~oa, ~oa_meaning,
  "title", "Journal name.",
  "display_name", "Journal name.",
  "publisher", "Publisher name.",
  "host_organization_name", "Publisher or hosting organisation name.",
  "ISSN", "Journal ISSNs.",
  "issn", "Journal ISSNs, without print/online labels.",
  "issn-type", "ISSNs with print or electronic labels.",
  "issn", "Same identifiers; no medium labels in this list.",
  "counts", "DOI counts: total, current and backfile publications.",
  "works_count", "Number of works recorded by OpenAlex.",
  "breakdowns", "DOI counts by publication (issued) year.",
  "counts_by_year", "Works published and citations received in each year.",
  "subjects", "Subject labels supplied by Crossref.",
  "topics", "OpenAlex topic labels and associated work counts.",
  "coverage", "Share of records containing each kind of metadata, for current and backfile publications.",
  "", "",
  "coverage-type", "The same coverage measures grouped by all, current and backfile publications.",
  "", "",
  "flags", "Yes/no indicators for deposited content and metadata.",
  "", "",
  "last-status-check-time", "Crossref status-check timestamp.",
  "", ""
)
cr_fields <- field_map$cr
stopifnot(setequal(cr_fields, names(crossref$journals[[1]])),
          all(paste0("crossref_", cr_fields) %in% names(example)),
          all(paste0("openalex_", field_map$oa[field_map$oa != ""]) %in%
                names(example)))
```

```{r}
#| label: extract-field-examples
#| code-fold: true
#| code-summary: "R code: select entries from nested fields"
doi_counts <- fromJSON(example$crossref_counts)
cr_years <- fromJSON(example$crossref_breakdowns)[["dois-by-issued-year"]]
oa_years <- fromJSON(example$openalex_counts_by_year)
oa_topics <- fromJSON(example$openalex_topics)
cr_coverage <- fromJSON(example$crossref_coverage)
coverage <- fromJSON(example[["crossref_coverage-type"]])$all
cr_flags <- fromJSON(example$crossref_flags)

# Use the same year for the DOI and work counts.
comparison_year <- 2025
cr_year_count <- cr_years[cr_years[, 1] == comparison_year, 2]
oa_year <- oa_years[oa_years$year == comparison_year, ]
stopifnot(length(cr_year_count) == 1, nrow(oa_year) == 1)

# These excerpts are computed from the saved values; the full values remain available.
excerpts <- list(
  crossref_breakdowns = list(`dois-by-issued-year` =
    list(c(comparison_year, cr_year_count))),
  openalex_counts_by_year = oa_year,
  openalex_topics = head(oa_topics[c("display_name", "count")], 3),
  crossref_coverage = cr_coverage[c("abstracts-current", "references-current",
                                   "orcids-current")],
  `crossref_coverage-type` = list(all = coverage[c("abstracts", "references",
                                                  "orcids")]),
  crossref_flags = cr_flags[c("deposits-abstracts-current",
                             "deposits-references-current",
                             "deposits-orcids-current")]
)
```

```{r}
#| label: format-field-comparison
#| code-fold: true
#| code-summary: "R code: display the comparison"
field_cell <- function(prefix, field, meaning) {
  if (field == "") return("")
  paste0("<code>", prefix, field, "</code><br>",
         htmltools::htmlEscape(meaning))
}
value_cell <- function(column) {
  if (column == "openalex_") return("")
  value <- example[[column]][[1]]
  if (is.na(value)) return("<code>NA</code> (not supplied)")
  value <- as.character(value)
  full <- if (jsonlite::validate(value)) prettify(value, indent = 2) else value
  full <- paste0("<pre class='field-value'>",
                 htmltools::htmlEscape(full), "</pre>")
  if (is.null(excerpts[[column]])) return(full)
  preview <- toJSON(excerpts[[column]], auto_unbox = TRUE, pretty = TRUE)
  paste0("<pre class='field-value'>", htmltools::htmlEscape(preview),
         "</pre><details><summary>Full saved value</summary>",
         full, "</details>")
}
comparison <- data.frame(
  `Crossref column and meaning` = mapply(field_cell, "crossref_",
                                         field_map$cr, field_map$cr_meaning),
  `Crossref example` = vapply(paste0("crossref_", field_map$cr),
                              value_cell, character(1)),
  `OpenAlex column and meaning` = mapply(field_cell, "openalex_",
                                         field_map$oa, field_map$oa_meaning),
  `OpenAlex example` = vapply(paste0("openalex_", field_map$oa),
                              value_cell, character(1)),
  check.names = FALSE
)
```

### Fields providing the same or related information

::: {.field-comparison}
```{r}
#| label: show-related-fields
#| results: asis
knitr::kable(comparison[field_map$oa != "", ], format = "html", escape = FALSE,
             row.names = FALSE)
```
:::

### Crossref fields without a counterpart in this OpenAlex table

The OpenAlex cells are left blank where no corresponding source field exists.

::: {.field-comparison}
```{r}
#| label: show-additional-fields
#| results: asis
knitr::kable(comparison[field_map$oa == "", ], format = "html", escape = FALSE,
             row.names = FALSE)
```
:::

**The title and both ISSNs agree.** Crossref adds the print/electronic labels
and the publisher name, **`r example$crossref_publisher`**, which is missing
from the related OpenAlex field for this journal. For subject information,
Crossref supplies an empty list, while OpenAlex supplies topic labels.

**Both sources describe publication volume, but count their own records.**
Crossref has `r fmt(doi_counts[["total-dois"]])` DOI registrations and OpenAlex
has `r fmt(example$openalex_works_count)` works. In `r comparison_year`, the
corresponding figures are `r fmt(cr_year_count)` and `r fmt(oa_year$works_count)`.
These figures show a difference between the databases; identifying the
individual publications behind it would require a publication-level comparison.
Crossref's *current* category covers the snapshot year and the preceding two
years of publication; *backfile* covers earlier publications.

**Crossref's main additional information concerns what publishers deposit.**
For this journal, across all years, `r round(100 * coverage$abstracts)`% of
Crossref records contain abstracts, `r round(100 * coverage$references)`%
contain references and `r round(100 * coverage$orcids, 1)`% contain ORCIDs.
The zero for references means references were not deposited with Crossref;
it does not mean the papers contain no references. `coverage`, `coverage-type`
and `flags` express related deposit information as proportions or yes/no
indicators. The status-check timestamp is a record of Crossref's check, not
an additional characteristic of the journal's research.

Field meanings follow the [Crossref journal schema](https://api.crossref.org/swagger-docs),
[Crossref coverage definitions](https://www.crossref.org/documentation/reports/participation-reports/)
and [OpenAlex Source fields](https://help.openalex.org/data/sources/attributes/).

## 2. Records added by Crossref

**Agreed task:** Find records with Crossref information but no OpenAlex match,
and show their number and examples.

We select records with one Crossref match and no OpenAlex match in the saved
ISSN search. Records with several possible OpenAlex matches are counted separately.

```{r}
#| label: crossref-only-records
crossref_only <- journals |>
  filter(crossref_match_status == "unique",
         openalex_match_status == "unmatched")
n_uncertain <- sum(journals$crossref_match_status == "unique" &
                   journals$openalex_match_status == "ambiguous")

examples <- crossref_only |>
  transmute(Journal = crossref_title,
            ISSN = gsub("|", ", ", crossref_matched_issns, fixed = TRUE),
            `Registered DOIs` = vapply(crossref_counts,
              function(x) as.numeric(fromJSON(x)$`total-dois`), numeric(1)))
knitr::kable(examples, format.args = list(big.mark = ","))
```

Crossref adds a match where OpenAlex found none for **`r nrow(crossref_only)` of the
`r fmt(n_valid)` records with a usable ISSN**
(`r sprintf("%.3f", 100 * nrow(crossref_only) / n_valid)`%). Its additional reach
is therefore small in this PKP list; its main contribution is the extra
information it provides about records already found in both sources.
Another **`r fmt(n_uncertain)` records** have one Crossref match but several
possible OpenAlex matches. They need a separate identity check.
The `r fmt(nrow(journals) - n_valid)` records without a usable ISSN were not searched.

## 3. Empty lists and missing values

**Agreed task:** Check an actual empty-list example and show how it should be treated.

Here we inspect `subjects`, the Crossref subject-label field, for the first
record listed above.

```{r}
#| label: inspect-empty-list
example_issn <- strsplit(crossref_only$crossref_matched_issns[1],
                        "|", fixed = TRUE)[[1]][1]
raw_example <- Filter(function(x) example_issn %in% x$ISSN,
                      crossref$journals)[[1]]
raw_example$subjects
```

The empty list means that **no subject labels were supplied**. We treat it as
missing subject information in the analysis:

```{r}
#| label: subject-missingness
matched <- journals |> filter(crossref_match_status == "unique")
matched <- matched |>
  mutate(subjects_clean = na_if(crossref_subjects, "[]"))

knitr::kable(data.frame(
  `Subject information` = c("Supplied", "Missing"),
  Records = c(sum(!is.na(matched$subjects_clean)),
              sum(is.na(matched$subjects_clean))),
  check.names = FALSE), format.args = list(big.mark = ","))
```

All **`r fmt(nrow(matched))` records with one Crossref match** have no subject labels
in this saved data. Crossref's `subjects` field therefore does not add subject
classification here; OpenAlex's topic information is the useful starting point.

## 4. BigQuery access

**Agreed task:** Test whether a query now runs and identify the permission needed.

```{r}
#| label: bigquery-test
query_test <- system2(params$bq_cli, c(
  "--headless=true", "--project_id=insyspo", "--location=US",
  "--dataset_id=insyspo:userdb_yann_jyal", "query",
  "--use_legacy_sql=false", "--maximum_bytes_billed=1",
  shQuote("SELECT 1 AS access_test")), stdout = TRUE, stderr = TRUE)
```

```{r}
#| label: record-query-result
#| include: false
query_status <- attr(query_test, "status")
if (is.null(query_status)) query_status <- 0L
missing_jobs_permission <- any(grepl("bigquery.jobs.create", query_test, fixed = TRUE))
query_sentence <- if (query_status == 0L) {
  "The constant query succeeded in insyspo; the query-job access test now passes."
} else if (missing_jobs_permission) {
  "The SELECT 1 test still fails in insyspo; the account needs BigQuery Job User (roles/bigquery.jobUser), which includes bigquery.jobs.create."
} else {
  "The query test did not complete; its error needs checking before attributing it to a permission."
}
writeLines(query_test, file.path(output_dir, "bigquery-test.txt"))
stopifnot(query_status == 0L || missing_jobs_permission)
write.csv(examples, file.path(output_dir, "crossref-only-examples.csv"), row.names = FALSE)
write_json(list(
  rendered_at_utc = format(Sys.time(), "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
  input_file = params$data_file,
  input_sha256 = digest::digest(file = params$data_file, algo = "sha256"),
  crossref_snapshot = crossref$retrieved_at,
  pkp_rows = nrow(journals), valid_issn_rows = n_valid,
  crossref_fields = cr_fields,
  crossref_only = nrow(crossref_only), crossref_with_uncertain_openalex = n_uncertain,
  subject_missing = sum(is.na(matched$subjects_clean)),
  bigquery_exit_status = query_status, missing_jobs_permission = missing_jobs_permission,
  R = R.version.string), file.path(output_dir, "run-summary.json"),
  auto_unbox = TRUE, pretty = TRUE)
```

`r query_sentence`

---

Tasks follow the [meeting discussion](https://github.com/YannJY02/InvisibleResearch/blob/c06f31f0d517dddafd779e5cc38a6bf589d3c039/meeting-reports/2026-09-14-openalex-baseline-and-access.md)
(02:13–08:29 and 34:23–34:31), alongside the
[field-review request](https://github.com/invisibleinfo/invisible-research/issues/5#issuecomment-5553748839).
