---
title: "Nature: OpenAlex and Crossref journal fields"
subtitle: "Meeting update · Full field comparison · 15 September 2026"
lang: en
format:
  html:
    theme: cosmo
    page-layout: full
    toc: true
    toc-depth: 2
    code-fold: show
    code-summary: "R code"
    code-copy: true
    embed-resources: true
    css: crossref-meeting-report.css
execute:
  echo: true
  message: false
  warning: false
  error: false
  cache: false
params:
  refresh_nature: false
  data_file: "../artifacts/ojs_journal_enrichment_simple_ver/pkp-ojs-multisource-enriched-simple_ver.parquet"
  crossref_cache: "../artifacts/ojs_journal_enrichment/full-v7/crossref-journals.rds"
  bq_cli: "/Users/yann.jy/.local/google-cloud-sdk/bin/bq"
---

```{r}
#| label: read-saved-data
library(arrow)
library(dplyr)
library(jsonlite)
library(httr2)
fmt <- function(x) format(x, big.mark = ",", scientific = FALSE, trim = TRUE)

journals <- read_parquet(params$data_file)
crossref <- readRDS(params$crossref_cache)
```

```{r}
#| label: display-settings
#| include: false
n_valid <- sum(journals$identifier_status == "valid")
snapshot_date <- substr(crossref$retrieved_at, 1, 10)
output_dir <- "../artifacts/crossref-meeting-report"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
stopifnot(nrow(journals) == 98273, !"admin_email" %in% names(journals))
```

The field comparison uses **Nature**, retrieved directly from both journal
APIs. Tasks 2 and 3 use the existing **PKP V7 list enriched with OpenAlex and
Crossref**: `r fmt(nrow(journals))` records, including `r fmt(n_valid)` with a
usable ISSN, with Crossref data collected on `r snapshot_date`.

## 1. Nature: all journal fields, expanded

**Agreed task:** Show every OpenAlex and Crossref journal field and subfield,
explain its meaning, and compare actual values for Nature. Put related fields
first, followed by fields available on only one side.

```{r}
#| label: retrieve-nature
nature_dir <- file.path(output_dir, "nature")
dir.create(nature_dir, recursive = TRUE, showWarnings = FALSE)
urls <- c(crossref = "https://api.crossref.org/journals/0028-0836",
          openalex = "https://api.openalex.org/sources/issn:0028-0836")
files <- setNames(file.path(nature_dir, paste0(names(urls), ".json")), names(urls))
if (params$refresh_nature || !all(file.exists(files))) {
  responses <- lapply(urls, function(url) {
    request(url) |>
      req_user_agent("InvisibleResearch journal field comparison") |>
      req_timeout(60) |> req_perform()
  })
  for (source in names(responses)) {
    writeBin(resp_body_raw(responses[[source]]), files[[source]])
  }
  write_json(list(retrieved_at_utc = format(Sys.time(),
    "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"), urls = as.list(urls)),
    file.path(nature_dir, "retrieval.json"), auto_unbox = TRUE, pretty = TRUE)
}
cr <- read_json(files[["crossref"]])$message
oa <- read_json(files[["openalex"]])
nature_retrieval <- read_json(file.path(nature_dir, "retrieval.json"))
snapshot_year <- as.integer(substr(nature_retrieval$retrieved_at_utc, 1, 4))
current_period <- paste0(snapshot_year - 2L, "–", snapshot_year)
stopifnot(cr$title == "Nature", oa$display_name == "Nature",
          oa$type == "journal", setequal(unlist(cr$ISSN), unlist(oa$issn)))
```

Both records identify **Nature**, with ISSNs **0028-0836** and **1476-4687**.
The responses were saved at **`r nature_retrieval$retrieved_at_utc`**.
This code retrieves the complete records without selecting a subset of fields;
subsequent renders reuse these snapshots. Set `refresh_nature: true` to fetch
them again.

```{r}
#| label: expand-every-field
#| code-fold: true
#| code-summary: "R code: expand every object and list without dropping empty values"
flatten_fields <- function(x, path = "", item = NA_integer_, context = "") {
  if (is.list(x) && length(x) && !is.null(names(x))) {
    return(bind_rows(lapply(names(x), function(key) {
      next_path <- if (nzchar(path)) paste(path, key, sep = ".") else key
      flatten_fields(x[[key]], next_path, item, context)
    })))
  }
  if (is.list(x) && length(x)) {
    return(bind_rows(lapply(seq_along(x), function(i) {
      entry <- x[[i]]
      if (path == "breakdowns.dois-by-issued-year") {
        return(bind_rows(lapply(seq_along(entry), function(j) {
          flatten_fields(entry[[j]], paste0(path, "[][", j, "]"),
                         i, paste("Year", entry[[1]]))
        })))
      }
      label <- paste("Item", i)
      if (is.list(entry) && !is.null(names(entry))) {
        for (key in c("year", "display_name", "currency", "type")) {
          if (!is.null(entry[[key]])) {
            label <- paste(key, entry[[key]], sep = ": ")
            break
          }
        }
      }
      flatten_fields(entry, paste0(path, "[]"), i, label)
    })))
  }
  value <- if (is.null(x)) "null" else if (is.list(x)) "[]" else
    if (is.logical(x)) tolower(as.character(x)) else as.character(x)
  tibble::tibble(path, item, context, value)
}
cr_flat <- flatten_fields(cr)
oa_flat <- flatten_fields(oa)
cr_fields <- unique(cr_flat$path)
oa_fields <- unique(oa_flat$path)
```

Crossref's **`r length(cr)` top-level fields** expand to
**`r length(cr_fields)` field paths**; OpenAlex's **`r length(oa)`** expand to
**`r length(oa_fields)`**. Every path has its own row below. `[]` in a field
name means a repeated list entry. For lists, the example and its label identify
the same entry across subfields; **all values** opens the remaining entries.
Annual comparisons use **2025**. A displayed `null` or `[]` is an actual empty
API value; a blank counterpart cell means there is no direct matching field.

```{r}
#| label: explain-field-paths
#| code-fold: true
#| code-summary: "R code: field explanations, kept directly in this QMD"
metrics <- c(
  abstracts = "an abstract", references = "a reference list",
  affiliations = "at least one contributor affiliation",
  orcids = "at least one author ORCID",
  licenses = "licence information", funders = "funder information",
  `award-numbers` = "a funding award number", `ror-ids` = "a ROR organisation ID",
  `affiliation-ror-ids` = "a contributor-affiliation ROR ID",
  `funder-ror-ids` = "a funder ROR ID",
  `resource-links` = "full-text resource links",
  `similarity-checking` = "a Similarity Check full-text URL",
  `update-policies` = "Crossmark update-policy information",
  descriptions = "description metadata (a separate API category from abstracts)"
)
periods <- c(all = "all publication years",
  current = paste(current_period, "publications"),
  backfile = paste("publications before", snapshot_year - 2L))
cr_definitions <- c(
  title = "Journal title.", publisher = "Registered publisher name.",
  `ISSN[]` = "Journal ISSNs.",
  `issn-type[].value` = "ISSN, paired with its medium label.",
  `issn-type[].type` = "Print or electronic medium for the paired ISSN.",
  `counts.total-dois` = "All DOI records registered for this journal.",
  `counts.current-dois` = paste0("DOI records for ", periods[["current"]], "."),
  `counts.backfile-dois` = paste0("DOI records for ", periods[["backfile"]], "."),
  `breakdowns.dois-by-issued-year[][1]` = "Publication (issued) year; first element of each pair.",
  `breakdowns.dois-by-issued-year[][2]` = "DOI count for that year; second element of the same pair.",
  subjects = "Retired subject classification; the API retains an empty list.",
  `last-status-check-time` = "Crossref status-check time, in Unix milliseconds.",
  `flags.deposits` = "Whether the journal has deposited records.",
  `flags.deposits-articles` = "Whether it has deposited article records."
)
describe_cr <- function(path) {
  if (path %in% names(cr_definitions)) return(unname(cr_definitions[path]))
  if (startsWith(path, "coverage-type.")) {
    parts <- strsplit(path, ".", fixed = TRUE)[[1]]
    period <- parts[2]; metric <- parts[3]
    if (metric == "last-status-check-time") return(paste(
      "Status-check time for", periods[[period]], "(Unix milliseconds)."))
  } else {
    key <- sub("^(coverage\\.|flags\\.deposits-)", "", path)
    period <- sub(".*-(current|backfile)$", "\\1", key)
    metric <- sub("-(current|backfile)$", "", key)
  }
  meaning <- metrics[[metric]]
  if (startsWith(path, "flags.")) {
    paste0("Whether ", meaning, " was deposited for ", periods[[period]], ".")
  } else {
    paste0("Share of records with ", meaning, "; ", periods[[period]], ".")
  }
}
oa_definitions <- c(
  id = "OpenAlex source identifier.", issn_l = "Linking ISSN joining medium variants.",
  `issn[]` = "Journal ISSNs.", display_name = "Journal title.",
  host_organization = "Hosting publisher's OpenAlex ID.",
  host_organization_name = "Hosting publisher or imprint name.",
  `host_organization_lineage[]` = "Host and parent publisher IDs.",
  works_count = "Works attributed to this journal in OpenAlex.",
  oa_works_count = "Open-access works.", cited_by_count = "Citations received by the journal's works.",
  `summary_stats.2yr_mean_citedness` = "Mean cumulative citations for recent publications; OpenAlex's measure, not JCR's impact factor.",
  `summary_stats.h_index` = "h works have each received at least h citations.",
  `summary_stats.i10_index` = "Works with at least 10 citations.",
  is_oa = "Fully open-access journal flag.", is_in_doaj = "DOAJ inclusion.",
  is_in_doaj_since_year = "Year DOAJ inclusion began.",
  is_high_oa_rate = "High open-access share flag.",
  is_high_oa_rate_since_year = "Year that high-share status began.",
  is_in_scielo = "SciELO inclusion.", is_ojs = "Open Journal Systems platform flag.",
  is_preprint_repository = "Preprint-repository flag.",
  is_core = "CWTS Core source flag; not Web of Science inclusion.",
  oa_flip_year = "Year of conversion to open access.",
  first_publication_year = "Earliest publication year in this OpenAlex record.",
  last_publication_year = "Latest publication year in this OpenAlex record.",
  `ids.openalex` = "Repeated OpenAlex ID.", `ids.issn_l` = "Repeated linking ISSN.",
  `ids.issn[]` = "Repeated ISSN list.", `ids.mag` = "Legacy Microsoft Academic Graph ID.",
  `ids.wikidata` = "Wikidata identifier.", homepage_url = "Journal homepage.",
  `apc_prices[].price` = "Listed article processing charge.",
  `apc_prices[].currency` = "Currency of that charge.",
  apc_usd = "Listed charge in USD.",
  `apc_usd_by_year[].year` = "Year of the listed charge.",
  `apc_usd_by_year[].price` = "That year's listed charge in USD.",
  country_code = "Journal's home country code.",
  societies = "Associated scholarly societies; Nature returns an empty list.",
  alternate_titles = "Alternative journal titles; Nature returns an empty list.",
  type = "Source type.",
  `counts_by_year[].year` = "Publication year of the counted works.",
  `counts_by_year[].works_count` = "Works published in that year.",
  `counts_by_year[].oa_works_count` = "Open-access works published in that year.",
  `counts_by_year[].cited_by_count` = "Cumulative citations to works published in that year.",
  works_api_url = "Link to this journal's works endpoint.",
  updated_date = "Last OpenAlex record update (UTC).",
  created_date = "Date the OpenAlex record was created."
)
for (group in c("topics", "topic_share")) {
  topic_definitions <- c(id = "Topic ID.", display_name = "Topic name.",
    `subfield.id` = "Subfield ID.", `subfield.display_name` = "Subfield name.",
    `field.id` = "Field ID.", `field.display_name` = "Field name.",
    `domain.id` = "Domain ID.", `domain.display_name` = "Domain name.")
  statistic <- if (group == "topics") c(count = "Journal works associated with this topic.") else
    c(value = "Journal's share of this topic across all sources; not its internal topic mix.")
  topic_definitions <- c(topic_definitions, statistic)
  names(topic_definitions) <- paste0(group, "[].", names(topic_definitions))
  oa_definitions <- c(oa_definitions, topic_definitions)
}
stopifnot(all(oa_fields %in% names(oa_definitions)))
cr_meanings <- setNames(vapply(cr_fields, describe_cr, character(1)), cr_fields)
```

```{r}
#| label: pair-related-fields
field_pairs <- tibble::tribble(
  ~cr, ~oa,
  "title", "display_name",
  "publisher", "host_organization_name",
  "ISSN[]", "issn[]",
  "issn-type[].value", "ids.issn[]",
  "counts.total-dois", "works_count",
  "breakdowns.dois-by-issued-year[][1]", "counts_by_year[].year",
  "breakdowns.dois-by-issued-year[][2]", "counts_by_year[].works_count",
  "subjects", "topics[].display_name"
)
```

```{r}
#| label: format-expanded-comparison
#| code-fold: true
#| code-summary: "R code: show real values and check both sources are fully covered"
escape <- function(x) as.character(htmltools::htmlEscape(x))
field_cell <- function(field, meanings) {
  if (!nzchar(field)) return("")
  paste0("<code>", escape(field), "</code><br>", escape(meanings[[field]]))
}
value_cell <- function(field, flat) {
  if (!nzchar(field)) return("")
  values <- flat[flat$path == field, ]
  stopifnot(nrow(values) > 0)
  lines <- ifelse(nzchar(values$context),
                  paste(values$context, values$value, sep = " → "), values$value)
  full <- paste0("<pre class='field-value'>", escape(paste(lines, collapse = "\n")), "</pre>")
  if (nrow(values) <= 3) return(full)
  chosen <- if (any(grepl("^(Year |year: )2025$", values$context))) {
    which(grepl("^(Year |year: )2025$", values$context))[1]
  } else 1L
  paste0("<pre class='field-value'>", escape(lines[chosen]), "</pre>",
         "<details><summary>All ", nrow(values), " values</summary>", full, "</details>")
}
show_comparison <- function(pairs) {
  table <- data.frame(
    `Crossref field and meaning` = vapply(pairs$cr, field_cell, character(1), cr_meanings),
    `Nature in Crossref` = vapply(pairs$cr, value_cell, character(1), cr_flat),
    `OpenAlex field and meaning` = vapply(pairs$oa, field_cell, character(1), oa_definitions),
    `Nature in OpenAlex` = vapply(pairs$oa, value_cell, character(1), oa_flat),
    check.names = FALSE)
  cat("<div class='field-comparison'>")
  print(knitr::kable(table, format = "html", escape = FALSE, row.names = FALSE))
  cat("</div>\n")
}
cr_remaining <- setdiff(cr_fields, field_pairs$cr)
oa_remaining <- setdiff(oa_fields, field_pairs$oa)
all_pairs <- bind_rows(field_pairs,
  tibble::tibble(cr = cr_remaining, oa = ""),
  tibble::tibble(cr = "", oa = oa_remaining))
stopifnot(setequal(all_pairs$cr[all_pairs$cr != ""], cr_fields),
          setequal(all_pairs$oa[all_pairs$oa != ""], oa_fields),
          !anyDuplicated(all_pairs$cr[all_pairs$cr != ""]),
          !anyDuplicated(all_pairs$oa[all_pairs$oa != ""]))
write.csv(cr_flat, file.path(nature_dir, "crossref-expanded-values.csv"), row.names = FALSE)
write.csv(oa_flat, file.path(nature_dir, "openalex-expanded-values.csv"), row.names = FALSE)
write.csv(all_pairs, file.path(nature_dir, "field-pairs.csv"), row.names = FALSE)
comparison_year <- 2025L
cr_year_count <- Filter(function(x) x[[1]] == comparison_year,
                        cr$breakdowns$`dois-by-issued-year`)[[1]][[2]]
oa_year <- Filter(function(x) x$year == comparison_year, oa$counts_by_year)[[1]]
coverage <- cr$`coverage-type`$current
```

### Same or related information

```{r}
#| label: show-related-fields
#| results: asis
show_comparison(field_pairs)
```

**Both sources agree on Nature's title and ISSNs.** The publisher names differ:
Crossref records **`r cr$publisher`**, whereas OpenAlex records
**`r oa$host_organization_name`**. This is a useful example of why a publisher
comparison needs more than exact text matching.

**The publication counts are close, but not identical.** Crossref records
`r fmt(cr$counts[["total-dois"]])` DOIs and OpenAlex records `r fmt(oa$works_count)`
works. For 2025, the counts are `r fmt(cr_year_count)` and
`r fmt(oa_year$works_count)`. Each figure describes the records held by that
database. Crossref's retired `subjects` list is empty; OpenAlex supplies topics
and a hierarchy of subfields, fields and domains.

### Crossref fields without a direct OpenAlex journal counterpart

**What Crossref adds is a view of the information deposited for a journal.**
For Nature's `r current_period` records, `r round(100 * coverage$abstracts, 1)`% contain
abstracts, `r round(100 * coverage$references, 1)`% contain reference lists,
`r round(100 * coverage$orcids, 1)`% contain an ORCID, and
`r round(100 * coverage$affiliations, 1)`% contain contributor affiliations.
These describe what is available through Crossref, rather than everything
printed in the papers. They can help us compare metadata provision across
journals; they do not by themselves measure research quality.

The metadata coverage measures under `coverage` and `coverage-type` are
**fractions from 0 to 1**; `last-status-check-time` is a timestamp.
*Current* means `r current_period` publications, *backfile* means earlier publications,
and *all* combines them. These periods follow the `r snapshot_year` snapshot.
`coverage-type` repeats the period-specific measures and adds the all-years
view. `flags` expresses deposit activity as yes/no values. Repeated measures
are shown because they are separate API fields, but are not separate findings.

```{r}
#| label: show-crossref-fields
#| results: asis
#| code-fold: true
#| code-summary: "R code: display all remaining Crossref fields"
cr_groups <- list(
  "Identifiers, period totals and record timing" = cr_remaining[
    !grepl("^(coverage\\.|coverage-type\\.|flags\\.)", cr_remaining)],
  "Metadata coverage: current and backfile" = sort(cr_remaining[startsWith(cr_remaining, "coverage.")]),
  "Metadata coverage: all years, current and backfile" = sort(cr_remaining[startsWith(cr_remaining, "coverage-type.")]),
  "Deposit flags" = sort(cr_remaining[startsWith(cr_remaining, "flags.")]))
for (group in names(cr_groups)) {
  cat("\n\n#### ", group, "\n\n", sep = "")
  show_comparison(tibble::tibble(cr = cr_groups[[group]], oa = ""))
}
```

Crossref's licence coverage measures whether licence information is deposited.
It does not state that every paper is open access. Its print/electronic ISSN
labels add detail to the shared identifiers. The current/backfile DOI totals
are precomputed groupings; OpenAlex's annual counts could be grouped over the
same years, but have no directly equivalent total fields.

### OpenAlex fields without a direct Crossref journal counterpart

**OpenAlex adds topic profiles, citation measures, open-access indicators,
listed publication charges, and publisher and country identifiers.** These
describe other aspects of a journal, beyond Crossref's metadata-deposit measures.
For Nature, OpenAlex records `r fmt(oa$oa_works_count)` open-access works while
`is_oa` is `r tolower(as.character(oa$is_oa))`: some works being open access does
not make the whole journal fully open access. The recorded USD charge is a
list price, not evidence of what authors paid.

```{r}
#| label: show-openalex-fields
#| results: asis
#| code-fold: true
#| code-summary: "R code: display all remaining OpenAlex fields"
oa_groups <- list(
  "Identity, publisher and journal description" = oa_remaining[
    grepl("^(id$|issn_l$|ids\\.|host_|homepage_url$|country_code$|societies$|alternate_titles$|type$)", oa_remaining)],
  "Publication and citation measures" = oa_remaining[
    grepl("^(oa_works_count$|cited_by_count$|summary_stats\\.|first_publication_year$|last_publication_year$|counts_by_year)", oa_remaining)],
  "Open access, index flags and publication charges" = oa_remaining[
    grepl("^(is_|oa_flip_year$|apc_)", oa_remaining)],
  "Topics and their subject hierarchy" = oa_remaining[startsWith(oa_remaining, "topics[]")],
  "The journal's share of topics across sources" = oa_remaining[startsWith(oa_remaining, "topic_share[]")],
  "API link and record dates" = oa_remaining[oa_remaining %in% c("works_api_url", "updated_date", "created_date")])
stopifnot(setequal(unlist(oa_groups), oa_remaining),
          !anyDuplicated(unlist(oa_groups)),
          setequal(unlist(cr_groups), cr_remaining))
for (group in names(oa_groups)) {
  cat("\n\n#### ", group, "\n\n", sep = "")
  show_comparison(tibble::tibble(cr = "", oa = oa_groups[[group]]))
}
```

`topics.count` counts this journal's works associated with a topic.
`topic_share.value` instead describes this journal's share of that topic
across all sources. The latter is not the percentage of Nature's own papers
on that topic. `is_core` refers to the **CWTS Core sources list**, so it cannot
serve as our Web of Science inclusion label.

The annual citation field groups cumulative citations by the **publication
year of the cited works**. The recent-publication mean in `summary_stats`
also uses OpenAlex's own calculation. These meanings follow the
[current OpenAlex source-building code](https://github.com/ourresearch/openalex-walden/blob/a157a4e5d9ad7809e83c66e66697874e5d94e9f7/notebooks/sources/CreateSourcesApi.ipynb).
Nature's `first_publication_year` is `r oa$first_publication_year` in this
response, but the annual list starts at
`r min(vapply(oa$counts_by_year, function(x) x$year, numeric(1)))`.
These two fields disagree; the earlier value should not be used as the
journal's founding year without checking it.

### Fields with no Nature value to expand

An empty list cannot supply an example of its child fields. The documented
fields below complete the schema explanation without inventing Nature values.
Their OpenAlex cells explicitly state why there is no value; Crossref cells
remain blank.

```{r}
#| label: show-unavailable-schema-fields
#| results: asis
#| code-fold: true
#| code-summary: "R code: distinguish empty lists from fields not returned"
not_returned <- tibble::tribble(
  ~field, ~meaning, ~value,
  "societies[].organization", "Associated scholarly society name.", "No item: societies = []",
  "societies[].url", "Associated society's URL.", "No item: societies = []",
  "alternate_titles[]", "An alternative or former title.", "No item: alternate_titles = []",
  "abbreviated_title", "Standard abbreviated journal title.", "Field not returned in this Nature response",
  "ids.fatcat", "Fatcat catalogue identifier.", "Field not returned in this Nature response")
unavailable_table <- knitr::kable(data.frame(
  `Crossref field and meaning` = "", `Nature in Crossref` = "",
  `OpenAlex field and meaning` = paste0("<code>", escape(not_returned$field),
    "</code><br>", escape(not_returned$meaning)),
  `Nature in OpenAlex` = escape(not_returned$value), check.names = FALSE),
  format = "html", escape = FALSE, row.names = FALSE)
cat(paste0("<div class='field-comparison'>", unavailable_table, "</div>"))
```

**For the meeting:** the clearest Crossref addition is the extent to which
different kinds of metadata are supplied. OpenAlex already supplies a broader
description of journal topics, publication activity and open access. This
Nature example shows the distinction and the field values; whether those
patterns hold across journals remains a separate analysis.

Field explanations use the [Crossref journal schema](https://api.crossref.org/swagger-docs),
[Crossref coverage definitions](https://www.crossref.org/documentation/reports/participation-reports/),
[OpenAlex Source dictionary](https://help.openalex.org/data/sources/attributes/)
and [OpenAlex source construction](https://github.com/ourresearch/openalex-walden/blob/main/notebooks/sources/CreateSources.ipynb).


## 2. Records added by Crossref

**Agreed task:** Find records with Crossref information but no OpenAlex match,
and show their number and examples.

We select records with one Crossref match and no OpenAlex match in the saved
ISSN search. Records with several possible OpenAlex matches are counted separately.

```{r}
#| label: crossref-only-records
crossref_only <- journals |>
  filter(crossref_match_status == "unique",
         openalex_match_status == "unmatched")
n_uncertain <- sum(journals$crossref_match_status == "unique" &
                   journals$openalex_match_status == "ambiguous")

examples <- crossref_only |>
  transmute(Journal = crossref_title,
            ISSN = gsub("|", ", ", crossref_matched_issns, fixed = TRUE),
            `Registered DOIs` = vapply(crossref_counts,
              function(x) as.numeric(fromJSON(x)$`total-dois`), numeric(1)))
knitr::kable(examples, format.args = list(big.mark = ","))
```

Crossref adds a match where OpenAlex found none for **`r nrow(crossref_only)` of the
`r fmt(n_valid)` records with a usable ISSN**
(`r sprintf("%.3f", 100 * nrow(crossref_only) / n_valid)`%). Its additional reach
is therefore small in this PKP list; its main contribution is the extra
information it provides about records already found in both sources.
Another **`r fmt(n_uncertain)` records** have one Crossref match but several
possible OpenAlex matches. They need a separate identity check.
The `r fmt(nrow(journals) - n_valid)` records without a usable ISSN were not searched.

## 3. Empty lists and missing values

**Agreed task:** Check an actual empty-list example and show how it should be treated.

Here we inspect `subjects`, the Crossref subject-label field, for the first
record listed above.

```{r}
#| label: inspect-empty-list
example_issn <- strsplit(crossref_only$crossref_matched_issns[1],
                        "|", fixed = TRUE)[[1]][1]
raw_example <- Filter(function(x) example_issn %in% x$ISSN,
                      crossref$journals)[[1]]
raw_example$subjects
```

Crossref [retired its old subject codes on 23 April 2024](https://www.crossref.org/deprecated/#2024-april-23-subject-codes-removed-from-rest-api),
leaving the fields empty for compatibility. The empty list is already present
in the saved API response, before merging. We represent this unavailable
classification as missing in the analysis:

```{r}
#| label: subject-missingness
raw_nonempty_subjects <- sum(vapply(crossref$journals,
  function(journal) length(journal$subjects) > 0, logical(1)))
matched <- journals |> filter(crossref_match_status == "unique")
matched <- matched |>
  mutate(subjects_clean = na_if(crossref_subjects, "[]"))

knitr::kable(data.frame(
  `Subject information` = c("Supplied", "Missing"),
  Records = c(sum(!is.na(matched$subjects_clean)),
              sum(is.na(matched$subjects_clean))),
  check.names = FALSE), format.args = list(big.mark = ","))
```

The original cache contains `r fmt(length(crossref$journals))` journal records,
of which **`r fmt(raw_nonempty_subjects)` have nonempty subject labels**.
All **`r fmt(nrow(matched))` records with one Crossref match** likewise have empty
subject labels in the merged data. This is consistent with the retired Crossref
classification, rather than evidence that the journals have no subject or that
the merge removed their labels. These legacy codes came from a Scopus-based
experiment; they were not subject labels deposited by publishers.
[Crossref's explanation](https://www.crossref.org/blog/subject-codes-incomplete-and-unreliable-have-got-to-go/)
also makes clear that the empty fields are intentional. OpenAlex's `topics` is
a separate field and is populated in the example above.

## 4. BigQuery access

**Agreed task:** Test whether a query now runs and identify the permission needed.

```{r}
#| label: bigquery-test
query_test <- system2(params$bq_cli, c(
  "--headless=true", "--project_id=insyspo", "--location=US",
  "--dataset_id=insyspo:userdb_yann_jyal", "query",
  "--use_legacy_sql=false", "--maximum_bytes_billed=1",
  shQuote("SELECT 1 AS access_test")), stdout = TRUE, stderr = TRUE)
```

```{r}
#| label: record-query-result
#| include: false
query_status <- attr(query_test, "status")
if (is.null(query_status)) query_status <- 0L
missing_jobs_permission <- any(grepl("bigquery.jobs.create", query_test, fixed = TRUE))
query_sentence <- if (query_status == 0L) {
  "The constant query succeeded in insyspo; the query-job access test now passes."
} else if (missing_jobs_permission) {
  "The SELECT 1 test still fails in insyspo; the account needs BigQuery Job User (roles/bigquery.jobUser), which includes bigquery.jobs.create."
} else {
  "The query test did not complete; its error needs checking before attributing it to a permission."
}
writeLines(query_test, file.path(output_dir, "bigquery-test.txt"))
stopifnot(query_status == 0L || missing_jobs_permission)
write.csv(examples, file.path(output_dir, "crossref-only-examples.csv"), row.names = FALSE)
write_json(list(
  rendered_at_utc = format(Sys.time(), "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
  input_file = params$data_file,
  input_sha256 = digest::digest(file = params$data_file, algo = "sha256"),
  crossref_snapshot = crossref$retrieved_at,
  pkp_rows = nrow(journals), valid_issn_rows = n_valid,
  crossref_fields = cr_fields,
  nature_retrieved_at_utc = nature_retrieval$retrieved_at_utc,
  nature_id = oa$id, nature_issns = unlist(oa$issn),
  nature_crossref_top_level_fields = length(cr),
  nature_openalex_top_level_fields = length(oa),
  nature_crossref_expanded_fields = length(cr_fields),
  nature_openalex_expanded_fields = length(oa_fields),
  nature_crossref_leaf_values = nrow(cr_flat),
  nature_openalex_leaf_values = nrow(oa_flat),
  nature_related_pairs = nrow(field_pairs),
  nature_crossref_only_fields = length(cr_remaining),
  nature_openalex_only_fields = length(oa_remaining),
  nature_sha256 = lapply(files, function(file) digest::digest(file = file, algo = "sha256")),
  crossref_only = nrow(crossref_only), crossref_with_uncertain_openalex = n_uncertain,
  subject_missing = sum(is.na(matched$subjects_clean)),
  bigquery_exit_status = query_status, missing_jobs_permission = missing_jobs_permission,
  R = R.version.string), file.path(output_dir, "run-summary.json"),
  auto_unbox = TRUE, pretty = TRUE)
```

`r query_sentence`

---

Tasks follow the [meeting discussion](https://github.com/YannJY02/InvisibleResearch/blob/c06f31f0d517dddafd779e5cc38a6bf589d3c039/meeting-reports/2026-09-14-openalex-baseline-and-access.md)
(02:13–08:29 and 34:23–34:31), alongside the
[field-review request](https://github.com/invisibleinfo/invisible-research/issues/5#issuecomment-5553748839).
