---
title: "What does Crossref add to our journal data?"
subtitle: "Meeting update · 15 September 2026"
lang: en
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    code-fold: show
    code-summary: "R code"
    code-copy: true
    embed-resources: true
    css: crossref-meeting-report.css
execute:
  echo: true
  message: false
  warning: false
  error: false
  cache: false
params:
  data_file: "../artifacts/ojs_journal_enrichment_simple_ver/pkp-ojs-multisource-enriched-simple_ver.parquet"
  crossref_cache: "../artifacts/ojs_journal_enrichment/full-v7/crossref-journals.rds"
  bq_cli: "/Users/yann.jy/.local/google-cloud-sdk/bin/bq"
---

```{r}
#| label: read-saved-data
library(arrow)
library(dplyr)
library(jsonlite)
fmt <- function(x) format(x, big.mark = ",", scientific = FALSE, trim = TRUE)

journals <- read_parquet(params$data_file)
crossref <- readRDS(params$crossref_cache)
```

```{r}
#| label: display-settings
#| include: false
n_valid <- sum(journals$identifier_status == "valid")
snapshot_date <- substr(crossref$retrieved_at, 1, 10)
output_dir <- "../artifacts/crossref-meeting-report"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
stopifnot(nrow(journals) == 98273, !"admin_email" %in% names(journals))
```

We use the existing **PKP V7 list enriched with OpenAlex and Crossref**:
`r fmt(nrow(journals))` records, including `r fmt(n_valid)` with a usable ISSN.
The saved Crossref data were collected on `r snapshot_date`. The results below
are calculated from that merged file when this report runs.

## 1. Crossref fields

**Agreed task:** Check which Crossref fields add useful information beyond OpenAlex.

We check the fields in the saved data and group them by what they describe.
The short descriptions are supplied in `crossref-field-notes.csv`.

```{r}
#| label: compare-fields
field_notes <- read.csv("crossref-field-notes.csv", check.names = FALSE)
cr_fields <- unlist(strsplit(field_notes$crossref_fields, ";", fixed = TRUE))
oa_fields <- unlist(strsplit(field_notes$openalex_fields, ";", fixed = TRUE))
stopifnot(setequal(cr_fields, names(crossref$journals[[1]])),
          all(paste0("crossref_", cr_fields) %in% names(journals)),
          all(paste0("openalex_", oa_fields[nzchar(oa_fields)]) %in%
                names(journals)))

field_notes$crossref_fields <- gsub(";", ", ", field_notes$crossref_fields)
field_notes$openalex_fields <- gsub(";", ", ", field_notes$openalex_fields)
field_notes$openalex_fields[field_notes$openalex_fields == ""] <-
  "No equivalent in this table"
knitr::kable(field_notes, col.names = c(
  "Crossref fields", "Related OpenAlex fields", "What they tell us"))
```

**The main additions are DOI registration counts and information about what
publishers deposit with Crossref.** Titles and identifiers largely serve the
same purpose in both sources. Crossref also distinguishes print and online ISSNs.

For a concrete example, take the first record with one match in each source:

```{r}
#| label: field-example
example <- journals |>
  filter(openalex_match_status == "unique",
         crossref_match_status == "unique") |>
  slice(1)
doi_counts <- fromJSON(example$crossref_counts)
coverage <- fromJSON(example[["crossref_coverage-type"]])$all

example_values <- data.frame(
  Measure = c("Publications recorded by OpenAlex",
              "DOIs registered with Crossref",
              "Crossref records with an abstract",
              "Crossref records with references"),
  Value = c(fmt(example$openalex_works_count), fmt(doi_counts$`total-dois`),
            paste0(round(100 * coverage$abstracts), "%"),
            paste0(round(100 * coverage$references), "%")))
knitr::kable(example_values)
```

The example is **`r example$crossref_title`**. OpenAlex records
`r fmt(example$openalex_works_count)` publications; Crossref has
`r fmt(doi_counts[["total-dois"]])` DOI registrations. The counts describe each
database's records, so a difference does not by itself identify an error.
Crossref additionally reports abstracts for `r round(100 * coverage$abstracts)`%
of this journal's Crossref records, across all years, and references for
`r round(100 * coverage$references)`%.
The reference figure describes what was deposited with Crossref, not whether
the publications themselves contain references.

## 2. Records added by Crossref

**Agreed task:** Find records with Crossref information but no OpenAlex match,
and show their number and examples.

We select records with one Crossref match and no OpenAlex match in the saved
ISSN search. Records with several possible OpenAlex matches are counted separately.

```{r}
#| label: crossref-only-records
crossref_only <- journals |>
  filter(crossref_match_status == "unique",
         openalex_match_status == "unmatched")
n_uncertain <- sum(journals$crossref_match_status == "unique" &
                   journals$openalex_match_status == "ambiguous")

examples <- crossref_only |>
  transmute(Journal = crossref_title,
            ISSN = gsub("|", ", ", crossref_matched_issns, fixed = TRUE),
            `Registered DOIs` = vapply(crossref_counts,
              function(x) as.numeric(fromJSON(x)$`total-dois`), numeric(1)))
knitr::kable(examples, format.args = list(big.mark = ","))
```

Crossref adds a match where OpenAlex found none for **`r nrow(crossref_only)` of the
`r fmt(n_valid)` records with a usable ISSN**
(`r sprintf("%.3f", 100 * nrow(crossref_only) / n_valid)`%). Its additional reach
is therefore small in this PKP list; its main contribution is the extra
information it provides about records already found in both sources.
Another **`r fmt(n_uncertain)` records** have one Crossref match but several
possible OpenAlex matches. They need a separate identity check.
The `r fmt(nrow(journals) - n_valid)` records without a usable ISSN were not searched.

## 3. Empty lists and missing values

**Agreed task:** Check an actual empty-list example and show how it should be treated.

Here we inspect `subjects`, the Crossref subject-label field, for the first
record listed above.

```{r}
#| label: inspect-empty-list
example_issn <- strsplit(crossref_only$crossref_matched_issns[1],
                        "|", fixed = TRUE)[[1]][1]
raw_example <- Filter(function(x) example_issn %in% x$ISSN,
                      crossref$journals)[[1]]
raw_example$subjects
```

The empty list means that **no subject labels were supplied**. We treat it as
missing subject information in the analysis:

```{r}
#| label: subject-missingness
matched <- journals |> filter(crossref_match_status == "unique")
matched <- matched |>
  mutate(subjects_clean = na_if(crossref_subjects, "[]"))

knitr::kable(data.frame(
  `Subject information` = c("Supplied", "Missing"),
  Records = c(sum(!is.na(matched$subjects_clean)),
              sum(is.na(matched$subjects_clean))),
  check.names = FALSE), format.args = list(big.mark = ","))
```

All **`r fmt(nrow(matched))` records with one Crossref match** have no subject labels
in this saved data. Crossref's `subjects` field therefore does not add subject
classification here; OpenAlex's topic information is the useful starting point.

## 4. BigQuery access

**Agreed task:** Test whether a query now runs and identify the permission needed.

```{r}
#| label: bigquery-test
query_test <- system2(params$bq_cli, c(
  "--headless=true", "--project_id=insyspo", "--location=US",
  "--dataset_id=insyspo:userdb_yann_jyal", "query",
  "--use_legacy_sql=false", "--maximum_bytes_billed=1",
  shQuote("SELECT 1 AS access_test")), stdout = TRUE, stderr = TRUE)
```

```{r}
#| label: record-query-result
#| include: false
query_status <- attr(query_test, "status")
if (is.null(query_status)) query_status <- 0L
missing_jobs_permission <- any(grepl("bigquery.jobs.create", query_test, fixed = TRUE))
query_sentence <- if (query_status == 0L) {
  "The constant query succeeded in insyspo; the query-job access test now passes."
} else if (missing_jobs_permission) {
  "The SELECT 1 test still fails in insyspo; the account needs BigQuery Job User (roles/bigquery.jobUser), which includes bigquery.jobs.create."
} else {
  "The query test did not complete; its error needs checking before attributing it to a permission."
}
writeLines(query_test, file.path(output_dir, "bigquery-test.txt"))
stopifnot(query_status == 0L || missing_jobs_permission)
write.csv(examples, file.path(output_dir, "crossref-only-examples.csv"), row.names = FALSE)
write_json(list(
  rendered_at_utc = format(Sys.time(), "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
  input_file = params$data_file,
  input_sha256 = digest::digest(file = params$data_file, algo = "sha256"),
  crossref_snapshot = crossref$retrieved_at,
  pkp_rows = nrow(journals), valid_issn_rows = n_valid,
  crossref_fields = cr_fields,
  crossref_only = nrow(crossref_only), crossref_with_uncertain_openalex = n_uncertain,
  subject_missing = sum(is.na(matched$subjects_clean)),
  bigquery_exit_status = query_status, missing_jobs_permission = missing_jobs_permission,
  R = R.version.string), file.path(output_dir, "run-summary.json"),
  auto_unbox = TRUE, pretty = TRUE)
```

`r query_sentence`

---

Tasks follow the [meeting discussion](https://github.com/YannJY02/InvisibleResearch/blob/c06f31f0d517dddafd779e5cc38a6bf589d3c039/meeting-reports/2026-09-14-openalex-baseline-and-access.md)
(02:13–08:29 and 34:23–34:31), alongside the
[field-review request](https://github.com/invisibleinfo/invisible-research/issues/5#issuecomment-5553748839).
