---
title: "PKP/OJS multi-source journal enrichment"
format: html
execute:
  echo: true
  warning: false
  message: false
---

This notebook left-joins OpenAlex and Crossref journal metadata onto PKP OJS
rows by exact ISSN. By default it runs a **purposive validation sample** of ten
source-authentic PKP rows selected to exercise missing, unmatched, consistent,
and inconsistent match branches. It is not a representative sample and must
not be used to estimate full-cohort coverage.

Set `OPENALEX_API_KEY` in the environment or the repository-root `.env` file
for either mode. An existing environment value takes precedence. To run the
same functions on the complete pinned Beacon file, also set
`OJS_ENRICHMENT_MODE=full`, `DATA_ROOT`, and `CROSSREF_MAILTO`. This notebook
validates sample mode only; it does not execute the full run here.

## Packages and paths

This setup chunk loads the required R packages, resolves sample or full mode,
and defines input and output paths. It uses an already exported
`OPENALEX_API_KEY` when available and otherwise reads the repository-root
`.env`; the selected mode is validated before any data or API request is made.
Artifacts are separated by run mode while source-response caches are shared.

```{r}
#| label: packages-and-paths

library(dplyr)
library(httr2)
library(jsonlite)
library(purrr)
library(readr)
library(tibble)

run_mode <- tolower(Sys.getenv("OJS_ENRICHMENT_MODE", "sample"))
stopifnot(run_mode %in% c("sample", "full"))
project_env_path <- file.path("..", "..", "..", ".env")
if (!nzchar(Sys.getenv("OPENALEX_API_KEY")) && file.exists(project_env_path)) {
  readRenviron(project_env_path)
}
if (!nzchar(Sys.getenv("OPENALEX_API_KEY"))) {
  stop("Set OPENALEX_API_KEY before running this notebook.", call. = FALSE)
}

fixture_path <- file.path(
  "..",
  "data",
  "ojs-journal-enrichment-sample.csv"
)
data_root <- Sys.getenv("DATA_ROOT")
if (run_mode == "full" && !nzchar(data_root)) {
  stop("Set DATA_ROOT before running full mode.", call. = FALSE)
}
input_path <- if (run_mode == "sample") {
  fixture_path
} else {
  file.path(data_root, "raw", "pkp-beacon-v6", "beacon.csv")
}

artifact_dir <- file.path("..", "artifacts", "ojs_journal_enrichment")
run_dir <- file.path(artifact_dir, run_mode)
dir.create(run_dir, recursive = TRUE, showWarnings = FALSE)

output_path <- file.path(run_dir, "pkp-ojs-multisource-enriched.csv")
metadata_path <- file.path(run_dir, "run-metadata.json")
```

## ISSN preparation

Exact matching depends on checksum-valid ISSNs rather than raw identifier
strings. This chunk removes display punctuation, validates the ISSN check
digit, parses rows that contain multiple identifiers, and applies the same
normalization to identifiers returned by external sources. Missing and invalid
identifiers therefore remain distinguishable from valid but unmatched ISSNs.

```{r}
#| label: issn-preparation

normalize_issn <- function(value) {
  compact <- gsub("[^0-9X]", "", toupper(trimws(as.character(value))))
  if (!grepl("^[0-9]{7}[0-9X]$", compact)) return(NA_character_)

  digits <- as.integer(strsplit(substr(compact, 1, 7), "")[[1]])
  check_value <- (11 - sum(digits * 8:2) %% 11) %% 11
  expected <- if (check_value == 10) "X" else as.character(check_value)
  if (substr(compact, 8, 8) != expected) return(NA_character_)

  paste0(substr(compact, 1, 4), "-", substr(compact, 5, 8))
}

parse_issns <- function(value) {
  if (is.na(value) || !nzchar(value)) return(character())

  tokens <- unlist(strsplit(value, "[[:space:],;|]+"))
  normalized <- vapply(tokens[nzchar(tokens)], normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}

source_issns <- function(source) {
  values <- c(unlist(source$issn, use.names = FALSE), source$issn_l)
  normalized <- vapply(values, normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}
```

## Source retrieval

OpenAlex accepts up to 100 OR values in one filter, whereas Crossref journal
records are requested one ISSN at a time. Both retrieval functions use bounded
timeouts and retries. OpenAlex pagination follows the returned cursor, and a
Crossref `404` is treated as an unmatched journal rather than an API failure.
The OpenAlex request deliberately omits `select`, retaining every top-level
Source field before later filtering.

```{r}
#| label: source-retrieval

fetch_openalex_batch <- function(
  issns,
  api_key = Sys.getenv("OPENALEX_API_KEY")
) {
  stopifnot(length(issns) >= 1L, length(issns) <= 100L, nzchar(api_key))

  results <- list()
  cursor <- "*"
  repeat {
    payload <- request("https://api.openalex.org/sources") |>
      req_url_query(
        filter = paste0("issn:", paste(issns, collapse = "|")),
        per_page = 100,
        cursor = cursor,
        api_key = api_key
      ) |>
      req_timeout(seconds = 30) |>
      req_retry(max_tries = 3) |>
      req_perform() |>
      resp_body_json(simplifyVector = FALSE)

    results <- c(results, payload$results)
    next_cursor <- payload$meta$next_cursor
    if (!length(payload$results) || is.null(next_cursor) ||
        identical(next_cursor, cursor)) break
    cursor <- next_cursor
  }
  results
}

fetch_crossref_journal <- function(
  issn,
  mailto = Sys.getenv("CROSSREF_MAILTO")
) {
  agent <- if (nzchar(mailto)) {
    paste0("InvisibleResearch/1.0 (mailto:", mailto, ")")
  } else {
    "InvisibleResearch/1.0"
  }
  crossref_request <- request(
    paste0("https://api.crossref.org/journals/", issn)
  ) |>
    req_headers(Accept = "application/json", `User-Agent` = agent) |>
    req_timeout(seconds = 30) |>
    req_error(
      is_error = \(response) {
        resp_status(response) != 404L && resp_is_error(response)
      }
    ) |>
    req_retry(max_tries = 3)
  if (nzchar(mailto)) {
    crossref_request <- req_url_query(crossref_request, mailto = mailto)
  }

  response <- req_perform(crossref_request)
  if (resp_status(response) == 404L) return(NULL)
  resp_body_json(response, simplifyVector = FALSE)$message
}
```

## Multi-source enrichment

The ignored cache files are owned by this QMD and store complete source
responses by ISSN. They are intentionally shared between sample and full modes
so a full run can reuse the sample lookups; each run records their paths,
checksums, and hit/miss counts. OpenAlex checkpoints after each batch; Crossref
checkpoints every 100 new ISSNs. API errors remain run-specific so the next run
retries them, while their per-ISSN messages are written to run metadata.

For each PKP row, the function normalizes all supplied ISSNs, retrieves missing
source records, and deduplicates candidates by source-specific identity. It
classifies each source independently as `not_attempted`, `api_error`,
`unmatched`, `consistent`, or `inconsistent`; only a single consistent
candidate supplies flattened source fields. Complete candidate objects and
match provenance remain attached before list columns are serialized for CSV.

```{r}
#| label: enrichment-function

enrich_ojs_rows <- function(ojs_rows, run_mode, artifact_dir, output_path) {
  ojs_rows <- ojs_rows |>
    mutate(
      .input_issns = map(issn, parse_issns),
      identifier_status = case_when(
        !nzchar(issn) ~ "missing",
        lengths(.input_issns) == 0L ~ "invalid",
        TRUE ~ "valid"
      )
    )
  tokens <- sort(unique(unlist(ojs_rows$.input_issns)))

  openalex_cache_path <- file.path(
    artifact_dir,
    "openalex-source-cache.rds"
  )
  crossref_cache_path <- file.path(
    artifact_dir,
    "crossref-journal-cache.rds"
  )

  openalex_cache <- if (file.exists(openalex_cache_path)) {
    readRDS(openalex_cache_path)
  } else {
    list()
  }
  crossref_cache <- if (file.exists(crossref_cache_path)) {
    readRDS(crossref_cache_path)
  } else {
    list()
  }
  openalex_entries_before <- length(openalex_cache)
  crossref_entries_before <- length(crossref_cache)

  openalex_errors <- character()
  openalex_cached_tokens <- intersect(tokens, names(openalex_cache))
  missing_openalex <- setdiff(tokens, names(openalex_cache))
  batches <- split(
    missing_openalex,
    ceiling(seq_along(missing_openalex) / 100)
  )
  for (batch in batches) {
    attempt <- tryCatch(
      list(sources = fetch_openalex_batch(batch)),
      error = \(error) list(error = conditionMessage(error))
    )
    if (!is.null(attempt$error)) {
      openalex_errors <- c(
        openalex_errors,
        setNames(rep(attempt$error, length(batch)), batch)
      )
      next
    }

    for (token in batch) {
      candidates <- keep(
        attempt$sources,
        \(source) token %in% source_issns(source)
      )
      openalex_cache[[token]] <- list(
        status = if (length(candidates)) "matched" else "unmatched",
        candidates = candidates
      )
    }
    saveRDS(openalex_cache, openalex_cache_path)
  }

  crossref_mailto <- Sys.getenv("CROSSREF_MAILTO")
  if (run_mode == "full" && !nzchar(crossref_mailto)) {
    stop("Set CROSSREF_MAILTO before running full mode.", call. = FALSE)
  }
  crossref_rate <- if (nzchar(crossref_mailto)) 9 else 4
  crossref_errors <- character()
  crossref_cached_tokens <- intersect(tokens, names(crossref_cache))
  missing_crossref <- setdiff(tokens, names(crossref_cache))
  for (index in seq_along(missing_crossref)) {
    token <- missing_crossref[[index]]
    attempt <- tryCatch(
      list(journal = fetch_crossref_journal(token, crossref_mailto)),
      error = \(error) list(error = conditionMessage(error))
    )
    if (!is.null(attempt$error)) {
      crossref_errors[[token]] <- attempt$error
    } else {
      crossref_cache[[token]] <- if (is.null(attempt$journal)) {
        list(status = "unmatched")
      } else {
        list(status = "matched", journal = attempt$journal)
      }
    }
    if (index %% 100L == 0L || index == length(missing_crossref)) {
      saveRDS(crossref_cache, crossref_cache_path)
    }
    Sys.sleep(1 / crossref_rate)
  }

  classify_openalex <- function(row_tokens) {
    matched_tokens <- keep(row_tokens, \(token) {
      identical(openalex_cache[[token]]$status, "matched")
    })
    candidates <- unlist(
      map(openalex_cache[matched_tokens], "candidates"),
      recursive = FALSE
    )
    candidate_ids <- if (length(candidates)) {
      map_chr(candidates, "id")
    } else {
      character()
    }
    candidates <- candidates[!duplicated(candidate_ids)]
    candidate_ids <- unique(candidate_ids)

    status <- case_when(
      length(row_tokens) == 0L ~ "not_attempted",
      any(row_tokens %in% names(openalex_errors)) ~ "api_error",
      length(candidate_ids) == 0L ~ "unmatched",
      length(candidate_ids) == 1L ~ "consistent",
      TRUE ~ "inconsistent"
    )
    list(
      status = status,
      matched_issns = matched_tokens,
      candidate_ids = candidate_ids,
      candidates = candidates,
      source = if (status == "consistent") candidates[[1]] else NULL
    )
  }

  classify_crossref <- function(row_tokens) {
    matched_tokens <- keep(row_tokens, \(token) {
      identical(crossref_cache[[token]]$status, "matched")
    })
    journals <- map(matched_tokens, \(token) {
      crossref_cache[[token]]$journal
    })
    candidate_keys <- map_chr(journals, \(journal) {
      issns <- vapply(journal$ISSN, normalize_issn, character(1))
      paste(sort(unique(issns[!is.na(issns)])), collapse = "|")
    })
    valid_candidates <- nzchar(candidate_keys)
    journals <- journals[valid_candidates]
    candidate_keys <- candidate_keys[valid_candidates]
    journals <- journals[!duplicated(candidate_keys)]
    candidate_keys <- unique(candidate_keys)

    status <- case_when(
      length(row_tokens) == 0L ~ "not_attempted",
      any(row_tokens %in% names(crossref_errors)) ~ "api_error",
      length(candidate_keys) == 0L ~ "unmatched",
      length(candidate_keys) == 1L ~ "consistent",
      TRUE ~ "inconsistent"
    )
    list(
      status = status,
      matched_issns = matched_tokens,
      candidate_ids = candidate_keys,
      candidates = journals,
      source = if (status == "consistent") journals[[1]] else NULL
    )
  }

  openalex_matches <- map(ojs_rows$.input_issns, classify_openalex)
  crossref_matches <- map(ojs_rows$.input_issns, classify_crossref)
  openalex_fields <- sort(unique(unlist(map(
    openalex_matches,
    \(match) unlist(map(match$candidates, names), use.names = FALSE)
  ))))
  crossref_fields <- sort(unique(unlist(map(
    crossref_matches,
    \(match) unlist(map(match$candidates, names), use.names = FALSE)
  ))))

  enriched <- map_dfr(seq_len(nrow(ojs_rows)), \(row_number) {
    row <- select(ojs_rows[row_number, ], -.input_issns)
    row_tokens <- ojs_rows$.input_issns[[row_number]]
    openalex <- openalex_matches[[row_number]]
    crossref <- crossref_matches[[row_number]]
    provenance <- tibble(
      openalex_input_issns = paste(row_tokens, collapse = "|"),
      openalex_matched_issns = paste(openalex$matched_issns, collapse = "|"),
      openalex_match_status = openalex$status,
      openalex_candidate_ids = paste(openalex$candidate_ids, collapse = "|"),
      openalex_candidate_records = list(openalex$candidates),
      crossref_input_issns = paste(row_tokens, collapse = "|"),
      crossref_matched_issns = paste(crossref$matched_issns, collapse = "|"),
      crossref_match_status = crossref$status,
      crossref_candidate_issn_sets = paste(
        crossref$candidate_ids,
        collapse = ";"
      ),
      crossref_candidate_records = list(crossref$candidates)
    )
    source_cells <- c(
      set_names(
        map(openalex_fields, \(field) list(openalex$source[[field]])),
        paste0("openalex_", openalex_fields)
      ),
      set_names(
        map(crossref_fields, \(field) list(crossref$source[[field]])),
        paste0("crossref_", crossref_fields)
      )
    )
    bind_cols(row, provenance, as_tibble(source_cells))
  })

  as_csv_cell <- function(value) {
    if (is.null(value) || !length(value)) return("")
    if (is.atomic(value) && length(value) == 1L) return(as.character(value))
    as.character(toJSON(
      value,
      auto_unbox = TRUE,
      null = "null",
      na = "null",
      digits = NA
    ))
  }
  csv_output <- enriched |>
    mutate(across(where(is.list), \(column) map_chr(column, as_csv_cell)))
  write_csv(csv_output, output_path, na = "")

  list(
    data = enriched,
    csv = csv_output,
    openalex_fields = openalex_fields,
    crossref_fields = crossref_fields,
    openalex_errors = openalex_errors,
    crossref_errors = crossref_errors,
    cache = list(
      openalex = list(
        file = normalizePath(openalex_cache_path, mustWork = FALSE),
        md5 = if (file.exists(openalex_cache_path)) {
          unname(tools::md5sum(openalex_cache_path))
        } else {
          NA_character_
        },
        entries_before = openalex_entries_before,
        entries_after = length(openalex_cache),
        requested_issns = length(tokens),
        hits = length(openalex_cached_tokens),
        misses = length(missing_openalex)
      ),
      crossref = list(
        file = normalizePath(crossref_cache_path, mustWork = FALSE),
        md5 = if (file.exists(crossref_cache_path)) {
          unname(tools::md5sum(crossref_cache_path))
        } else {
          NA_character_
        },
        entries_before = crossref_entries_before,
        entries_after = length(crossref_cache),
        requested_issns = length(tokens),
        hits = length(crossref_cached_tokens),
        misses = length(missing_crossref)
      )
    )
  )
}
```

## Load PKP/OJS input and run

The sample fixture was extracted from PKP Beacon V6 file `13173372` on
2026-07-17. The source file MD5 is
`9f43fa051c7ed1cc45d8592593542011`; the fixture MD5 is
`b5fb99c10f38c2a5c3e7e6e0d4aa2212`. The fixture retains the original PKP
columns except `admin_email`.

This chunk verifies the selected input checksum and expected row counts before
enrichment, retains only OJS rows, and checks the composite PKP row identity for
duplicates. It then runs the shared enrichment function and writes run metadata
that records inputs, outputs, source names, cache provenance, and API errors.

```{r}
#| label: run-enrichment

stopifnot(file.exists(input_path))
input_md5 <- unname(tools::md5sum(input_path))
if (run_mode == "sample") {
  stopifnot(input_md5 == "b5fb99c10f38c2a5c3e7e6e0d4aa2212")
} else {
  stopifnot(input_md5 == "9f43fa051c7ed1cc45d8592593542011")
}

raw_input <- read_csv(
  input_path,
  col_types = cols(.default = col_character()),
  na = character(),
  show_col_types = FALSE
)
ojs_rows <- raw_input |>
  filter(tolower(application) == "ojs") |>
  select(-any_of("admin_email"))
row_identity <- paste(
  ojs_rows$oai_url,
  ojs_rows$repository_name,
  ojs_rows$set_spec,
  sep = "\r"
)
stopifnot(!anyDuplicated(row_identity))
if (run_mode == "sample") {
  stopifnot(nrow(ojs_rows) == 10L)
} else {
  stopifnot(nrow(raw_input) == 87170L, nrow(ojs_rows) == 86282L)
}

result <- enrich_ojs_rows(ojs_rows, run_mode, artifact_dir, output_path)

write_json(
  list(
    producer = "analysis/ojs_journal_enrichment.qmd",
    run_mode = run_mode,
    input_file = normalizePath(input_path),
    input_md5 = input_md5,
    input_rows = nrow(raw_input),
    ojs_rows = nrow(ojs_rows),
    distinct_valid_issns = length(unique(unlist(map(ojs_rows$issn, parse_issns)))),
    sources = c("PKP OJS", "OpenAlex", "Crossref"),
    output_file = normalizePath(output_path),
    cache = result$cache,
    openalex_api_errors = length(result$openalex_errors),
    openalex_api_error_messages = as.list(result$openalex_errors),
    crossref_api_errors = length(result$crossref_errors),
    crossref_api_error_messages = as.list(result$crossref_errors),
    created_at = format(Sys.time(), tz = "UTC", usetz = TRUE)
  ),
  metadata_path,
  auto_unbox = TRUE,
  pretty = TRUE
)
```

## Checks

These assertions form the executable notebook contract. They verify that the
retrieval and enrichment functions exist, one output row is retained per OJS
row, required artifacts are written, and complete namespaced source fields and
candidate records remain available. In sample mode, they also require the
purposive fixture to exercise the planned match branches; they do not establish
representative full-cohort coverage.

```{r}
#| label: contract-check

stopifnot(
  exists("fetch_openalex_batch", mode = "function"),
  exists("fetch_crossref_journal", mode = "function"),
  exists("enrich_ojs_rows", mode = "function"),
  nrow(result$data) == nrow(ojs_rows),
  file.exists(output_path),
  file.exists(metadata_path),
  "openalex_candidate_records" %in% names(result$data),
  "crossref_candidate_records" %in% names(result$data),
  all(paste0("openalex_", result$openalex_fields) %in% names(result$data)),
  all(paste0("crossref_", result$crossref_fields) %in% names(result$data))
)
if (run_mode == "sample") {
  stopifnot(
    "consistent" %in% result$data$openalex_match_status,
    "unmatched" %in% result$data$openalex_match_status,
    "inconsistent" %in% result$data$openalex_match_status,
    "not_attempted" %in% result$data$openalex_match_status,
    "not_attempted" %in% result$data$crossref_match_status
  )
}
```

## Sample review after all fields are retained

The table below is a review surface only. It uses scalar identity, match, title, country, count, and publisher fields so that rows remain readable; JSON candidate records and nested source fields remain available in the CSV and are intentionally excluded from this preview.

```{r}
#| label: selected-review-fields

result$csv |>
  select(any_of(c(
    "context_name",
    "openalex_input_issns",
    "identifier_status",
    "openalex_match_status",
    "openalex_matched_issns",
    "openalex_display_name",
    "openalex_country_code",
    "openalex_works_count",
    "crossref_match_status",
    "crossref_matched_issns",
    "crossref_title",
    "crossref_publisher"
  ))) |>
  knitr::kable()
```

## Complete merged column list

This final audit table enumerates the exact merged schema produced by the run.
It supports inspection of retained source fields and provenance columns but
does not select variables for substantive analysis.

```{r}
#| label: complete-column-list

tibble(column = names(result$csv)) |>
  knitr::kable()
```
