---
title: "PKP/OJS multi-source journal enrichment"
format: html
execute:
  echo: true
  warning: false
  message: false
---

This notebook adds the complete available OpenAlex and Crossref journal responses to ten real PKP OJS rows from the pinned V7 release. It matches records by exact ISSN. The ten rows were chosen to include missing ISSNs, no matches, one match, partial identifier coverage, and multiple matches. This sample checks whether the code works; it does not measure coverage across the full dataset.

Set `OPENALEX_API_KEY` in the environment or the repository-root `.env` file before rendering. Sample mode is the only mode implemented here; full-cohort retrieval remains a separate, explicit step.

## Packages and paths

This chunk installs only missing CRAN packages, loads them, downloads the pinned PKP V7 input when needed, verifies its official checksum, and keeps every local artifact under this analysis owner.

```{r}
#| label: packages-and-paths

required_packages <- c(
  "dplyr",
  "httr2",
  "jsonlite",
  "knitr",
  "purrr",
  "readr",
  "rmarkdown",
  "tibble"
)
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]
if (length(missing_packages)) {
  install.packages(missing_packages, repos = "https://cloud.r-project.org")
}

library(dplyr)
library(httr2)
library(purrr)
library(readr)
library(tibble)

project_env_path <- file.path("..", "..", "..", ".env")
if (!nzchar(Sys.getenv("OPENALEX_API_KEY")) && file.exists(project_env_path)) {
  readRenviron(project_env_path)
}
if (!nzchar(Sys.getenv("OPENALEX_API_KEY"))) {
  stop("Set OPENALEX_API_KEY before running this notebook.", call. = FALSE)
}

pkp_version <- 7L
pkp_url <- paste0(
  "https://dataverse.harvard.edu/api/access/datafile/14084919",
  "?format=original"
)
pkp_md5 <- "3a4ad8ae1ebfcc2b991aaf55b2d82c92"
artifact_dir <- file.path("..", "artifacts", "ojs_journal_enrichment")
input_dir <- file.path(artifact_dir, "pkp-v7")
input_path <- file.path(input_dir, "beacon.csv")
run_dir <- file.path(artifact_dir, "sample")
output_path <- file.path(run_dir, "pkp-ojs-multisource-enriched.csv")
metadata_path <- file.path(run_dir, "run-metadata.json")
openalex_cache_path <- file.path(run_dir, "openalex-sources.rds")
crossref_cache_path <- file.path(run_dir, "crossref-journals.rds")

dir.create(input_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(run_dir, recursive = TRUE, showWarnings = FALSE)

download_pkp_input <- function(path) {
  if (file.exists(path)) return(invisible(path))

  download_path <- paste0(path, ".download")
  on.exit(unlink(download_path), add = TRUE)
  download.file(pkp_url, download_path, mode = "wb", method = "libcurl")
  if (unname(tools::md5sum(download_path)) != pkp_md5) {
    stop("Downloaded PKP V7 file failed checksum verification.", call. = FALSE)
  }
  if (!file.rename(download_path, path)) {
    stop("Could not move the verified PKP V7 file into place.", call. = FALSE)
  }
  invisible(path)
}

download_pkp_input(input_path)
if (unname(tools::md5sum(input_path)) != pkp_md5) {
  stop("Local PKP V7 file failed checksum verification.", call. = FALSE)
}
```

## ISSN preparation

ISSNs may contain spaces, punctuation, or more than one value. This chunk converts each valid ISSN to the same `XXXX-XXXX` format and checks its final digit. It keeps missing, invalid, and valid-but-unmatched ISSNs as separate cases.

```{r}
#| label: issn-preparation

normalize_issn <- function(value) {
  compact <- gsub("[^0-9X]", "", toupper(trimws(as.character(value))))
  if (!grepl("^[0-9]{7}[0-9X]$", compact)) return(NA_character_)

  digits <- as.integer(strsplit(substr(compact, 1, 7), "")[[1]])
  check_value <- (11 - sum(digits * 8:2) %% 11) %% 11
  expected <- if (check_value == 10) "X" else as.character(check_value)
  if (substr(compact, 8, 8) != expected) return(NA_character_)

  paste0(substr(compact, 1, 4), "-", substr(compact, 5, 8))
}

parse_issns <- function(value) {
  if (
    is.na(value) ||
    !nzchar(trimws(value)) ||
    toupper(trimws(value)) == "NA"
  ) {
    return(character())
  }

  tokens <- unlist(strsplit(value, "[[:space:],;|]+"))
  normalized <- vapply(tokens[nzchar(tokens)], normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}

source_issns <- function(source) {
  values <- c(unlist(source$issn, use.names = FALSE), source$issn_l)
  normalized <- vapply(values, normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}
```

## Source retrieval

The sample sends all valid ISSNs to OpenAlex in one request. It asks Crossref about each ISSN separately and waits between requests. A Crossref `404` means that no journal record was found.

Sample responses are cached beneath the sample run directory. Full-cohort batching, pagination, retry provenance, and checkpointing are intentionally deferred to the full-mode ticket.

```{r}
#| label: source-retrieval

fetch_openalex_sources <- function(
  issns,
  api_key = Sys.getenv("OPENALEX_API_KEY")
) {
  stopifnot(length(issns) <= 100L, nzchar(api_key))

  request("https://api.openalex.org/sources") |>
    req_url_query(
      filter = paste0("issn:", paste(issns, collapse = "|")),
      per_page = 100,
      api_key = api_key
    ) |>
    req_timeout(seconds = 30) |>
    req_perform() |>
    resp_body_json(simplifyVector = FALSE) |>
    pluck("results")
}

fetch_crossref_journal <- function(issn) {
  response <- request(paste0("https://api.crossref.org/journals/", issn)) |>
    req_timeout(seconds = 30) |>
    req_error(
      is_error = \(response) {
        resp_status(response) != 404L && resp_is_error(response)
      }
    ) |>
    req_perform()
  if (resp_status(response) == 404L) return(NULL)
  resp_body_json(response, simplifyVector = FALSE)$message
}

```

## Load the PKP/OJS sample

The pinned PKP V7 file is Dataverse file `14084919`, released on 2026-07-18. This chunk checks the complete 98,273-row baseline, schema, row identity, and identifier counts before selecting ten fixed source rows in a deliberate display order.

```{r}
#| label: load-sample

raw_input <- read_csv(
  input_path,
  col_types = cols(.default = col_character()),
  na = character(),
  show_col_types = FALSE
)

expected_schema <- c(
  "context_name",
  "issn",
  "oai_url",
  "application",
  "country",
  "region",
  "total_record_count",
  "record_count_2025",
  "earliest_datestamp",
  "repository_name",
  "set_spec",
  "stats_id",
  "best_doaj_url",
  "last_completed_update",
  "first_beacon",
  "last_beacon",
  "last_oai_response",
  "unresponsive_endpoint",
  "unresponsive_context"
)
raw_row_identity <- paste(
  raw_input$oai_url,
  raw_input$repository_name,
  raw_input$set_spec,
  sep = "\r"
)

stopifnot(
  nrow(raw_input) == 98273L,
  identical(names(raw_input), expected_schema),
  all(tolower(raw_input$application) == "ojs"),
  !anyDuplicated(raw_row_identity)
)

all_input_issns <- map(raw_input$issn, parse_issns)
missing_identifier <- toupper(trimws(raw_input$issn)) == "NA"
stopifnot(
  sum(missing_identifier) == 26189L,
  sum(lengths(all_input_issns) > 0L) == 72084L,
  length(unique(unlist(all_input_issns))) == 103017L
)

sample_spec <- tribble(
  ~context_name, ~oai_url, ~repository_name, ~set_spec,
  "Kuwait Journal of Science",
  "https://journalskuwait.org/kjs/index.php/index/oai",
  "Kuwait Journal of Science",
  "KJS",
  "Cakrawala",
  "https://cakrawalajournal.org/index.php/index/oai",
  "CAKRAWALA",
  "cakrawala",
  "Sintesa: Jurnal Ilmu Pendidikan",
  "https://sintesa.stkip-arrahmaniyah.ac.id/index.php/index/oai",
  "Sintesa: Jurnal Ilmu Pendidikan STKIP Arrahmaniyah Depok",
  "sintesa",
  "CARAKA: Jurnal Teologi Biblika dan Praktika",
  "https://ojs.sttibc.ac.id/index.php/index/oai",
  "CARAKA:Jurnal Teologi",
  "ibc",
  "Jurnal Tata Kelola dan Akuntabilitas Keuangan Negara",
  "https://jurnal.bpk.go.id/index.php/index/oai",
  "Open Journal Systems",
  "TAKEN",
  "JOEEL (Journal of English Education and Literature)",
  "https://journal.stkippamanetalino.ac.id/index.php/index/oai",
  "Rumah Jurnal STKIP Pamane Talino Ngabang",
  "bahasa-inggris",
  "Malaysian Journal of Paediatrics and Child Health",
  "https://mpaeds.my/journals/index.php/index/oai",
  "NA",
  "MJPCH",
  "Jami Scientific Research Quarterly Journal",
  "https://journals.jami.edu.af/index.php/index/oai",
  "Jami University Press",
  "jsrqj",
  "Journal of Polymer & Composites",
  "https://engineeringjournals.stmjournals.in/index.php/index/oai",
  "Engineering Journals",
  "JoPC",
  "JURNAL TEKNIK PERTAMBANGAN",
  "https://e-journal.upr.ac.id/index.php/index/oai",
  "Journal Online Universitas Palangka Raya",
  "JTP"
)
sample_identity <- paste(
  sample_spec$oai_url,
  sample_spec$repository_name,
  sample_spec$set_spec,
  sep = "\r"
)
sample_positions <- match(sample_identity, raw_row_identity)
stopifnot(!anyNA(sample_positions))

ojs_rows <- raw_input[sample_positions, ] |>
  select(-any_of("admin_email"))
stopifnot(
  nrow(ojs_rows) == 10L,
  identical(ojs_rows$context_name, sample_spec$context_name)
)
```

## Multi-source enrichment

For each PKP row, the sample looks for OpenAlex and Crossref records with the same ISSN. If the same source record appears more than once, it is counted once. Every returned top-level API field becomes a dataframe column; nested objects and arrays are stored as JSON in their field columns, and the complete candidate lists are also retained as lossless JSON.

### Sample matching helpers

These helpers label each row as having no usable ISSN, no match, one match, or multiple matches. They also identify Crossref records by their ISSNs and convert every candidate field into a dataframe-safe value without filtering the response schema.

```{r}
#| label: sample-matching-helpers

classify_status <- function(input_issns, candidate_ids) {
  case_when(
    length(input_issns) == 0L ~ "not_attempted",
    length(candidate_ids) == 0L ~ "unmatched",
    length(candidate_ids) == 1L ~ "consistent",
    TRUE ~ "inconsistent"
  )
}

crossref_key <- function(journal) {
  issns <- vapply(journal$ISSN, normalize_issn, character(1))
  paste(sort(unique(issns[!is.na(issns)])), collapse = "|")
}

deduplicate_candidates <- function(candidates, identity) {
  candidate_ids <- vapply(candidates, identity, character(1))
  keep <- !duplicated(candidate_ids)
  list(
    candidates = candidates[keep],
    candidate_ids = candidate_ids[keep]
  )
}

match_exact_issns <- function(
  input_issns,
  candidates,
  candidate_issns,
  candidate_identity
) {
  matched <- keep(candidates, \(candidate) {
    any(input_issns %in% candidate_issns(candidate))
  })
  deduplicated <- deduplicate_candidates(matched, candidate_identity)
  matched_issns <- intersect(
    input_issns,
    unique(unlist(map(deduplicated$candidates, candidate_issns)))
  )

  c(
    list(
      status = classify_status(
        input_issns,
        deduplicated$candidate_ids
      ),
      matched_issns = matched_issns
    ),
    deduplicated
  )
}

serialize_json <- function(value) {
  as.character(jsonlite::toJSON(
    value,
    auto_unbox = TRUE,
    null = "null",
    na = "null",
    digits = NA
  ))
}

dataframe_value <- function(value) {
  if (is.null(value) || !length(value)) return(NA)
  if (is.atomic(value) && length(value) == 1L) return(value)
  serialize_json(value)
}

candidate_prefix <- function(prefix, index) {
  if (index == 1L) prefix else paste0(prefix, "_candidate_", index)
}

candidate_field_names <- function(match, prefix) {
  unlist(imap(match$candidates, \(candidate, index) {
    paste(candidate_prefix(prefix, index), names(candidate), sep = "_")
  }), use.names = FALSE)
}

expand_candidate_fields <- function(match, prefix) {
  cells <- list()
  for (index in seq_along(match$candidates)) {
    candidate <- match$candidates[[index]]
    values <- map(candidate, dataframe_value)
    names(values) <- paste(
      candidate_prefix(prefix, index),
      names(candidate),
      sep = "_"
    )
    cells <- c(cells, values)
  }
  cells[[paste0(prefix, "_candidates__json")]] <- serialize_json(
    match$candidates
  )
  as_tibble(cells, .name_repair = "check_unique")
}
```

### Sample enrichment orchestration

This chunk collects the unique sample ISSNs, retrieves or reuses source records, matches them to the PKP rows, adds every returned source field, and writes one CSV row for each PKP row. The first candidate uses the ordinary source prefix; additional candidates use numbered prefixes so inconsistent matches remain complete and auditable.

```{r}
#| label: sample-enrichment

input_issns <- map(ojs_rows$issn, parse_issns)
tokens <- sort(unique(unlist(input_issns)))

openalex_cache_hit <- file.exists(openalex_cache_path)
openalex_sources <- if (openalex_cache_hit) {
  readRDS(openalex_cache_path)
} else {
  sources <- fetch_openalex_sources(tokens)
  saveRDS(sources, openalex_cache_path)
  sources
}

crossref_cache_hit <- file.exists(crossref_cache_path)
crossref_journals <- if (crossref_cache_hit) {
  readRDS(crossref_cache_path)
} else {
  journals <- set_names(
    map(tokens, \(token) {
      journal <- fetch_crossref_journal(token)
      Sys.sleep(0.25)
      journal
    }),
    tokens
  )
  saveRDS(journals, crossref_cache_path)
  journals
}
crossref_candidates <- unname(compact(crossref_journals))

openalex_matches <- map(input_issns, \(issns) {
  match_exact_issns(
    issns,
    openalex_sources,
    source_issns,
    \(source) source$id
  )
})

crossref_matches <- map(input_issns, \(issns) {
  match_exact_issns(
    issns,
    crossref_candidates,
    \(journal) {
      issns <- vapply(journal$ISSN, normalize_issn, character(1))
      unique(issns[!is.na(issns)])
    },
    crossref_key
  )
})

openalex_data <- map_dfr(
  openalex_matches,
  expand_candidate_fields,
  prefix = "openalex"
)
crossref_data <- map_dfr(
  crossref_matches,
  expand_candidate_fields,
  prefix = "crossref"
)

result <- ojs_rows |>
  mutate(
    identifier_status = case_when(
      is.na(issn) |
        !nzchar(trimws(issn)) |
        toupper(trimws(issn)) == "NA" ~ "missing",
      lengths(input_issns) == 0L ~ "invalid",
      TRUE ~ "valid"
    ),
    openalex_input_issns = map_chr(input_issns, paste, collapse = "|"),
    openalex_matched_issns = map_chr(
      openalex_matches,
      \(match) paste(match$matched_issns, collapse = "|")
    ),
    openalex_match_status = map_chr(openalex_matches, "status"),
    openalex_candidate_ids = map_chr(
      openalex_matches,
      \(match) paste(match$candidate_ids, collapse = "|")
    ),
    crossref_input_issns = map_chr(input_issns, paste, collapse = "|"),
    crossref_matched_issns = map_chr(
      crossref_matches,
      \(match) paste(match$matched_issns, collapse = "|")
    ),
    crossref_match_status = map_chr(crossref_matches, "status"),
    crossref_candidate_issn_sets = map_chr(
      crossref_matches,
      \(match) paste(match$candidate_ids, collapse = ";")
    )
  ) |>
  bind_cols(openalex_data, crossref_data)

write_csv(result, output_path, na = "")

run_metadata <- list(
  producer = "analysis/ojs_journal_enrichment.qmd",
  run_mode = "sample",
  created_at = format(Sys.time(), tz = "UTC", usetz = TRUE),
  input = list(
    version = pkp_version,
    dataverse_file_id = 14084919L,
    md5 = pkp_md5,
    rows = nrow(raw_input),
    valid_identifier_rows = sum(lengths(all_input_issns) > 0L),
    missing_identifier_rows = sum(missing_identifier),
    distinct_valid_issns = length(unique(unlist(all_input_issns)))
  ),
  sample_rows = nrow(result),
  output = list(
    file = basename(output_path),
    md5 = unname(tools::md5sum(output_path))
  ),
  cache = list(
    openalex = list(
      file = basename(openalex_cache_path),
      reused = openalex_cache_hit
    ),
    crossref = list(
      file = basename(crossref_cache_path),
      reused = crossref_cache_hit
    )
  ),
  runtime = list(
    r = R.version.string,
    quarto = system2("quarto", "--version", stdout = TRUE)[1],
    packages = as.list(set_names(
      vapply(
        required_packages,
        \(package) as.character(packageVersion(package)),
        character(1)
      ),
      required_packages
    ))
  )
)
jsonlite::write_json(
  run_metadata,
  metadata_path,
  auto_unbox = TRUE,
  pretty = TRUE
)
```

## Checks

These checks confirm that every input row appears in the output, every returned top-level API field is retained, the lossless candidate JSON matches the fetched records, the CSV file was written, and the sample contains the expected match outcomes.

```{r}
#| label: contract-check

expected_openalex_columns <- unique(unlist(map(
  openalex_matches,
  candidate_field_names,
  prefix = "openalex"
)))
expected_crossref_columns <- unique(unlist(map(
  crossref_matches,
  candidate_field_names,
  prefix = "crossref"
)))
expected_openalex_status <- c(
  "consistent",
  "consistent",
  "unmatched",
  "consistent",
  "consistent",
  "consistent",
  "consistent",
  "consistent",
  "not_attempted",
  "inconsistent"
)
metadata_text <- paste(readLines(metadata_path, warn = FALSE), collapse = "\n")
artifact_root <- normalizePath(artifact_dir)

stopifnot(
  identical(pkp_version, 7L),
  identical(unname(tools::md5sum(input_path)), pkp_md5),
  nrow(raw_input) == 98273L,
  identical(names(raw_input), expected_schema),
  !anyDuplicated(raw_row_identity),
  sum(missing_identifier) == 26189L,
  sum(lengths(all_input_issns) > 0L) == 72084L,
  length(unique(unlist(all_input_issns))) == 103017L,
  identical(
    paste(
      ojs_rows$oai_url,
      ojs_rows$repository_name,
      ojs_rows$set_spec,
      sep = "\r"
    ),
    sample_identity
  ),
  sum(result$identifier_status == "missing") == 1L,
  nrow(result) == 10L,
  startsWith(normalizePath(input_path), artifact_root),
  startsWith(normalizePath(output_path), artifact_root),
  startsWith(normalizePath(metadata_path), artifact_root),
  file.exists(metadata_path),
  !grepl("OPENALEX_API_KEY|CROSSREF_MAILTO", metadata_text),
  all(nzchar(unlist(run_metadata$runtime))),
  all(vapply(c(
    "normalize_issn",
    "match_exact_issns",
    "deduplicate_candidates",
    "expand_candidate_fields",
    "serialize_json"
  ), exists, logical(1), mode = "function")),
  exists("fetch_openalex_sources", mode = "function"),
  exists("fetch_crossref_journal", mode = "function"),
  nrow(result) == nrow(ojs_rows),
  file.exists(output_path),
  all(c(
    "openalex_match_status",
    "openalex_id",
    "openalex_summary_stats",
    "openalex_topics",
    "openalex_candidate_2_id",
    "openalex_candidates__json",
    "crossref_match_status",
    "crossref_title",
    "crossref_counts",
    "crossref_coverage",
    "crossref_candidates__json"
  ) %in% names(result)),
  all(expected_openalex_columns %in% names(result)),
  all(expected_crossref_columns %in% names(result)),
  identical(
    result$openalex_candidates__json,
    map_chr(openalex_matches, \(match) serialize_json(match$candidates))
  ),
  identical(
    result$crossref_candidates__json,
    map_chr(crossref_matches, \(match) serialize_json(match$candidates))
  ),
  identical(result$openalex_match_status, expected_openalex_status),
  result$openalex_matched_issns[
    result$context_name == "Jami Scientific Research Quarterly Journal"
  ] == "2710-4990",
  "not_attempted" %in% result$crossref_match_status
)
```

## Complete merged dataframe

This table displays all OJS, OpenAlex, and Crossref columns in the merged sample dataframe.

```{r}
#| label: complete-sample-dataframe

result |>
  knitr::kable(format = "html")
```

## Complete merged column list

This table lists every column in the CSV so collaborators can review its structure.

```{r}
#| label: complete-column-list

tibble(column = names(result)) |>
  knitr::kable()
```
