---
title: "PKP/OJS multi-source journal enrichment"
format: html
execute:
  echo: true
  warning: false
  message: false
---

This notebook adds current OpenAlex Source data to PKP OJS rows from the pinned V7 release by exact ISSN. Sample mode is the default and also checks Crossref against ten deliberately selected rows. Full mode must be started with `render_full.sh`; it checkpoints V7 OpenAlex responses and preserves all 98,273 PKP rows.

Set `OPENALEX_API_KEY` in the environment or the repository-root `.env` file before rendering. A normal render and `render_sample.sh` always use sample mode.

## Packages and paths

This chunk installs only missing CRAN packages, loads them, downloads the pinned PKP V7 input when needed, verifies its official checksum, and keeps every local artifact under this analysis owner.

```{r}
#| label: packages-and-paths

required_packages <- c(
  "dplyr",
  "httr2",
  "jsonlite",
  "knitr",
  "purrr",
  "readr",
  "rmarkdown",
  "tibble"
)
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]
if (length(missing_packages)) {
  install.packages(missing_packages, repos = "https://cloud.r-project.org")
}

library(dplyr)
library(httr2)
library(purrr)
library(readr)
library(tibble)

run_mode <- Sys.getenv("OJS_ENRICHMENT_MODE", unset = "sample")
stopifnot(run_mode %in% c("sample", "full"))

project_env_path <- file.path("..", "..", "..", ".env")
if (!nzchar(Sys.getenv("OPENALEX_API_KEY")) && file.exists(project_env_path)) {
  readRenviron(project_env_path)
}
if (!nzchar(Sys.getenv("OPENALEX_API_KEY"))) {
  stop("Set OPENALEX_API_KEY before running this notebook.", call. = FALSE)
}

pkp_version <- 7L
pkp_url <- paste0(
  "https://dataverse.harvard.edu/api/access/datafile/14084919",
  "?format=original"
)
pkp_md5 <- "3a4ad8ae1ebfcc2b991aaf55b2d82c92"
artifact_dir <- file.path("..", "artifacts", "ojs_journal_enrichment")
input_dir <- file.path(artifact_dir, "pkp-v7")
input_path <- file.path(input_dir, "beacon.csv")
run_dir <- file.path(
  artifact_dir,
  if (run_mode == "full") "full-v7" else "sample"
)
output_path <- file.path(
  run_dir,
  if (run_mode == "full") {
    "pkp-ojs-openalex-enriched.parquet"
  } else {
    "pkp-ojs-multisource-enriched.csv"
  }
)
metadata_path <- file.path(run_dir, "run-metadata.json")
staging_path <- file.path(run_dir, "pkp-ojs-openalex-enriched.ndjson")
parquet_schema_path <- file.path(run_dir, "parquet-schema.json")
source_output_path <- file.path(run_dir, "openalex-sources.parquet")
source_staging_path <- file.path(run_dir, "openalex-sources.ndjson")
source_schema_path <- file.path(run_dir, "openalex-sources-schema.json")
parquet_validation_path <- file.path(run_dir, "parquet-validation.json")
openalex_cache_path <- file.path(run_dir, "openalex-sources.rds")
crossref_cache_path <- file.path(run_dir, "crossref-journals.rds")
openalex_batch_dir <- file.path(run_dir, "openalex-batches")

dir.create(input_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(run_dir, recursive = TRUE, showWarnings = FALSE)
if (run_mode == "full") {
  dir.create(openalex_batch_dir, recursive = TRUE, showWarnings = FALSE)
}

curl <- Sys.which("curl")
if (!nzchar(curl)) {
  stop("curl is required to download the pinned PKP V7 input.", call. = FALSE)
}

download_pkp_input <- function(path) {
  if (file.exists(path)) return(invisible(path))

  download_path <- paste0(path, ".download")
  on.exit(unlink(download_path), add = TRUE)
  status <- system2(
    curl,
    c(
      "-fL",
      "--retry", "5",
      "--retry-all-errors",
      "--retry-delay", "2",
      "--continue-at", "-",
      "-o", shQuote(download_path),
      shQuote(pkp_url)
    )
  )
  if (status != 0L) {
    stop("Could not download PKP V7.", call. = FALSE)
  }
  if (unname(tools::md5sum(download_path)) != pkp_md5) {
    stop("Downloaded PKP V7 file failed checksum verification.", call. = FALSE)
  }
  if (!file.rename(download_path, path)) {
    stop("Could not move the verified PKP V7 file into place.", call. = FALSE)
  }
  invisible(path)
}

download_pkp_input(input_path)
if (unname(tools::md5sum(input_path)) != pkp_md5) {
  stop("Local PKP V7 file failed checksum verification.", call. = FALSE)
}
```

## ISSN preparation

ISSNs may contain spaces, punctuation, or more than one value. This chunk converts each valid ISSN to the same `XXXX-XXXX` format and checks its final digit. It keeps missing, invalid, and valid-but-unmatched ISSNs as separate cases.

```{r}
#| label: issn-preparation

normalize_issn <- function(value) {
  compact <- gsub("[^0-9X]", "", toupper(trimws(as.character(value))))
  if (!grepl("^[0-9]{7}[0-9X]$", compact)) return(NA_character_)

  digits <- as.integer(strsplit(substr(compact, 1, 7), "")[[1]])
  check_value <- (11 - sum(digits * 8:2) %% 11) %% 11
  expected <- if (check_value == 10) "X" else as.character(check_value)
  if (substr(compact, 8, 8) != expected) return(NA_character_)

  paste0(substr(compact, 1, 4), "-", substr(compact, 5, 8))
}

is_missing_identifier <- function(value) {
  is.na(value) |
    !nzchar(trimws(value)) |
    toupper(trimws(value)) == "NA"
}

parse_issns <- function(value) {
  if (is_missing_identifier(value)) return(character())

  tokens <- unlist(strsplit(value, "[[:space:],;|]+"))
  normalized <- vapply(tokens[nzchar(tokens)], normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}

source_issns <- function(source) {
  values <- c(unlist(source$issn, use.names = FALSE), source$issn_l)
  normalized <- vapply(values, normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}
```

## Source retrieval

The sample sends all valid ISSNs to OpenAlex in one request. Full mode splits the 103,017 valid ISSNs into deterministic groups of at most 100, requests the maximum page size, follows every cursor, and atomically checkpoints each completed V7 batch. A failed batch remains an API-error state and is retried on the next run.

Sample mode asks Crossref about each ISSN separately and waits between requests. A Crossref `404` means that no journal record was found.

```{r}
#| label: source-retrieval

fetch_openalex_sources <- function(
  issns,
  api_key = Sys.getenv("OPENALEX_API_KEY")
) {
  stopifnot(length(issns) <= 100L, nzchar(api_key))

  request("https://api.openalex.org/sources") |>
    req_url_query(
      filter = paste0("issn:", paste(issns, collapse = "|")),
      per_page = 100,
      api_key = api_key
    ) |>
    req_timeout(seconds = 30) |>
    req_perform() |>
    resp_body_json(simplifyVector = FALSE) |>
    pluck("results")
}

perform_openalex_page <- function(
  issns,
  cursor,
  api_key = Sys.getenv("OPENALEX_API_KEY"),
  max_attempts = 4L
) {
  stopifnot(length(issns) <= 100L, nzchar(api_key))

  for (attempt in seq_len(max_attempts)) {
    response <- tryCatch(
      request("https://api.openalex.org/sources") |>
        req_url_query(
          filter = paste0("issn:", paste(issns, collapse = "|")),
          per_page = 200,
          cursor = cursor,
          api_key = api_key
        ) |>
        req_timeout(seconds = 30) |>
        req_error(is_error = \(response) FALSE) |>
        req_perform(),
      error = identity
    )

    if (!inherits(response, "condition")) {
      status <- resp_status(response)
      if (status < 400L) {
        return(list(
          ok = TRUE,
          attempts = attempt,
          body = resp_body_json(response, simplifyVector = FALSE)
        ))
      }
      failure <- paste0("http_", status)
      retryable <- status %in% c(403L, 429L, 500L, 502L, 503L, 504L)
    } else {
      failure <- "transport_error"
      retryable <- TRUE
    }

    if (!retryable || attempt == max_attempts) {
      return(list(
        ok = FALSE,
        attempts = attempt,
        failure = failure
      ))
    }
    Sys.sleep(min(2^(attempt - 1L), 8))
  }
}

fetch_openalex_batch <- function(issns) {
  cursor <- "*"
  seen_cursors <- character()
  pages <- list()
  page_log <- list()
  attempts <- 0L

  repeat {
    if (cursor %in% seen_cursors) {
      return(list(
        status = "api_error",
        issns = issns,
        sources = unlist(pages, recursive = FALSE),
        pages = page_log,
        attempts = attempts,
        failure = "repeated_cursor",
        pagination_complete = FALSE
      ))
    }
    seen_cursors <- c(seen_cursors, cursor)

    page <- perform_openalex_page(issns, cursor)
    attempts <- attempts + page$attempts
    if (!page$ok) {
      return(list(
        status = "api_error",
        issns = issns,
        sources = unlist(pages, recursive = FALSE),
        pages = page_log,
        attempts = attempts,
        failure = page$failure,
        pagination_complete = FALSE
      ))
    }

    results <- if (is.null(page$body$results)) list() else page$body$results
    next_cursor <- page$body$meta$next_cursor
    pages[[length(pages) + 1L]] <- results
    page_log[[length(page_log) + 1L]] <- list(
      result_count = length(results),
      attempts = page$attempts,
      retrieved_at = format(Sys.time(), tz = "UTC", usetz = TRUE)
    )
    if (is.null(next_cursor) || !nzchar(next_cursor)) break
    cursor <- next_cursor
  }

  list(
    status = "complete",
    issns = issns,
    sources = unlist(pages, recursive = FALSE),
    pages = page_log,
    attempts = attempts,
    failure = NULL,
    pagination_complete = TRUE
  )
}

save_rds_atomic <- function(value, path) {
  temporary <- paste0(path, ".tmp")
  saveRDS(value, temporary)
  if (!file.rename(temporary, path)) {
    unlink(temporary)
    stop("Could not checkpoint OpenAlex batch.", call. = FALSE)
  }
  invisible(path)
}

fetch_crossref_journal <- function(issn) {
  response <- request(paste0("https://api.crossref.org/journals/", issn)) |>
    req_timeout(seconds = 30) |>
    req_error(
      is_error = \(response) {
        resp_status(response) != 404L && resp_is_error(response)
      }
    ) |>
    req_perform()
  if (resp_status(response) == 404L) return(NULL)
  resp_body_json(response, simplifyVector = FALSE)$message
}

```

## Load the PKP V7 rows

The pinned PKP V7 file is Dataverse file `14084919`, released on 2026-07-18. This chunk checks the complete 98,273-row baseline, schema, row identity, and identifier counts before either retaining the full cohort or selecting ten fixed source rows in a deliberate display order.

```{r}
#| label: load-sample

raw_input <- read_csv(
  input_path,
  col_types = cols(.default = col_character()),
  na = character(),
  show_col_types = FALSE
)

expected_schema <- c(
  "context_name",
  "issn",
  "oai_url",
  "application",
  "country",
  "region",
  "total_record_count",
  "record_count_2025",
  "earliest_datestamp",
  "repository_name",
  "set_spec",
  "stats_id",
  "best_doaj_url",
  "last_completed_update",
  "first_beacon",
  "last_beacon",
  "last_oai_response",
  "unresponsive_endpoint",
  "unresponsive_context"
)
row_identity <- function(data) {
  paste(data$oai_url, data$repository_name, data$set_spec, sep = "\r")
}
raw_row_identity <- row_identity(raw_input)

stopifnot(
  nrow(raw_input) == 98273L,
  identical(names(raw_input), expected_schema),
  all(tolower(raw_input$application) == "ojs"),
  !anyDuplicated(raw_row_identity)
)

all_input_issns <- map(raw_input$issn, parse_issns)
missing_identifier <- is_missing_identifier(raw_input$issn)
stopifnot(
  sum(missing_identifier) == 26189L,
  sum(lengths(all_input_issns) > 0L) == 72084L,
  length(unique(unlist(all_input_issns))) == 103017L
)

sample_spec <- tribble(
  ~context_name, ~oai_url, ~repository_name, ~set_spec,
  "Kuwait Journal of Science",
  "https://journalskuwait.org/kjs/index.php/index/oai",
  "Kuwait Journal of Science",
  "KJS",
  "Cakrawala",
  "https://cakrawalajournal.org/index.php/index/oai",
  "CAKRAWALA",
  "cakrawala",
  "Sintesa: Jurnal Ilmu Pendidikan",
  "https://sintesa.stkip-arrahmaniyah.ac.id/index.php/index/oai",
  "Sintesa: Jurnal Ilmu Pendidikan STKIP Arrahmaniyah Depok",
  "sintesa",
  "CARAKA: Jurnal Teologi Biblika dan Praktika",
  "https://ojs.sttibc.ac.id/index.php/index/oai",
  "CARAKA:Jurnal Teologi",
  "ibc",
  "Jurnal Tata Kelola dan Akuntabilitas Keuangan Negara",
  "https://jurnal.bpk.go.id/index.php/index/oai",
  "Open Journal Systems",
  "TAKEN",
  "JOEEL (Journal of English Education and Literature)",
  "https://journal.stkippamanetalino.ac.id/index.php/index/oai",
  "Rumah Jurnal STKIP Pamane Talino Ngabang",
  "bahasa-inggris",
  "Malaysian Journal of Paediatrics and Child Health",
  "https://mpaeds.my/journals/index.php/index/oai",
  "NA",
  "MJPCH",
  "Jami Scientific Research Quarterly Journal",
  "https://journals.jami.edu.af/index.php/index/oai",
  "Jami University Press",
  "jsrqj",
  "Journal of Polymer & Composites",
  "https://engineeringjournals.stmjournals.in/index.php/index/oai",
  "Engineering Journals",
  "JoPC",
  "JURNAL TEKNIK PERTAMBANGAN",
  "https://e-journal.upr.ac.id/index.php/index/oai",
  "Journal Online Universitas Palangka Raya",
  "JTP"
)
sample_identity <- row_identity(sample_spec)
sample_positions <- match(sample_identity, raw_row_identity)
stopifnot(!anyNA(sample_positions))

sample_rows <- raw_input[sample_positions, ] |>
  select(-any_of("admin_email"))
stopifnot(
  nrow(sample_rows) == 10L,
  identical(sample_rows$context_name, sample_spec$context_name)
)

ojs_rows <- if (run_mode == "full") {
  raw_input |> select(-any_of("admin_email"))
} else {
  sample_rows
}
```

## Multi-source enrichment

For each PKP row, the sample looks for OpenAlex and Crossref records with the same ISSN. If the same source record appears more than once, it is counted once. Sample mode expands returned fields for review. Full mode keeps candidate IDs in the one-row-per-PKP master and indexes each unique Source ID to its complete V7 checkpoint response in a companion Parquet catalog.

### Sample matching helpers

These helpers label each row as having no usable ISSN, no match, one match, or multiple matches. They also identify Crossref records by their ISSNs and convert every candidate field into a dataframe-safe value without filtering the response schema.

```{r}
#| label: sample-matching-helpers

classify_status <- function(input_issns, candidate_ids) {
  case_when(
    length(input_issns) == 0L ~ "not_attempted",
    length(candidate_ids) == 0L ~ "unmatched",
    length(candidate_ids) == 1L ~ "consistent",
    TRUE ~ "inconsistent"
  )
}

crossref_key <- function(journal) {
  issns <- vapply(journal$ISSN, normalize_issn, character(1))
  paste(sort(unique(issns[!is.na(issns)])), collapse = "|")
}

deduplicate_candidates <- function(candidates, identity) {
  candidate_ids <- vapply(candidates, identity, character(1))
  keep <- !duplicated(candidate_ids)
  list(
    candidates = candidates[keep],
    candidate_ids = candidate_ids[keep]
  )
}

index_candidates_by_issn <- function(candidates, candidate_issns) {
  index <- new.env(hash = TRUE, parent = emptyenv())
  for (candidate_position in seq_along(candidates)) {
    for (issn in candidate_issns(candidates[[candidate_position]])) {
      positions <- if (exists(issn, index, inherits = FALSE)) {
        get(issn, index, inherits = FALSE)
      } else {
        integer()
      }
      assign(issn, c(positions, candidate_position), index)
    }
  }
  index
}

candidates_for_issns <- function(candidates, index, issns) {
  positions <- unique(unlist(map(issns, \(issn) {
    if (exists(issn, index, inherits = FALSE)) {
      get(issn, index, inherits = FALSE)
    } else {
      integer()
    }
  })))
  candidates[positions]
}

match_exact_issns <- function(
  input_issns,
  candidates,
  candidate_issns,
  candidate_identity
) {
  matched <- keep(candidates, \(candidate) {
    any(input_issns %in% candidate_issns(candidate))
  })
  deduplicated <- deduplicate_candidates(matched, candidate_identity)
  matched_issns <- intersect(
    input_issns,
    unique(unlist(map(deduplicated$candidates, candidate_issns)))
  )

  c(
    list(
      status = classify_status(
        input_issns,
        deduplicated$candidate_ids
      ),
      matched_issns = matched_issns
    ),
    deduplicated
  )
}

serialize_json <- function(value) {
  as.character(jsonlite::toJSON(
    value,
    auto_unbox = TRUE,
    null = "null",
    na = "null",
    digits = NA
  ))
}

dataframe_value <- function(value) {
  if (is.null(value) || !length(value)) return(NA)
  if (is.atomic(value) && length(value) == 1L) return(value)
  serialize_json(value)
}

candidate_prefix <- function(prefix, index) {
  if (index == 1L) prefix else paste0(prefix, "_candidate_", index)
}

candidate_field_names <- function(match, prefix) {
  unlist(imap(match$candidates, \(candidate, index) {
    paste(candidate_prefix(prefix, index), names(candidate), sep = "_")
  }), use.names = FALSE)
}

expand_candidate_fields <- function(match, prefix) {
  cells <- list()
  for (index in seq_along(match$candidates)) {
    candidate <- match$candidates[[index]]
    values <- map(candidate, dataframe_value)
    names(values) <- paste(
      candidate_prefix(prefix, index),
      names(candidate),
      sep = "_"
    )
    cells <- c(cells, values)
  }
  cells[[paste0(prefix, "_candidates__json")]] <- serialize_json(
    match$candidates
  )
  as_tibble(cells, .name_repair = "check_unique")
}
```

### Enrichment orchestration

Both modes use the same exact-ISSN matcher. Full mode indexes candidates by normalized ISSN so matching 98,273 rows does not repeatedly scan the complete Source response set. Candidate IDs connect the compact master to the Source catalog and lossless V7 checkpoints, so inconsistent identities remain complete without duplicating large Source JSON.

```{r}
#| label: enrichment

input_issns <- map(ojs_rows$issn, parse_issns)
tokens <- sort(unique(unlist(input_issns)))

if (run_mode == "full") {
  token_batches <- split(tokens, ceiling(seq_along(tokens) / 100L))
  process_batch <- function(batch_index) {
    batch <- token_batches[[batch_index]]
    checkpoint_path <- file.path(
      openalex_batch_dir,
      sprintf("batch-%04d.rds", batch_index)
    )
    checkpoint <- if (file.exists(checkpoint_path)) {
      tryCatch(readRDS(checkpoint_path), error = \(error) NULL)
    } else {
      NULL
    }
    reusable <- !is.null(checkpoint) &&
      identical(checkpoint$pkp_version, pkp_version) &&
      identical(checkpoint$input_md5, pkp_md5) &&
      identical(checkpoint$issns, batch) &&
      identical(checkpoint$status, "complete") &&
      isTRUE(checkpoint$pagination_complete)

    if (reusable) {
      checkpoint$reused <- TRUE
      return(checkpoint)
    }

    checkpoint <- fetch_openalex_batch(batch)
    checkpoint$pkp_version <- pkp_version
    checkpoint$input_md5 <- pkp_md5
    checkpoint$batch_index <- as.integer(batch_index)
    checkpoint$reused <- FALSE
    save_rds_atomic(checkpoint, checkpoint_path)
    checkpoint
  }
  openalex_cluster <- parallel::makeCluster(4L)
  invisible(parallel::clusterEvalQ(openalex_cluster, library(httr2)))
  parallel::clusterExport(
    openalex_cluster,
    c(
      "token_batches",
      "openalex_batch_dir",
      "pkp_version",
      "pkp_md5",
      "perform_openalex_page",
      "fetch_openalex_batch",
      "save_rds_atomic",
      "process_batch"
    ),
    envir = environment()
  )
  batch_results <- tryCatch(
    parallel::parLapplyLB(
      openalex_cluster,
      seq_along(token_batches),
      process_batch
    ),
    finally = parallel::stopCluster(openalex_cluster)
  )
  openalex_cache_hit <- all(map_lgl(batch_results, "reused"))
  openalex_sources <- unlist(
    map(batch_results, "sources"),
    recursive = FALSE
  )
  failed_issns <- unique(unlist(map(
    keep(batch_results, \(batch) batch$status != "complete"),
    "issns"
  )))
} else {
  token_batches <- list(tokens)
  batch_results <- list()
  failed_issns <- character()
  openalex_cache_hit <- file.exists(openalex_cache_path)
  openalex_sources <- if (openalex_cache_hit) {
    readRDS(openalex_cache_path)
  } else {
    sources <- fetch_openalex_sources(tokens)
    saveRDS(sources, openalex_cache_path)
    sources
  }
}
openalex_sources <- deduplicate_candidates(
  openalex_sources,
  \(source) source$id
)$candidates
openalex_index <- index_candidates_by_issn(
  openalex_sources,
  source_issns
)

openalex_matches <- map(input_issns, \(issns) {
  match <- match_exact_issns(
    issns,
    candidates_for_issns(openalex_sources, openalex_index, issns),
    source_issns,
    \(source) source$id
  )
  if (any(issns %in% failed_issns)) match$status <- "api_error"
  match
})
output_status_counts <- table(map_chr(openalex_matches, "status"))

base_result <- ojs_rows |>
  mutate(
    identifier_status = case_when(
      is_missing_identifier(issn) ~ "missing",
      lengths(input_issns) == 0L ~ "invalid",
      TRUE ~ "valid"
    ),
    openalex_input_issns = map_chr(input_issns, paste, collapse = "|"),
    openalex_matched_issns = map_chr(
      openalex_matches,
      \(match) paste(match$matched_issns, collapse = "|")
    ),
    openalex_match_status = map_chr(openalex_matches, "status"),
    openalex_candidate_count = map_int(
      openalex_matches,
      \(match) length(match$candidate_ids)
    ),
    openalex_candidate_ids = map_chr(
      openalex_matches,
      \(match) paste(match$candidate_ids, collapse = "|")
    )
  )

if (run_mode == "sample") {
  openalex_data <- map_dfr(
    openalex_matches,
    expand_candidate_fields,
    prefix = "openalex"
  )
  result <- bind_cols(base_result, openalex_data)
  crossref_cache_hit <- file.exists(crossref_cache_path)
  crossref_journals <- if (crossref_cache_hit) {
    readRDS(crossref_cache_path)
  } else {
    journals <- set_names(
      map(tokens, \(token) {
        journal <- fetch_crossref_journal(token)
        Sys.sleep(0.25)
        journal
      }),
      tokens
    )
    saveRDS(journals, crossref_cache_path)
    journals
  }
  crossref_candidates <- unname(compact(crossref_journals))
  crossref_index <- index_candidates_by_issn(
    crossref_candidates,
    \(journal) {
      issns <- vapply(journal$ISSN, normalize_issn, character(1))
      unique(issns[!is.na(issns)])
    }
  )
  crossref_matches <- map(input_issns, \(issns) {
    match_exact_issns(
      issns,
      candidates_for_issns(
        crossref_candidates,
        crossref_index,
        issns
      ),
      \(journal) {
        issns <- vapply(journal$ISSN, normalize_issn, character(1))
        unique(issns[!is.na(issns)])
      },
      crossref_key
    )
  })
  crossref_data <- map_dfr(
    crossref_matches,
    expand_candidate_fields,
    prefix = "crossref"
  )
  result <- result |>
    mutate(
      crossref_input_issns = map_chr(input_issns, paste, collapse = "|"),
      crossref_matched_issns = map_chr(
        crossref_matches,
        \(match) paste(match$matched_issns, collapse = "|")
      ),
      crossref_match_status = map_chr(crossref_matches, "status"),
      crossref_candidate_issn_sets = map_chr(
        crossref_matches,
        \(match) paste(match$candidate_ids, collapse = ";")
      )
    ) |>
    bind_cols(crossref_data)
  write_csv(result, output_path, na = "")
  output_rows <- nrow(result)
  output_columns <- names(result)
  review_result <- result
} else {
  crossref_cache_hit <- FALSE
  crossref_matches <- list()
  source_fields <- sort(unique(unlist(map(openalex_sources, names))))
  output_columns <- names(base_result)
  source_output_columns <- c("openalex_source_id", "checkpoint_file")
  checkpoint_by_source <- new.env(hash = TRUE, parent = emptyenv())
  for (batch in batch_results) {
    checkpoint_file <- sprintf("batch-%04d.rds", batch$batch_index)
    for (source in batch$sources) {
      if (!exists(source$id, checkpoint_by_source, inherits = FALSE)) {
        assign(source$id, checkpoint_file, checkpoint_by_source)
      }
    }
  }

  build_master_chunk <- function(indices) {
    chunk <- as.data.frame(base_result[indices, ], stringsAsFactors = FALSE)
    chunk[] <- lapply(chunk, as.character)
    chunk[output_columns]
  }

  build_source_chunk <- function(indices) {
    source_ids <- unname(map_chr(openalex_sources[indices], "id"))
    data.frame(
      openalex_source_id = source_ids,
      checkpoint_file = unname(vapply(
        source_ids,
        get,
        character(1),
        envir = checkpoint_by_source,
        inherits = FALSE
      )),
      stringsAsFactors = FALSE
    )
  }

  write_ndjson_chunks <- function(path, rows, builder) {
    unlink(path)
    connection <- file(path, open = "wb")
    tryCatch(
      {
        for (start in seq.int(1L, rows, by = 1000L)) {
          indices <- start:min(start + 999L, rows)
          jsonlite::stream_out(
            builder(indices),
            connection,
            verbose = FALSE
          )
        }
      },
      finally = close(connection)
    )
  }

  uv <- Sys.which("uv")
  if (!nzchar(uv)) {
    stop("uv is required to write the full Parquet artifacts.", call. = FALSE)
  }
  convert_to_parquet <- function(input, schema, output, expected_rows) {
    conversion <- system2(
      uv,
      c(
        "run",
        "--with", "pyarrow==21.0.0",
        "python", "ndjson_to_parquet.py", "convert",
        shQuote(input),
        shQuote(schema),
        shQuote(output),
        "--expected-rows", expected_rows
      ),
      stdout = TRUE,
      stderr = TRUE
    )
    if (!is.null(attr(conversion, "status"))) {
      stop(paste(conversion, collapse = "\n"), call. = FALSE)
    }
    jsonlite::fromJSON(tail(conversion, 1L))
  }

  write_ndjson_chunks(staging_path, nrow(base_result), build_master_chunk)
  jsonlite::write_json(
    list(columns = output_columns),
    parquet_schema_path,
    auto_unbox = TRUE,
    pretty = TRUE
  )
  parquet_contract <- convert_to_parquet(
    staging_path,
    parquet_schema_path,
    output_path,
    nrow(base_result)
  )
  unlink(staging_path)

  write_ndjson_chunks(
    source_staging_path,
    length(openalex_sources),
    build_source_chunk
  )
  jsonlite::write_json(
    list(
      columns = source_output_columns,
      source_fields = source_fields
    ),
    source_schema_path,
    auto_unbox = TRUE,
    pretty = TRUE
  )
  source_parquet_contract <- convert_to_parquet(
    source_staging_path,
    source_schema_path,
    source_output_path,
    length(openalex_sources)
  )
  unlink(source_staging_path)

  jsonlite::write_json(
    list(
      master_rows = nrow(base_result),
      status_counts = as.list(output_status_counts)
    ),
    parquet_validation_path,
    auto_unbox = TRUE,
    pretty = TRUE
  )
  validation <- system2(
    uv,
    c(
      "run",
      "--with", "pyarrow==21.0.0",
      "python", "ndjson_to_parquet.py", "validate",
      shQuote(output_path),
      shQuote(source_output_path),
      shQuote(openalex_batch_dir),
      shQuote(parquet_validation_path)
    ),
    stdout = TRUE,
    stderr = TRUE
  )
  if (!is.null(attr(validation, "status"))) {
    stop(paste(validation, collapse = "\n"), call. = FALSE)
  }
  parquet_validation <- jsonlite::fromJSON(tail(validation, 1L))

  output_rows <- parquet_contract$rows
  review_result <- build_master_chunk(sample_positions)
}

run_metadata <- list(
  producer = "analysis/ojs_journal_enrichment.qmd",
  run_mode = run_mode,
  created_at = format(Sys.time(), tz = "UTC", usetz = TRUE),
  input = list(
    version = pkp_version,
    dataverse_file_id = 14084919L,
    md5 = pkp_md5,
    rows = nrow(raw_input),
    valid_identifier_rows = sum(lengths(all_input_issns) > 0L),
    missing_identifier_rows = sum(missing_identifier),
    distinct_valid_issns = length(unique(unlist(all_input_issns)))
  ),
  output = list(
    file = basename(output_path),
    md5 = unname(tools::md5sum(output_path)),
    format = if (run_mode == "full") "parquet" else "csv",
    rows = output_rows,
    columns = length(output_columns),
    status_counts = as.list(output_status_counts)
  ),
  source_catalog = if (run_mode == "full") {
    list(
      file = basename(source_output_path),
      md5 = unname(tools::md5sum(source_output_path)),
      rows = source_parquet_contract$rows,
      columns = source_parquet_contract$columns,
      retained_top_level_fields = source_fields
    )
  } else {
    NULL
  },
  parquet_validation = if (run_mode == "full") parquet_validation else NULL,
  openalex = if (run_mode == "full") {
    list(
      batch_size = 100L,
      per_page = 200L,
      batches = length(batch_results),
      completed_batches = sum(map_chr(batch_results, "status") == "complete"),
      reused_batches = sum(map_lgl(batch_results, "reused")),
      pages = sum(map_int(batch_results, \(batch) length(batch$pages))),
      attempts = sum(map_int(batch_results, "attempts")),
      failed_batches = sum(map_chr(batch_results, "status") != "complete")
    )
  } else {
    list(
      cache_file = basename(openalex_cache_path),
      reused = openalex_cache_hit
    )
  },
  crossref = if (run_mode == "sample") {
    list(
      cache_file = basename(crossref_cache_path),
      reused = crossref_cache_hit
    )
  } else {
    list(status = "not_attempted")
  },
  runtime = list(
    r = R.version.string,
    quarto = system2("quarto", "--version", stdout = TRUE)[1],
    curl = system2(curl, "--version", stdout = TRUE)[1],
    pyarrow = if (run_mode == "full") {
      source_parquet_contract$pyarrow
    } else {
      "not used"
    },
    packages = as.list(set_names(
      vapply(
        required_packages,
        \(package) as.character(packageVersion(package)),
        character(1)
      ),
      required_packages
    ))
  )
)
jsonlite::write_json(
  run_metadata,
  metadata_path,
  auto_unbox = TRUE,
  pretty = TRUE
)
```

## Checks

These checks confirm that every input row appears in the master, every returned top-level API field remains in the lossless Source catalog, candidate IDs reconcile, credentials stay private, and full-mode pagination completed without unresolved API failures.

```{r}
#| label: contract-check

metadata_text <- paste(readLines(metadata_path, warn = FALSE), collapse = "\n")
artifact_root <- normalizePath(artifact_dir)

stopifnot(
  identical(pkp_version, 7L),
  identical(unname(tools::md5sum(input_path)), pkp_md5),
  nrow(raw_input) == 98273L,
  identical(names(raw_input), expected_schema),
  !anyDuplicated(raw_row_identity),
  sum(missing_identifier) == 26189L,
  sum(lengths(all_input_issns) > 0L) == 72084L,
  length(unique(unlist(all_input_issns))) == 103017L,
  startsWith(normalizePath(input_path), artifact_root),
  startsWith(normalizePath(output_path), artifact_root),
  startsWith(normalizePath(metadata_path), artifact_root),
  file.exists(metadata_path),
  !grepl("OPENALEX_API_KEY|CROSSREF_MAILTO", metadata_text),
  all(nzchar(unlist(run_metadata$runtime))),
  all(vapply(c(
    "normalize_issn",
    "is_missing_identifier",
    "row_identity",
    "match_exact_issns",
    "deduplicate_candidates",
    "index_candidates_by_issn",
    "candidates_for_issns",
    "expand_candidate_fields",
    "serialize_json"
  ), exists, logical(1), mode = "function")),
  exists("fetch_openalex_sources", mode = "function"),
  exists("fetch_crossref_journal", mode = "function"),
  output_rows == nrow(ojs_rows),
  file.exists(output_path),
  all(c(
    "openalex_match_status",
    "openalex_candidate_count",
    "openalex_candidate_ids"
  ) %in% output_columns),
  identical(
    base_result$openalex_candidate_count,
    map_int(openalex_matches, \(match) length(match$candidate_ids))
  ),
  all(base_result$openalex_match_status %in% c(
    "consistent",
    "inconsistent",
    "unmatched",
    "not_attempted",
    "api_error"
  )),
  sum(unlist(run_metadata$output$status_counts)) == output_rows,
  !"admin_email" %in% output_columns
)

if (run_mode == "sample") {
  expected_openalex_columns <- unique(unlist(map(
    openalex_matches,
    candidate_field_names,
    prefix = "openalex"
  )))
  expected_openalex_status <- c(
    "consistent",
    "consistent",
    "unmatched",
    "consistent",
    "consistent",
    "consistent",
    "consistent",
    "consistent",
    "not_attempted",
    "inconsistent"
  )
  expected_crossref_columns <- unique(unlist(map(
    crossref_matches,
    candidate_field_names,
    prefix = "crossref"
  )))
  stopifnot(
    identical(row_identity(ojs_rows), sample_identity),
    sum(result$identifier_status == "missing") == 1L,
    nrow(result) == 10L,
    all(c(
      "openalex_id",
      "openalex_summary_stats",
      "openalex_topics",
      "openalex_candidates__json",
      "openalex_candidate_2_id",
      "crossref_match_status",
      "crossref_title",
      "crossref_counts",
      "crossref_coverage",
      "crossref_candidates__json"
    ) %in% names(result)),
    all(expected_openalex_columns %in% names(result)),
    all(expected_crossref_columns %in% names(result)),
    identical(
      result$openalex_candidates__json,
      map_chr(openalex_matches, \(match) serialize_json(match$candidates))
    ),
    identical(
      result$crossref_candidates__json,
      map_chr(crossref_matches, \(match) serialize_json(match$candidates))
    ),
    identical(result$openalex_match_status, expected_openalex_status),
    result$openalex_matched_issns[
      result$context_name == "Jami Scientific Research Quarterly Journal"
    ] == "2710-4990",
    "not_attempted" %in% result$crossref_match_status
  )
} else {
  stopifnot(
    identical(row_identity(ojs_rows), raw_row_identity),
    output_rows == 98273L,
    sum(base_result$identifier_status == "missing") == 26189L,
    length(tokens) == 103017L,
    length(token_batches) == 1031L,
    max(lengths(token_batches)) <= 100L,
    all(map_lgl(batch_results, \(batch) {
      identical(batch$pkp_version, 7L) &&
        identical(batch$input_md5, pkp_md5) &&
        identical(batch$status, "complete") &&
        isTRUE(batch$pagination_complete) &&
        length(batch$pages) >= 1L
    })),
    run_metadata$openalex$completed_batches == length(token_batches),
    run_metadata$openalex$failed_batches == 0L,
    !any(base_result$openalex_match_status == "api_error"),
    identical(run_metadata$output$format, "parquet"),
    parquet_contract$columns == length(output_columns),
    source_parquet_contract$rows == length(openalex_sources),
    parquet_validation$master_rows == 98273L,
    parquet_validation$unique_identities == 98273L,
    parquet_validation$catalog_rows == length(openalex_sources),
    !anyDuplicated(map_chr(openalex_sources, "id")),
    all(unique(unlist(map(openalex_matches, "candidate_ids"))) %in%
      map_chr(openalex_sources, "id")),
    length(run_metadata$source_catalog$retained_top_level_fields) ==
      length(source_fields),
    file.exists(source_output_path),
    !file.exists(staging_path),
    !file.exists(source_staging_path)
  )
}
```

## Complete merged dataframe

This table displays all columns for the ten purposive review rows. In full mode the complete 98,273-row result stays in the master Parquet file while the report remains inspectable.

```{r}
#| label: complete-sample-dataframe

review_result |>
  knitr::kable(format = "html")
```

## Complete merged column list

This table lists every column in the master artifact so collaborators can review its structure.

```{r}
#| label: complete-column-list

tibble(column = output_columns) |>
  knitr::kable()
```
