---
title: "PKP/OJS multi-source journal enrichment"
format: html
execute:
  echo: true
  warning: true
  message: false
---

This notebook joins the pinned PKP OJS V7 release to retrieved OpenAlex Source
and Crossref journal snapshots by exact ISSN. Sample mode is the default. Full
mode must be started with `render_full.sh`; it preserves all 98,273 PKP rows and
keeps source candidate IDs in one compact compressed CSV. Complete source
objects remain in the OpenAlex checkpoints and Crossref cache.

Set `OPENALEX_API_KEY` before retrieving uncached OpenAlex data. Full mode also
requires `CROSSREF_MAILTO` when a Crossref page is missing. A cache-only rerun
needs neither value. A normal render and `render_sample.sh` always use sample
mode.

## Packages and paths

This chunk verifies the documented R environment, downloads the pinned PKP V7 input when needed, verifies its official checksum, and keeps every local artifact under this analysis owner.

```{r}
#| label: packages-and-paths

library(dplyr)
library(httr2)
library(purrr)
library(readr)
library(tibble)

analysis_packages <- c(
  "curl", "dplyr", "DT", "httr2", "jsonlite", "knitr", "purrr", "readr",
  "tibble"
)
run_mode <- Sys.getenv("OJS_ENRICHMENT_MODE", unset = "sample")
stopifnot(run_mode %in% c("sample", "full"))

project_env_path <- file.path("..", "..", "..", ".env")
if (file.exists(project_env_path)) {
  readRenviron(project_env_path)
}
crossref_mailto <- Sys.getenv("CROSSREF_MAILTO")

pkp_version <- 7L
pkp_url <- "https://dataverse.harvard.edu/api/access/datafile/14084919?format=original"
pkp_md5 <- "3a4ad8ae1ebfcc2b991aaf55b2d82c92"
artifact_dir <- file.path("..", "artifacts", "ojs_journal_enrichment")
input_dir <- file.path(artifact_dir, "pkp-v7")
input_path <- file.path(input_dir, "beacon.csv")
run_dir <- file.path(
  artifact_dir,
  if (run_mode == "full") "full-v7" else "sample"
)
output_path <- file.path(
  run_dir,
  if (run_mode == "full") {
    "pkp-ojs-multisource-enriched.csv.gz"
  } else {
    "pkp-ojs-multisource-enriched.csv"
  }
)
metadata_path <- file.path(run_dir, "run-metadata.json")
openalex_cache_path <- file.path(run_dir, "openalex-sources.rds")
crossref_cache_path <- file.path(run_dir, "crossref-journals.rds")
openalex_batch_dir <- file.path(run_dir, "openalex-batches")
crossref_page_dir <- file.path(run_dir, "crossref-pages")

dir.create(input_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(run_dir, recursive = TRUE, showWarnings = FALSE)
if (run_mode == "full") {
  dir.create(openalex_batch_dir, recursive = TRUE, showWarnings = FALSE)
  dir.create(crossref_page_dir, recursive = TRUE, showWarnings = FALSE)
}

download_pkp_input <- function(path) {
  if (file.exists(path)) return(invisible(path))

  download_path <- paste0(path, ".download")
  on.exit(unlink(download_path), add = TRUE)
  tryCatch(
    download.file(pkp_url, download_path, mode = "wb", quiet = FALSE),
    error = \(error) {
      stop("Could not download PKP V7.", call. = FALSE)
    }
  )
  if (unname(tools::md5sum(download_path)) != pkp_md5) {
    stop("Downloaded PKP V7 file failed checksum verification.", call. = FALSE)
  }
  if (!file.rename(download_path, path)) {
    stop("Could not move the verified PKP V7 file into place.", call. = FALSE)
  }
  invisible(path)
}

download_pkp_input(input_path)
if (unname(tools::md5sum(input_path)) != pkp_md5) {
  stop("Local PKP V7 file failed checksum verification.", call. = FALSE)
}
```

## ISSN preparation

ISSNs may contain spaces, punctuation, or more than one value. This chunk converts each valid ISSN to the same `XXXX-XXXX` format and checks its final digit. It keeps missing, invalid, and valid-but-unmatched ISSNs as separate cases.

```{r}
#| label: issn-preparation

normalize_issn <- function(value) {
  if (length(value) != 1L || is.na(value)) return(NA_character_)
  compact <- gsub("[^0-9X]", "", toupper(trimws(as.character(value))))
  if (!grepl("^[0-9]{7}[0-9X]$", compact)) return(NA_character_)

  digits <- as.integer(strsplit(substr(compact, 1, 7), "")[[1]])
  check_value <- (11 - sum(digits * 8:2) %% 11) %% 11
  expected <- if (check_value == 10) "X" else as.character(check_value)
  if (substr(compact, 8, 8) != expected) return(NA_character_)

  paste0(substr(compact, 1, 4), "-", substr(compact, 5, 8))
}

normalize_issn_set <- function(values) {
  normalized <- vapply(values, normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}

is_missing_value <- function(value) {
  is.na(value) |
    !nzchar(trimws(value)) |
    toupper(trimws(value)) == "NA"
}

parse_issns <- function(value) {
  if (is_missing_value(value)) return(character())

  tokens <- unlist(strsplit(value, "[[:space:],;|]+"))
  normalize_issn_set(tokens[nzchar(tokens)])
}

source_issns <- function(source) {
  normalize_issn_set(unlist(source$issn, use.names = FALSE))
}
```

## Source retrieval

Both modes use the same paginated OpenAlex retrieval. Full mode splits the
103,017 valid ISSNs into groups of at most 100 and caches each completed group.
It downloads the complete Crossref journal directory in sequential 1,000-record
cursor pages and caches the completed directory.

Sample mode asks Crossref about each ISSN separately and waits between requests. A Crossref `404` means that no journal record was found.

```{r}
#| label: openalex-retrieval

retryable_response <- function(response) {
  resp_status(response) %in% c(403L, 408L, 429L, 500L, 502L, 503L, 504L)
}

fetch_openalex_page <- function(
  issns,
  cursor,
  api_key = Sys.getenv("OPENALEX_API_KEY")
) {
  if (!nzchar(api_key)) {
    stop("Set OPENALEX_API_KEY before retrieving OpenAlex data.", call. = FALSE)
  }
  stopifnot(length(issns) <= 100L)

  body <- request("https://api.openalex.org/sources") |>
    req_url_query(
      filter = paste0("issn:", paste(issns, collapse = "|")),
      per_page = 200,
      cursor = cursor,
      api_key = api_key
    ) |>
    req_timeout(seconds = 30) |>
    req_retry(
      max_tries = 4L,
      retry_on_failure = TRUE,
      is_transient = retryable_response
    ) |>
    req_perform() |>
    resp_body_json(simplifyVector = FALSE)

  if (
    !is.list(body$results) ||
      !is.list(body$meta) ||
      length(body$meta$count) != 1L ||
      !(is.null(body$meta$next_cursor) ||
        length(body$meta$next_cursor) == 1L)
  ) {
    stop("OpenAlex response contract changed.", call. = FALSE)
  }
  body
}

fetch_openalex_batch <- function(issns) {
  cursor <- "*"
  seen_cursors <- character()
  sources <- list()
  total_results <- NULL

  repeat {
    if (cursor %in% seen_cursors) {
      stop("OpenAlex repeated a cursor.", call. = FALSE)
    }
    seen_cursors <- c(seen_cursors, cursor)

    page <- fetch_openalex_page(issns, cursor)
    if (is.null(total_results)) {
      total_results <- as.integer(page$meta$count)
    }
    next_cursor <- page$meta$next_cursor
    sources <- c(sources, page$results)
    if (is.null(next_cursor) || !nzchar(next_cursor)) break
    cursor <- next_cursor
  }

  if (length(sources) != total_results) {
    stop("OpenAlex result count did not reconcile.", call. = FALSE)
  }
  sources
}
```

### Crossref retrieval

```{r}
#| label: crossref-retrieval

save_rds_atomic <- function(value, path) {
  temporary <- paste0(path, ".tmp")
  saveRDS(value, temporary)
  if (!file.rename(temporary, path)) {
    unlink(temporary)
    stop("Could not save checkpoint atomically.", call. = FALSE)
  }
  invisible(path)
}

read_rds_or_null <- function(path) {
  if (!file.exists(path)) return(NULL)
  tryCatch(readRDS(path), error = \(error) NULL)
}

file_set_md5 <- function(paths) {
  manifest <- tempfile()
  on.exit(unlink(manifest), add = TRUE)
  paths <- sort(paths)
  writeLines(paste(basename(paths), unname(tools::md5sum(paths))), manifest)
  unname(tools::md5sum(manifest))
}

fetch_crossref_journal <- function(issn) {
  response <- request(paste0("https://api.crossref.org/journals/", issn)) |>
    req_user_agent(
      "InvisibleResearch journal enrichment; https://github.com/YannJY02/InvisibleResearch"
    ) |>
    req_timeout(seconds = 30) |>
    req_retry(
      max_tries = 4L,
      retry_on_failure = TRUE,
      is_transient = retryable_response
    ) |>
    req_error(
      is_error = \(response) {
        resp_status(response) != 404L && resp_is_error(response)
      }
    ) |>
    req_perform()
  if (resp_status(response) == 404L) return(NULL)
  resp_body_json(response, simplifyVector = FALSE)$message
}

fetch_crossref_page <- function(cursor, mailto) {
  stopifnot(nzchar(cursor), nzchar(mailto))

  body <- request("https://api.crossref.org/journals") |>
    req_url_query(rows = 1000, cursor = cursor, mailto = mailto) |>
    req_user_agent(
      "InvisibleResearch journal enrichment; https://github.com/YannJY02/InvisibleResearch"
    ) |>
    req_timeout(seconds = 60) |>
    req_retry(
      max_tries = 4L,
      retry_on_failure = TRUE,
      is_transient = retryable_response
    ) |>
    req_perform() |>
    resp_body_json(simplifyVector = FALSE) |>
    pluck("message")

  if (!is.list(body$items) || length(body[["total-results"]]) != 1L) {
    stop("Crossref response contract changed.", call. = FALSE)
  }
  body
}

fetch_crossref_directory <- function() {
  cursor <- "*"
  journals <- list()
  previous_items <- NULL
  total_results <- NULL
  page_index <- 1L

  repeat {
    checkpoint_path <- file.path(
      crossref_page_dir,
      sprintf("page-%04d.rds", page_index)
    )
    checkpoint <- read_rds_or_null(checkpoint_path)
    reusable <- is.list(checkpoint) &&
      identical(checkpoint$cursor, cursor) &&
      length(checkpoint$total_results) == 1L &&
      is.list(checkpoint$items)

    if (!reusable) {
      if (!grepl("^[^@[:space:]]+@[^@[:space:]]+$", crossref_mailto)) {
        stop("Set CROSSREF_MAILTO before retrieving Crossref data.", call. = FALSE)
      }
      page <- fetch_crossref_page(cursor, crossref_mailto)
      checkpoint <- list(
        cursor = cursor,
        total_results = as.integer(page[["total-results"]]),
        next_cursor = page[["next-cursor"]],
        items = page$items,
        retrieved_at = format(Sys.time(), tz = "UTC", usetz = TRUE)
      )
      save_rds_atomic(checkpoint, checkpoint_path)
    }

    if (!is.null(previous_items) && identical(checkpoint$items, previous_items)) {
      stop("Crossref repeated a page.", call. = FALSE)
    }
    if (is.null(total_results)) {
      total_results <- checkpoint$total_results
    } else if (!identical(total_results, checkpoint$total_results)) {
      stop("Crossref total-results changed during pagination.", call. = FALSE)
    }
    journals <- c(journals, checkpoint$items)
    previous_items <- checkpoint$items
    if (length(journals) > total_results) {
      stop("Crossref pagination exceeded total-results.", call. = FALSE)
    }

    if (length(checkpoint$items) < 1000L) break

    cursor <- checkpoint$next_cursor
    if (is.null(cursor) || !nzchar(cursor)) {
      stop("Crossref returned a full page without a next cursor.", call. = FALSE)
    }
    page_index <- page_index + 1L
    if (!reusable) Sys.sleep(0.2)
  }

  if (length(journals) != total_results) {
    stop(
      paste0(
        "Crossref directory row count mismatch: ",
        length(journals), " != ", total_results
      ),
      call. = FALSE
    )
  }
  list(journals = journals, total_results = total_results)
}

```

## Load the PKP V7 rows

The pinned PKP V7 file is Dataverse file `14084919`, released on 2026-07-18. This chunk checks the complete 98,273-row baseline, schema, row identity, and identifier counts before either retaining the full cohort or selecting ten fixed source rows in a deliberate display order.

```{r}
#| label: load-pkp-v7

raw_input <- read_csv(
  input_path,
  col_types = cols(.default = col_character()),
  na = character(),
  show_col_types = FALSE
)

expected_schema <- c(
  "context_name",
  "issn",
  "oai_url",
  "application",
  "country",
  "region",
  "total_record_count",
  "record_count_2025",
  "earliest_datestamp",
  "repository_name",
  "set_spec",
  "stats_id",
  "best_doaj_url",
  "last_completed_update",
  "first_beacon",
  "last_beacon",
  "last_oai_response",
  "unresponsive_endpoint",
  "unresponsive_context"
)
row_identity <- function(data) {
  paste(data$oai_url, data$repository_name, data$set_spec, sep = "\r")
}
raw_row_identity <- row_identity(raw_input)

stopifnot(
  nrow(raw_input) == 98273L,
  identical(names(raw_input), expected_schema),
  all(tolower(raw_input$application) == "ojs"),
  !anyDuplicated(raw_row_identity)
)

all_input_issns <- map(raw_input$issn, parse_issns)
missing_identifier <- is_missing_value(raw_input$issn)
stopifnot(
  sum(missing_identifier) == 26189L,
  sum(!is_missing_value(raw_input$best_doaj_url)) == 7149L,
  sum(lengths(all_input_issns) > 0L) == 72084L,
  length(unique(unlist(all_input_issns))) == 103017L
)

sample_spec <- tribble(
  ~context_name, ~oai_url, ~repository_name, ~set_spec,
  "Kuwait Journal of Science",
  "https://journalskuwait.org/kjs/index.php/index/oai",
  "Kuwait Journal of Science",
  "KJS",
  "Cakrawala",
  "https://cakrawalajournal.org/index.php/index/oai",
  "CAKRAWALA",
  "cakrawala",
  "Sintesa: Jurnal Ilmu Pendidikan",
  "https://sintesa.stkip-arrahmaniyah.ac.id/index.php/index/oai",
  "Sintesa: Jurnal Ilmu Pendidikan STKIP Arrahmaniyah Depok",
  "sintesa",
  "CARAKA: Jurnal Teologi Biblika dan Praktika",
  "https://ojs.sttibc.ac.id/index.php/index/oai",
  "CARAKA:Jurnal Teologi",
  "ibc",
  "Jurnal Tata Kelola dan Akuntabilitas Keuangan Negara",
  "https://jurnal.bpk.go.id/index.php/index/oai",
  "Open Journal Systems",
  "TAKEN",
  "JOEEL (Journal of English Education and Literature)",
  "https://journal.stkippamanetalino.ac.id/index.php/index/oai",
  "Rumah Jurnal STKIP Pamane Talino Ngabang",
  "bahasa-inggris",
  "Malaysian Journal of Paediatrics and Child Health",
  "https://mpaeds.my/journals/index.php/index/oai",
  "NA",
  "MJPCH",
  "Jami Scientific Research Quarterly Journal",
  "https://journals.jami.edu.af/index.php/index/oai",
  "Jami University Press",
  "jsrqj",
  "Journal of Polymer & Composites",
  "https://engineeringjournals.stmjournals.in/index.php/index/oai",
  "Engineering Journals",
  "JoPC",
  "JURNAL TEKNIK PERTAMBANGAN",
  "https://e-journal.upr.ac.id/index.php/index/oai",
  "Journal Online Universitas Palangka Raya",
  "JTP"
)
sample_identity <- row_identity(sample_spec)
sample_positions <- match(sample_identity, raw_row_identity)
stopifnot(!anyNA(sample_positions))

sample_rows <- raw_input[sample_positions, ] |>
  select(-any_of("admin_email"))
stopifnot(
  nrow(sample_rows) == 10L,
  identical(sample_rows$context_name, sample_spec$context_name)
)

ojs_rows <- if (run_mode == "full") {
  raw_input |> select(-any_of("admin_email"))
} else {
  sample_rows
}
```

## Multi-source enrichment

For each PKP row, the notebook looks for OpenAlex and Crossref records with the
same checksum-valid normalized ISSN. Sample mode expands complete candidates
for review. Full mode keeps only candidate identities in the compact master;
complete records remain in the source caches.

### Matching helpers

These helpers label each row as having no usable ISSN, no match, one match, or multiple matches. They also identify Crossref records by their ISSNs and convert every candidate field into a dataframe-safe value without filtering the response schema.

```{r}
#| label: matching-helpers

classify_status <- function(input_issns, candidate_ids) {
  case_when(
    length(input_issns) == 0L ~ "not_attempted",
    length(candidate_ids) == 0L ~ "unmatched",
    length(candidate_ids) == 1L ~ "unique",
    TRUE ~ "ambiguous"
  )
}

crossref_issns <- function(journal) {
  normalize_issn_set(journal$ISSN)
}

crossref_key <- function(journal) {
  paste(sort(crossref_issns(journal)), collapse = "|")
}

deduplicate_candidates <- function(candidates, identity) {
  candidate_ids <- vapply(candidates, identity, character(1))
  keep <- !duplicated(candidate_ids)
  list(
    candidates = candidates[keep],
    candidate_ids = candidate_ids[keep]
  )
}

index_candidates_by_issn <- function(candidates, candidate_issns) {
  index <- new.env(hash = TRUE, parent = emptyenv())
  for (candidate_position in seq_along(candidates)) {
    for (issn in candidate_issns(candidates[[candidate_position]])) {
      positions <- get0(issn, index, inherits = FALSE, ifnotfound = integer())
      assign(issn, c(positions, candidate_position), index)
    }
  }
  index
}

candidates_for_issns <- function(candidates, index, issns) {
  positions <- unique(unlist(map(issns, \(issn) {
    get0(issn, index, inherits = FALSE, ifnotfound = integer())
  })))
  candidates[positions]
}

match_indexed_ids <- function(input_issns, candidate_ids, index, issns_by_id) {
  matched_ids <- candidates_for_issns(candidate_ids, index, input_issns)
  candidate_issns <- unique(unlist(issns_by_id[matched_ids]))
  list(
    status = classify_status(input_issns, matched_ids),
    matched_issns = base::intersect(input_issns, candidate_issns),
    candidate_ids = matched_ids
  )
}

match_exact_issns <- function(
  input_issns,
  candidates,
  candidate_issns,
  candidate_identity,
  deduplicate = TRUE
) {
  matched <- keep(candidates, \(candidate) {
    any(input_issns %in% candidate_issns(candidate))
  })
  resolved <- if (deduplicate) {
    deduplicate_candidates(matched, candidate_identity)
  } else {
    list(
      candidates = matched,
      candidate_ids = map_chr(matched, candidate_identity)
    )
  }
  matched_issns <- base::intersect(
    input_issns,
    unique(unlist(map(resolved$candidates, candidate_issns)))
  )

  c(
    list(
      status = classify_status(
        input_issns,
        resolved$candidate_ids
      ),
      matched_issns = matched_issns
    ),
    resolved
  )
}

serialize_json <- function(value) {
  as.character(jsonlite::toJSON(
    value,
    auto_unbox = TRUE,
    null = "null",
    na = "null",
    digits = NA
  ))
}

dataframe_value <- function(value) {
  if (is.null(value)) return(NA_character_)
  if (is.atomic(value) && length(value) == 1L) {
    return(as.character(value))
  }
  serialize_json(value)
}

expand_candidate_fields <- function(matches, prefix, fields) {
  candidates <- map(
    matches,
    \(match) {
      if (length(match$candidates) == 1L) match$candidates[[1]] else list()
    }
  )
  cells <- set_names(
    map(fields, \(field) {
      map_chr(candidates, \(candidate) {
        dataframe_value(candidate[[field]])
      })
    }),
    paste(prefix, fields, sep = "_")
  )
  cells[[paste0(prefix, "_candidates__json")]] <- map_chr(
    matches,
    \(match) {
      if (length(match$candidates) > 1L) {
        serialize_json(match$candidates)
      } else {
        NA_character_
      }
    }
  )
  as_tibble(cells, .name_repair = "check_unique")
}

```

### OpenAlex enrichment

Both modes use the same exact-ISSN rule. Full mode indexes candidates by normalized ISSN so matching 98,273 rows does not repeatedly scan either complete response set.

```{r}
#| label: openalex-enrichment

input_issns <- map(ojs_rows$issn, parse_issns)
tokens <- sort(unique(unlist(input_issns)))

if (run_mode == "full") {
  token_batches <- split(tokens, ceiling(seq_along(tokens) / 100L))
  # ponytail: retrieval is sequential because it is an occasional research run;
  # add bounded parallelism only if fresh full retrieval becomes routine.
  batch_results <- imap(token_batches, \(batch, batch_index) {
    checkpoint_path <- file.path(
      openalex_batch_dir,
      sprintf("batch-%04d.rds", as.integer(batch_index))
    )
    checkpoint <- read_rds_or_null(checkpoint_path)
    if (
      is.list(checkpoint) &&
      identical(checkpoint$issns, batch) &&
        is.list(checkpoint$sources)
    ) {
      return(checkpoint)
    }

    checkpoint <- list(
      issns = batch,
      retrieved_at = format(Sys.time(), tz = "UTC", usetz = TRUE),
      sources = fetch_openalex_batch(batch)
    )
    save_rds_atomic(checkpoint, checkpoint_path)
    checkpoint
  })
  openalex_sources <- unlist(
    map(batch_results, "sources"),
    recursive = FALSE
  )
} else {
  openalex_cache <- read_rds_or_null(openalex_cache_path)
  openalex_cache_is_valid <- is.list(openalex_cache) &&
    identical(openalex_cache$issns, tokens) &&
    is.list(openalex_cache$sources) &&
    length(openalex_cache$retrieved_at) == 1L
  if (openalex_cache_is_valid) {
    openalex_result <- openalex_cache
  } else {
    openalex_result <- list(
      issns = tokens,
      retrieved_at = format(Sys.time(), tz = "UTC", usetz = TRUE),
      sources = fetch_openalex_batch(tokens)
    )
    save_rds_atomic(openalex_result, openalex_cache_path)
  }
  openalex_sources <- openalex_result$sources
}
openalex_sources <- deduplicate_candidates(
  openalex_sources,
  \(source) source$id
)$candidates
if (run_mode == "full") {
  openalex_candidate_ids <- map_chr(openalex_sources, "id")
  openalex_issns_by_id <- set_names(
    map(openalex_sources, source_issns),
    openalex_candidate_ids
  )
  openalex_index <- index_candidates_by_issn(
    openalex_candidate_ids,
    \(candidate_id) openalex_issns_by_id[[candidate_id]]
  )
  openalex_matches <- map(
    input_issns,
    match_indexed_ids,
    candidate_ids = openalex_candidate_ids,
    index = openalex_index,
    issns_by_id = openalex_issns_by_id
  )
} else {
  openalex_index <- index_candidates_by_issn(
    openalex_sources,
    source_issns
  )
  openalex_matches <- map(input_issns, \(issns) {
    match_exact_issns(
      issns,
      candidates_for_issns(openalex_sources, openalex_index, issns),
      source_issns,
      \(source) source$id
    )
  })
}
openalex_status_counts <- table(map_chr(openalex_matches, "status"))

base_result <- ojs_rows |>
  mutate(
    identifier_status = case_when(
      is_missing_value(issn) ~ "missing",
      lengths(input_issns) == 0L ~ "invalid",
      TRUE ~ "valid"
    ),
    openalex_input_issns = map_chr(input_issns, paste, collapse = "|"),
    openalex_matched_issns = map_chr(
      openalex_matches,
      \(match) paste(match$matched_issns, collapse = "|")
    ),
    openalex_match_status = map_chr(openalex_matches, "status"),
    openalex_candidate_count = map_int(
      openalex_matches,
      \(match) length(match$candidate_ids)
    ),
    openalex_candidate_ids = map_chr(
      openalex_matches,
      \(match) paste(match$candidate_ids, collapse = "|")
    )
  )
```

### Crossref enrichment and compact master

Crossref is matched by the same normalized exact ISSNs. Sample mode expands
complete fields for the ten-row review. Full mode keeps one row per PKP record
with candidate IDs that resolve to complete objects in the source caches.

```{r}
#| label: crossref-enrichment

crossref_cache <- read_rds_or_null(crossref_cache_path)
crossref_cache_is_valid <- is.list(crossref_cache) &&
  is.list(crossref_cache$journals) &&
  length(crossref_cache$retrieved_at) == 1L &&
  if (run_mode == "full") {
    identical(crossref_cache$total_results, length(crossref_cache$journals))
  } else {
    identical(crossref_cache$issns, tokens)
  }

if (crossref_cache_is_valid) {
  crossref_result <- crossref_cache
  rm(crossref_cache)
} else {
  crossref_result <- if (run_mode == "sample") {
    list(
      issns = tokens,
      retrieved_at = format(Sys.time(), tz = "UTC", usetz = TRUE),
      journals = compact(map(tokens, \(token) {
        journal <- fetch_crossref_journal(token)
        Sys.sleep(0.25)
        journal
      }))
    )
  } else {
    c(
      list(retrieved_at = format(Sys.time(), tz = "UTC", usetz = TRUE)),
      fetch_crossref_directory()
    )
  }
  save_rds_atomic(crossref_result, crossref_cache_path)
  if (run_mode == "full") unlink(crossref_page_dir, recursive = TRUE)
}

if (run_mode == "sample") {
  crossref_candidates <- deduplicate_candidates(
    unname(compact(crossref_result$journals)),
    serialize_json
  )$candidates
} else {
  candidate_issns <- map(crossref_result$journals, crossref_issns)
  candidate_ids <- map_chr(
    candidate_issns,
    \(issns) paste(sort(issns), collapse = "|")
  )
  keep_candidate <- nzchar(candidate_ids) & !duplicated(candidate_ids)
  crossref_candidate_ids <- candidate_ids[keep_candidate]
  crossref_issns_by_id <- set_names(
    candidate_issns[keep_candidate],
    crossref_candidate_ids
  )
  rm(crossref_result, candidate_issns, candidate_ids, keep_candidate)
  gc()
}

if (run_mode == "sample") {
  crossref_index <- index_candidates_by_issn(
    crossref_candidates,
    crossref_issns
  )
  crossref_matches <- map(input_issns, \(issns) {
    match_exact_issns(
      issns,
      candidates_for_issns(
        crossref_candidates,
        crossref_index,
        issns
      ),
      crossref_issns,
      crossref_key
    )
  })
} else {
  crossref_index <- index_candidates_by_issn(
    crossref_candidate_ids,
    \(candidate_id) crossref_issns_by_id[[candidate_id]]
  )
  crossref_matches <- map(
    input_issns,
    match_indexed_ids,
    candidate_ids = crossref_candidate_ids,
    index = crossref_index,
    issns_by_id = crossref_issns_by_id
  )
}
crossref_status_counts <- table(map_chr(crossref_matches, "status"))

result <- base_result |>
  mutate(
    crossref_input_issns = map_chr(input_issns, paste, collapse = "|"),
    crossref_matched_issns = map_chr(
      crossref_matches,
      \(match) paste(match$matched_issns, collapse = "|")
    ),
    crossref_match_status = map_chr(crossref_matches, "status"),
    crossref_candidate_count = map_int(
      crossref_matches,
      \(match) length(match$candidate_ids)
    ),
    crossref_candidate_issn_sets = map_chr(
      crossref_matches,
      \(match) paste(match$candidate_ids, collapse = ";")
    )
  )

if (run_mode == "sample") {
  openalex_fields <- sort(unique(unlist(map(openalex_sources, names))))
  crossref_fields <- sort(unique(unlist(map(
    crossref_candidates,
    names
  ))))
  openalex_data <- expand_candidate_fields(
    openalex_matches,
    prefix = "openalex",
    fields = openalex_fields
  )
  crossref_data <- expand_candidate_fields(
    crossref_matches,
    prefix = "crossref",
    fields = crossref_fields
  )
  result <- bind_cols(result, openalex_data, crossref_data)
}

output_data <- result
output_rows <- nrow(result)
output_columns <- names(result)
```

### Validate and promote

The master is written to a temporary file, read back, and promoted only after
its rows, match states, and every serialized cell pass.

```{r}
#| label: pre-promotion-validation

candidate_json_is_valid <- function(matches, json_values) {
  map2_lgl(matches, json_values, \(match, json_value) {
    candidate_count <- length(match$candidates)
    if (candidate_count <= 1L) return(is.na(json_value))
    parsed <- tryCatch(
      jsonlite::fromJSON(json_value, simplifyVector = FALSE),
      error = \(error) NULL
    )
    is.list(parsed) && length(parsed) == candidate_count
  })
}

stopifnot(
  !"admin_email" %in% output_columns,
  output_rows == nrow(ojs_rows),
  all(c(
    "openalex_match_status",
    "openalex_candidate_count",
    "openalex_candidate_ids",
    "crossref_match_status",
    "crossref_candidate_count",
    "crossref_candidate_issn_sets"
  ) %in% output_columns),
  sum(openalex_status_counts) == output_rows,
  sum(crossref_status_counts) == output_rows,
  identical(
    result$openalex_match_status,
    map2_chr(input_issns, openalex_matches, \(issns, match) {
      classify_status(issns, match$candidate_ids)
    })
  ),
  identical(
    result$crossref_match_status,
    map2_chr(input_issns, crossref_matches, \(issns, match) {
      classify_status(issns, match$candidate_ids)
    })
  )
)

if (run_mode == "sample") {
  stopifnot(
    identical(row_identity(ojs_rows), sample_identity),
    sum(result$identifier_status == "missing") == 1L,
    nrow(result) == 10L,
    !anyDuplicated(map_chr(crossref_candidates, serialize_json)),
    all(paste("openalex", openalex_fields, sep = "_") %in% output_columns),
    all(paste("crossref", crossref_fields, sep = "_") %in% output_columns),
    all(candidate_json_is_valid(
      openalex_matches,
      output_data$openalex_candidates__json
    )),
    all(candidate_json_is_valid(
      crossref_matches,
      output_data$crossref_candidates__json
    ))
  )
} else {
  stopifnot(
    identical(row_identity(ojs_rows), raw_row_identity),
    output_rows == 98273L,
    sum(base_result$identifier_status == "missing") == 26189L,
    length(tokens) == 103017L,
    length(token_batches) == 1031L,
    max(lengths(token_batches)) <= 100L,
    all(result$crossref_match_status %in% c(
      "unique",
      "ambiguous",
      "unmatched",
      "not_attempted"
    )),
    !any(endsWith(output_columns, "candidates__json")),
    endsWith(output_path, ".csv.gz")
  )
}

temporary_output_path <- paste0(
  output_path,
  ".tmp",
  if (endsWith(output_path, ".gz")) ".gz" else ""
)
unlink(temporary_output_path)
write_csv(output_data, temporary_output_path, na = "")

observed_output <- read_csv(
  temporary_output_path,
  col_types = cols(.default = col_character()),
  na = character(),
  show_col_types = FALSE
)
expected_output <- as_tibble(map(output_data, \(column) {
  values <- as.character(column)
  values[is.na(values)] <- ""
  values
}))
observed_identity <- row_identity(observed_output)
stopifnot(
  identical(as.data.frame(observed_output), as.data.frame(expected_output)),
  identical(observed_identity, row_identity(ojs_rows)),
  !anyDuplicated(observed_identity)
)

openalex_batch_files <- if (run_mode == "full") {
  list.files(
    openalex_batch_dir,
    pattern = "^batch-[0-9]{4}\\.rds$",
    full.names = TRUE
  )
} else {
  character()
}
run_metadata <- list(
  run_mode = run_mode,
  input = list(
    version = pkp_version,
    md5 = pkp_md5
  ),
  output = list(
    file = basename(output_path),
    md5 = unname(tools::md5sum(temporary_output_path)),
    rows = output_rows,
    columns = length(output_columns)
  ),
  openalex_checkpoints = if (run_mode == "full") {
    list(
      count = length(openalex_batch_files),
      md5 = file_set_md5(openalex_batch_files)
    )
  } else {
    NULL
  },
  runtime = list(
    r = R.version.string,
    quarto = system2("quarto", "--version", stdout = TRUE)[1],
    curl = curl::curl_version()$version,
    packages = as.list(setNames(
      vapply(analysis_packages, \(package) {
        as.character(packageVersion(package))
      }, character(1)),
      analysis_packages
    ))
  )
)

metadata_temporary_path <- paste0(metadata_path, ".tmp")
unlink(metadata_temporary_path)
jsonlite::write_json(
  run_metadata,
  metadata_temporary_path,
  auto_unbox = TRUE,
  pretty = TRUE
)
promote_artifact_pair <- function(temporary_paths, final_paths) {
  backup_paths <- paste0(final_paths, ".previous")
  if (any(file.exists(backup_paths))) {
    stop("Resolve stale .previous artifacts before promotion.", call. = FALSE)
  }

  backed_up <- promoted <- rep(FALSE, length(final_paths))
  complete <- FALSE
  on.exit({
    if (!complete) {
      unlink(final_paths[promoted])
      for (position in which(backed_up)) {
        file.rename(backup_paths[position], final_paths[position])
      }
    }
  }, add = TRUE)

  for (position in seq_along(final_paths)) {
    if (file.exists(final_paths[position])) {
      backed_up[position] <- file.rename(
        final_paths[position],
        backup_paths[position]
      )
      if (!backed_up[position]) stop("Could not back up an artifact.")
    }
    if (!file.rename(temporary_paths[position], final_paths[position])) {
      stop("Could not promote validated artifacts.", call. = FALSE)
    }
    promoted[position] <- TRUE
  }
  complete <- TRUE
  unlink(backup_paths)
}

promote_artifact_pair(
  c(temporary_output_path, metadata_temporary_path),
  c(output_path, metadata_path)
)
```

## Ten-row validation review

```{r}
#| label: validation-review

review_result <- if (run_mode == "full") result[sample_positions, ] else result
review_result |>
  select(
    context_name,
    issn,
    identifier_status,
    openalex_match_status,
    openalex_candidate_count,
    crossref_match_status,
    crossref_candidate_count
  ) |>
  DT::datatable(
    filter = "top",
    rownames = FALSE,
    options = list(pageLength = 10)
  )

tibble(column = output_columns) |>
  knitr::kable()
```
