---
title: "PKP/OJS multi-source journal enrichment"
format: html
execute:
  echo: true
  warning: false
  message: false
---

This notebook adds current OpenAlex Source and Crossref journal data to PKP OJS rows from the pinned V7 release by exact ISSN. Sample mode is the default. Full mode must be started with `render_full.sh`; it checkpoints both source directories and preserves all 98,273 PKP rows in a compact Parquet master.

Set `OPENALEX_API_KEY` in the environment or the repository-root `.env` file before rendering. Full mode also requires `CROSSREF_MAILTO` as process configuration. A normal render and `render_sample.sh` always use sample mode.

## Packages and paths

This chunk installs only missing CRAN packages, loads them, downloads the pinned PKP V7 input when needed, verifies its official checksum, and keeps every local artifact under this analysis owner.

```{r}
#| label: packages-and-paths

required_packages <- c(
  "countrycode",
  "dplyr",
  "ggplot2",
  "httr2",
  "jsonlite",
  "knitr",
  "purrr",
  "readr",
  "reactable",
  "rmarkdown",
  "tibble"
)
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]
if (length(missing_packages)) {
  install.packages(missing_packages, repos = "https://cloud.r-project.org")
}

library(dplyr)
library(ggplot2)
library(httr2)
library(purrr)
library(readr)
library(tibble)

run_mode <- Sys.getenv("OJS_ENRICHMENT_MODE", unset = "sample")
stopifnot(run_mode %in% c("sample", "full"))

project_env_path <- file.path("..", "..", "..", ".env")
if (!nzchar(Sys.getenv("OPENALEX_API_KEY")) && file.exists(project_env_path)) {
  readRenviron(project_env_path)
}
if (!nzchar(Sys.getenv("OPENALEX_API_KEY"))) {
  stop("Set OPENALEX_API_KEY before running this notebook.", call. = FALSE)
}
crossref_mailto <- Sys.getenv("CROSSREF_MAILTO")
if (
  run_mode == "full" &&
    !grepl("^[^@[:space:]]+@[^@[:space:]]+$", crossref_mailto)
) {
  stop(
    "Set CROSSREF_MAILTO to a contact email before a full render.",
    call. = FALSE
  )
}

pkp_version <- 7L
pkp_url <- paste0(
  "https://dataverse.harvard.edu/api/access/datafile/14084919",
  "?format=original"
)
pkp_md5 <- "3a4ad8ae1ebfcc2b991aaf55b2d82c92"
artifact_dir <- file.path("..", "artifacts", "ojs_journal_enrichment")
input_dir <- file.path(artifact_dir, "pkp-v7")
input_path <- file.path(input_dir, "beacon.csv")
run_dir <- file.path(
  artifact_dir,
  if (run_mode == "full") "full-v7" else "sample"
)
output_path <- file.path(
  run_dir,
  if (run_mode == "full") {
    "pkp-ojs-multisource-enriched.parquet"
  } else {
    "pkp-ojs-multisource-enriched.csv"
  }
)
metadata_path <- file.path(run_dir, "run-metadata.json")
audit_path <- file.path(run_dir, "pkp-openalex-disagreement-audit.csv")
audit_summary_path <- file.path(
  run_dir,
  "pkp-openalex-disagreement-summary.csv"
)
staging_path <- file.path(run_dir, "pkp-ojs-multisource-enriched.ndjson")
parquet_schema_path <- file.path(run_dir, "parquet-schema.json")
source_output_path <- file.path(run_dir, "openalex-sources.parquet")
source_staging_path <- file.path(run_dir, "openalex-sources.ndjson")
source_schema_path <- file.path(run_dir, "openalex-sources-schema.json")
crossref_output_path <- file.path(run_dir, "crossref-journals.parquet")
crossref_staging_path <- file.path(run_dir, "crossref-journals.ndjson")
crossref_schema_path <- file.path(run_dir, "crossref-journals-schema.json")
parquet_validation_path <- file.path(run_dir, "parquet-validation.json")
openalex_cache_path <- file.path(run_dir, "openalex-sources.rds")
crossref_cache_path <- file.path(run_dir, "crossref-journals.rds")
openalex_batch_dir <- file.path(run_dir, "openalex-batches")
crossref_page_dir <- file.path(run_dir, "crossref-pages")

dir.create(input_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(run_dir, recursive = TRUE, showWarnings = FALSE)
if (run_mode == "full") {
  dir.create(openalex_batch_dir, recursive = TRUE, showWarnings = FALSE)
  dir.create(crossref_page_dir, recursive = TRUE, showWarnings = FALSE)
}

curl <- Sys.which("curl")
if (!nzchar(curl)) {
  stop("curl is required to download the pinned PKP V7 input.", call. = FALSE)
}

download_pkp_input <- function(path) {
  if (file.exists(path)) return(invisible(path))

  download_path <- paste0(path, ".download")
  on.exit(unlink(download_path), add = TRUE)
  status <- system2(
    curl,
    c(
      "-fL",
      "--retry", "5",
      "--retry-all-errors",
      "--retry-delay", "2",
      "--continue-at", "-",
      "-o", shQuote(download_path),
      shQuote(pkp_url)
    )
  )
  if (status != 0L) {
    stop("Could not download PKP V7.", call. = FALSE)
  }
  if (unname(tools::md5sum(download_path)) != pkp_md5) {
    stop("Downloaded PKP V7 file failed checksum verification.", call. = FALSE)
  }
  if (!file.rename(download_path, path)) {
    stop("Could not move the verified PKP V7 file into place.", call. = FALSE)
  }
  invisible(path)
}

download_pkp_input(input_path)
if (unname(tools::md5sum(input_path)) != pkp_md5) {
  stop("Local PKP V7 file failed checksum verification.", call. = FALSE)
}
```

## ISSN preparation

ISSNs may contain spaces, punctuation, or more than one value. This chunk converts each valid ISSN to the same `XXXX-XXXX` format and checks its final digit. It keeps missing, invalid, and valid-but-unmatched ISSNs as separate cases.

```{r}
#| label: issn-preparation

normalize_issn <- function(value) {
  if (length(value) != 1L || is.na(value)) return(NA_character_)
  compact <- gsub("[^0-9X]", "", toupper(trimws(as.character(value))))
  if (!grepl("^[0-9]{7}[0-9X]$", compact)) return(NA_character_)

  digits <- as.integer(strsplit(substr(compact, 1, 7), "")[[1]])
  check_value <- (11 - sum(digits * 8:2) %% 11) %% 11
  expected <- if (check_value == 10) "X" else as.character(check_value)
  if (substr(compact, 8, 8) != expected) return(NA_character_)

  paste0(substr(compact, 1, 4), "-", substr(compact, 5, 8))
}

is_missing_identifier <- function(value) {
  is.na(value) |
    !nzchar(trimws(value)) |
    toupper(trimws(value)) == "NA"
}

parse_issns <- function(value) {
  if (is_missing_identifier(value)) return(character())

  tokens <- unlist(strsplit(value, "[[:space:],;|]+"))
  normalized <- vapply(tokens[nzchar(tokens)], normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}

source_issns <- function(source) {
  values <- c(unlist(source$issn, use.names = FALSE), source$issn_l)
  normalized <- vapply(values, normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}
```

## Source retrieval

The sample sends all valid ISSNs to OpenAlex in one request. Full mode splits the 103,017 valid ISSNs into deterministic groups of at most 100, requests the maximum page size, follows every cursor, and atomically checkpoints each completed V7 batch. It also downloads the complete Crossref journals directory in sequential 1,000-record cursor pages, sleeping between requests and checkpointing each page. Failed responses remain explicit and are retried on the next run.

Sample mode asks Crossref about each ISSN separately and waits between requests. A Crossref `404` means that no journal record was found.

```{r}
#| label: source-retrieval

fetch_openalex_sources <- function(
  issns,
  api_key = Sys.getenv("OPENALEX_API_KEY")
) {
  stopifnot(length(issns) <= 100L, nzchar(api_key))

  request("https://api.openalex.org/sources") |>
    req_url_query(
      filter = paste0("issn:", paste(issns, collapse = "|")),
      per_page = 100,
      api_key = api_key
    ) |>
    req_timeout(seconds = 30) |>
    req_perform() |>
    resp_body_json(simplifyVector = FALSE) |>
    pluck("results")
}

perform_openalex_page <- function(
  issns,
  cursor,
  api_key = Sys.getenv("OPENALEX_API_KEY"),
  max_attempts = 4L
) {
  stopifnot(length(issns) <= 100L, nzchar(api_key))

  for (attempt in seq_len(max_attempts)) {
    response <- tryCatch(
      request("https://api.openalex.org/sources") |>
        req_url_query(
          filter = paste0("issn:", paste(issns, collapse = "|")),
          per_page = 200,
          cursor = cursor,
          api_key = api_key
        ) |>
        req_timeout(seconds = 30) |>
        req_error(is_error = \(response) FALSE) |>
        req_perform(),
      error = identity
    )

    if (!inherits(response, "condition")) {
      status <- resp_status(response)
      if (status < 400L) {
        return(list(
          ok = TRUE,
          attempts = attempt,
          body = resp_body_json(response, simplifyVector = FALSE)
        ))
      }
      failure <- paste0("http_", status)
      retryable <- status %in% c(403L, 429L, 500L, 502L, 503L, 504L)
    } else {
      failure <- "transport_error"
      retryable <- TRUE
    }

    if (!retryable || attempt == max_attempts) {
      return(list(
        ok = FALSE,
        attempts = attempt,
        failure = failure
      ))
    }
    Sys.sleep(min(2^(attempt - 1L), 8))
  }
}

fetch_openalex_batch <- function(issns) {
  cursor <- "*"
  seen_cursors <- character()
  pages <- list()
  page_log <- list()
  attempts <- 0L

  repeat {
    if (cursor %in% seen_cursors) {
      return(list(
        status = "api_error",
        issns = issns,
        sources = unlist(pages, recursive = FALSE),
        pages = page_log,
        attempts = attempts,
        failure = "repeated_cursor",
        pagination_complete = FALSE
      ))
    }
    seen_cursors <- c(seen_cursors, cursor)

    page <- perform_openalex_page(issns, cursor)
    attempts <- attempts + page$attempts
    if (!page$ok) {
      return(list(
        status = "api_error",
        issns = issns,
        sources = unlist(pages, recursive = FALSE),
        pages = page_log,
        attempts = attempts,
        failure = page$failure,
        pagination_complete = FALSE
      ))
    }

    results <- if (is.null(page$body$results)) list() else page$body$results
    next_cursor <- page$body$meta$next_cursor
    pages[[length(pages) + 1L]] <- results
    page_log[[length(page_log) + 1L]] <- list(
      result_count = length(results),
      attempts = page$attempts,
      retrieved_at = format(Sys.time(), tz = "UTC", usetz = TRUE)
    )
    if (is.null(next_cursor) || !nzchar(next_cursor)) break
    cursor <- next_cursor
  }

  list(
    status = "complete",
    issns = issns,
    sources = unlist(pages, recursive = FALSE),
    pages = page_log,
    attempts = attempts,
    failure = NULL,
    pagination_complete = TRUE
  )
}

save_rds_atomic <- function(value, path) {
  temporary <- paste0(path, ".tmp")
  saveRDS(value, temporary)
  if (!file.rename(temporary, path)) {
    unlink(temporary)
    stop("Could not save checkpoint atomically.", call. = FALSE)
  }
  invisible(path)
}

fetch_crossref_journal <- function(issn) {
  response <- request(paste0("https://api.crossref.org/journals/", issn)) |>
    req_timeout(seconds = 30) |>
    req_error(
      is_error = \(response) {
        resp_status(response) != 404L && resp_is_error(response)
      }
    ) |>
    req_perform()
  if (resp_status(response) == 404L) return(NULL)
  resp_body_json(response, simplifyVector = FALSE)$message
}

perform_crossref_page <- function(
  cursor,
  mailto,
  max_attempts = 4L
) {
  stopifnot(nzchar(cursor), nzchar(mailto))

  for (attempt in seq_len(max_attempts)) {
    response <- tryCatch(
      request("https://api.crossref.org/journals") |>
        req_url_query(
          rows = 1000,
          cursor = cursor,
          mailto = mailto
        ) |>
        req_user_agent(
          "InvisibleResearch journal enrichment; https://github.com/YannJY02/InvisibleResearch"
        ) |>
        req_timeout(seconds = 60) |>
        req_error(is_error = \(response) FALSE) |>
        req_perform(),
      error = identity
    )

    if (!inherits(response, "condition")) {
      status <- resp_status(response)
      if (status < 400L) {
        body <- tryCatch(
          resp_body_json(response, simplifyVector = FALSE)$message,
          error = identity
        )
        if (
          !inherits(body, "condition") &&
            is.list(body) &&
            is.list(body$items) &&
            length(body[["total-results"]]) == 1L
        ) {
          return(list(ok = TRUE, attempts = attempt, body = body))
        }
        failure <- "parse_error"
        retryable <- TRUE
      } else {
        failure <- paste0("http_", status)
        retryable <- status %in% c(
          403L, 408L, 429L, 500L, 502L, 503L, 504L
        )
      }
    } else {
      failure <- "transport_error"
      retryable <- TRUE
    }

    if (!retryable || attempt == max_attempts) {
      return(list(
        ok = FALSE,
        attempts = attempt,
        failure = failure
      ))
    }
    Sys.sleep(min(2^(attempt - 1L), 8))
  }
}

fetch_crossref_directory <- function() {
  cursor <- "*"
  page_index <- 1L
  page_summaries <- list()
  page_files <- character()
  total_results <- NULL

  repeat {
    page_file <- sprintf("page-%04d.rds", page_index)
    checkpoint_path <- file.path(crossref_page_dir, page_file)
    checkpoint <- if (file.exists(checkpoint_path)) {
      tryCatch(readRDS(checkpoint_path), error = \(error) NULL)
    } else {
      NULL
    }
    reusable <- !is.null(checkpoint) &&
      identical(checkpoint$page_index, page_index) &&
      identical(checkpoint$cursor, cursor) &&
      identical(checkpoint$rows_per_page, 1000L) &&
      identical(checkpoint$status, "complete") &&
      identical(checkpoint$item_count, length(checkpoint$items))

    if (reusable) {
      checkpoint$reused <- TRUE
    } else {
      page <- perform_crossref_page(cursor, crossref_mailto)
      checkpoint <- list(
        status = if (page$ok) "complete" else "api_error",
        page_index = page_index,
        cursor = cursor,
        rows_per_page = 1000L,
        total_results = if (page$ok) {
          as.integer(page$body[["total-results"]])
        } else {
          NA_integer_
        },
        next_cursor = if (page$ok) page$body[["next-cursor"]] else NULL,
        items = if (page$ok) page$body$items else list(),
        item_count = if (page$ok) length(page$body$items) else 0L,
        attempts = page$attempts,
        failure = if (page$ok) NULL else page$failure,
        pagination_complete = page$ok && length(page$body$items) < 1000L,
        retrieved_at = format(Sys.time(), tz = "UTC", usetz = TRUE),
        reused = FALSE
      )
      save_rds_atomic(checkpoint, checkpoint_path)
    }

    if (checkpoint$status != "complete") {
      stop(
        paste0(
          "Crossref page ", page_index,
          " failed explicitly: ", checkpoint$failure
        ),
        call. = FALSE
      )
    }
    if (is.null(total_results)) {
      total_results <- checkpoint$total_results
    } else if (!identical(total_results, checkpoint$total_results)) {
      stop(
        "Crossref total-results changed during cursor pagination.",
        call. = FALSE
      )
    }

    page_files <- c(page_files, page_file)
    page_summaries[[length(page_summaries) + 1L]] <- checkpoint[
      setdiff(names(checkpoint), c("cursor", "next_cursor", "items"))
    ]
    retrieved_rows <- sum(map_int(page_summaries, "item_count"))
    if (retrieved_rows > total_results) {
      stop(
        "Crossref pagination exceeded total-results.",
        call. = FALSE
      )
    }
    if (isTRUE(checkpoint$pagination_complete)) break
    if (
      is.null(checkpoint$next_cursor) ||
        !nzchar(checkpoint$next_cursor)
    ) {
      stop(
        "Crossref returned a full page without a next cursor.",
        call. = FALSE
      )
    }

    cursor <- checkpoint$next_cursor
    page_index <- page_index + 1L
    Sys.sleep(0.2)
  }

  directory_rows <- sum(map_int(page_summaries, "item_count"))
  if (directory_rows != total_results) {
    stop(
      paste0(
        "Crossref directory row count mismatch: ",
        directory_rows, " != ", total_results
      ),
      call. = FALSE
    )
  }
  list(
    status = "complete",
    rows_per_page = 1000L,
    directory_rows = directory_rows,
    pages = page_summaries,
    page_files = page_files
  )
}

```

## Load the PKP V7 rows

The pinned PKP V7 file is Dataverse file `14084919`, released on 2026-07-18. This chunk checks the complete 98,273-row baseline, schema, row identity, and identifier counts before either retaining the full cohort or selecting ten fixed source rows in a deliberate display order.

```{r}
#| label: load-sample

raw_input <- read_csv(
  input_path,
  col_types = cols(.default = col_character()),
  na = character(),
  show_col_types = FALSE
)

expected_schema <- c(
  "context_name",
  "issn",
  "oai_url",
  "application",
  "country",
  "region",
  "total_record_count",
  "record_count_2025",
  "earliest_datestamp",
  "repository_name",
  "set_spec",
  "stats_id",
  "best_doaj_url",
  "last_completed_update",
  "first_beacon",
  "last_beacon",
  "last_oai_response",
  "unresponsive_endpoint",
  "unresponsive_context"
)
row_identity <- function(data) {
  paste(data$oai_url, data$repository_name, data$set_spec, sep = "\r")
}
raw_row_identity <- row_identity(raw_input)

stopifnot(
  nrow(raw_input) == 98273L,
  identical(names(raw_input), expected_schema),
  all(tolower(raw_input$application) == "ojs"),
  !anyDuplicated(raw_row_identity)
)

all_input_issns <- map(raw_input$issn, parse_issns)
missing_identifier <- is_missing_identifier(raw_input$issn)
stopifnot(
  sum(missing_identifier) == 26189L,
  sum(lengths(all_input_issns) > 0L) == 72084L,
  length(unique(unlist(all_input_issns))) == 103017L
)

sample_spec <- tribble(
  ~context_name, ~oai_url, ~repository_name, ~set_spec,
  "Kuwait Journal of Science",
  "https://journalskuwait.org/kjs/index.php/index/oai",
  "Kuwait Journal of Science",
  "KJS",
  "Cakrawala",
  "https://cakrawalajournal.org/index.php/index/oai",
  "CAKRAWALA",
  "cakrawala",
  "Sintesa: Jurnal Ilmu Pendidikan",
  "https://sintesa.stkip-arrahmaniyah.ac.id/index.php/index/oai",
  "Sintesa: Jurnal Ilmu Pendidikan STKIP Arrahmaniyah Depok",
  "sintesa",
  "CARAKA: Jurnal Teologi Biblika dan Praktika",
  "https://ojs.sttibc.ac.id/index.php/index/oai",
  "CARAKA:Jurnal Teologi",
  "ibc",
  "Jurnal Tata Kelola dan Akuntabilitas Keuangan Negara",
  "https://jurnal.bpk.go.id/index.php/index/oai",
  "Open Journal Systems",
  "TAKEN",
  "JOEEL (Journal of English Education and Literature)",
  "https://journal.stkippamanetalino.ac.id/index.php/index/oai",
  "Rumah Jurnal STKIP Pamane Talino Ngabang",
  "bahasa-inggris",
  "Malaysian Journal of Paediatrics and Child Health",
  "https://mpaeds.my/journals/index.php/index/oai",
  "NA",
  "MJPCH",
  "Jami Scientific Research Quarterly Journal",
  "https://journals.jami.edu.af/index.php/index/oai",
  "Jami University Press",
  "jsrqj",
  "Journal of Polymer & Composites",
  "https://engineeringjournals.stmjournals.in/index.php/index/oai",
  "Engineering Journals",
  "JoPC",
  "JURNAL TEKNIK PERTAMBANGAN",
  "https://e-journal.upr.ac.id/index.php/index/oai",
  "Journal Online Universitas Palangka Raya",
  "JTP"
)
sample_identity <- row_identity(sample_spec)
sample_positions <- match(sample_identity, raw_row_identity)
stopifnot(!anyNA(sample_positions))

sample_rows <- raw_input[sample_positions, ] |>
  select(-any_of("admin_email"))
stopifnot(
  nrow(sample_rows) == 10L,
  identical(sample_rows$context_name, sample_spec$context_name)
)

ojs_rows <- if (run_mode == "full") {
  raw_input |> select(-any_of("admin_email"))
} else {
  sample_rows
}
```

## Multi-source enrichment

For each PKP row, the notebook looks for OpenAlex and Crossref records with the same checksum-valid normalized ISSN. If the same source identity appears more than once, it is counted once. Sample mode expands returned fields for review. Full mode keeps compact candidate identities in the one-row-per-PKP master and indexes both sources to complete checkpoint responses in companion Parquet catalogs.

### Sample matching helpers

These helpers label each row as having no usable ISSN, no match, one match, or multiple matches. They also identify Crossref records by their ISSNs and convert every candidate field into a dataframe-safe value without filtering the response schema.

```{r}
#| label: sample-matching-helpers

classify_status <- function(input_issns, candidate_ids) {
  case_when(
    length(input_issns) == 0L ~ "not_attempted",
    length(candidate_ids) == 0L ~ "unmatched",
    length(candidate_ids) == 1L ~ "consistent",
    TRUE ~ "inconsistent"
  )
}

crossref_key <- function(journal) {
  issns <- vapply(journal$ISSN, normalize_issn, character(1))
  paste(sort(unique(issns[!is.na(issns)])), collapse = "|")
}

crossref_catalog_from_pages <- function(page_files) {
  seen_keys <- new.env(hash = TRUE, parent = emptyenv())
  catalog_pages <- vector("list", length(page_files))
  source_fields <- character()

  for (page_position in seq_along(page_files)) {
    page_file <- page_files[[page_position]]
    checkpoint <- readRDS(file.path(crossref_page_dir, page_file))
    if (checkpoint$status != "complete") {
      stop("Crossref catalog encountered an incomplete page.", call. = FALSE)
    }
    source_fields <- union(
      source_fields,
      unlist(map(checkpoint$items, names), use.names = FALSE)
    )
    page_rows <- imap(checkpoint$items, \(journal, record_index) {
      candidate_key <- crossref_key(journal)
      if (
        !nzchar(candidate_key) ||
          exists(candidate_key, seen_keys, inherits = FALSE)
      ) {
        return(NULL)
      }
      assign(candidate_key, TRUE, seen_keys)
      tibble(
        crossref_candidate_issn_set = candidate_key,
        checkpoint_file = page_file,
        record_index = as.character(record_index)
      )
    })
    catalog_pages[[page_position]] <- bind_rows(compact(page_rows))
  }

  list(
    rows = bind_rows(catalog_pages),
    source_fields = sort(source_fields)
  )
}

validate_crossref_catalog_pointers <- function(catalog, page_files) {
  catalog_rows_by_page <- split(
    seq_len(nrow(catalog)),
    catalog$checkpoint_file
  )
  checkpoint_keys <- character()

  for (page_file in page_files) {
    checkpoint <- readRDS(file.path(crossref_page_dir, page_file))
    page_keys <- vapply(
      checkpoint$items,
      crossref_key,
      character(1)
    )
    checkpoint_keys <- union(
      checkpoint_keys,
      page_keys[nzchar(page_keys)]
    )

    catalog_rows <- catalog_rows_by_page[[page_file]]
    if (is.null(catalog_rows)) next
    record_indices <- as.integer(catalog$record_index[catalog_rows])
    if (
      anyNA(record_indices) ||
        any(record_indices < 1L | record_indices > length(page_keys))
    ) {
      stop(
        "Crossref catalog contains an invalid record index.",
        call. = FALSE
      )
    }
    if (!identical(
      unname(catalog$crossref_candidate_issn_set[catalog_rows]),
      unname(page_keys[record_indices])
    )) {
      stop(
        "Crossref catalog pointer does not resolve to its ISSN set.",
        call. = FALSE
      )
    }
  }

  if (!setequal(
    catalog$crossref_candidate_issn_set,
    checkpoint_keys
  )) {
    stop(
      "Crossref catalog does not cover every checkpoint candidate.",
      call. = FALSE
    )
  }
  TRUE
}

index_crossref_catalog <- function(catalog) {
  index <- new.env(hash = TRUE, parent = emptyenv())
  for (candidate_key in catalog$crossref_candidate_issn_set) {
    for (issn in strsplit(candidate_key, "|", fixed = TRUE)[[1]]) {
      candidate_keys <- if (exists(issn, index, inherits = FALSE)) {
        get(issn, index, inherits = FALSE)
      } else {
        character()
      }
      assign(issn, c(candidate_keys, candidate_key), index)
    }
  }
  index
}

match_crossref_catalog <- function(input_issns, index) {
  candidate_keys <- as.character(unique(unlist(map(input_issns, \(issn) {
    if (exists(issn, index, inherits = FALSE)) {
      get(issn, index, inherits = FALSE)
    } else {
      character()
    }
  }))))
  candidate_issns <- unique(unlist(strsplit(
    candidate_keys,
    "|",
    fixed = TRUE
  )))
  list(
    status = classify_status(input_issns, candidate_keys),
    matched_issns = intersect(input_issns, candidate_issns),
    candidate_ids = candidate_keys
  )
}

deduplicate_candidates <- function(candidates, identity) {
  candidate_ids <- vapply(candidates, identity, character(1))
  keep <- !duplicated(candidate_ids)
  list(
    candidates = candidates[keep],
    candidate_ids = candidate_ids[keep]
  )
}

index_candidates_by_issn <- function(candidates, candidate_issns) {
  index <- new.env(hash = TRUE, parent = emptyenv())
  for (candidate_position in seq_along(candidates)) {
    for (issn in candidate_issns(candidates[[candidate_position]])) {
      positions <- if (exists(issn, index, inherits = FALSE)) {
        get(issn, index, inherits = FALSE)
      } else {
        integer()
      }
      assign(issn, c(positions, candidate_position), index)
    }
  }
  index
}

candidates_for_issns <- function(candidates, index, issns) {
  positions <- unique(unlist(map(issns, \(issn) {
    if (exists(issn, index, inherits = FALSE)) {
      get(issn, index, inherits = FALSE)
    } else {
      integer()
    }
  })))
  candidates[positions]
}

match_exact_issns <- function(
  input_issns,
  candidates,
  candidate_issns,
  candidate_identity
) {
  matched <- keep(candidates, \(candidate) {
    any(input_issns %in% candidate_issns(candidate))
  })
  deduplicated <- deduplicate_candidates(matched, candidate_identity)
  matched_issns <- intersect(
    input_issns,
    unique(unlist(map(deduplicated$candidates, candidate_issns)))
  )

  c(
    list(
      status = classify_status(
        input_issns,
        deduplicated$candidate_ids
      ),
      matched_issns = matched_issns
    ),
    deduplicated
  )
}

serialize_json <- function(value) {
  as.character(jsonlite::toJSON(
    value,
    auto_unbox = TRUE,
    null = "null",
    na = "null",
    digits = NA
  ))
}

dataframe_value <- function(value) {
  if (is.null(value) || !length(value)) return(NA)
  if (is.atomic(value) && length(value) == 1L) return(value)
  serialize_json(value)
}

candidate_prefix <- function(prefix, index) {
  if (index == 1L) prefix else paste0(prefix, "_candidate_", index)
}

candidate_field_names <- function(match, prefix) {
  unlist(imap(match$candidates, \(candidate, index) {
    paste(candidate_prefix(prefix, index), names(candidate), sep = "_")
  }), use.names = FALSE)
}

expand_candidate_fields <- function(match, prefix) {
  cells <- list()
  for (index in seq_along(match$candidates)) {
    candidate <- match$candidates[[index]]
    values <- map(candidate, dataframe_value)
    names(values) <- paste(
      candidate_prefix(prefix, index),
      names(candidate),
      sep = "_"
    )
    cells <- c(cells, values)
  }
  cells[[paste0(prefix, "_candidates__json")]] <- serialize_json(
    match$candidates
  )
  as_tibble(cells, .name_repair = "check_unique")
}

normalize_title <- function(value) {
  if (length(value) != 1L || is.na(value) || !nzchar(trimws(value))) {
    return(NA_character_)
  }
  transliterated <- iconv(value, from = "", to = "ASCII//TRANSLIT")
  if (is.na(transliterated)) transliterated <- value
  gsub("[^[:alnum:]]", "", tolower(transliterated))
}

source_value <- function(source, field) {
  value <- source[[field]]
  if (is.null(value) || !length(value) || is.na(value[[1]])) {
    return(NA_character_)
  }
  as.character(value[[1]])
}

comparison_state <- function(left, right) {
  if (is.na(left) || is.na(right)) return("not_observed")
  if (identical(left, right)) "same" else "different"
}

category_summary <- function(
  values,
  categories,
  dimension,
  denominator_name
) {
  counts <- table(factor(values, levels = categories))
  denominator <- length(values)
  tibble(
    dimension = dimension,
    category = names(counts),
    denominator_name = denominator_name,
    denominator = denominator,
    count = as.integer(counts),
    percent = round(100 * count / denominator, 3)
  )
}
```

### Enrichment orchestration

Both modes use the same exact-ISSN rule. Full mode indexes candidates by normalized ISSN so matching 98,273 rows does not repeatedly scan either complete response set. Candidate identities connect the compact master to source catalogs and lossless checkpoints, so inconsistent identities remain complete without duplicating large JSON.

```{r}
#| label: enrichment

input_issns <- map(ojs_rows$issn, parse_issns)
tokens <- sort(unique(unlist(input_issns)))

if (run_mode == "full") {
  token_batches <- split(tokens, ceiling(seq_along(tokens) / 100L))
  process_batch <- function(batch_index) {
    batch <- token_batches[[batch_index]]
    checkpoint_path <- file.path(
      openalex_batch_dir,
      sprintf("batch-%04d.rds", batch_index)
    )
    checkpoint <- if (file.exists(checkpoint_path)) {
      tryCatch(readRDS(checkpoint_path), error = \(error) NULL)
    } else {
      NULL
    }
    reusable <- !is.null(checkpoint) &&
      identical(checkpoint$pkp_version, pkp_version) &&
      identical(checkpoint$input_md5, pkp_md5) &&
      identical(checkpoint$issns, batch) &&
      identical(checkpoint$status, "complete") &&
      isTRUE(checkpoint$pagination_complete)

    if (reusable) {
      checkpoint$reused <- TRUE
      return(checkpoint)
    }

    checkpoint <- fetch_openalex_batch(batch)
    checkpoint$pkp_version <- pkp_version
    checkpoint$input_md5 <- pkp_md5
    checkpoint$batch_index <- as.integer(batch_index)
    checkpoint$reused <- FALSE
    save_rds_atomic(checkpoint, checkpoint_path)
    checkpoint
  }
  openalex_cluster <- parallel::makeCluster(4L)
  invisible(parallel::clusterEvalQ(openalex_cluster, library(httr2)))
  parallel::clusterExport(
    openalex_cluster,
    c(
      "token_batches",
      "openalex_batch_dir",
      "pkp_version",
      "pkp_md5",
      "perform_openalex_page",
      "fetch_openalex_batch",
      "save_rds_atomic",
      "process_batch"
    ),
    envir = environment()
  )
  batch_results <- tryCatch(
    parallel::parLapplyLB(
      openalex_cluster,
      seq_along(token_batches),
      process_batch
    ),
    finally = parallel::stopCluster(openalex_cluster)
  )
  openalex_cache_hit <- all(map_lgl(batch_results, "reused"))
  openalex_sources <- unlist(
    map(batch_results, "sources"),
    recursive = FALSE
  )
  failed_issns <- unique(unlist(map(
    keep(batch_results, \(batch) batch$status != "complete"),
    "issns"
  )))
} else {
  token_batches <- list(tokens)
  batch_results <- list()
  failed_issns <- character()
  openalex_cache_hit <- file.exists(openalex_cache_path)
  openalex_sources <- if (openalex_cache_hit) {
    readRDS(openalex_cache_path)
  } else {
    sources <- fetch_openalex_sources(tokens)
    saveRDS(sources, openalex_cache_path)
    sources
  }
}
openalex_sources <- deduplicate_candidates(
  openalex_sources,
  \(source) source$id
)$candidates
openalex_index <- index_candidates_by_issn(
  openalex_sources,
  source_issns
)

openalex_matches <- map(input_issns, \(issns) {
  match <- match_exact_issns(
    issns,
    candidates_for_issns(openalex_sources, openalex_index, issns),
    source_issns,
    \(source) source$id
  )
  if (any(issns %in% failed_issns)) match$status <- "api_error"
  match
})
openalex_status_counts <- table(map_chr(openalex_matches, "status"))

base_result <- ojs_rows |>
  mutate(
    identifier_status = case_when(
      is_missing_identifier(issn) ~ "missing",
      lengths(input_issns) == 0L ~ "invalid",
      TRUE ~ "valid"
    ),
    openalex_input_issns = map_chr(input_issns, paste, collapse = "|"),
    openalex_matched_issns = map_chr(
      openalex_matches,
      \(match) paste(match$matched_issns, collapse = "|")
    ),
    openalex_match_status = map_chr(openalex_matches, "status"),
    openalex_candidate_count = map_int(
      openalex_matches,
      \(match) length(match$candidate_ids)
    ),
    openalex_candidate_ids = map_chr(
      openalex_matches,
      \(match) paste(match$candidate_ids, collapse = "|")
    )
  )

if (run_mode == "sample") {
  openalex_data <- map_dfr(
    openalex_matches,
    expand_candidate_fields,
    prefix = "openalex"
  )
  result <- bind_cols(base_result, openalex_data)
  crossref_cache_hit <- file.exists(crossref_cache_path)
  crossref_journals <- if (crossref_cache_hit) {
    readRDS(crossref_cache_path)
  } else {
    journals <- set_names(
      map(tokens, \(token) {
        journal <- fetch_crossref_journal(token)
        Sys.sleep(0.25)
        journal
      }),
      tokens
    )
    saveRDS(journals, crossref_cache_path)
    journals
  }
  crossref_candidates <- unname(compact(crossref_journals))
  crossref_index <- index_candidates_by_issn(
    crossref_candidates,
    \(journal) {
      issns <- vapply(journal$ISSN, normalize_issn, character(1))
      unique(issns[!is.na(issns)])
    }
  )
  crossref_matches <- map(input_issns, \(issns) {
    match_exact_issns(
      issns,
      candidates_for_issns(
        crossref_candidates,
        crossref_index,
        issns
      ),
      \(journal) {
        issns <- vapply(journal$ISSN, normalize_issn, character(1))
        unique(issns[!is.na(issns)])
      },
      crossref_key
    )
  })
  crossref_status_counts <- table(map_chr(crossref_matches, "status"))
  crossref_data <- map_dfr(
    crossref_matches,
    expand_candidate_fields,
    prefix = "crossref"
  )
  result <- result |>
    mutate(
      crossref_input_issns = map_chr(input_issns, paste, collapse = "|"),
      crossref_matched_issns = map_chr(
        crossref_matches,
        \(match) paste(match$matched_issns, collapse = "|")
      ),
      crossref_match_status = map_chr(crossref_matches, "status"),
      crossref_candidate_issn_sets = map_chr(
        crossref_matches,
        \(match) paste(match$candidate_ids, collapse = ";")
      )
    ) |>
    bind_cols(crossref_data)
  write_csv(result, output_path, na = "")
  output_rows <- nrow(result)
  output_columns <- names(result)
  review_result <- result
} else {
  crossref_directory <- fetch_crossref_directory()
  crossref_catalog_contract <- crossref_catalog_from_pages(
    crossref_directory$page_files
  )
  crossref_catalog <- crossref_catalog_contract$rows
  crossref_catalog_valid <- validate_crossref_catalog_pointers(
    crossref_catalog,
    crossref_directory$page_files
  )
  crossref_index <- index_crossref_catalog(crossref_catalog)
  crossref_matches <- map(
    input_issns,
    match_crossref_catalog,
    index = crossref_index
  )
  crossref_status_counts <- table(map_chr(crossref_matches, "status"))
  base_result <- base_result |>
    mutate(
      crossref_input_issns = map_chr(input_issns, paste, collapse = "|"),
      crossref_matched_issns = map_chr(
        crossref_matches,
        \(match) paste(match$matched_issns, collapse = "|")
      ),
      crossref_match_status = map_chr(crossref_matches, "status"),
      crossref_candidate_count = map_int(
        crossref_matches,
        \(match) length(match$candidate_ids)
      ),
      crossref_candidate_issn_sets = map_chr(
        crossref_matches,
        \(match) paste(match$candidate_ids, collapse = ";")
      )
    )

  openalex_fields <- sort(unique(unlist(map(openalex_sources, names))))
  output_columns <- names(base_result)
  source_output_columns <- c("openalex_source_id", "checkpoint_file")
  crossref_output_columns <- names(crossref_catalog)
  checkpoint_by_source <- new.env(hash = TRUE, parent = emptyenv())
  for (batch in batch_results) {
    checkpoint_file <- sprintf("batch-%04d.rds", batch$batch_index)
    for (source in batch$sources) {
      if (!exists(source$id, checkpoint_by_source, inherits = FALSE)) {
        assign(source$id, checkpoint_file, checkpoint_by_source)
      }
    }
  }

  build_master_chunk <- function(indices) {
    chunk <- as.data.frame(base_result[indices, ], stringsAsFactors = FALSE)
    chunk[] <- lapply(chunk, as.character)
    chunk[output_columns]
  }

  build_source_chunk <- function(indices) {
    source_ids <- unname(map_chr(openalex_sources[indices], "id"))
    data.frame(
      openalex_source_id = source_ids,
      checkpoint_file = unname(vapply(
        source_ids,
        get,
        character(1),
        envir = checkpoint_by_source,
        inherits = FALSE
      )),
      stringsAsFactors = FALSE
    )
  }

  build_crossref_chunk <- function(indices) {
    chunk <- as.data.frame(
      crossref_catalog[indices, ],
      stringsAsFactors = FALSE
    )
    chunk[] <- lapply(chunk, as.character)
    chunk[crossref_output_columns]
  }

  write_ndjson_chunks <- function(path, rows, builder) {
    unlink(path)
    connection <- file(path, open = "wb")
    tryCatch(
      {
        for (start in seq.int(1L, rows, by = 1000L)) {
          indices <- start:min(start + 999L, rows)
          jsonlite::stream_out(
            builder(indices),
            connection,
            verbose = FALSE
          )
        }
      },
      finally = close(connection)
    )
  }

  uv <- Sys.which("uv")
  if (!nzchar(uv)) {
    stop("uv is required to write the full Parquet artifacts.", call. = FALSE)
  }
  convert_to_parquet <- function(input, schema, output, expected_rows) {
    conversion <- system2(
      uv,
      c(
        "run",
        "--with", "pyarrow==21.0.0",
        "python", "ndjson_to_parquet.py", "convert",
        shQuote(input),
        shQuote(schema),
        shQuote(output),
        "--expected-rows", expected_rows
      ),
      stdout = TRUE,
      stderr = TRUE
    )
    if (!is.null(attr(conversion, "status"))) {
      stop(paste(conversion, collapse = "\n"), call. = FALSE)
    }
    jsonlite::fromJSON(tail(conversion, 1L))
  }

  write_ndjson_chunks(staging_path, nrow(base_result), build_master_chunk)
  jsonlite::write_json(
    list(columns = output_columns),
    parquet_schema_path,
    auto_unbox = TRUE,
    pretty = TRUE
  )
  parquet_contract <- convert_to_parquet(
    staging_path,
    parquet_schema_path,
    output_path,
    nrow(base_result)
  )
  unlink(staging_path)

  write_ndjson_chunks(
    source_staging_path,
    length(openalex_sources),
    build_source_chunk
  )
  jsonlite::write_json(
    list(
      columns = source_output_columns,
      source_fields = openalex_fields
    ),
    source_schema_path,
    auto_unbox = TRUE,
    pretty = TRUE
  )
  source_parquet_contract <- convert_to_parquet(
    source_staging_path,
    source_schema_path,
    source_output_path,
    length(openalex_sources)
  )
  unlink(source_staging_path)

  write_ndjson_chunks(
    crossref_staging_path,
    nrow(crossref_catalog),
    build_crossref_chunk
  )
  jsonlite::write_json(
    list(
      columns = crossref_output_columns,
      source_fields = crossref_catalog_contract$source_fields
    ),
    crossref_schema_path,
    auto_unbox = TRUE,
    pretty = TRUE
  )
  crossref_parquet_contract <- convert_to_parquet(
    crossref_staging_path,
    crossref_schema_path,
    crossref_output_path,
    nrow(crossref_catalog)
  )
  unlink(crossref_staging_path)

  jsonlite::write_json(
    list(
      master_rows = nrow(base_result),
      status_counts = list(
        openalex = as.list(openalex_status_counts),
        crossref = as.list(crossref_status_counts)
      ),
      openalex_source_checkpoints = as.list(set_names(
        vapply(
          map_chr(openalex_sources, "id"),
          get,
          character(1),
          envir = checkpoint_by_source,
          inherits = FALSE
        ),
        map_chr(openalex_sources, "id")
      )),
      crossref_source_checkpoints = as.list(set_names(
        paste(
          crossref_catalog$checkpoint_file,
          crossref_catalog$record_index,
          sep = "#"
        ),
        crossref_catalog$crossref_candidate_issn_set
      ))
    ),
    parquet_validation_path,
    auto_unbox = TRUE,
    pretty = TRUE
  )
  validation <- system2(
    uv,
    c(
      "run",
      "--with", "pyarrow==21.0.0",
      "python", "ndjson_to_parquet.py", "validate",
      shQuote(output_path),
      shQuote(source_output_path),
      shQuote(openalex_batch_dir),
      shQuote(crossref_output_path),
      shQuote(crossref_page_dir),
      shQuote(input_path),
      shQuote(parquet_validation_path)
    ),
    stdout = TRUE,
    stderr = TRUE
  )
  if (!is.null(attr(validation, "status"))) {
    stop(paste(validation, collapse = "\n"), call. = FALSE)
  }
  parquet_validation <- jsonlite::fromJSON(tail(validation, 1L))

  output_rows <- parquet_contract$rows
  review_result <- build_master_chunk(sample_positions)
}

selected_openalex_sources <- map(openalex_matches, \(match) {
  if (length(match$candidates) == 1L) match$candidates[[1]] else NULL
})
openalex_source_types <- map_chr(
  selected_openalex_sources,
  source_value,
  field = "type"
)
openalex_source_issn_sets <- map_chr(selected_openalex_sources, \(source) {
  paste(sort(source_issns(source)), collapse = "|")
})
input_issn_sets <- map_chr(input_issns, \(issns) {
  paste(sort(issns), collapse = "|")
})
matched_issn_sets <- map_chr(openalex_matches, \(match) {
  paste(sort(match$matched_issns), collapse = "|")
})

identity_outcome <- case_when(
  base_result$identifier_status != "valid" ~ "no_valid_issn",
  base_result$openalex_match_status == "api_error" ~ "api_error",
  base_result$openalex_candidate_count == 0L ~ "valid_issn_unmatched",
  base_result$openalex_candidate_count > 1L ~ "conflicting_source_ids",
  is.na(openalex_source_types) |
    tolower(openalex_source_types) != "journal" ~ "non_journal_source",
  input_issn_sets != matched_issn_sets ~ "partial_issn_agreement",
  TRUE ~ "consistent"
)
identity_matched <- identity_outcome %in% c(
  "consistent",
  "partial_issn_agreement"
)

pkp_normalized_titles <- map_chr(ojs_rows$context_name, normalize_title)
openalex_normalized_titles <- map(selected_openalex_sources, \(source) {
  titles <- c(
    source_value(source, "display_name"),
    unlist(source$alternate_titles, use.names = FALSE)
  )
  normalized <- unique(na.omit(map_chr(titles, normalize_title)))
  normalized[nzchar(normalized)]
})
title_comparison <- map2_chr(
  pkp_normalized_titles,
  openalex_normalized_titles,
  \(pkp_title, openalex_titles) {
    if (!length(openalex_titles) || is.na(pkp_title)) return("not_observed")
    if (pkp_title %in% openalex_titles) "same" else "different"
  }
)
title_comparison[!identity_matched] <- "not_observed"

issn_set_comparison <- map2_chr(
  input_issn_sets,
  openalex_source_issn_sets,
  comparison_state
)
issn_set_comparison[!identity_matched] <- "not_observed"

openalex_ojs_evidence <- map_chr(
  selected_openalex_sources,
  source_value,
  field = "is_ojs"
)
ojs_evidence_comparison <- case_when(
  !identity_matched | is.na(openalex_ojs_evidence) ~ "not_observed",
  openalex_ojs_evidence == "TRUE" ~ "same",
  TRUE ~ "different"
)

pkp_doaj_evidence <- !is.na(ojs_rows$best_doaj_url) &
  nzchar(trimws(ojs_rows$best_doaj_url))
openalex_doaj_evidence <- map_chr(
  selected_openalex_sources,
  source_value,
  field = "is_in_doaj"
)
doaj_evidence_comparison <- case_when(
  !identity_matched | is.na(openalex_doaj_evidence) ~ "not_observed",
  pkp_doaj_evidence & openalex_doaj_evidence == "TRUE" ~ "both_observed",
  pkp_doaj_evidence ~ "pkp_only",
  openalex_doaj_evidence == "TRUE" ~ "openalex_only",
  TRUE ~ "neither_observed"
)

pkp_country_code <- suppressWarnings(countrycode::countrycode(
  ojs_rows$country,
  origin = "country.name",
  destination = "iso2c"
))
openalex_country_code <- toupper(map_chr(
  selected_openalex_sources,
  source_value,
  field = "country_code"
))
country_evidence_comparison <- map2_chr(
  pkp_country_code,
  openalex_country_code,
  comparison_state
)
country_evidence_comparison[!identity_matched] <- "not_observed"

audit_base <- if (run_mode == "sample") result else base_result
audit <- audit_base |>
  transmute(
    oai_url,
    repository_name,
    set_spec,
    context_name,
    issn,
    identifier_status,
    openalex_input_issns,
    openalex_matched_issns,
    openalex_match_status,
    openalex_candidate_count,
    openalex_candidate_ids,
    crossref_match_status,
    crossref_candidate_count = if ("crossref_candidate_count" %in% names(
      audit_base
    )) {
      crossref_candidate_count
    } else {
      map_int(crossref_matches, \(match) length(match$candidate_ids))
    },
    crossref_candidate_issn_sets,
    identity_outcome,
    openalex_source_type = openalex_source_types,
    openalex_display_name = map_chr(
      selected_openalex_sources,
      source_value,
      field = "display_name"
    ),
    pkp_normalized_title = pkp_normalized_titles,
    openalex_normalized_titles = map_chr(
      openalex_normalized_titles,
      paste,
      collapse = "|"
    ),
    title_comparison,
    pkp_issn_set = input_issn_sets,
    openalex_issn_set = openalex_source_issn_sets,
    issn_set_comparison,
    ojs_evidence_comparison,
    pkp_doaj_evidence,
    openalex_doaj_evidence,
    doaj_evidence_comparison,
    pkp_country = country,
    pkp_country_code,
    openalex_country_code,
    country_evidence_comparison
  )

identity_categories <- c(
  "consistent",
  "partial_issn_agreement",
  "conflicting_source_ids",
  "non_journal_source",
  "valid_issn_unmatched",
  "no_valid_issn",
  "api_error"
)
comparison_categories <- c("same", "different", "not_observed")
matched_denominator <- sum(identity_matched)
status_categories <- c(
  "consistent",
  "inconsistent",
  "unmatched",
  "not_attempted",
  "api_error"
)
source_status_counts <- as_tibble(as.data.frame(table(
  openalex = factor(
    audit$openalex_match_status,
    levels = status_categories
  ),
  crossref = factor(
    audit$crossref_match_status,
    levels = status_categories
  )
))) |>
  transmute(
    dimension = "source_status_crosstab",
    category = paste0("openalex=", openalex, "|crossref=", crossref),
    denominator_name = "all_pkp_v7_rows",
    denominator = nrow(audit),
    count = as.integer(Freq),
    percent = round(100 * count / denominator, 3)
  )
metadata_disagreement_counts <- c(
  normalized_title = sum(title_comparison == "different"),
  issn_set = sum(issn_set_comparison == "different"),
  ojs_evidence = sum(ojs_evidence_comparison == "different"),
  doaj_evidence = sum(doaj_evidence_comparison %in% c(
    "pkp_only",
    "openalex_only"
  )),
  country_evidence = sum(country_evidence_comparison == "different")
)
metadata_disagreement_summary <- tibble(
  dimension = "metadata_disagreement",
  category = names(metadata_disagreement_counts),
  denominator_name = "identity_matched_openalex_journals",
  denominator = matched_denominator,
  count = as.integer(metadata_disagreement_counts),
  percent = round(100 * count / denominator, 3)
)
audit_summary <- bind_rows(
  category_summary(
    identity_outcome,
    identity_categories,
    "identity_outcome",
    "all_pkp_v7_rows"
  ),
  category_summary(
    title_comparison[identity_matched],
    comparison_categories,
    "normalized_title",
    "identity_matched_openalex_journals"
  ),
  category_summary(
    issn_set_comparison[identity_matched],
    comparison_categories,
    "issn_set",
    "identity_matched_openalex_journals"
  ),
  category_summary(
    ojs_evidence_comparison[identity_matched],
    comparison_categories,
    "ojs_evidence",
    "identity_matched_openalex_journals"
  ),
  category_summary(
    doaj_evidence_comparison[identity_matched],
    c(
      "both_observed",
      "pkp_only",
      "openalex_only",
      "neither_observed",
      "not_observed"
    ),
    "doaj_evidence",
    "identity_matched_openalex_journals"
  ),
  category_summary(
    country_evidence_comparison[identity_matched],
    comparison_categories,
    "country_evidence",
    "identity_matched_openalex_journals"
  ),
  metadata_disagreement_summary,
  source_status_counts
)
write_csv(audit, audit_path, na = "")
write_csv(audit_summary, audit_summary_path, na = "")

openalex_retrieval_times <- if (run_mode == "full") {
  unlist(map(batch_results, \(batch) map_chr(batch$pages, "retrieved_at")))
} else {
  character()
}
crossref_retrieval_times <- if (run_mode == "full") {
  map_chr(crossref_directory$pages, "retrieved_at")
} else {
  character()
}

run_metadata <- list(
  producer = "analysis/ojs_journal_enrichment.qmd",
  run_mode = run_mode,
  created_at = format(Sys.time(), tz = "UTC", usetz = TRUE),
  input = list(
    version = pkp_version,
    dataverse_file_id = 14084919L,
    md5 = pkp_md5,
    rows = nrow(raw_input),
    valid_identifier_rows = sum(lengths(all_input_issns) > 0L),
    missing_identifier_rows = sum(missing_identifier),
    distinct_valid_issns = length(unique(unlist(all_input_issns)))
  ),
  output = list(
    file = basename(output_path),
    md5 = unname(tools::md5sum(output_path)),
    format = if (run_mode == "full") "parquet" else "csv",
    rows = output_rows,
    columns = length(output_columns),
    status_counts = list(
      openalex = as.list(openalex_status_counts),
      crossref = as.list(crossref_status_counts)
    )
  ),
  disagreement_audit = list(
    file = basename(audit_path),
    md5 = unname(tools::md5sum(audit_path)),
    rows = nrow(audit),
    columns = ncol(audit),
    summary_file = basename(audit_summary_path),
    summary_md5 = unname(tools::md5sum(audit_summary_path)),
    summary_rows = nrow(audit_summary),
    identity_matched_rows = matched_denominator
  ),
  source_catalog = if (run_mode == "full") {
    list(
      file = basename(source_output_path),
      md5 = unname(tools::md5sum(source_output_path)),
      rows = source_parquet_contract$rows,
      columns = source_parquet_contract$columns,
      retained_top_level_fields = openalex_fields
    )
  } else {
    NULL
  },
  crossref_catalog = if (run_mode == "full") {
    list(
      file = basename(crossref_output_path),
      md5 = unname(tools::md5sum(crossref_output_path)),
      rows = crossref_parquet_contract$rows,
      columns = crossref_parquet_contract$columns,
      retained_top_level_fields =
        crossref_catalog_contract$source_fields
    )
  } else {
    NULL
  },
  parquet_validation = if (run_mode == "full") parquet_validation else NULL,
  retrieval = if (run_mode == "full") {
    list(
      openalex = list(
        first = min(openalex_retrieval_times),
        last = max(openalex_retrieval_times)
      ),
      crossref = list(
        first = min(crossref_retrieval_times),
        last = max(crossref_retrieval_times)
      )
    )
  } else {
    NULL
  },
  openalex = if (run_mode == "full") {
    list(
      batch_size = 100L,
      per_page = 200L,
      batches = length(batch_results),
      completed_batches = sum(map_chr(batch_results, "status") == "complete"),
      reused_batches = sum(map_lgl(batch_results, "reused")),
      pages = sum(map_int(batch_results, \(batch) length(batch$pages))),
      attempts = sum(map_int(batch_results, "attempts")),
      failed_batches = sum(map_chr(batch_results, "status") != "complete")
    )
  } else {
    list(
      cache_file = basename(openalex_cache_path),
      reused = openalex_cache_hit
    )
  },
  crossref = if (run_mode == "full") {
    list(
      status = crossref_directory$status,
      rows_per_page = crossref_directory$rows_per_page,
      directory_rows = crossref_directory$directory_rows,
      pages = length(crossref_directory$pages),
      completed_pages = sum(
        map_chr(crossref_directory$pages, "status") == "complete"
      ),
      reused_pages = sum(map_lgl(
        crossref_directory$pages,
        "reused"
      )),
      attempts = sum(map_int(crossref_directory$pages, "attempts")),
      failed_pages = sum(
        map_chr(crossref_directory$pages, "status") != "complete"
      )
    )
  } else {
    list(
      cache_file = basename(crossref_cache_path),
      reused = crossref_cache_hit
    )
  },
  runtime = list(
    r = R.version.string,
    quarto = system2("quarto", "--version", stdout = TRUE)[1],
    curl = system2(curl, "--version", stdout = TRUE)[1],
    pyarrow = if (run_mode == "full") {
      source_parquet_contract$pyarrow
    } else {
      "not used"
    },
    packages = as.list(set_names(
      vapply(
        required_packages,
        \(package) as.character(packageVersion(package)),
        character(1)
      ),
      required_packages
    ))
  )
)
jsonlite::write_json(
  run_metadata,
  metadata_path,
  auto_unbox = TRUE,
  pretty = TRUE
)
```

## Checks

These checks confirm that every input row appears in the master, every returned top-level API field remains in the lossless source checkpoints, both catalogs reconcile, credentials stay private, and full-mode pagination completed without unresolved API failures.

```{r}
#| label: contract-check

metadata_text <- paste(readLines(metadata_path, warn = FALSE), collapse = "\n")
artifact_root <- normalizePath(artifact_dir)
metadata_dimensions <- c(
  "normalized_title",
  "issn_set",
  "ojs_evidence",
  "doaj_evidence",
  "country_evidence"
)
metadata_reconciliation <- audit_summary |>
  filter(dimension %in% metadata_dimensions) |>
  group_by(dimension) |>
  summarise(
    denominator = unique(denominator),
    count = sum(count),
    .groups = "drop"
  )

stopifnot(
  identical(pkp_version, 7L),
  is.na(normalize_issn(NULL)),
  identical(unname(tools::md5sum(input_path)), pkp_md5),
  nrow(raw_input) == 98273L,
  identical(names(raw_input), expected_schema),
  !anyDuplicated(raw_row_identity),
  sum(missing_identifier) == 26189L,
  sum(lengths(all_input_issns) > 0L) == 72084L,
  length(unique(unlist(all_input_issns))) == 103017L,
  startsWith(normalizePath(input_path), artifact_root),
  startsWith(normalizePath(output_path), artifact_root),
  startsWith(normalizePath(metadata_path), artifact_root),
  startsWith(normalizePath(audit_path), artifact_root),
  startsWith(normalizePath(audit_summary_path), artifact_root),
  file.exists(metadata_path),
  file.exists(audit_path),
  file.exists(audit_summary_path),
  !grepl("OPENALEX_API_KEY|CROSSREF_MAILTO", metadata_text),
  all(nzchar(unlist(run_metadata$runtime))),
  all(vapply(c(
    "normalize_issn",
    "is_missing_identifier",
    "row_identity",
    "match_exact_issns",
    "deduplicate_candidates",
    "index_candidates_by_issn",
    "candidates_for_issns",
    "crossref_catalog_from_pages",
    "validate_crossref_catalog_pointers",
    "index_crossref_catalog",
    "match_crossref_catalog",
    "expand_candidate_fields",
    "serialize_json",
    "normalize_title",
    "source_value",
    "comparison_state",
    "category_summary"
  ), exists, logical(1), mode = "function")),
  exists("fetch_openalex_sources", mode = "function"),
  exists("fetch_crossref_journal", mode = "function"),
  exists("fetch_crossref_directory", mode = "function"),
  output_rows == nrow(ojs_rows),
  file.exists(output_path),
  all(c(
    "openalex_match_status",
    "openalex_candidate_count",
    "openalex_candidate_ids"
  ) %in% output_columns),
  identical(
    base_result$openalex_candidate_count,
    map_int(openalex_matches, \(match) length(match$candidate_ids))
  ),
  all(base_result$openalex_match_status %in% c(
    "consistent",
    "inconsistent",
    "unmatched",
    "not_attempted",
    "api_error"
  )),
  sum(unlist(run_metadata$output$status_counts$openalex)) == output_rows,
  sum(unlist(run_metadata$output$status_counts$crossref)) == output_rows,
  !"admin_email" %in% output_columns,
  nrow(audit) == output_rows,
  identical(audit$identity_outcome, identity_outcome),
  identical(
    pkp_doaj_evidence,
    !is_missing_identifier(ojs_rows$best_doaj_url)
  ),
  setequal(
    audit_summary$category[audit_summary$dimension == "identity_outcome"],
    identity_categories
  ),
  sum(audit_summary$count[
    audit_summary$dimension == "identity_outcome"
  ]) == output_rows,
  nrow(metadata_reconciliation) == length(metadata_dimensions),
  all(metadata_reconciliation$denominator == matched_denominator),
  all(metadata_reconciliation$count == matched_denominator),
  sum(audit_summary$count[
    audit_summary$dimension == "source_status_crosstab"
  ]) == output_rows,
  all(audit_summary$denominator[
    audit_summary$dimension == "source_status_crosstab"
  ] == output_rows),
  !"admin_email" %in% names(audit),
  all(is.finite(audit_summary$percent))
)

if (run_mode == "sample") {
  expected_openalex_columns <- unique(unlist(map(
    openalex_matches,
    candidate_field_names,
    prefix = "openalex"
  )))
  expected_openalex_status <- c(
    "consistent",
    "consistent",
    "unmatched",
    "consistent",
    "consistent",
    "consistent",
    "consistent",
    "consistent",
    "not_attempted",
    "inconsistent"
  )
  expected_crossref_columns <- unique(unlist(map(
    crossref_matches,
    candidate_field_names,
    prefix = "crossref"
  )))
  expected_identity_outcome <- c(
    "consistent",
    "consistent",
    "valid_issn_unmatched",
    "consistent",
    "consistent",
    "consistent",
    "consistent",
    "partial_issn_agreement",
    "no_valid_issn",
    "conflicting_source_ids"
  )
  stopifnot(
    identical(row_identity(ojs_rows), sample_identity),
    sum(result$identifier_status == "missing") == 1L,
    nrow(result) == 10L,
    all(c(
      "openalex_id",
      "openalex_summary_stats",
      "openalex_topics",
      "openalex_candidates__json",
      "openalex_candidate_2_id",
      "crossref_match_status",
      "crossref_title",
      "crossref_counts",
      "crossref_coverage",
      "crossref_candidates__json"
    ) %in% names(result)),
    all(expected_openalex_columns %in% names(result)),
    all(expected_crossref_columns %in% names(result)),
    identical(
      result$openalex_candidates__json,
      map_chr(openalex_matches, \(match) serialize_json(match$candidates))
    ),
    identical(
      result$crossref_candidates__json,
      map_chr(crossref_matches, \(match) serialize_json(match$candidates))
    ),
    identical(result$openalex_match_status, expected_openalex_status),
    identical(identity_outcome, expected_identity_outcome),
    result$openalex_matched_issns[
      result$context_name == "Jami Scientific Research Quarterly Journal"
    ] == "2710-4990",
    "not_attempted" %in% result$crossref_match_status
  )
} else {
  stopifnot(
    identical(row_identity(ojs_rows), raw_row_identity),
    output_rows == 98273L,
    sum(base_result$identifier_status == "missing") == 26189L,
    length(tokens) == 103017L,
    length(token_batches) == 1031L,
    max(lengths(token_batches)) <= 100L,
    all(map_lgl(batch_results, \(batch) {
      identical(batch$pkp_version, 7L) &&
        identical(batch$input_md5, pkp_md5) &&
        identical(batch$status, "complete") &&
        isTRUE(batch$pagination_complete) &&
        length(batch$pages) >= 1L
    })),
    run_metadata$openalex$completed_batches == length(token_batches),
    run_metadata$openalex$failed_batches == 0L,
    !any(base_result$openalex_match_status == "api_error"),
    !any(identity_outcome == "api_error"),
    all(nzchar(unlist(run_metadata$retrieval))),
    run_metadata$disagreement_audit$rows == 98273L,
    identical(
      unname(tools::md5sum(audit_path)),
      run_metadata$disagreement_audit$md5
    ),
    identical(
      unname(tools::md5sum(audit_summary_path)),
      run_metadata$disagreement_audit$summary_md5
    ),
    identical(crossref_directory$status, "complete"),
    identical(
      match_crossref_catalog(
        character(),
        new.env(hash = TRUE, parent = emptyenv())
      )$status,
      "not_attempted"
    ),
    identical(crossref_directory$rows_per_page, 1000L),
    crossref_directory$directory_rows ==
      sum(map_int(crossref_directory$pages, "item_count")),
    all(map_int(
      crossref_directory$pages[-length(crossref_directory$pages)],
      "item_count"
    ) == 1000L),
    last(crossref_directory$pages)$item_count < 1000L,
    all(map_chr(crossref_directory$pages, "status") == "complete"),
    run_metadata$crossref$completed_pages ==
      length(crossref_directory$pages),
    run_metadata$crossref$failed_pages == 0L,
    all(c(
      "crossref_match_status",
      "crossref_candidate_count",
      "crossref_candidate_issn_sets"
    ) %in% output_columns),
    identical(
      base_result$crossref_candidate_count,
      map_int(crossref_matches, \(match) length(match$candidate_ids))
    ),
    all(base_result$crossref_match_status %in% c(
      "consistent",
      "inconsistent",
      "unmatched",
      "not_attempted"
    )),
    !any(base_result$crossref_match_status == "api_error"),
    endsWith(output_path, ".csv.gz"),
    all(c(
      "openalex_display_name",
      "crossref_title"
    ) %in% names(review_result)),
    identical(run_metadata$output$format, "parquet"),
    parquet_contract$columns == length(output_columns),
    source_parquet_contract$rows == length(openalex_sources),
    crossref_parquet_contract$rows == nrow(crossref_catalog),
    isTRUE(crossref_catalog_valid),
    parquet_validation$master_rows == 98273L,
    parquet_validation$unique_identities == 98273L,
    parquet_validation$openalex_catalog_rows == length(openalex_sources),
    parquet_validation$crossref_catalog_rows == nrow(crossref_catalog),
    !anyDuplicated(map_chr(openalex_sources, "id")),
    !anyDuplicated(crossref_catalog$crossref_candidate_issn_set),
    all(unique(unlist(map(openalex_matches, "candidate_ids"))) %in%
      map_chr(openalex_sources, "id")),
    all(unique(unlist(map(crossref_matches, "candidate_ids"))) %in%
      crossref_catalog$crossref_candidate_issn_set),
    length(run_metadata$source_catalog$retained_top_level_fields) ==
      length(openalex_fields),
    length(
      run_metadata$crossref_catalog$retained_top_level_fields
    ) == length(crossref_catalog_contract$source_fields),
    file.exists(source_output_path),
    file.exists(crossref_output_path),
    !file.exists(staging_path),
    !file.exists(source_staging_path),
    !file.exists(crossref_staging_path)
  )
}
```

## Identity outcomes

These categories describe exact-ISSN identity evidence. They do not establish
which source is correct and do not explain why a difference exists.

```{r}
#| label: identity-outcome-chart
#| fig-width: 9
#| fig-height: 5

audit_summary |>
  filter(dimension == "identity_outcome") |>
  ggplot(aes(x = reorder(category, count), y = count)) +
  geom_col() +
  geom_text(
    aes(label = sprintf("%s (%.2f%%)", count, percent)),
    hjust = -0.05,
    size = 3
  ) +
  coord_flip(clip = "off") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  labs(x = NULL, y = "PKP V7 rows") +
  theme_minimal()
```

## Metadata comparisons

Metadata comparisons are limited to PKP rows with one exact-ISSN OpenAlex
journal identity. Title comparison uses normalized PKP titles against both the
OpenAlex display name and alternate titles. PKP country is inferred metadata,
not authoritative publisher location. PKP and OpenAlex DOAJ fields are
source-specific positive evidence; absence is not proof that a journal is not
in DOAJ.

```{r}
#| label: metadata-disagreement-chart
#| fig-width: 9
#| fig-height: 4.5

metadata_disagreement_summary |>
  ggplot(aes(x = reorder(category, count), y = count)) +
  geom_col() +
  geom_text(
    aes(label = sprintf("%s (%.2f%%)", count, percent)),
    hjust = -0.05,
    size = 3
  ) +
  coord_flip(clip = "off") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  labs(x = NULL, y = "Identity-matched OpenAlex journals") +
  theme_minimal()
```

## OpenAlex and Crossref match states

Source absence remains a source-specific `unmatched` or `not_attempted` state;
it is not collapsed into a generic metadata failure.

```{r}
#| label: source-status-heatmap
#| fig-width: 9
#| fig-height: 6

audit_summary |>
  filter(dimension == "source_status_crosstab") |>
  mutate(
    openalex = sub("\\|.*$", "", sub("^openalex=", "", category)),
    crossref = sub("^.*\\|crossref=", "", category)
  ) |>
  ggplot(aes(x = openalex, y = crossref, fill = count)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%s\n%.2f%%", count, percent)), size = 3) +
  labs(x = "OpenAlex match state", y = "Crossref match state", fill = "Rows") +
  theme_minimal()
```

## Purposive ten-row review

This searchable, filterable table uses the fixed V7 sample in its deliberate
display order. It includes every compact master field and omits only columns
whose values are long JSON; complete OpenAlex and Crossref objects remain
available through the source catalogs and checkpoint pointers.

```{r}
#| label: interactive-sample-dataframe

is_long_json_column <- function(column) {
  values <- na.omit(as.character(column))
  any(
    nchar(values) > 200L &
      grepl("^[[:space:]]*[\\[{]", values)
  )
}
long_json_columns <- names(review_result)[
  endsWith(names(review_result), "__json") |
    map_lgl(review_result, is_long_json_column)
]
review_table <- review_result[
  setdiff(names(review_result), long_json_columns)
]

htmltools::div(
  style = "overflow-x: auto;",
  reactable::reactable(
    review_table,
    searchable = TRUE,
    filterable = TRUE,
    pagination = TRUE,
    defaultPageSize = 10,
    showPageSizeOptions = TRUE,
    resizable = TRUE,
    compact = TRUE,
    fullWidth = FALSE
  )
)
```

## Complete merged column list

This table lists every column in the compact master artifact. Long source
objects live in checkpoints instead of being duplicated across PKP rows.

```{r}
#| label: complete-column-list

tibble(column = output_columns) |>
  knitr::kable()
```
