---
title: "PKP/OJS multi-source journal enrichment"
format: html
execute:
  echo: true
  warning: false
  message: false
---

This notebook adds selected OpenAlex and Crossref journal details to ten real PKP OJS rows. It matches records by exact ISSN. The ten rows were chosen to include missing ISSNs, no matches, one match, and multiple matches. This sample checks whether the code works; it does not measure coverage across the full dataset.

Set `OPENALEX_API_KEY` in the environment or the repository-root `.env` file before rendering. This notebook runs the ten-row sample. Before a formal full-cohort run, replace the sections marked `FULL-COHORT REPLACEMENT` with reviewed batching, caching, checkpointing, and error-provenance code.

## Packages and paths

This chunk loads the required R packages and sets the sample input and output paths. It first looks for `OPENALEX_API_KEY` in the current environment and then in the repository-root `.env` file.

The commented lines mark the path replacement required for a formal full-cohort run.

```{r}
#| label: packages-and-paths

library(dplyr)
library(httr2)
library(purrr)
library(readr)
library(tibble)

project_env_path <- file.path("..", "..", "..", ".env")
if (!nzchar(Sys.getenv("OPENALEX_API_KEY")) && file.exists(project_env_path)) {
  readRenviron(project_env_path)
}
if (!nzchar(Sys.getenv("OPENALEX_API_KEY"))) {
  stop("Set OPENALEX_API_KEY before running this notebook.", call. = FALSE)
}

input_path <- file.path(
  "..",
  "data",
  "ojs-journal-enrichment-sample.csv"
)
artifact_dir <- file.path("..", "artifacts", "ojs_journal_enrichment")
run_dir <- file.path(artifact_dir, "sample")
dir.create(run_dir, recursive = TRUE, showWarnings = FALSE)

output_path <- file.path(run_dir, "pkp-ojs-multisource-enriched.csv")

# FULL-COHORT REPLACEMENT:
# data_root <- Sys.getenv("DATA_ROOT")
# input_path <- file.path(data_root, "raw", "pkp-beacon-v6", "beacon.csv")
# run_dir <- file.path(artifact_dir, "full")
```

## ISSN preparation

ISSNs may contain spaces, punctuation, or more than one value. This chunk converts each valid ISSN to the same `XXXX-XXXX` format and checks its final digit. It keeps missing, invalid, and valid-but-unmatched ISSNs as separate cases.

```{r}
#| label: issn-preparation

normalize_issn <- function(value) {
  compact <- gsub("[^0-9X]", "", toupper(trimws(as.character(value))))
  if (!grepl("^[0-9]{7}[0-9X]$", compact)) return(NA_character_)

  digits <- as.integer(strsplit(substr(compact, 1, 7), "")[[1]])
  check_value <- (11 - sum(digits * 8:2) %% 11) %% 11
  expected <- if (check_value == 10) "X" else as.character(check_value)
  if (substr(compact, 8, 8) != expected) return(NA_character_)

  paste0(substr(compact, 1, 4), "-", substr(compact, 5, 8))
}

parse_issns <- function(value) {
  if (is.na(value) || !nzchar(value)) return(character())

  tokens <- unlist(strsplit(value, "[[:space:],;|]+"))
  normalized <- vapply(tokens[nzchar(tokens)], normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}

source_issns <- function(source) {
  values <- c(unlist(source$issn, use.names = FALSE), source$issn_l)
  normalized <- vapply(values, normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}
```

## Source retrieval

The sample sends all valid ISSNs to OpenAlex in one request. It asks Crossref about each ISSN separately and waits between requests. A Crossref `404` means that no journal record was found.

For a formal full-cohort run, replace this chunk. The full version must divide requests into small groups, read every page of results, identify the project to Crossref, retry temporary failures, save progress, and record which ISSNs failed.

```{r}
#| label: source-retrieval

fetch_openalex_sources <- function(
  issns,
  api_key = Sys.getenv("OPENALEX_API_KEY")
) {
  stopifnot(length(issns) <= 100L, nzchar(api_key))

  request("https://api.openalex.org/sources") |>
    req_url_query(
      filter = paste0("issn:", paste(issns, collapse = "|")),
      per_page = 100,
      api_key = api_key
    ) |>
    req_timeout(seconds = 30) |>
    req_perform() |>
    resp_body_json(simplifyVector = FALSE) |>
    pluck("results")
}

fetch_crossref_journal <- function(issn) {
  response <- request(paste0("https://api.crossref.org/journals/", issn)) |>
    req_timeout(seconds = 30) |>
    req_error(
      is_error = \(response) {
        resp_status(response) != 404L && resp_is_error(response)
      }
    ) |>
    req_perform()
  if (resp_status(response) == 404L) return(NULL)
  resp_body_json(response, simplifyVector = FALSE)$message
}

# FULL-COHORT REPLACEMENT:
# - send OpenAlex ISSNs in groups of at most 100 and read every result page;
# - identify Crossref requests with CROSSREF_MAILTO and use its request rate;
# - retry temporary failures, save responses, and record failed ISSNs.
```

## Load the PKP/OJS sample

The sample file was taken from PKP Beacon V6 file `13173372` on 2026-07-17. The MD5 value confirms that the file has not changed. This chunk reads the file, keeps its ten OJS rows, and checks that the same PKP row does not appear twice.

For a formal full-cohort run, replace the sample path, MD5 value, and expected row count with the approved full-dataset values.

```{r}
#| label: load-sample

stopifnot(
  file.exists(input_path),
  unname(tools::md5sum(input_path)) == "b5fb99c10f38c2a5c3e7e6e0d4aa2212"
)

raw_input <- read_csv(
  input_path,
  col_types = cols(.default = col_character()),
  na = character(),
  show_col_types = FALSE
)
ojs_rows <- raw_input |>
  filter(tolower(application) == "ojs") |>
  select(-any_of("admin_email"))
row_identity <- paste(
  ojs_rows$oai_url,
  ojs_rows$repository_name,
  ojs_rows$set_spec,
  sep = "\r"
)
stopifnot(nrow(ojs_rows) == 10L, !anyDuplicated(row_identity))

# FULL-COHORT REPLACEMENT:
# stopifnot(unname(tools::md5sum(input_path)) ==
#   "9f43fa051c7ed1cc45d8592593542011")
# stopifnot(nrow(raw_input) == 87170L, nrow(ojs_rows) == 86282L)
```

## Multi-source enrichment

For each PKP row, the sample looks for OpenAlex and Crossref records with the same ISSN. If the same source record appears more than once, it is counted once. The result uses simple CSV columns that are easy to review.

### Sample matching helpers

These helpers label each row as having no usable ISSN, no match, one match, or multiple matches. They also identify Crossref records by their ISSNs and copy source details when there is exactly one match.

```{r}
#| label: sample-matching-helpers

classify_status <- function(input_issns, candidate_ids) {
  case_when(
    length(input_issns) == 0L ~ "not_attempted",
    length(candidate_ids) == 0L ~ "unmatched",
    length(candidate_ids) == 1L ~ "consistent",
    TRUE ~ "inconsistent"
  )
}

crossref_key <- function(journal) {
  issns <- vapply(journal$ISSN, normalize_issn, character(1))
  paste(sort(unique(issns[!is.na(issns)])), collapse = "|")
}

consistent_field <- function(match, field) {
  if (match$status != "consistent") return(NA_character_)
  value <- pluck(match$candidates, 1, field, .default = NA_character_)
  if (!length(value)) NA_character_ else as.character(value[[1]])
}
```

### Sample enrichment orchestration

This chunk collects the unique sample ISSNs, retrieves source records, matches them to the PKP rows, adds the selected fields, and writes one CSV row for each PKP row.

For a formal full-cohort run, replace this chunk with the full version described in the comments at its end.

```{r}
#| label: sample-enrichment

input_issns <- map(ojs_rows$issn, parse_issns)
tokens <- sort(unique(unlist(input_issns)))

openalex_sources <- fetch_openalex_sources(tokens)
crossref_journals <- set_names(
  map(tokens, \(token) {
    journal <- fetch_crossref_journal(token)
    Sys.sleep(0.25)
    journal
  }),
  tokens
)

openalex_matches <- map(input_issns, \(issns) {
  candidates <- keep(openalex_sources, \(source) {
    any(issns %in% source_issns(source))
  })
  candidate_ids <- map_chr(candidates, "id")
  candidates <- candidates[!duplicated(candidate_ids)]
  candidate_ids <- unique(candidate_ids)

  list(
    status = classify_status(issns, candidate_ids),
    matched_issns = intersect(
      issns,
      unique(unlist(map(candidates, source_issns)))
    ),
    candidate_ids = candidate_ids,
    candidates = candidates
  )
})

crossref_matches <- map(input_issns, \(issns) {
  matched_issns <- keep(issns, \(issn) {
    !is.null(crossref_journals[[issn]])
  })
  candidates <- unname(crossref_journals[matched_issns])
  candidate_ids <- map_chr(candidates, crossref_key)
  candidates <- candidates[!duplicated(candidate_ids)]
  candidate_ids <- unique(candidate_ids)

  list(
    status = classify_status(issns, candidate_ids),
    matched_issns = matched_issns,
    candidate_ids = candidate_ids,
    candidates = candidates
  )
})

result <- ojs_rows |>
  mutate(
    identifier_status = case_when(
      is.na(issn) | !nzchar(issn) ~ "missing",
      lengths(input_issns) == 0L ~ "invalid",
      TRUE ~ "valid"
    ),
    openalex_input_issns = map_chr(input_issns, paste, collapse = "|"),
    openalex_matched_issns = map_chr(
      openalex_matches,
      \(match) paste(match$matched_issns, collapse = "|")
    ),
    openalex_match_status = map_chr(openalex_matches, "status"),
    openalex_candidate_ids = map_chr(
      openalex_matches,
      \(match) paste(match$candidate_ids, collapse = "|")
    ),
    openalex_display_name = map_chr(
      openalex_matches,
      consistent_field,
      field = "display_name"
    ),
    openalex_country_code = map_chr(
      openalex_matches,
      consistent_field,
      field = "country_code"
    ),
    openalex_works_count = map_chr(
      openalex_matches,
      consistent_field,
      field = "works_count"
    ),
    crossref_input_issns = map_chr(input_issns, paste, collapse = "|"),
    crossref_matched_issns = map_chr(
      crossref_matches,
      \(match) paste(match$matched_issns, collapse = "|")
    ),
    crossref_match_status = map_chr(crossref_matches, "status"),
    crossref_candidate_issn_sets = map_chr(
      crossref_matches,
      \(match) paste(match$candidate_ids, collapse = ";")
    ),
    crossref_title = map_chr(
      crossref_matches,
      consistent_field,
      field = "title"
    ),
    crossref_publisher = map_chr(
      crossref_matches,
      consistent_field,
      field = "publisher"
    )
  )

write_csv(result, output_path, na = "")

# FULL-COHORT REPLACEMENT:
# - send OpenAlex requests in groups and read every result page;
# - save API responses so an interrupted run can continue;
# - retry temporary failures and record failed ISSNs;
# - keep the complete source fields and matched records;
# - record file checksums and run details in the metadata file.
```

## Checks

These checks confirm that every input row appears in the output, the required columns exist, the CSV file was written, and the sample contains the expected match outcomes.

```{r}
#| label: contract-check

stopifnot(
  exists("fetch_openalex_sources", mode = "function"),
  exists("fetch_crossref_journal", mode = "function"),
  nrow(result) == nrow(ojs_rows),
  file.exists(output_path),
  all(c(
    "openalex_match_status",
    "openalex_display_name",
    "crossref_match_status",
    "crossref_title"
  ) %in% names(result)),
  all(c(
    "consistent",
    "unmatched",
    "inconsistent",
    "not_attempted"
  ) %in% result$openalex_match_status),
  "not_attempted" %in% result$crossref_match_status
)
```

## Complete merged dataframe

This table displays all OJS, OpenAlex, and Crossref columns in the merged sample dataframe.

```{r}
#| label: complete-sample-dataframe

result |>
  knitr::kable(format = "html")
```

## Complete merged column list

This table lists every column in the CSV so collaborators can review its structure.

```{r}
#| label: complete-column-list

tibble(column = names(result)) |>
  knitr::kable()
```
