---
title: "PKP/OJS multi-source journal enrichment"
format: html
execute:
  echo: true
  warning: true
  message: false
---

This notebook joins the pinned PKP OJS V7 release to retrieved OpenAlex Source
and Crossref journal snapshots by exact ISSN. Sample mode is the default. Full
mode must be started with `render_full.sh`; it preserves all 98,273 PKP rows and
every returned top-level source field in one wide compressed CSV.

Set `OPENALEX_API_KEY` before retrieving uncached OpenAlex data. Full mode also
requires `CROSSREF_MAILTO` when a Crossref page is missing. A cache-only rerun
needs neither value. A normal render and `render_sample.sh` always use sample
mode.

## Packages and paths

This chunk verifies the documented R environment, downloads the pinned PKP V7 input when needed, verifies its official checksum, and keeps every local artifact under this analysis owner.

```{r}
#| label: packages-and-paths

required_r_version <- "4.5.2"
required_quarto_version <- "1.6.32"
quarto <- Sys.which("quarto")
quarto_version <- if (nzchar(quarto)) {
  trimws(system2(quarto, "--version", stdout = TRUE)[1])
} else {
  ""
}
if (
  as.character(getRversion()) != required_r_version ||
    quarto_version != required_quarto_version
) {
  stop(
    paste(
      "This analysis requires R",
      required_r_version,
      "and Quarto",
      required_quarto_version
    ),
    call. = FALSE
  )
}

required_packages <- c(
  "dplyr",
  "httr2",
  "jsonlite",
  "knitr",
  "purrr",
  "readr",
  "reactable",
  "rmarkdown",
  "tibble"
)
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]
if (length(missing_packages)) {
  stop(
    paste(
      "Install the documented R environment before rendering. Missing:",
      paste(missing_packages, collapse = ", ")
    ),
    call. = FALSE
  )
}
lockfile <- jsonlite::read_json("renv.lock", simplifyVector = FALSE)
required_package_versions <- vapply(
  required_packages,
  \(package) lockfile$Packages[[package]]$Version,
  character(1)
)
names(required_package_versions) <- required_packages
installed_package_versions <- vapply(
  required_packages,
  \(package) as.character(packageVersion(package)),
  character(1)
)
if (!identical(installed_package_versions, required_package_versions)) {
  stop(
    "Installed package versions differ from the documented environment.",
    call. = FALSE
  )
}

library(dplyr)
library(httr2)
library(purrr)
library(readr)
library(tibble)

run_mode <- Sys.getenv("OJS_ENRICHMENT_MODE", unset = "sample")
stopifnot(run_mode %in% c("sample", "full"))

project_env_path <- file.path("..", "..", "..", ".env")
if (file.exists(project_env_path)) {
  readRenviron(project_env_path)
}
crossref_mailto <- Sys.getenv("CROSSREF_MAILTO")

pkp_version <- 7L
pkp_url <- paste0(
  "https://dataverse.harvard.edu/api/access/datafile/14084919",
  "?format=original"
)
pkp_md5 <- "3a4ad8ae1ebfcc2b991aaf55b2d82c92"
artifact_dir <- file.path("..", "artifacts", "ojs_journal_enrichment")
input_dir <- file.path(artifact_dir, "pkp-v7")
input_path <- file.path(input_dir, "beacon.csv")
run_dir <- file.path(
  artifact_dir,
  if (run_mode == "full") "full-v7" else "sample"
)
output_path <- file.path(
  run_dir,
  if (run_mode == "full") {
    "pkp-ojs-multisource-enriched.csv.gz"
  } else {
    "pkp-ojs-multisource-enriched.csv"
  }
)
metadata_path <- file.path(run_dir, "run-metadata.json")
openalex_cache_path <- file.path(run_dir, "openalex-sources.rds")
crossref_cache_path <- file.path(run_dir, "crossref-journals.rds")
openalex_batch_dir <- file.path(run_dir, "openalex-batches")
crossref_page_dir <- file.path(run_dir, "crossref-pages")
cache_contract_version <- 1L

dir.create(input_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(run_dir, recursive = TRUE, showWarnings = FALSE)
if (run_mode == "full") {
  dir.create(openalex_batch_dir, recursive = TRUE, showWarnings = FALSE)
  dir.create(crossref_page_dir, recursive = TRUE, showWarnings = FALSE)
}

download_pkp_input <- function(path) {
  if (file.exists(path)) return(invisible(path))

  download_path <- paste0(path, ".download")
  on.exit(unlink(download_path), add = TRUE)
  tryCatch(
    download.file(pkp_url, download_path, mode = "wb", quiet = FALSE),
    error = \(error) {
      stop("Could not download PKP V7.", call. = FALSE)
    }
  )
  if (unname(tools::md5sum(download_path)) != pkp_md5) {
    stop("Downloaded PKP V7 file failed checksum verification.", call. = FALSE)
  }
  if (!file.rename(download_path, path)) {
    stop("Could not move the verified PKP V7 file into place.", call. = FALSE)
  }
  invisible(path)
}

download_pkp_input(input_path)
if (unname(tools::md5sum(input_path)) != pkp_md5) {
  stop("Local PKP V7 file failed checksum verification.", call. = FALSE)
}
```

## ISSN preparation

ISSNs may contain spaces, punctuation, or more than one value. This chunk converts each valid ISSN to the same `XXXX-XXXX` format and checks its final digit. It keeps missing, invalid, and valid-but-unmatched ISSNs as separate cases.

```{r}
#| label: issn-preparation

normalize_issn <- function(value) {
  if (length(value) != 1L || is.na(value)) return(NA_character_)
  compact <- gsub("[^0-9X]", "", toupper(trimws(as.character(value))))
  if (!grepl("^[0-9]{7}[0-9X]$", compact)) return(NA_character_)

  digits <- as.integer(strsplit(substr(compact, 1, 7), "")[[1]])
  check_value <- (11 - sum(digits * 8:2) %% 11) %% 11
  expected <- if (check_value == 10) "X" else as.character(check_value)
  if (substr(compact, 8, 8) != expected) return(NA_character_)

  paste0(substr(compact, 1, 4), "-", substr(compact, 5, 8))
}

is_missing_value <- function(value) {
  is.na(value) |
    !nzchar(trimws(value)) |
    toupper(trimws(value)) == "NA"
}

parse_issns <- function(value) {
  if (is_missing_value(value)) return(character())

  tokens <- unlist(strsplit(value, "[[:space:],;|]+"))
  normalized <- vapply(tokens[nzchar(tokens)], normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}

source_issns <- function(source) {
  values <- unlist(source$issn, use.names = FALSE)
  normalized <- vapply(values, normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}
```

## Source retrieval

Both modes use the same paginated OpenAlex batch retrieval. Full mode splits the
103,017 valid ISSNs into deterministic groups of at most 100 and atomically
checkpoints each group. It also downloads the complete Crossref journal
directory in sequential 1,000-record cursor pages. Exhausted requests stop the
run without replacing the previous formal output; the next run resumes from
completed checkpoints.

Sample mode asks Crossref about each ISSN separately and waits between requests. A Crossref `404` means that no journal record was found.

```{r}
#| label: openalex-retrieval

retryable_response <- function(response) {
  resp_status(response) %in% c(403L, 408L, 429L, 500L, 502L, 503L, 504L)
}

fetch_openalex_page <- function(
  issns,
  cursor,
  api_key = Sys.getenv("OPENALEX_API_KEY")
) {
  if (!nzchar(api_key)) {
    stop("Set OPENALEX_API_KEY before retrieving OpenAlex data.", call. = FALSE)
  }
  stopifnot(length(issns) <= 100L)

  body <- request("https://api.openalex.org/sources") |>
    req_url_query(
      filter = paste0("issn:", paste(issns, collapse = "|")),
      per_page = 200,
      cursor = cursor,
      api_key = api_key
    ) |>
    req_timeout(seconds = 30) |>
    req_retry(
      max_tries = 4L,
      retry_on_failure = TRUE,
      is_transient = retryable_response
    ) |>
    req_perform() |>
    resp_body_json(simplifyVector = FALSE)

  if (
    !is.list(body$results) ||
      !is.list(body$meta) ||
      length(body$meta$count) != 1L ||
      !(is.null(body$meta$next_cursor) ||
        length(body$meta$next_cursor) == 1L)
  ) {
    stop("OpenAlex response contract changed.", call. = FALSE)
  }
  body
}

fetch_openalex_batch <- function(issns) {
  cursor <- "*"
  seen_cursors <- character()
  pages <- list()
  page_log <- list()
  total_results <- NULL

  repeat {
    if (cursor %in% seen_cursors) {
      stop("OpenAlex repeated a cursor.", call. = FALSE)
    }
    seen_cursors <- c(seen_cursors, cursor)

    page <- fetch_openalex_page(issns, cursor)
    results <- page$results
    observed_total <- as.integer(page$meta$count)
    if (is.null(total_results)) {
      total_results <- observed_total
    } else if (!identical(total_results, observed_total)) {
      stop("OpenAlex result count changed during pagination.", call. = FALSE)
    }
    next_cursor <- page$meta$next_cursor
    pages[[length(pages) + 1L]] <- results
    page_log[[length(page_log) + 1L]] <- list(
      result_count = length(results),
      retrieved_at = format(Sys.time(), tz = "UTC", usetz = TRUE)
    )
    if (is.null(next_cursor) || !nzchar(next_cursor)) break
    cursor <- next_cursor
  }

  sources <- unlist(pages, recursive = FALSE)
  if (length(sources) != total_results) {
    stop("OpenAlex result count did not reconcile.", call. = FALSE)
  }

  list(
    status = "complete",
    issns = issns,
    sources = sources,
    pages = page_log,
    pagination_complete = TRUE
  )
}
```

### Crossref retrieval

```{r}
#| label: crossref-retrieval

save_rds_atomic <- function(value, path) {
  temporary <- paste0(path, ".tmp")
  saveRDS(value, temporary)
  if (!file.rename(temporary, path)) {
    unlink(temporary)
    stop("Could not save checkpoint atomically.", call. = FALSE)
  }
  invisible(path)
}

read_rds_or_null <- function(path) {
  if (!file.exists(path)) return(NULL)
  tryCatch(readRDS(path), error = \(error) NULL)
}

fetch_crossref_journal <- function(issn) {
  response <- request(paste0("https://api.crossref.org/journals/", issn)) |>
    req_user_agent(
      "InvisibleResearch journal enrichment; https://github.com/YannJY02/InvisibleResearch"
    ) |>
    req_timeout(seconds = 30) |>
    req_retry(
      max_tries = 4L,
      retry_on_failure = TRUE,
      is_transient = retryable_response
    ) |>
    req_error(
      is_error = \(response) {
        resp_status(response) != 404L && resp_is_error(response)
      }
    ) |>
    req_perform()
  if (resp_status(response) == 404L) return(NULL)
  resp_body_json(response, simplifyVector = FALSE)$message
}

fetch_crossref_page <- function(cursor, mailto) {
  stopifnot(nzchar(cursor), nzchar(mailto))

  body <- request("https://api.crossref.org/journals") |>
    req_url_query(rows = 1000, cursor = cursor, mailto = mailto) |>
    req_user_agent(
      "InvisibleResearch journal enrichment; https://github.com/YannJY02/InvisibleResearch"
    ) |>
    req_timeout(seconds = 60) |>
    req_retry(
      max_tries = 4L,
      retry_on_failure = TRUE,
      is_transient = retryable_response
    ) |>
    req_perform() |>
    resp_body_json(simplifyVector = FALSE) |>
    pluck("message")

  if (!is.list(body$items) || length(body[["total-results"]]) != 1L) {
    stop("Crossref response contract changed.", call. = FALSE)
  }
  body
}

fetch_crossref_directory <- function() {
  cursor <- "*"
  page_index <- 1L
  page_summaries <- list()
  page_files <- character()
  total_results <- NULL

  repeat {
    page_file <- sprintf("page-%04d.rds", page_index)
    checkpoint_path <- file.path(crossref_page_dir, page_file)
    checkpoint <- read_rds_or_null(checkpoint_path)
    reusable <- is.list(checkpoint) &&
      identical(checkpoint$contract_version, cache_contract_version) &&
      identical(checkpoint$page_index, page_index) &&
      identical(checkpoint$cursor, cursor) &&
      identical(checkpoint$rows_per_page, 1000L) &&
      identical(checkpoint$status, "complete") &&
      is.list(checkpoint$items) &&
      identical(checkpoint$item_count, length(checkpoint$items))

    if (reusable) {
      checkpoint$reused <- TRUE
    } else {
      if (!grepl("^[^@[:space:]]+@[^@[:space:]]+$", crossref_mailto)) {
        stop(
          "Set CROSSREF_MAILTO before retrieving Crossref data.",
          call. = FALSE
        )
      }
      page <- fetch_crossref_page(cursor, crossref_mailto)
      checkpoint <- list(
        status = "complete",
        contract_version = cache_contract_version,
        page_index = page_index,
        cursor = cursor,
        rows_per_page = 1000L,
        total_results = as.integer(page[["total-results"]]),
        next_cursor = page[["next-cursor"]],
        items = page$items,
        item_count = length(page$items),
        pagination_complete = length(page$items) < 1000L,
        retrieved_at = format(Sys.time(), tz = "UTC", usetz = TRUE),
        reused = FALSE
      )
      save_rds_atomic(checkpoint, checkpoint_path)
    }

    if (is.null(total_results)) {
      total_results <- checkpoint$total_results
    } else if (!identical(total_results, checkpoint$total_results)) {
      stop(
        "Crossref total-results changed during cursor pagination.",
        call. = FALSE
      )
    }

    page_files <- c(page_files, page_file)
    page_summaries[[length(page_summaries) + 1L]] <- checkpoint[
      setdiff(names(checkpoint), c("cursor", "next_cursor", "items"))
    ]
    retrieved_rows <- sum(map_int(page_summaries, "item_count"))
    if (retrieved_rows > total_results) {
      stop(
        "Crossref pagination exceeded total-results.",
        call. = FALSE
      )
    }
    if (isTRUE(checkpoint$pagination_complete)) break
    if (
      is.null(checkpoint$next_cursor) ||
        !nzchar(checkpoint$next_cursor)
    ) {
      stop(
        "Crossref returned a full page without a next cursor.",
        call. = FALSE
      )
    }

    cursor <- checkpoint$next_cursor
    page_index <- page_index + 1L
    Sys.sleep(0.2)
  }

  directory_rows <- sum(map_int(page_summaries, "item_count"))
  if (directory_rows != total_results) {
    stop(
      paste0(
        "Crossref directory row count mismatch: ",
        directory_rows, " != ", total_results
      ),
      call. = FALSE
    )
  }
  list(
    status = "complete",
    rows_per_page = 1000L,
    directory_rows = directory_rows,
    pages = page_summaries,
    page_files = page_files
  )
}

```

## Load the PKP V7 rows

The pinned PKP V7 file is Dataverse file `14084919`, released on 2026-07-18. This chunk checks the complete 98,273-row baseline, schema, row identity, and identifier counts before either retaining the full cohort or selecting ten fixed source rows in a deliberate display order.

```{r}
#| label: load-pkp-v7

raw_input <- read_csv(
  input_path,
  col_types = cols(.default = col_character()),
  na = character(),
  show_col_types = FALSE
)

expected_schema <- c(
  "context_name",
  "issn",
  "oai_url",
  "application",
  "country",
  "region",
  "total_record_count",
  "record_count_2025",
  "earliest_datestamp",
  "repository_name",
  "set_spec",
  "stats_id",
  "best_doaj_url",
  "last_completed_update",
  "first_beacon",
  "last_beacon",
  "last_oai_response",
  "unresponsive_endpoint",
  "unresponsive_context"
)
row_identity <- function(data) {
  paste(data$oai_url, data$repository_name, data$set_spec, sep = "\r")
}
raw_row_identity <- row_identity(raw_input)

stopifnot(
  nrow(raw_input) == 98273L,
  identical(names(raw_input), expected_schema),
  all(tolower(raw_input$application) == "ojs"),
  !anyDuplicated(raw_row_identity)
)

all_input_issns <- map(raw_input$issn, parse_issns)
missing_identifier <- is_missing_value(raw_input$issn)
stopifnot(
  sum(missing_identifier) == 26189L,
  sum(!is_missing_value(raw_input$best_doaj_url)) == 7149L,
  sum(lengths(all_input_issns) > 0L) == 72084L,
  length(unique(unlist(all_input_issns))) == 103017L
)

sample_spec <- tribble(
  ~context_name, ~oai_url, ~repository_name, ~set_spec,
  "Kuwait Journal of Science",
  "https://journalskuwait.org/kjs/index.php/index/oai",
  "Kuwait Journal of Science",
  "KJS",
  "Cakrawala",
  "https://cakrawalajournal.org/index.php/index/oai",
  "CAKRAWALA",
  "cakrawala",
  "Sintesa: Jurnal Ilmu Pendidikan",
  "https://sintesa.stkip-arrahmaniyah.ac.id/index.php/index/oai",
  "Sintesa: Jurnal Ilmu Pendidikan STKIP Arrahmaniyah Depok",
  "sintesa",
  "CARAKA: Jurnal Teologi Biblika dan Praktika",
  "https://ojs.sttibc.ac.id/index.php/index/oai",
  "CARAKA:Jurnal Teologi",
  "ibc",
  "Jurnal Tata Kelola dan Akuntabilitas Keuangan Negara",
  "https://jurnal.bpk.go.id/index.php/index/oai",
  "Open Journal Systems",
  "TAKEN",
  "JOEEL (Journal of English Education and Literature)",
  "https://journal.stkippamanetalino.ac.id/index.php/index/oai",
  "Rumah Jurnal STKIP Pamane Talino Ngabang",
  "bahasa-inggris",
  "Malaysian Journal of Paediatrics and Child Health",
  "https://mpaeds.my/journals/index.php/index/oai",
  "NA",
  "MJPCH",
  "Jami Scientific Research Quarterly Journal",
  "https://journals.jami.edu.af/index.php/index/oai",
  "Jami University Press",
  "jsrqj",
  "Journal of Polymer & Composites",
  "https://engineeringjournals.stmjournals.in/index.php/index/oai",
  "Engineering Journals",
  "JoPC",
  "JURNAL TEKNIK PERTAMBANGAN",
  "https://e-journal.upr.ac.id/index.php/index/oai",
  "Journal Online Universitas Palangka Raya",
  "JTP"
)
sample_identity <- row_identity(sample_spec)
sample_positions <- match(sample_identity, raw_row_identity)
stopifnot(!anyNA(sample_positions))

sample_rows <- raw_input[sample_positions, ] |>
  select(-any_of("admin_email"))
stopifnot(
  nrow(sample_rows) == 10L,
  identical(sample_rows$context_name, sample_spec$context_name)
)

ojs_rows <- if (run_mode == "full") {
  raw_input |> select(-any_of("admin_email"))
} else {
  sample_rows
}
```

## Multi-source enrichment

For each PKP row, the notebook looks for OpenAlex and Crossref records with the same checksum-valid normalized ISSN. If the same source identity appears more than once, it is counted once. Both modes expand every returned top-level field into the one-row-per-PKP master and preserve nested values and complete candidate records as JSON cells.

### Matching helpers

These helpers label each row as having no usable ISSN, no match, one match, or multiple matches. They also identify Crossref records by their ISSNs and convert every candidate field into a dataframe-safe value without filtering the response schema.

```{r}
#| label: matching-helpers

classify_status <- function(input_issns, candidate_ids) {
  case_when(
    length(input_issns) == 0L ~ "not_attempted",
    length(candidate_ids) == 0L ~ "unmatched",
    length(candidate_ids) == 1L ~ "unique",
    TRUE ~ "ambiguous"
  )
}

crossref_key <- function(journal) {
  issns <- vapply(journal$ISSN, normalize_issn, character(1))
  paste(sort(unique(issns[!is.na(issns)])), collapse = "|")
}

crossref_candidates_from_pages <- function(page_files) {
  candidates <- list()

  for (page_file in page_files) {
    checkpoint <- readRDS(file.path(crossref_page_dir, page_file))
    if (checkpoint$status != "complete") {
      stop("Crossref enrichment encountered an incomplete page.", call. = FALSE)
    }
    for (journal in checkpoint$items) {
      candidates[[length(candidates) + 1L]] <- journal
    }
  }

  candidates
}

deduplicate_candidates <- function(candidates, identity) {
  candidate_ids <- vapply(candidates, identity, character(1))
  keep <- !duplicated(candidate_ids)
  list(
    candidates = candidates[keep],
    candidate_ids = candidate_ids[keep]
  )
}

index_candidates_by_issn <- function(candidates, candidate_issns) {
  index <- new.env(hash = TRUE, parent = emptyenv())
  for (candidate_position in seq_along(candidates)) {
    for (issn in candidate_issns(candidates[[candidate_position]])) {
      positions <- if (exists(issn, index, inherits = FALSE)) {
        get(issn, index, inherits = FALSE)
      } else {
        integer()
      }
      assign(issn, c(positions, candidate_position), index)
    }
  }
  index
}

candidates_for_issns <- function(candidates, index, issns) {
  positions <- unique(unlist(map(issns, \(issn) {
    if (exists(issn, index, inherits = FALSE)) {
      get(issn, index, inherits = FALSE)
    } else {
      integer()
    }
  })))
  candidates[positions]
}

match_exact_issns <- function(
  input_issns,
  candidates,
  candidate_issns,
  candidate_identity,
  deduplicate = TRUE
) {
  matched <- keep(candidates, \(candidate) {
    any(input_issns %in% candidate_issns(candidate))
  })
  resolved <- if (deduplicate) {
    deduplicate_candidates(matched, candidate_identity)
  } else {
    list(
      candidates = matched,
      candidate_ids = map_chr(matched, candidate_identity)
    )
  }
  matched_issns <- intersect(
    input_issns,
    unique(unlist(map(resolved$candidates, candidate_issns)))
  )

  c(
    list(
      status = classify_status(
        input_issns,
        resolved$candidate_ids
      ),
      matched_issns = matched_issns
    ),
    resolved
  )
}

serialize_json <- function(value) {
  as.character(jsonlite::toJSON(
    value,
    auto_unbox = TRUE,
    null = "null",
    na = "null",
    digits = NA
  ))
}

deduplicate_identical_candidates <- function(candidates) {
  fingerprints <- map_chr(candidates, serialize_json)
  candidates[!duplicated(fingerprints)]
}

dataframe_value <- function(value) {
  if (is.null(value)) return(NA_character_)
  if (is.atomic(value) && length(value) == 1L) {
    return(as.character(value))
  }
  serialize_json(value)
}

expand_candidate_fields <- function(matches, prefix, fields) {
  candidates <- map(
    matches,
    \(match) {
      if (length(match$candidates) == 1L) match$candidates[[1]] else list()
    }
  )
  cells <- set_names(
    map(fields, \(field) {
      map_chr(candidates, \(candidate) {
        dataframe_value(candidate[[field]])
      })
    }),
    paste(prefix, fields, sep = "_")
  )
  cells[[paste0(prefix, "_candidates__json")]] <- map_chr(
    matches,
    \(match) {
      if (length(match$candidates) > 1L) {
        serialize_json(match$candidates)
      } else {
        NA_character_
      }
    }
  )
  as_tibble(cells, .name_repair = "check_unique")
}

```

### OpenAlex enrichment

Both modes use the same exact-ISSN rule. Full mode indexes candidates by normalized ISSN so matching 98,273 rows does not repeatedly scan either complete response set.

```{r}
#| label: openalex-enrichment

input_issns <- map(ojs_rows$issn, parse_issns)
tokens <- sort(unique(unlist(input_issns)))

if (run_mode == "full") {
  token_batches <- split(tokens, ceiling(seq_along(tokens) / 100L))
  process_batch <- function(batch_index) {
    batch <- token_batches[[batch_index]]
    checkpoint_path <- file.path(
      openalex_batch_dir,
      sprintf("batch-%04d.rds", batch_index)
    )
    checkpoint <- read_rds_or_null(checkpoint_path)
    reusable <- is.list(checkpoint) &&
      identical(checkpoint$contract_version, cache_contract_version) &&
      identical(checkpoint$pkp_version, pkp_version) &&
      identical(checkpoint$input_md5, pkp_md5) &&
      identical(checkpoint$batch_index, as.integer(batch_index)) &&
      identical(checkpoint$issns, batch) &&
      identical(checkpoint$status, "complete") &&
      is.list(checkpoint$sources) &&
      is.list(checkpoint$pages) &&
      isTRUE(checkpoint$pagination_complete)

    if (reusable) {
      checkpoint$reused <- TRUE
      return(checkpoint)
    }

    checkpoint <- fetch_openalex_batch(batch)
    checkpoint$contract_version <- cache_contract_version
    checkpoint$pkp_version <- pkp_version
    checkpoint$input_md5 <- pkp_md5
    checkpoint$batch_index <- as.integer(batch_index)
    checkpoint$reused <- FALSE
    save_rds_atomic(checkpoint, checkpoint_path)
    checkpoint
  }
  openalex_cluster <- parallel::makeCluster(4L)
  invisible(parallel::clusterEvalQ(openalex_cluster, library(httr2)))
  parallel::clusterExport(
    openalex_cluster,
    c(
      "token_batches",
      "openalex_batch_dir",
      "pkp_version",
      "pkp_md5",
      "cache_contract_version",
      "retryable_response",
      "fetch_openalex_page",
      "fetch_openalex_batch",
      "save_rds_atomic",
      "read_rds_or_null",
      "process_batch"
    ),
    envir = environment()
  )
  batch_results <- tryCatch(
    parallel::parLapplyLB(
      openalex_cluster,
      seq_along(token_batches),
      process_batch
    ),
    finally = parallel::stopCluster(openalex_cluster)
  )
  openalex_cache_hit <- all(map_lgl(batch_results, "reused"))
  openalex_sources <- unlist(
    map(batch_results, "sources"),
    recursive = FALSE
  )
  openalex_snapshot_times <- unlist(map(
    batch_results,
    \(batch) map_chr(batch$pages, "retrieved_at")
  ))
} else {
  openalex_cache <- read_rds_or_null(openalex_cache_path)
  openalex_cache_hit <- is.list(openalex_cache) &&
    identical(openalex_cache$contract_version, cache_contract_version) &&
    identical(openalex_cache$pkp_version, pkp_version) &&
    identical(openalex_cache$input_md5, pkp_md5) &&
    identical(openalex_cache$issns, tokens) &&
    is.list(openalex_cache$sources) &&
    length(openalex_cache$retrieved_at) >= 1L
  openalex_sources <- if (openalex_cache_hit) {
    openalex_cache$sources
  } else {
    batch <- fetch_openalex_batch(tokens)
    sources <- batch$sources
    save_rds_atomic(
      list(
        contract_version = cache_contract_version,
        pkp_version = pkp_version,
        input_md5 = pkp_md5,
        issns = tokens,
        retrieved_at = map_chr(batch$pages, "retrieved_at"),
        sources = sources
      ),
      openalex_cache_path
    )
    sources
  }
  openalex_snapshot_times <- if (openalex_cache_hit) {
    openalex_cache$retrieved_at
  } else {
    map_chr(batch$pages, "retrieved_at")
  }
}
openalex_sources <- deduplicate_candidates(
  openalex_sources,
  \(source) source$id
)$candidates
openalex_index <- index_candidates_by_issn(
  openalex_sources,
  source_issns
)

openalex_matches <- map(input_issns, \(issns) {
  match_exact_issns(
    issns,
    candidates_for_issns(openalex_sources, openalex_index, issns),
    source_issns,
    \(source) source$id
  )
})
openalex_status_counts <- table(map_chr(openalex_matches, "status"))

base_result <- ojs_rows |>
  mutate(
    identifier_status = case_when(
      is_missing_value(issn) ~ "missing",
      lengths(input_issns) == 0L ~ "invalid",
      TRUE ~ "valid"
    ),
    openalex_input_issns = map_chr(input_issns, paste, collapse = "|"),
    openalex_matched_issns = map_chr(
      openalex_matches,
      \(match) paste(match$matched_issns, collapse = "|")
    ),
    openalex_match_status = map_chr(openalex_matches, "status"),
    openalex_candidate_count = map_int(
      openalex_matches,
      \(match) length(match$candidate_ids)
    ),
    openalex_candidate_ids = map_chr(
      openalex_matches,
      \(match) paste(match$candidate_ids, collapse = "|")
    )
  )
```

### Crossref enrichment and wide master

Crossref is matched by the same normalized exact ISSNs. Both modes write every
returned top-level OpenAlex and Crossref field to the same one-row-per-PKP
master. Nested and repeated values are JSON-encoded; all ambiguous candidate
records remain in the candidate JSON column.

```{r}
#| label: crossref-enrichment

if (run_mode == "sample") {
  crossref_cache <- read_rds_or_null(crossref_cache_path)
  crossref_cache_hit <- is.list(crossref_cache) &&
    identical(crossref_cache$contract_version, cache_contract_version) &&
    identical(crossref_cache$pkp_version, pkp_version) &&
    identical(crossref_cache$input_md5, pkp_md5) &&
    identical(crossref_cache$issns, tokens) &&
    is.list(crossref_cache$journals) &&
    length(crossref_cache$retrieved_at) == 1L
  crossref_journals <- if (crossref_cache_hit) {
    crossref_cache$journals
  } else {
    journals <- set_names(
      map(tokens, \(token) {
        journal <- fetch_crossref_journal(token)
        Sys.sleep(0.25)
        journal
      }),
      tokens
    )
    crossref_retrieved_at <- format(Sys.time(), tz = "UTC", usetz = TRUE)
    save_rds_atomic(
      list(
        contract_version = cache_contract_version,
        pkp_version = pkp_version,
        input_md5 = pkp_md5,
        issns = tokens,
        retrieved_at = crossref_retrieved_at,
        journals = journals
      ),
      crossref_cache_path
    )
    journals
  }
  crossref_candidates <- deduplicate_identical_candidates(
    unname(compact(crossref_journals))
  )
  crossref_snapshot_times <- if (crossref_cache_hit) {
    crossref_cache$retrieved_at
  } else {
    crossref_retrieved_at
  }
} else {
  crossref_directory <- fetch_crossref_directory()
  crossref_candidates <- crossref_candidates_from_pages(
    crossref_directory$page_files
  )
  crossref_cache_hit <- all(map_lgl(crossref_directory$pages, "reused"))
  crossref_snapshot_times <- map_chr(
    crossref_directory$pages,
    "retrieved_at"
  )
}

crossref_issns <- function(journal) {
  issns <- vapply(journal$ISSN, normalize_issn, character(1))
  unique(issns[!is.na(issns)])
}
crossref_index <- index_candidates_by_issn(
  crossref_candidates,
  crossref_issns
)
crossref_matches <- map(input_issns, \(issns) {
  match_exact_issns(
    issns,
    candidates_for_issns(
      crossref_candidates,
      crossref_index,
      issns
    ),
    crossref_issns,
    crossref_key,
    deduplicate = FALSE
  )
})
crossref_status_counts <- table(map_chr(crossref_matches, "status"))

# ponytail: in-memory wide assembly fits V7; stream chunks if cohort growth
# makes memory material.
openalex_fields <- sort(unique(unlist(map(openalex_sources, names))))
crossref_fields <- sort(unique(unlist(map(crossref_candidates, names))))
openalex_data <- expand_candidate_fields(
  openalex_matches,
  prefix = "openalex",
  fields = openalex_fields
)
crossref_data <- expand_candidate_fields(
  crossref_matches,
  prefix = "crossref",
  fields = crossref_fields
)
result <- base_result |>
  mutate(
    crossref_input_issns = map_chr(input_issns, paste, collapse = "|"),
    crossref_matched_issns = map_chr(
      crossref_matches,
      \(match) paste(match$matched_issns, collapse = "|")
    ),
    crossref_match_status = map_chr(crossref_matches, "status"),
    crossref_candidate_count = map_int(
      crossref_matches,
      \(match) length(match$candidate_ids)
    ),
    crossref_candidate_issn_sets = map_chr(
      crossref_matches,
      \(match) paste(match$candidate_ids, collapse = ";")
    )
  ) |>
  bind_cols(openalex_data, crossref_data)

output_data <- result
output_rows <- nrow(result)
output_columns <- names(result)
review_result <- if (run_mode == "full") {
  result[sample_positions, ]
} else {
  result
}
```

### Validate and promote

The formal master and metadata are promoted only after retrieval, matching,
privacy, and full-cell round-trip checks pass. Validation failures leave the
previous pair untouched; ordinary promotion failures roll back both files.

```{r}
#| label: pre-promotion-validation

has_source_field <- function(prefix, field) {
  paste(prefix, field, sep = "_") %in% output_columns
}
candidate_json_is_valid <- function(matches, json_values) {
  map2_lgl(matches, json_values, \(match, json_value) {
    candidate_count <- length(match$candidates)
    if (candidate_count <= 1L) return(is.na(json_value))
    parsed <- tryCatch(
      jsonlite::fromJSON(json_value, simplifyVector = FALSE),
      error = \(error) NULL
    )
    is.list(parsed) && length(parsed) == candidate_count
  })
}
artifact_root <- normalizePath(artifact_dir)

stopifnot(
  startsWith(normalizePath(input_path), artifact_root),
  startsWith(normalizePath(output_path, mustWork = FALSE), artifact_root),
  startsWith(normalizePath(metadata_path, mustWork = FALSE), artifact_root),
  !"admin_email" %in% output_columns,
  output_rows == nrow(ojs_rows),
  all(nzchar(c(openalex_snapshot_times, crossref_snapshot_times))),
  all(c(
    "openalex_match_status",
    "openalex_candidate_count",
    "openalex_candidate_ids",
    "openalex_candidates__json",
    "crossref_match_status",
    "crossref_candidate_count",
    "crossref_candidate_issn_sets",
    "crossref_candidates__json"
  ) %in% output_columns),
  all(map_lgl(openalex_fields, \(field) {
    has_source_field("openalex", field)
  })),
  all(map_lgl(crossref_fields, \(field) {
    has_source_field("crossref", field)
  })),
  sum(openalex_status_counts) == output_rows,
  sum(crossref_status_counts) == output_rows,
  identical(
    result$openalex_match_status,
    map2_chr(input_issns, openalex_matches, \(issns, match) {
      classify_status(issns, match$candidate_ids)
    })
  ),
  identical(
    result$crossref_match_status,
    map2_chr(input_issns, crossref_matches, \(issns, match) {
      classify_status(issns, match$candidate_ids)
    })
  ),
  all(candidate_json_is_valid(
    openalex_matches,
    output_data$openalex_candidates__json
  )),
  all(candidate_json_is_valid(
    crossref_matches,
    output_data$crossref_candidates__json
  ))
)

if (run_mode == "sample") {
  stopifnot(
    identical(row_identity(ojs_rows), sample_identity),
    sum(result$identifier_status == "missing") == 1L,
    nrow(result) == 10L,
    !anyDuplicated(map_chr(crossref_candidates, serialize_json))
  )
} else {
  stopifnot(
    identical(row_identity(ojs_rows), raw_row_identity),
    output_rows == 98273L,
    sum(base_result$identifier_status == "missing") == 26189L,
    length(tokens) == 103017L,
    length(token_batches) == 1031L,
    max(lengths(token_batches)) <= 100L,
    all(map_lgl(batch_results, \(batch) {
      identical(batch$pkp_version, 7L) &&
        identical(batch$input_md5, pkp_md5) &&
        identical(batch$status, "complete") &&
        isTRUE(batch$pagination_complete) &&
        length(batch$pages) >= 1L
    })),
    all(result$crossref_match_status %in% c(
      "unique",
      "ambiguous",
      "unmatched",
      "not_attempted"
    )),
    endsWith(output_path, ".csv.gz")
  )
}

temporary_output_path <- paste0(
  output_path,
  ".tmp",
  if (endsWith(output_path, ".gz")) ".gz" else ""
)
unlink(temporary_output_path)
write_csv(output_data, temporary_output_path, na = "")

observed_output <- read_csv(
  temporary_output_path,
  col_types = cols(.default = col_character()),
  na = character(),
  show_col_types = FALSE
)
expected_output <- as_tibble(map(output_data, \(column) {
  values <- as.character(column)
  values[is.na(values)] <- ""
  values
}))
observed_identity <- row_identity(observed_output)
stopifnot(
  identical(as.data.frame(observed_output), as.data.frame(expected_output)),
  identical(observed_identity, row_identity(ojs_rows)),
  !anyDuplicated(observed_identity)
)

run_metadata <- list(
  producer = "analysis/ojs_journal_enrichment.qmd",
  run_mode = run_mode,
  created_at = format(Sys.time(), tz = "UTC", usetz = TRUE),
  code = list(
    qmd_md5 = unname(tools::md5sum(knitr::current_input(dir = TRUE))),
    renv_lock_md5 = unname(tools::md5sum("renv.lock")),
    cache_contract_version = cache_contract_version
  ),
  input = list(
    version = pkp_version,
    dataverse_file_id = 14084919L,
    md5 = pkp_md5,
    rows = nrow(raw_input),
    valid_identifier_rows = sum(lengths(all_input_issns) > 0L),
    missing_identifier_rows = sum(missing_identifier),
    distinct_valid_issns = length(unique(unlist(all_input_issns)))
  ),
  output = list(
    file = basename(output_path),
    md5 = unname(tools::md5sum(temporary_output_path)),
    format = if (run_mode == "full") "csv.gz" else "csv",
    rows = output_rows,
    columns = length(output_columns),
    retained_top_level_fields = list(
      openalex = openalex_fields,
      crossref = crossref_fields
    ),
    status_counts = list(
      openalex = as.list(openalex_status_counts),
      crossref = as.list(crossref_status_counts)
    )
  ),
  retrieval = list(
    openalex = list(
      first = min(openalex_snapshot_times),
      last = max(openalex_snapshot_times)
    ),
    crossref = list(
      first = min(crossref_snapshot_times),
      last = max(crossref_snapshot_times)
    )
  ),
  openalex = if (run_mode == "full") {
    list(
      records = length(openalex_sources),
      batches = length(batch_results),
      reused_batches = sum(map_lgl(batch_results, "reused"))
    )
  } else {
    list(records = length(openalex_sources), reused = openalex_cache_hit)
  },
  crossref = if (run_mode == "full") {
    list(
      records = length(crossref_candidates),
      pages = length(crossref_directory$pages),
      rows_per_page = crossref_directory$rows_per_page,
      reused_pages = sum(map_lgl(crossref_directory$pages, "reused"))
    )
  } else {
    list(records = length(crossref_candidates), reused = crossref_cache_hit)
  },
  runtime = list(
    r = R.version.string,
    quarto = system2("quarto", "--version", stdout = TRUE)[1],
    locale = Sys.getlocale(),
    packages = as.list(installed_package_versions)
  )
)

metadata_temporary_path <- paste0(metadata_path, ".tmp")
unlink(metadata_temporary_path)
jsonlite::write_json(
  run_metadata,
  metadata_temporary_path,
  auto_unbox = TRUE,
  pretty = TRUE
)
metadata_text <- paste(
  readLines(metadata_temporary_path, warn = FALSE),
  collapse = "\n"
)
private_values <- c(Sys.getenv("OPENALEX_API_KEY"), crossref_mailto)
private_values <- private_values[nzchar(private_values)]
stopifnot(
  !grepl("OPENALEX_API_KEY|CROSSREF_MAILTO", metadata_text),
  !any(vapply(
    private_values,
    \(value) grepl(value, metadata_text, fixed = TRUE),
    logical(1)
  ))
)

promote_artifact_pair <- function(temporary_paths, final_paths) {
  backup_paths <- paste0(final_paths, ".previous")
  if (any(file.exists(backup_paths))) {
    stop("Resolve stale .previous artifacts before promotion.", call. = FALSE)
  }

  backed_up <- promoted <- rep(FALSE, length(final_paths))
  complete <- FALSE
  on.exit({
    if (!complete) {
      unlink(final_paths[promoted])
      for (position in rev(which(backed_up))) {
        if (!file.rename(backup_paths[position], final_paths[position])) {
          warning("Could not restore a previous formal artifact.")
        }
      }
    }
  }, add = TRUE)

  for (position in seq_along(final_paths)) {
    if (file.exists(final_paths[position])) {
      if (!file.rename(final_paths[position], backup_paths[position])) {
        stop("Could not back up a previous formal artifact.", call. = FALSE)
      }
      backed_up[position] <- TRUE
    }
  }
  for (position in seq_along(final_paths)) {
    if (!file.rename(temporary_paths[position], final_paths[position])) {
      stop("Could not promote validated artifacts.", call. = FALSE)
    }
    promoted[position] <- TRUE
  }

  complete <- TRUE
  unlink(backup_paths[backed_up])
  invisible(final_paths)
}

promote_artifact_pair(
  c(temporary_output_path, metadata_temporary_path),
  c(output_path, metadata_path)
)
```

## Run summary

Counts use all rows in the selected run mode. They describe identifier and
candidate availability, not global OJS coverage.

```{r}
#| label: run-summary

bind_rows(
  tibble(
    source = "OpenAlex",
    status = names(openalex_status_counts),
    count = as.integer(openalex_status_counts)
  ),
  tibble(
    source = "Crossref",
    status = names(crossref_status_counts),
    count = as.integer(crossref_status_counts)
  )
) |>
  knitr::kable()
```

## Purposive ten-row review

This searchable, filterable table uses ten rows chosen to exercise known code
paths. It is not representative and cannot estimate coverage or source
differences. The table omits long JSON columns; those values remain in the
wide master.

```{r}
#| label: interactive-sample-dataframe

is_long_json_column <- function(column) {
  values <- na.omit(as.character(column))
  any(
    nchar(values) > 200L &
      grepl("^[[:space:]]*[\\[{]", values)
  )
}
long_json_columns <- names(review_result)[
  endsWith(names(review_result), "__json") |
    map_lgl(review_result, is_long_json_column)
]
review_table <- review_result[
  setdiff(names(review_result), long_json_columns)
]

htmltools::div(
  style = "overflow-x: auto;",
  reactable::reactable(
    review_table,
    searchable = TRUE,
    filterable = TRUE,
    pagination = TRUE,
    defaultPageSize = 10,
    showPageSizeOptions = TRUE,
    resizable = TRUE,
    compact = TRUE,
    fullWidth = FALSE
  )
)
```

## Complete merged column list

This table lists every column in the wide master artifact.

```{r}
#| label: complete-column-list

tibble(column = output_columns) |>
  knitr::kable()
```
