---
title: "PKP–OpenAlex disagreement analysis"
format: html
execute:
  echo: true
  warning: true
  message: false
---

This exploratory analysis reads the validated PKP V7 wide master produced by
`ojs_journal_enrichment.qmd`. It compares source-specific evidence without
treating either source as authoritative and without interpreting source absence
as a causal measure of journal invisibility.

## Load the wide master

```{r}
#| label: load-master

required_packages <- c(
  "countrycode",
  "dplyr",
  "ggplot2",
  "jsonlite",
  "purrr",
  "readr",
  "reactable",
  "tibble"
)
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]
if (length(missing_packages)) {
  stop(
    paste("Install the documented R environment. Missing:",
      paste(missing_packages, collapse = ", ")),
    call. = FALSE
  )
}

library(dplyr)
library(ggplot2)
library(purrr)
library(readr)
library(tibble)

enrichment_artifact_dir <- file.path(
  "..",
  "artifacts",
  "ojs_journal_enrichment"
)
master_path <- file.path(
  enrichment_artifact_dir,
  "full-v7",
  "pkp-ojs-multisource-enriched.csv.gz"
)
analysis_dir <- file.path(
  "..",
  "artifacts",
  "ojs_journal_disagreement_analysis"
)
audit_path <- file.path(
  analysis_dir,
  "pkp-openalex-disagreement-audit.csv.gz"
)
audit_summary_path <- file.path(
  analysis_dir,
  "pkp-openalex-disagreement-summary.csv"
)
dir.create(analysis_dir, recursive = TRUE, showWarnings = FALSE)

if (!file.exists(master_path)) {
  stop("Run render_full.sh before this analysis.", call. = FALSE)
}
master <- read_csv(
  master_path,
  col_types = cols(.default = col_character()),
  na = character(),
  show_col_types = FALSE
)

required_columns <- c(
  "oai_url",
  "repository_name",
  "set_spec",
  "context_name",
  "issn",
  "country",
  "best_doaj_url",
  "identifier_status",
  "openalex_input_issns",
  "openalex_matched_issns",
  "openalex_match_status",
  "openalex_candidate_count",
  "openalex_candidate_ids",
  "openalex_type",
  "openalex_display_name",
  "openalex_alternate_titles",
  "openalex_issn",
  "openalex_issn_l",
  "openalex_is_ojs",
  "openalex_is_in_doaj",
  "openalex_country_code",
  "crossref_match_status",
  "crossref_candidate_count",
  "crossref_candidate_issn_sets"
)
row_identity <- function(data) {
  paste(data$oai_url, data$repository_name, data$set_spec, sep = "\r")
}
stopifnot(
  nrow(master) == 98273L,
  all(required_columns %in% names(master)),
  !anyDuplicated(row_identity(master)),
  !"admin_email" %in% names(master)
)
```

## Classify identity and metadata differences

Only rows with one exact-ISSN OpenAlex journal identity enter the metadata
comparisons. Missing or ineligible evidence remains `not_observed`.

```{r}
#| label: classify-disagreements

is_missing_value <- function(value) {
  is.na(value) |
    !nzchar(trimws(value)) |
    toupper(trimws(value)) == "NA"
}

normalize_issn <- function(value) {
  if (length(value) != 1L || is.na(value)) return(NA_character_)
  compact <- gsub("[^0-9X]", "", toupper(trimws(as.character(value))))
  if (!grepl("^[0-9]{7}[0-9X]$", compact)) return(NA_character_)

  digits <- as.integer(strsplit(substr(compact, 1, 7), "")[[1]])
  check_value <- (11 - sum(digits * 8:2) %% 11) %% 11
  expected <- if (check_value == 10) "X" else as.character(check_value)
  if (substr(compact, 8, 8) != expected) return(NA_character_)

  paste0(substr(compact, 1, 4), "-", substr(compact, 5, 8))
}

json_values <- function(value) {
  if (is_missing_value(value)) return(character())
  parsed <- tryCatch(
    jsonlite::fromJSON(value, simplifyVector = TRUE),
    error = function(error) {
      stop("Wide master contains invalid source JSON.", call. = FALSE)
    }
  )
  as.character(unlist(parsed, use.names = FALSE))
}

pipe_values <- function(value) {
  if (is_missing_value(value)) return(character())
  strsplit(value, "|", fixed = TRUE)[[1]]
}

normalize_title <- function(value) {
  if (is_missing_value(value)) return(NA_character_)
  gsub("[[:punct:][:space:]]", "", tolower(enc2utf8(value)))
}

comparison_state <- function(left, right) {
  if (is_missing_value(left) || is_missing_value(right)) {
    return("not_observed")
  }
  if (identical(left, right)) "same" else "different"
}
stopifnot(
  comparison_state("", "US") == "not_observed",
  comparison_state("US", "") == "not_observed",
  comparison_state("US", "US") == "same",
  comparison_state("US", "GB") == "different"
)

category_summary <- function(
  values,
  categories,
  dimension,
  denominator_name
) {
  counts <- table(factor(values, levels = categories))
  denominator <- length(values)
  tibble(
    dimension = dimension,
    category = names(counts),
    denominator_name = denominator_name,
    denominator = denominator,
    count = as.integer(counts),
    percent = round(100 * count / denominator, 3)
  )
}

candidate_count <- as.integer(master$openalex_candidate_count)
input_issn_sets <- map_chr(master$openalex_input_issns, \(value) {
  paste(sort(pipe_values(value)), collapse = "|")
})
matched_issn_sets <- map_chr(master$openalex_matched_issns, \(value) {
  paste(sort(pipe_values(value)), collapse = "|")
})
openalex_source_issn_sets <- map2_chr(
  master$openalex_issn,
  master$openalex_issn_l,
  \(issns, issn_l) {
    values <- c(json_values(issns), issn_l)
    normalized <- vapply(values, normalize_issn, character(1))
    paste(sort(unique(normalized[!is.na(normalized)])), collapse = "|")
  }
)

identity_outcome <- case_when(
  master$identifier_status != "valid" ~ "no_valid_issn",
  master$openalex_match_status == "api_error" ~ "api_error",
  candidate_count == 0L ~ "valid_issn_unmatched",
  candidate_count > 1L ~ "conflicting_source_ids",
  is_missing_value(master$openalex_type) |
    tolower(master$openalex_type) != "journal" ~ "non_journal_source",
  input_issn_sets != matched_issn_sets ~ "partial_issn_agreement",
  TRUE ~ "consistent"
)
identity_matched <- identity_outcome %in% c(
  "consistent",
  "partial_issn_agreement"
)

pkp_normalized_titles <- map_chr(master$context_name, normalize_title)
openalex_normalized_titles <- map2(
  master$openalex_display_name,
  master$openalex_alternate_titles,
  \(display_name, alternate_titles) {
    titles <- c(display_name, json_values(alternate_titles))
    normalized <- unique(na.omit(map_chr(titles, normalize_title)))
    normalized[nzchar(normalized)]
  }
)
title_comparison <- map2_chr(
  pkp_normalized_titles,
  openalex_normalized_titles,
  \(pkp_title, openalex_titles) {
    if (!length(openalex_titles) || is.na(pkp_title)) return("not_observed")
    if (pkp_title %in% openalex_titles) "same" else "different"
  }
)
title_comparison[!identity_matched] <- "not_observed"

issn_set_comparison <- map2_chr(
  input_issn_sets,
  openalex_source_issn_sets,
  comparison_state
)
issn_set_comparison[!identity_matched] <- "not_observed"

ojs_evidence_comparison <- case_when(
  !identity_matched | is_missing_value(master$openalex_is_ojs) ~ "not_observed",
  master$openalex_is_ojs == "TRUE" ~ "same",
  TRUE ~ "different"
)

pkp_doaj_evidence <- !is_missing_value(master$best_doaj_url)
openalex_doaj_evidence <- master$openalex_is_in_doaj
doaj_evidence_comparison <- case_when(
  !identity_matched | is_missing_value(openalex_doaj_evidence) ~ "not_observed",
  pkp_doaj_evidence & openalex_doaj_evidence == "TRUE" ~ "both_observed",
  pkp_doaj_evidence ~ "pkp_only",
  openalex_doaj_evidence == "TRUE" ~ "openalex_only",
  TRUE ~ "neither_observed"
)

pkp_country_missing <- is_missing_value(master$country)
pkp_country_code <- countrycode::countrycode(
  master$country,
  origin = "country.name",
  destination = "iso2c",
  warn = FALSE
)
openalex_country_code <- toupper(master$openalex_country_code)
country_evidence_comparison <- map2_chr(
  pkp_country_code,
  openalex_country_code,
  comparison_state
)
country_evidence_comparison[!identity_matched] <- "not_observed"

audit <- master |>
  transmute(
    oai_url,
    repository_name,
    set_spec,
    context_name,
    issn,
    identifier_status,
    openalex_input_issns,
    openalex_matched_issns,
    openalex_match_status,
    openalex_candidate_count,
    openalex_candidate_ids,
    crossref_match_status,
    crossref_candidate_count,
    crossref_candidate_issn_sets,
    identity_outcome,
    openalex_source_type = openalex_type,
    openalex_display_name,
    pkp_normalized_title = pkp_normalized_titles,
    openalex_normalized_titles = map_chr(
      openalex_normalized_titles,
      paste,
      collapse = "|"
    ),
    title_comparison,
    pkp_issn_set = input_issn_sets,
    openalex_issn_set = openalex_source_issn_sets,
    issn_set_comparison,
    ojs_evidence_comparison,
    pkp_doaj_evidence,
    openalex_doaj_evidence,
    doaj_evidence_comparison,
    pkp_country = country,
    pkp_country_code,
    openalex_country_code,
    country_evidence_comparison
  )
```

## Summarize and validate

```{r}
#| label: summarize-and-validate

identity_categories <- c(
  "consistent",
  "partial_issn_agreement",
  "conflicting_source_ids",
  "non_journal_source",
  "valid_issn_unmatched",
  "no_valid_issn",
  "api_error"
)
comparison_categories <- c("same", "different", "not_observed")
status_categories <- c(
  "consistent",
  "inconsistent",
  "unmatched",
  "not_attempted",
  "api_error"
)
matched_denominator <- sum(identity_matched)

source_status_counts <- as_tibble(as.data.frame(table(
  openalex = factor(
    audit$openalex_match_status,
    levels = status_categories
  ),
  crossref = factor(
    audit$crossref_match_status,
    levels = status_categories
  )
))) |>
  transmute(
    dimension = "source_status_crosstab",
    category = paste0("openalex=", openalex, "|crossref=", crossref),
    denominator_name = "all_pkp_v7_rows",
    denominator = nrow(audit),
    count = as.integer(Freq),
    percent = round(100 * count / denominator, 3)
  )

metadata_disagreement_counts <- c(
  normalized_title = sum(title_comparison == "different"),
  issn_set = sum(issn_set_comparison == "different"),
  ojs_evidence = sum(ojs_evidence_comparison == "different"),
  doaj_evidence = sum(doaj_evidence_comparison %in% c(
    "pkp_only",
    "openalex_only"
  )),
  country_evidence = sum(country_evidence_comparison == "different")
)
metadata_disagreement_summary <- tibble(
  dimension = "metadata_disagreement",
  category = names(metadata_disagreement_counts),
  denominator_name = "identity_matched_openalex_journals",
  denominator = matched_denominator,
  count = as.integer(metadata_disagreement_counts),
  percent = round(100 * count / denominator, 3)
)

audit_summary <- bind_rows(
  category_summary(
    identity_outcome,
    identity_categories,
    "identity_outcome",
    "all_pkp_v7_rows"
  ),
  category_summary(
    title_comparison[identity_matched],
    comparison_categories,
    "normalized_title",
    "identity_matched_openalex_journals"
  ),
  category_summary(
    issn_set_comparison[identity_matched],
    comparison_categories,
    "issn_set",
    "identity_matched_openalex_journals"
  ),
  category_summary(
    ojs_evidence_comparison[identity_matched],
    comparison_categories,
    "ojs_evidence",
    "identity_matched_openalex_journals"
  ),
  category_summary(
    doaj_evidence_comparison[identity_matched],
    c(
      "both_observed",
      "pkp_only",
      "openalex_only",
      "neither_observed",
      "not_observed"
    ),
    "doaj_evidence",
    "identity_matched_openalex_journals"
  ),
  category_summary(
    country_evidence_comparison[identity_matched],
    comparison_categories,
    "country_evidence",
    "identity_matched_openalex_journals"
  ),
  metadata_disagreement_summary,
  source_status_counts
)

metadata_dimensions <- c(
  "normalized_title",
  "issn_set",
  "ojs_evidence",
  "doaj_evidence",
  "country_evidence"
)
metadata_reconciliation <- audit_summary |>
  filter(dimension %in% metadata_dimensions) |>
  group_by(dimension) |>
  summarise(
    denominator = unique(denominator),
    count = sum(count),
    .groups = "drop"
  )
stopifnot(
  nrow(audit) == nrow(master),
  identical(audit$identity_outcome, identity_outcome),
  sum(audit_summary$count[
    audit_summary$dimension == "identity_outcome"
  ]) == nrow(master),
  nrow(metadata_reconciliation) == length(metadata_dimensions),
  all(metadata_reconciliation$denominator == matched_denominator),
  all(metadata_reconciliation$count == matched_denominator),
  sum(audit_summary$count[
    audit_summary$dimension == "source_status_crosstab"
  ]) == nrow(master),
  all(is.finite(audit_summary$percent))
)

write_csv_atomic <- function(data, path) {
  temporary_path <- paste0(
    path,
    ".tmp",
    if (endsWith(path, ".gz")) ".gz" else ""
  )
  unlink(temporary_path)
  write_csv(data, temporary_path, na = "")
  observed <- read_csv(
    temporary_path,
    col_types = cols(.default = col_character()),
    na = character(),
    show_col_types = FALSE
  )
  stopifnot(
    nrow(observed) == nrow(data),
    identical(names(observed), names(data))
  )
  if (!file.rename(temporary_path, path)) {
    stop("Could not promote validated analysis output.", call. = FALSE)
  }
}
write_csv_atomic(audit, audit_path)
write_csv_atomic(audit_summary, audit_summary_path)
stopifnot(file.exists(audit_path), file.exists(audit_summary_path))
```

## Identity outcomes

These categories describe exact-ISSN identity evidence. They do not establish
which source is correct and do not explain why a difference exists.

```{r}
#| label: identity-outcome-chart
#| fig-width: 9
#| fig-height: 5

audit_summary |>
  filter(dimension == "identity_outcome") |>
  ggplot(aes(x = reorder(category, count), y = count)) +
  geom_col() +
  geom_text(
    aes(label = sprintf("%s (%.2f%%)", count, percent)),
    hjust = -0.05,
    size = 3
  ) +
  coord_flip(clip = "off") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  labs(x = NULL, y = "PKP V7 rows") +
  theme_minimal()
```

## Metadata comparisons

Title comparison uses the PKP title against the OpenAlex display name and
alternate titles. PKP country is inferred metadata, not authoritative
publisher location. Absent DOAJ evidence is not a negative assertion.

```{r}
#| label: metadata-disagreement-chart
#| fig-width: 9
#| fig-height: 4.5

metadata_disagreement_summary |>
  ggplot(aes(x = reorder(category, count), y = count)) +
  geom_col() +
  geom_text(
    aes(label = sprintf("%s (%.2f%%)", count, percent)),
    hjust = -0.05,
    size = 3
  ) +
  coord_flip(clip = "off") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  labs(x = NULL, y = "Identity-matched OpenAlex journals") +
  theme_minimal()
```

## OpenAlex and Crossref match states

```{r}
#| label: source-status-heatmap
#| fig-width: 9
#| fig-height: 6

audit_summary |>
  filter(dimension == "source_status_crosstab") |>
  mutate(
    openalex = sub("\\|.*$", "", sub("^openalex=", "", category)),
    crossref = sub("^.*\\|crossref=", "", category)
  ) |>
  ggplot(aes(x = openalex, y = crossref, fill = count)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%s\n%.2f%%", count, percent)), size = 3) +
  labs(x = "OpenAlex match state", y = "Crossref match state", fill = "Rows") +
  theme_minimal()
```

## Purposive ten-row review

```{r}
#| label: disagreement-review

sample_context_names <- c(
  "Kuwait Journal of Science",
  "Cakrawala",
  "Sintesa: Jurnal Ilmu Pendidikan",
  "CARAKA: Jurnal Teologi Biblika dan Praktika",
  "Jurnal Tata Kelola dan Akuntabilitas Keuangan Negara",
  "JOEEL (Journal of English Education and Literature)",
  "Malaysian Journal of Paediatrics and Child Health",
  "Jami Scientific Research Quarterly Journal",
  "Journal of Polymer & Composites",
  "JURNAL TEKNIK PERTAMBANGAN"
)
review_rows <- audit[
  match(sample_context_names, audit$context_name),
]
stopifnot(!anyNA(review_rows$context_name))

htmltools::div(
  style = "overflow-x: auto;",
  reactable::reactable(
    review_rows,
    searchable = TRUE,
    filterable = TRUE,
    pagination = TRUE,
    defaultPageSize = 10,
    resizable = TRUE,
    compact = TRUE,
    fullWidth = FALSE
  )
)
```
