---
title: "PKP/OJS multi-source journal enrichment"
format: html
execute:
  echo: true
  warning: true
  message: false
---

This notebook adds current OpenAlex Source and Crossref journal data to PKP OJS rows from the pinned V7 release by exact ISSN. Sample mode is the default. Full mode must be started with `render_full.sh`; it checkpoints both source directories and preserves all 98,273 PKP rows plus every returned top-level source field in one wide compressed CSV.

Set `OPENALEX_API_KEY` before retrieving uncached OpenAlex data. Full mode also requires `CROSSREF_MAILTO` when a Crossref page is missing. A cache-only reproduction needs neither value. A normal render and `render_sample.sh` always use sample mode.

## Packages and paths

This chunk verifies the documented R environment, downloads the pinned PKP V7 input when needed, verifies its official checksum, and keeps every local artifact under this analysis owner.

```{r}
#| label: packages-and-paths

required_r_version <- "4.5.2"
required_quarto_version <- "1.6.32"
quarto <- Sys.which("quarto")
quarto_version <- if (nzchar(quarto)) {
  trimws(system2(quarto, "--version", stdout = TRUE)[1])
} else {
  ""
}
if (
  as.character(getRversion()) != required_r_version ||
    quarto_version != required_quarto_version
) {
  stop(
    paste(
      "This analysis requires R",
      required_r_version,
      "and Quarto",
      required_quarto_version
    ),
    call. = FALSE
  )
}

required_packages <- c(
  "dplyr",
  "httr2",
  "jsonlite",
  "knitr",
  "purrr",
  "readr",
  "reactable",
  "rmarkdown",
  "tibble"
)
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]
if (length(missing_packages)) {
  stop(
    paste(
      "Install the documented R environment before rendering. Missing:",
      paste(missing_packages, collapse = ", ")
    ),
    call. = FALSE
  )
}
lockfile <- jsonlite::read_json("renv.lock", simplifyVector = FALSE)
required_package_versions <- vapply(
  required_packages,
  \(package) lockfile$Packages[[package]]$Version,
  character(1)
)
names(required_package_versions) <- required_packages
installed_package_versions <- vapply(
  required_packages,
  \(package) as.character(packageVersion(package)),
  character(1)
)
if (!identical(installed_package_versions, required_package_versions)) {
  stop(
    "Installed package versions differ from the documented environment.",
    call. = FALSE
  )
}

library(dplyr)
library(httr2)
library(purrr)
library(readr)
library(tibble)

run_mode <- Sys.getenv("OJS_ENRICHMENT_MODE", unset = "sample")
stopifnot(run_mode %in% c("sample", "full"))

project_env_path <- file.path("..", "..", "..", ".env")
if (file.exists(project_env_path)) {
  readRenviron(project_env_path)
}
crossref_mailto <- Sys.getenv("CROSSREF_MAILTO")

pkp_version <- 7L
pkp_url <- paste0(
  "https://dataverse.harvard.edu/api/access/datafile/14084919",
  "?format=original"
)
pkp_md5 <- "3a4ad8ae1ebfcc2b991aaf55b2d82c92"
artifact_dir <- file.path("..", "artifacts", "ojs_journal_enrichment")
input_dir <- file.path(artifact_dir, "pkp-v7")
input_path <- file.path(input_dir, "beacon.csv")
run_dir <- file.path(
  artifact_dir,
  if (run_mode == "full") "full-v7" else "sample"
)
output_path <- file.path(
  run_dir,
  if (run_mode == "full") {
    "pkp-ojs-multisource-enriched.csv.gz"
  } else {
    "pkp-ojs-multisource-enriched.csv"
  }
)
metadata_path <- file.path(run_dir, "run-metadata.json")
openalex_cache_path <- file.path(run_dir, "openalex-sources.rds")
crossref_cache_path <- file.path(run_dir, "crossref-journals.rds")
openalex_batch_dir <- file.path(run_dir, "openalex-batches")
crossref_page_dir <- file.path(run_dir, "crossref-pages")
cache_contract_version <- 1L

dir.create(input_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(run_dir, recursive = TRUE, showWarnings = FALSE)
if (run_mode == "full") {
  dir.create(openalex_batch_dir, recursive = TRUE, showWarnings = FALSE)
  dir.create(crossref_page_dir, recursive = TRUE, showWarnings = FALSE)
}

curl <- Sys.which("curl")
if (!nzchar(curl)) {
  stop("curl is required to download the pinned PKP V7 input.", call. = FALSE)
}

download_pkp_input <- function(path) {
  if (file.exists(path)) return(invisible(path))

  download_path <- paste0(path, ".download")
  on.exit(unlink(download_path), add = TRUE)
  status <- system2(
    curl,
    c(
      "-fL",
      "--retry", "5",
      "--retry-all-errors",
      "--retry-delay", "2",
      "--continue-at", "-",
      "-o", shQuote(download_path),
      shQuote(pkp_url)
    )
  )
  if (status != 0L) {
    stop("Could not download PKP V7.", call. = FALSE)
  }
  if (unname(tools::md5sum(download_path)) != pkp_md5) {
    stop("Downloaded PKP V7 file failed checksum verification.", call. = FALSE)
  }
  if (!file.rename(download_path, path)) {
    stop("Could not move the verified PKP V7 file into place.", call. = FALSE)
  }
  invisible(path)
}

download_pkp_input(input_path)
if (unname(tools::md5sum(input_path)) != pkp_md5) {
  stop("Local PKP V7 file failed checksum verification.", call. = FALSE)
}
```

## ISSN preparation

ISSNs may contain spaces, punctuation, or more than one value. This chunk converts each valid ISSN to the same `XXXX-XXXX` format and checks its final digit. It keeps missing, invalid, and valid-but-unmatched ISSNs as separate cases.

```{r}
#| label: issn-preparation

normalize_issn <- function(value) {
  if (length(value) != 1L || is.na(value)) return(NA_character_)
  compact <- gsub("[^0-9X]", "", toupper(trimws(as.character(value))))
  if (!grepl("^[0-9]{7}[0-9X]$", compact)) return(NA_character_)

  digits <- as.integer(strsplit(substr(compact, 1, 7), "")[[1]])
  check_value <- (11 - sum(digits * 8:2) %% 11) %% 11
  expected <- if (check_value == 10) "X" else as.character(check_value)
  if (substr(compact, 8, 8) != expected) return(NA_character_)

  paste0(substr(compact, 1, 4), "-", substr(compact, 5, 8))
}

is_missing_value <- function(value) {
  is.na(value) |
    !nzchar(trimws(value)) |
    toupper(trimws(value)) == "NA"
}

parse_issns <- function(value) {
  if (is_missing_value(value)) return(character())

  tokens <- unlist(strsplit(value, "[[:space:],;|]+"))
  normalized <- vapply(tokens[nzchar(tokens)], normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}

source_issns <- function(source) {
  values <- c(unlist(source$issn, use.names = FALSE), source$issn_l)
  normalized <- vapply(values, normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}
```

## Source retrieval

The sample sends all valid ISSNs to OpenAlex in one request. Full mode splits the 103,017 valid ISSNs into deterministic groups of at most 100, requests the maximum page size, follows every cursor, and atomically checkpoints each completed V7 batch. It also downloads the complete Crossref journals directory in sequential 1,000-record cursor pages, sleeping between requests and checkpointing each page. Failed responses remain explicit and are retried on the next run.

Sample mode asks Crossref about each ISSN separately and waits between requests. A Crossref `404` means that no journal record was found.

```{r}
#| label: source-retrieval

fetch_openalex_sources <- function(
  issns,
  api_key = Sys.getenv("OPENALEX_API_KEY")
) {
  if (!nzchar(api_key)) {
    stop("Set OPENALEX_API_KEY before retrieving OpenAlex data.", call. = FALSE)
  }
  stopifnot(length(issns) <= 100L)

  request("https://api.openalex.org/sources") |>
    req_url_query(
      filter = paste0("issn:", paste(issns, collapse = "|")),
      per_page = 100,
      api_key = api_key
    ) |>
    req_timeout(seconds = 30) |>
    req_perform() |>
    resp_body_json(simplifyVector = FALSE) |>
    pluck("results")
}

perform_openalex_page <- function(
  issns,
  cursor,
  api_key = Sys.getenv("OPENALEX_API_KEY"),
  max_attempts = 4L
) {
  if (!nzchar(api_key)) {
    stop("Set OPENALEX_API_KEY before retrieving OpenAlex data.", call. = FALSE)
  }
  stopifnot(length(issns) <= 100L)

  for (attempt in seq_len(max_attempts)) {
    response <- tryCatch(
      request("https://api.openalex.org/sources") |>
        req_url_query(
          filter = paste0("issn:", paste(issns, collapse = "|")),
          per_page = 200,
          cursor = cursor,
          api_key = api_key
        ) |>
        req_timeout(seconds = 30) |>
        req_error(is_error = \(response) FALSE) |>
        req_perform(),
      error = identity
    )

    if (!inherits(response, "condition")) {
      status <- resp_status(response)
      if (status < 400L) {
        return(list(
          ok = TRUE,
          attempts = attempt,
          body = resp_body_json(response, simplifyVector = FALSE)
        ))
      }
      failure <- paste0("http_", status)
      retryable <- status %in% c(403L, 429L, 500L, 502L, 503L, 504L)
    } else {
      failure <- "transport_error"
      retryable <- TRUE
    }

    if (!retryable || attempt == max_attempts) {
      return(list(
        ok = FALSE,
        attempts = attempt,
        failure = failure
      ))
    }
    Sys.sleep(min(2^(attempt - 1L), 8))
  }
}

fetch_openalex_batch <- function(issns) {
  cursor <- "*"
  seen_cursors <- character()
  pages <- list()
  page_log <- list()
  attempts <- 0L

  repeat {
    if (cursor %in% seen_cursors) {
      return(list(
        status = "api_error",
        issns = issns,
        sources = unlist(pages, recursive = FALSE),
        pages = page_log,
        attempts = attempts,
        failure = "repeated_cursor",
        pagination_complete = FALSE
      ))
    }
    seen_cursors <- c(seen_cursors, cursor)

    page <- perform_openalex_page(issns, cursor)
    attempts <- attempts + page$attempts
    if (!page$ok) {
      return(list(
        status = "api_error",
        issns = issns,
        sources = unlist(pages, recursive = FALSE),
        pages = page_log,
        attempts = attempts,
        failure = page$failure,
        pagination_complete = FALSE
      ))
    }

    results <- if (is.null(page$body$results)) list() else page$body$results
    next_cursor <- page$body$meta$next_cursor
    pages[[length(pages) + 1L]] <- results
    page_log[[length(page_log) + 1L]] <- list(
      result_count = length(results),
      attempts = page$attempts,
      retrieved_at = format(Sys.time(), tz = "UTC", usetz = TRUE)
    )
    if (is.null(next_cursor) || !nzchar(next_cursor)) break
    cursor <- next_cursor
  }

  list(
    status = "complete",
    issns = issns,
    sources = unlist(pages, recursive = FALSE),
    pages = page_log,
    attempts = attempts,
    failure = NULL,
    pagination_complete = TRUE
  )
}

save_rds_atomic <- function(value, path) {
  temporary <- paste0(path, ".tmp")
  saveRDS(value, temporary)
  if (!file.rename(temporary, path)) {
    unlink(temporary)
    stop("Could not save checkpoint atomically.", call. = FALSE)
  }
  invisible(path)
}

fetch_crossref_journal <- function(issn) {
  response <- request(paste0("https://api.crossref.org/journals/", issn)) |>
    req_timeout(seconds = 30) |>
    req_error(
      is_error = \(response) {
        resp_status(response) != 404L && resp_is_error(response)
      }
    ) |>
    req_perform()
  if (resp_status(response) == 404L) return(NULL)
  resp_body_json(response, simplifyVector = FALSE)$message
}

perform_crossref_page <- function(
  cursor,
  mailto,
  max_attempts = 4L
) {
  stopifnot(nzchar(cursor), nzchar(mailto))

  for (attempt in seq_len(max_attempts)) {
    response <- tryCatch(
      request("https://api.crossref.org/journals") |>
        req_url_query(
          rows = 1000,
          cursor = cursor,
          mailto = mailto
        ) |>
        req_user_agent(
          "InvisibleResearch journal enrichment; https://github.com/YannJY02/InvisibleResearch"
        ) |>
        req_timeout(seconds = 60) |>
        req_error(is_error = \(response) FALSE) |>
        req_perform(),
      error = identity
    )

    if (!inherits(response, "condition")) {
      status <- resp_status(response)
      if (status < 400L) {
        body <- tryCatch(
          resp_body_json(response, simplifyVector = FALSE)$message,
          error = identity
        )
        if (
          !inherits(body, "condition") &&
            is.list(body) &&
            is.list(body$items) &&
            length(body[["total-results"]]) == 1L
        ) {
          return(list(ok = TRUE, attempts = attempt, body = body))
        }
        failure <- "parse_error"
        retryable <- TRUE
      } else {
        failure <- paste0("http_", status)
        retryable <- status %in% c(
          403L, 408L, 429L, 500L, 502L, 503L, 504L
        )
      }
    } else {
      failure <- "transport_error"
      retryable <- TRUE
    }

    if (!retryable || attempt == max_attempts) {
      return(list(
        ok = FALSE,
        attempts = attempt,
        failure = failure
      ))
    }
    Sys.sleep(min(2^(attempt - 1L), 8))
  }
}

fetch_crossref_directory <- function() {
  cursor <- "*"
  page_index <- 1L
  page_summaries <- list()
  page_files <- character()
  total_results <- NULL

  repeat {
    page_file <- sprintf("page-%04d.rds", page_index)
    checkpoint_path <- file.path(crossref_page_dir, page_file)
    checkpoint <- if (file.exists(checkpoint_path)) {
      tryCatch(readRDS(checkpoint_path), error = \(error) NULL)
    } else {
      NULL
    }
    reusable <- !is.null(checkpoint) &&
      identical(
        if (is.null(checkpoint$contract_version)) {
          1L
        } else {
          checkpoint$contract_version
        },
        cache_contract_version
      ) &&
      identical(checkpoint$page_index, page_index) &&
      identical(checkpoint$cursor, cursor) &&
      identical(checkpoint$rows_per_page, 1000L) &&
      identical(checkpoint$status, "complete") &&
      identical(checkpoint$item_count, length(checkpoint$items))

    if (reusable) {
      checkpoint$reused <- TRUE
    } else {
      if (!grepl("^[^@[:space:]]+@[^@[:space:]]+$", crossref_mailto)) {
        stop(
          "Set CROSSREF_MAILTO before retrieving Crossref data.",
          call. = FALSE
        )
      }
      page <- perform_crossref_page(cursor, crossref_mailto)
      checkpoint <- list(
        status = if (page$ok) "complete" else "api_error",
        contract_version = cache_contract_version,
        page_index = page_index,
        cursor = cursor,
        rows_per_page = 1000L,
        total_results = if (page$ok) {
          as.integer(page$body[["total-results"]])
        } else {
          NA_integer_
        },
        next_cursor = if (page$ok) page$body[["next-cursor"]] else NULL,
        items = if (page$ok) page$body$items else list(),
        item_count = if (page$ok) length(page$body$items) else 0L,
        attempts = page$attempts,
        failure = if (page$ok) NULL else page$failure,
        pagination_complete = page$ok && length(page$body$items) < 1000L,
        retrieved_at = format(Sys.time(), tz = "UTC", usetz = TRUE),
        reused = FALSE
      )
      save_rds_atomic(checkpoint, checkpoint_path)
    }

    if (checkpoint$status != "complete") {
      stop(
        paste0(
          "Crossref page ", page_index,
          " failed explicitly: ", checkpoint$failure
        ),
        call. = FALSE
      )
    }
    if (is.null(total_results)) {
      total_results <- checkpoint$total_results
    } else if (!identical(total_results, checkpoint$total_results)) {
      stop(
        "Crossref total-results changed during cursor pagination.",
        call. = FALSE
      )
    }

    page_files <- c(page_files, page_file)
    page_summaries[[length(page_summaries) + 1L]] <- checkpoint[
      setdiff(names(checkpoint), c("cursor", "next_cursor", "items"))
    ]
    retrieved_rows <- sum(map_int(page_summaries, "item_count"))
    if (retrieved_rows > total_results) {
      stop(
        "Crossref pagination exceeded total-results.",
        call. = FALSE
      )
    }
    if (isTRUE(checkpoint$pagination_complete)) break
    if (
      is.null(checkpoint$next_cursor) ||
        !nzchar(checkpoint$next_cursor)
    ) {
      stop(
        "Crossref returned a full page without a next cursor.",
        call. = FALSE
      )
    }

    cursor <- checkpoint$next_cursor
    page_index <- page_index + 1L
    Sys.sleep(0.2)
  }

  directory_rows <- sum(map_int(page_summaries, "item_count"))
  if (directory_rows != total_results) {
    stop(
      paste0(
        "Crossref directory row count mismatch: ",
        directory_rows, " != ", total_results
      ),
      call. = FALSE
    )
  }
  list(
    status = "complete",
    rows_per_page = 1000L,
    directory_rows = directory_rows,
    pages = page_summaries,
    page_files = page_files
  )
}

```

## Load the PKP V7 rows

The pinned PKP V7 file is Dataverse file `14084919`, released on 2026-07-18. This chunk checks the complete 98,273-row baseline, schema, row identity, and identifier counts before either retaining the full cohort or selecting ten fixed source rows in a deliberate display order.

```{r}
#| label: load-sample

raw_input <- read_csv(
  input_path,
  col_types = cols(.default = col_character()),
  na = character(),
  show_col_types = FALSE
)

expected_schema <- c(
  "context_name",
  "issn",
  "oai_url",
  "application",
  "country",
  "region",
  "total_record_count",
  "record_count_2025",
  "earliest_datestamp",
  "repository_name",
  "set_spec",
  "stats_id",
  "best_doaj_url",
  "last_completed_update",
  "first_beacon",
  "last_beacon",
  "last_oai_response",
  "unresponsive_endpoint",
  "unresponsive_context"
)
row_identity <- function(data) {
  paste(data$oai_url, data$repository_name, data$set_spec, sep = "\r")
}
raw_row_identity <- row_identity(raw_input)

stopifnot(
  nrow(raw_input) == 98273L,
  identical(names(raw_input), expected_schema),
  all(tolower(raw_input$application) == "ojs"),
  !anyDuplicated(raw_row_identity)
)

all_input_issns <- map(raw_input$issn, parse_issns)
missing_identifier <- is_missing_value(raw_input$issn)
stopifnot(
  sum(missing_identifier) == 26189L,
  sum(!is_missing_value(raw_input$best_doaj_url)) == 7149L,
  sum(lengths(all_input_issns) > 0L) == 72084L,
  length(unique(unlist(all_input_issns))) == 103017L
)

sample_spec <- tribble(
  ~context_name, ~oai_url, ~repository_name, ~set_spec,
  "Kuwait Journal of Science",
  "https://journalskuwait.org/kjs/index.php/index/oai",
  "Kuwait Journal of Science",
  "KJS",
  "Cakrawala",
  "https://cakrawalajournal.org/index.php/index/oai",
  "CAKRAWALA",
  "cakrawala",
  "Sintesa: Jurnal Ilmu Pendidikan",
  "https://sintesa.stkip-arrahmaniyah.ac.id/index.php/index/oai",
  "Sintesa: Jurnal Ilmu Pendidikan STKIP Arrahmaniyah Depok",
  "sintesa",
  "CARAKA: Jurnal Teologi Biblika dan Praktika",
  "https://ojs.sttibc.ac.id/index.php/index/oai",
  "CARAKA:Jurnal Teologi",
  "ibc",
  "Jurnal Tata Kelola dan Akuntabilitas Keuangan Negara",
  "https://jurnal.bpk.go.id/index.php/index/oai",
  "Open Journal Systems",
  "TAKEN",
  "JOEEL (Journal of English Education and Literature)",
  "https://journal.stkippamanetalino.ac.id/index.php/index/oai",
  "Rumah Jurnal STKIP Pamane Talino Ngabang",
  "bahasa-inggris",
  "Malaysian Journal of Paediatrics and Child Health",
  "https://mpaeds.my/journals/index.php/index/oai",
  "NA",
  "MJPCH",
  "Jami Scientific Research Quarterly Journal",
  "https://journals.jami.edu.af/index.php/index/oai",
  "Jami University Press",
  "jsrqj",
  "Journal of Polymer & Composites",
  "https://engineeringjournals.stmjournals.in/index.php/index/oai",
  "Engineering Journals",
  "JoPC",
  "JURNAL TEKNIK PERTAMBANGAN",
  "https://e-journal.upr.ac.id/index.php/index/oai",
  "Journal Online Universitas Palangka Raya",
  "JTP"
)
sample_identity <- row_identity(sample_spec)
sample_positions <- match(sample_identity, raw_row_identity)
stopifnot(!anyNA(sample_positions))

sample_rows <- raw_input[sample_positions, ] |>
  select(-any_of("admin_email"))
stopifnot(
  nrow(sample_rows) == 10L,
  identical(sample_rows$context_name, sample_spec$context_name)
)

ojs_rows <- if (run_mode == "full") {
  raw_input |> select(-any_of("admin_email"))
} else {
  sample_rows
}
```

## Multi-source enrichment

For each PKP row, the notebook looks for OpenAlex and Crossref records with the same checksum-valid normalized ISSN. If the same source identity appears more than once, it is counted once. Both modes expand every returned top-level field into the one-row-per-PKP master and preserve nested values and complete candidate records as JSON cells.

### Sample matching helpers

These helpers label each row as having no usable ISSN, no match, one match, or multiple matches. They also identify Crossref records by their ISSNs and convert every candidate field into a dataframe-safe value without filtering the response schema.

```{r}
#| label: sample-matching-helpers

classify_status <- function(input_issns, candidate_ids) {
  case_when(
    length(input_issns) == 0L ~ "not_attempted",
    length(candidate_ids) == 0L ~ "unmatched",
    length(candidate_ids) == 1L ~ "consistent",
    TRUE ~ "inconsistent"
  )
}

crossref_key <- function(journal) {
  issns <- vapply(journal$ISSN, normalize_issn, character(1))
  paste(sort(unique(issns[!is.na(issns)])), collapse = "|")
}

crossref_candidates_from_pages <- function(page_files) {
  candidates <- list()

  for (page_file in page_files) {
    checkpoint <- readRDS(file.path(crossref_page_dir, page_file))
    if (checkpoint$status != "complete") {
      stop("Crossref enrichment encountered an incomplete page.", call. = FALSE)
    }
    for (journal in checkpoint$items) {
      candidates[[length(candidates) + 1L]] <- journal
    }
  }

  candidates
}

deduplicate_candidates <- function(candidates, identity) {
  candidate_ids <- vapply(candidates, identity, character(1))
  keep <- !duplicated(candidate_ids)
  list(
    candidates = candidates[keep],
    candidate_ids = candidate_ids[keep]
  )
}

index_candidates_by_issn <- function(candidates, candidate_issns) {
  index <- new.env(hash = TRUE, parent = emptyenv())
  for (candidate_position in seq_along(candidates)) {
    for (issn in candidate_issns(candidates[[candidate_position]])) {
      positions <- if (exists(issn, index, inherits = FALSE)) {
        get(issn, index, inherits = FALSE)
      } else {
        integer()
      }
      assign(issn, c(positions, candidate_position), index)
    }
  }
  index
}

candidates_for_issns <- function(candidates, index, issns) {
  positions <- unique(unlist(map(issns, \(issn) {
    if (exists(issn, index, inherits = FALSE)) {
      get(issn, index, inherits = FALSE)
    } else {
      integer()
    }
  })))
  candidates[positions]
}

match_exact_issns <- function(
  input_issns,
  candidates,
  candidate_issns,
  candidate_identity,
  deduplicate = TRUE
) {
  matched <- keep(candidates, \(candidate) {
    any(input_issns %in% candidate_issns(candidate))
  })
  resolved <- if (deduplicate) {
    deduplicate_candidates(matched, candidate_identity)
  } else {
    list(
      candidates = matched,
      candidate_ids = map_chr(matched, candidate_identity)
    )
  }
  matched_issns <- intersect(
    input_issns,
    unique(unlist(map(resolved$candidates, candidate_issns)))
  )

  c(
    list(
      status = classify_status(
        input_issns,
        resolved$candidate_ids
      ),
      matched_issns = matched_issns
    ),
    resolved
  )
}

serialize_json <- function(value) {
  as.character(jsonlite::toJSON(
    value,
    auto_unbox = TRUE,
    null = "null",
    na = "null",
    digits = NA
  ))
}

dataframe_value <- function(value) {
  if (is.null(value) || !length(value)) return(NA)
  if (is.atomic(value) && length(value) == 1L) return(value)
  serialize_json(value)
}

expand_candidate_fields <- function(matches, prefix, fields) {
  candidates <- map(
    matches,
    \(match) if (length(match$candidates)) match$candidates[[1]] else list()
  )
  cells <- set_names(
    map(fields, \(field) {
      map_chr(candidates, \(candidate) {
        as.character(dataframe_value(candidate[[field]]))
      })
    }),
    paste(prefix, fields, sep = "_")
  )
  cells[[paste0(prefix, "_candidates__json")]] <- map_chr(
    matches,
    \(match) {
      if (length(match$candidates) > 1L) {
        serialize_json(match$candidates)
      } else {
        NA_character_
      }
    }
  )
  as_tibble(cells, .name_repair = "check_unique")
}

```

### Enrichment orchestration

Both modes use the same exact-ISSN rule. Full mode indexes candidates by normalized ISSN so matching 98,273 rows does not repeatedly scan either complete response set.

```{r}
#| label: enrichment

input_issns <- map(ojs_rows$issn, parse_issns)
tokens <- sort(unique(unlist(input_issns)))

if (run_mode == "full") {
  token_batches <- split(tokens, ceiling(seq_along(tokens) / 100L))
  process_batch <- function(batch_index) {
    batch <- token_batches[[batch_index]]
    checkpoint_path <- file.path(
      openalex_batch_dir,
      sprintf("batch-%04d.rds", batch_index)
    )
    checkpoint <- if (file.exists(checkpoint_path)) {
      tryCatch(readRDS(checkpoint_path), error = \(error) NULL)
    } else {
      NULL
    }
    reusable <- !is.null(checkpoint) &&
      identical(
        if (is.null(checkpoint$contract_version)) {
          1L
        } else {
          checkpoint$contract_version
        },
        cache_contract_version
      ) &&
      identical(checkpoint$pkp_version, pkp_version) &&
      identical(checkpoint$input_md5, pkp_md5) &&
      identical(checkpoint$issns, batch) &&
      identical(checkpoint$status, "complete") &&
      isTRUE(checkpoint$pagination_complete)

    if (reusable) {
      checkpoint$reused <- TRUE
      return(checkpoint)
    }

    checkpoint <- fetch_openalex_batch(batch)
    checkpoint$contract_version <- cache_contract_version
    checkpoint$pkp_version <- pkp_version
    checkpoint$input_md5 <- pkp_md5
    checkpoint$batch_index <- as.integer(batch_index)
    checkpoint$reused <- FALSE
    save_rds_atomic(checkpoint, checkpoint_path)
    checkpoint
  }
  openalex_cluster <- parallel::makeCluster(4L)
  invisible(parallel::clusterEvalQ(openalex_cluster, library(httr2)))
  parallel::clusterExport(
    openalex_cluster,
    c(
      "token_batches",
      "openalex_batch_dir",
      "pkp_version",
      "pkp_md5",
      "cache_contract_version",
      "perform_openalex_page",
      "fetch_openalex_batch",
      "save_rds_atomic",
      "process_batch"
    ),
    envir = environment()
  )
  batch_results <- tryCatch(
    parallel::parLapplyLB(
      openalex_cluster,
      seq_along(token_batches),
      process_batch
    ),
    finally = parallel::stopCluster(openalex_cluster)
  )
  openalex_cache_hit <- all(map_lgl(batch_results, "reused"))
  openalex_sources <- unlist(
    map(batch_results, "sources"),
    recursive = FALSE
  )
  failed_issns <- unique(unlist(map(
    keep(batch_results, \(batch) batch$status != "complete"),
    "issns"
  )))
} else {
  token_batches <- list(tokens)
  batch_results <- list()
  failed_issns <- character()
  openalex_cache <- if (file.exists(openalex_cache_path)) {
    tryCatch(readRDS(openalex_cache_path), error = \(error) NULL)
  } else {
    NULL
  }
  openalex_cache_hit <- is.list(openalex_cache) &&
    identical(openalex_cache$contract_version, cache_contract_version) &&
    identical(openalex_cache$pkp_version, pkp_version) &&
    identical(openalex_cache$input_md5, pkp_md5) &&
    identical(openalex_cache$issns, tokens)
  openalex_sources <- if (openalex_cache_hit) {
    openalex_cache$sources
  } else {
    sources <- fetch_openalex_sources(tokens)
    save_rds_atomic(
      list(
        contract_version = cache_contract_version,
        pkp_version = pkp_version,
        input_md5 = pkp_md5,
        issns = tokens,
        retrieved_at = format(Sys.time(), tz = "UTC", usetz = TRUE),
        sources = sources
      ),
      openalex_cache_path
    )
    sources
  }
}
openalex_sources <- deduplicate_candidates(
  openalex_sources,
  \(source) source$id
)$candidates
openalex_index <- index_candidates_by_issn(
  openalex_sources,
  source_issns
)

openalex_matches <- map(input_issns, \(issns) {
  match <- match_exact_issns(
    issns,
    candidates_for_issns(openalex_sources, openalex_index, issns),
    source_issns,
    \(source) source$id
  )
  if (any(issns %in% failed_issns)) match$status <- "api_error"
  match
})
openalex_status_counts <- table(map_chr(openalex_matches, "status"))

base_result <- ojs_rows |>
  mutate(
    identifier_status = case_when(
      is_missing_value(issn) ~ "missing",
      lengths(input_issns) == 0L ~ "invalid",
      TRUE ~ "valid"
    ),
    openalex_input_issns = map_chr(input_issns, paste, collapse = "|"),
    openalex_matched_issns = map_chr(
      openalex_matches,
      \(match) paste(match$matched_issns, collapse = "|")
    ),
    openalex_match_status = map_chr(openalex_matches, "status"),
    openalex_candidate_count = map_int(
      openalex_matches,
      \(match) length(match$candidate_ids)
    ),
    openalex_candidate_ids = map_chr(
      openalex_matches,
      \(match) paste(match$candidate_ids, collapse = "|")
    )
  )
```

### Crossref enrichment and wide master

Crossref is matched by the same normalized exact ISSNs. Both modes write every
returned top-level OpenAlex and Crossref field to the same one-row-per-PKP
master. Nested and repeated values remain lossless JSON cells.

```{r}
#| label: crossref-enrichment

if (run_mode == "sample") {
  crossref_cache <- if (file.exists(crossref_cache_path)) {
    tryCatch(readRDS(crossref_cache_path), error = \(error) NULL)
  } else {
    NULL
  }
  crossref_cache_hit <- is.list(crossref_cache) &&
    identical(crossref_cache$contract_version, cache_contract_version) &&
    identical(crossref_cache$pkp_version, pkp_version) &&
    identical(crossref_cache$input_md5, pkp_md5) &&
    identical(crossref_cache$issns, tokens)
  crossref_journals <- if (crossref_cache_hit) {
    crossref_cache$journals
  } else {
    journals <- set_names(
      map(tokens, \(token) {
        journal <- fetch_crossref_journal(token)
        Sys.sleep(0.25)
        journal
      }),
      tokens
    )
    save_rds_atomic(
      list(
        contract_version = cache_contract_version,
        pkp_version = pkp_version,
        input_md5 = pkp_md5,
        issns = tokens,
        retrieved_at = format(Sys.time(), tz = "UTC", usetz = TRUE),
        journals = journals
      ),
      crossref_cache_path
    )
    journals
  }
  crossref_candidates <- unname(compact(crossref_journals))
} else {
  crossref_directory <- fetch_crossref_directory()
  crossref_candidates <- crossref_candidates_from_pages(
    crossref_directory$page_files
  )
  crossref_cache_hit <- all(map_lgl(crossref_directory$pages, "reused"))
}

crossref_issns <- function(journal) {
  issns <- vapply(journal$ISSN, normalize_issn, character(1))
  unique(issns[!is.na(issns)])
}
crossref_index <- index_candidates_by_issn(
  crossref_candidates,
  crossref_issns
)
crossref_matches <- map(input_issns, \(issns) {
  match_exact_issns(
    issns,
    candidates_for_issns(
      crossref_candidates,
      crossref_index,
      issns
    ),
    crossref_issns,
    crossref_key,
    deduplicate = FALSE
  )
})
crossref_status_counts <- table(map_chr(crossref_matches, "status"))

# ponytail: in-memory wide assembly fits V7; stream chunks if cohort growth
# makes memory material.
openalex_fields <- sort(unique(unlist(map(openalex_sources, names))))
crossref_fields <- sort(unique(unlist(map(crossref_candidates, names))))
openalex_data <- expand_candidate_fields(
  openalex_matches,
  prefix = "openalex",
  fields = openalex_fields
)
crossref_data <- expand_candidate_fields(
  crossref_matches,
  prefix = "crossref",
  fields = crossref_fields
)
result <- base_result |>
  mutate(
    crossref_input_issns = map_chr(input_issns, paste, collapse = "|"),
    crossref_matched_issns = map_chr(
      crossref_matches,
      \(match) paste(match$matched_issns, collapse = "|")
    ),
    crossref_match_status = map_chr(crossref_matches, "status"),
    crossref_candidate_count = map_int(
      crossref_matches,
      \(match) length(match$candidate_ids)
    ),
    crossref_candidate_issn_sets = map_chr(
      crossref_matches,
      \(match) paste(match$candidate_ids, collapse = ";")
    )
  ) |>
  bind_cols(openalex_data, crossref_data)

output_data <- result
output_rows <- nrow(result)
output_columns <- names(result)
review_result <- if (run_mode == "full") {
  result[sample_positions, ]
} else {
  result
}
```

### Validated master and provenance

The wide master is written to a temporary file, read back, and promoted only
after its schema, row count, and PKP identities pass.

```{r}
#| label: validated-master

temporary_output_path <- paste0(
  output_path,
  ".tmp",
  if (endsWith(output_path, ".gz")) ".gz" else ""
)
unlink(temporary_output_path)
write_csv(output_data, temporary_output_path, na = "")

observed_output <- read_csv(
  temporary_output_path,
  col_types = cols(.default = col_character()),
  na = character(),
  show_col_types = FALSE
)
observed_identity <- row_identity(observed_output)
stopifnot(
  nrow(observed_output) == nrow(output_data),
  identical(names(observed_output), names(output_data)),
  identical(observed_identity, row_identity(ojs_rows)),
  !anyDuplicated(observed_identity)
)
if (!file.rename(temporary_output_path, output_path)) {
  stop("Could not promote the validated wide master.", call. = FALSE)
}
artifact_validation <- list(
  rows = nrow(observed_output),
  columns = ncol(observed_output)
)

openalex_retrieval_times <- if (run_mode == "full") {
  unlist(map(batch_results, \(batch) map_chr(batch$pages, "retrieved_at")))
} else {
  character()
}
crossref_retrieval_times <- if (run_mode == "full") {
  map_chr(crossref_directory$pages, "retrieved_at")
} else {
  character()
}

run_metadata <- list(
  producer = "analysis/ojs_journal_enrichment.qmd",
  run_mode = run_mode,
  created_at = format(Sys.time(), tz = "UTC", usetz = TRUE),
  input = list(
    version = pkp_version,
    dataverse_file_id = 14084919L,
    md5 = pkp_md5,
    rows = nrow(raw_input),
    valid_identifier_rows = sum(lengths(all_input_issns) > 0L),
    missing_identifier_rows = sum(missing_identifier),
    distinct_valid_issns = length(unique(unlist(all_input_issns)))
  ),
  output = list(
    file = basename(output_path),
    md5 = unname(tools::md5sum(output_path)),
    format = if (run_mode == "full") "csv.gz" else "csv",
    rows = output_rows,
    columns = length(output_columns),
    retained_top_level_fields = list(
      openalex = openalex_fields,
      crossref = crossref_fields
    ),
    status_counts = list(
      openalex = as.list(openalex_status_counts),
      crossref = as.list(crossref_status_counts)
    )
  ),
  retrieval = if (run_mode == "full") {
    list(
      openalex = list(
        first = min(openalex_retrieval_times),
        last = max(openalex_retrieval_times)
      ),
      crossref = list(
        first = min(crossref_retrieval_times),
        last = max(crossref_retrieval_times)
      )
    )
  } else {
    NULL
  },
  openalex = if (run_mode == "full") {
    list(
      batch_size = 100L,
      per_page = 200L,
      batches = length(batch_results),
      completed_batches = sum(map_chr(batch_results, "status") == "complete"),
      reused_batches = sum(map_lgl(batch_results, "reused")),
      pages = sum(map_int(batch_results, \(batch) length(batch$pages))),
      attempts = sum(map_int(batch_results, "attempts")),
      failed_batches = sum(map_chr(batch_results, "status") != "complete")
    )
  } else {
    list(
      cache_file = basename(openalex_cache_path),
      reused = openalex_cache_hit
    )
  },
  crossref = if (run_mode == "full") {
    list(
      status = crossref_directory$status,
      rows_per_page = crossref_directory$rows_per_page,
      directory_rows = crossref_directory$directory_rows,
      pages = length(crossref_directory$pages),
      completed_pages = sum(
        map_chr(crossref_directory$pages, "status") == "complete"
      ),
      reused_pages = sum(map_lgl(crossref_directory$pages, "reused")),
      attempts = sum(map_int(crossref_directory$pages, "attempts")),
      failed_pages = sum(
        map_chr(crossref_directory$pages, "status") != "complete"
      )
    )
  } else {
    list(
      cache_file = basename(crossref_cache_path),
      reused = crossref_cache_hit
    )
  },
  runtime = list(
    r = R.version.string,
    quarto = system2("quarto", "--version", stdout = TRUE)[1],
    curl = system2(curl, "--version", stdout = TRUE)[1],
    locale = Sys.getlocale(),
    packages = as.list(installed_package_versions)
  )
)
metadata_temporary_path <- paste0(metadata_path, ".tmp")
jsonlite::write_json(
  run_metadata,
  metadata_temporary_path,
  auto_unbox = TRUE,
  pretty = TRUE
)
if (!file.rename(metadata_temporary_path, metadata_path)) {
  stop("Could not promote validated run metadata.", call. = FALSE)
}
```


## Checks

These checks confirm that the wide master preserves every input identity, every
returned top-level source field, explicit match states, and no private values.

```{r}
#| label: contract-check

metadata_text <- paste(readLines(metadata_path, warn = FALSE), collapse = "\n")
artifact_root <- normalizePath(artifact_dir)
private_values <- c(Sys.getenv("OPENALEX_API_KEY"), crossref_mailto)
private_values <- private_values[nzchar(private_values)]
metadata_contains_private_value <- any(vapply(
  private_values,
  \(value) grepl(value, metadata_text, fixed = TRUE),
  logical(1)
))
has_source_field <- function(prefix, field) {
  any(
    startsWith(output_columns, paste0(prefix, "_")) &
      endsWith(output_columns, paste0("_", field))
  )
}

stopifnot(
  identical(pkp_version, 7L),
  is.na(normalize_issn(NULL)),
  identical(unname(tools::md5sum(input_path)), pkp_md5),
  nrow(raw_input) == 98273L,
  identical(names(raw_input), expected_schema),
  !anyDuplicated(raw_row_identity),
  sum(missing_identifier) == 26189L,
  sum(lengths(all_input_issns) > 0L) == 72084L,
  length(unique(unlist(all_input_issns))) == 103017L,
  startsWith(normalizePath(input_path), artifact_root),
  startsWith(normalizePath(output_path), artifact_root),
  startsWith(normalizePath(metadata_path), artifact_root),
  file.exists(output_path),
  file.exists(metadata_path),
  !grepl("OPENALEX_API_KEY|CROSSREF_MAILTO", metadata_text),
  !metadata_contains_private_value,
  !"admin_email" %in% output_columns,
  all(nzchar(unlist(run_metadata$runtime))),
  output_rows == nrow(ojs_rows),
  identical(output_columns, names(observed_output)),
  all(c(
    "openalex_match_status",
    "openalex_candidate_count",
    "openalex_candidate_ids",
    "openalex_candidates__json",
    "crossref_match_status",
    "crossref_candidate_count",
    "crossref_candidate_issn_sets",
    "crossref_candidates__json"
  ) %in% output_columns),
  all(map_lgl(openalex_fields, \(field) {
    has_source_field("openalex", field)
  })),
  all(map_lgl(crossref_fields, \(field) {
    has_source_field("crossref", field)
  })),
  sum(unlist(run_metadata$output$status_counts$openalex)) == output_rows,
  sum(unlist(run_metadata$output$status_counts$crossref)) == output_rows,
  artifact_validation$rows == output_rows,
  artifact_validation$columns == length(output_columns)
)

if (run_mode == "sample") {
  expected_openalex_status <- c(
    "consistent",
    "consistent",
    "unmatched",
    "consistent",
    "consistent",
    "consistent",
    "consistent",
    "consistent",
    "not_attempted",
    "inconsistent"
  )
  stopifnot(
    identical(row_identity(ojs_rows), sample_identity),
    sum(result$identifier_status == "missing") == 1L,
    nrow(result) == 10L,
    identical(result$openalex_match_status, expected_openalex_status),
    result$openalex_matched_issns[
      result$context_name == "Jami Scientific Research Quarterly Journal"
    ] == "2710-4990",
    "not_attempted" %in% result$crossref_match_status
  )
} else {
  stopifnot(
    identical(row_identity(ojs_rows), raw_row_identity),
    output_rows == 98273L,
    sum(base_result$identifier_status == "missing") == 26189L,
    length(tokens) == 103017L,
    length(token_batches) == 1031L,
    max(lengths(token_batches)) <= 100L,
    all(map_lgl(batch_results, \(batch) {
      identical(batch$pkp_version, 7L) &&
        identical(batch$input_md5, pkp_md5) &&
        identical(batch$status, "complete") &&
        isTRUE(batch$pagination_complete) &&
        length(batch$pages) >= 1L
    })),
    run_metadata$openalex$completed_batches == length(token_batches),
    run_metadata$openalex$failed_batches == 0L,
    !any(result$openalex_match_status == "api_error"),
    all(nzchar(unlist(run_metadata$retrieval))),
    identical(crossref_directory$status, "complete"),
    identical(crossref_directory$rows_per_page, 1000L),
    crossref_directory$directory_rows ==
      sum(map_int(crossref_directory$pages, "item_count")),
    all(map_int(
      crossref_directory$pages[-length(crossref_directory$pages)],
      "item_count"
    ) == 1000L),
    last(crossref_directory$pages)$item_count < 1000L,
    all(map_chr(crossref_directory$pages, "status") == "complete"),
    run_metadata$crossref$completed_pages ==
      length(crossref_directory$pages),
    run_metadata$crossref$failed_pages == 0L,
    all(result$crossref_match_status %in% c(
      "consistent",
      "inconsistent",
      "unmatched",
      "not_attempted"
    )),
    endsWith(output_path, ".csv.gz"),
    identical(run_metadata$output$format, "csv.gz")
  )
}
```


## Purposive ten-row review

This searchable, filterable table uses the fixed V7 sample in its deliberate
display order. It includes every master field and omits only columns whose
values are long JSON; those values remain complete in the wide master.

```{r}
#| label: interactive-sample-dataframe

is_long_json_column <- function(column) {
  values <- na.omit(as.character(column))
  any(
    nchar(values) > 200L &
      grepl("^[[:space:]]*[\\[{]", values)
  )
}
long_json_columns <- names(review_result)[
  endsWith(names(review_result), "__json") |
    map_lgl(review_result, is_long_json_column)
]
review_table <- review_result[
  setdiff(names(review_result), long_json_columns)
]

htmltools::div(
  style = "overflow-x: auto;",
  reactable::reactable(
    review_table,
    searchable = TRUE,
    filterable = TRUE,
    pagination = TRUE,
    defaultPageSize = 10,
    showPageSizeOptions = TRUE,
    resizable = TRUE,
    compact = TRUE,
    fullWidth = FALSE
  )
)
```

## Complete merged column list

This table lists every column in the wide master artifact.

```{r}
#| label: complete-column-list

tibble(column = output_columns) |>
  knitr::kable()
```
