---
title: "PKP–OpenAlex disagreement analysis"
format: html
execute:
  echo: true
  warning: true
  message: false
---

This exploratory analysis reads the validated compact PKP V7 master and its
OpenAlex checkpoints. It compares source-specific evidence without treating
either source as authoritative and without interpreting source absence as a
causal measure of journal invisibility. It also separates valid-ISSN rows that
were unmatched from rows that could not be attempted because PKP supplied no
usable ISSN.

## Load the compact master

```{r}
#| label: load-master

required_packages <- c(
  "countrycode",
  "dplyr",
  "ggplot2",
  "jsonlite",
  "purrr",
  "readr",
  "reactable",
  "tibble"
)
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]
if (length(missing_packages)) {
  stop(
    paste("Install the documented R environment. Missing:",
      paste(missing_packages, collapse = ", ")),
    call. = FALSE
  )
}

library(dplyr)
library(ggplot2)
library(purrr)
library(readr)
library(tibble)

enrichment_artifact_dir <- file.path(
  "..",
  "artifacts",
  "ojs_journal_enrichment"
)
master_path <- file.path(
  enrichment_artifact_dir,
  "full-v7",
  "pkp-ojs-multisource-enriched.csv.gz"
)
metadata_path <- file.path(
  enrichment_artifact_dir,
  "full-v7",
  "run-metadata.json"
)
openalex_batch_dir <- file.path(
  enrichment_artifact_dir,
  "full-v7",
  "openalex-batches"
)
analysis_dir <- file.path(
  "..",
  "artifacts",
  "ojs_journal_disagreement_analysis"
)
audit_path <- file.path(
  analysis_dir,
  "pkp-openalex-disagreement-audit.csv.gz"
)
audit_summary_path <- file.path(
  analysis_dir,
  "pkp-openalex-disagreement-summary.csv"
)
dir.create(analysis_dir, recursive = TRUE, showWarnings = FALSE)

if (
  !file.exists(master_path) ||
    !file.exists(metadata_path) ||
    !dir.exists(openalex_batch_dir)
) {
  stop("Run render_full.sh before this analysis.", call. = FALSE)
}
metadata <- jsonlite::read_json(metadata_path, simplifyVector = TRUE)
file_set_md5 <- function(paths) {
  manifest <- tempfile()
  on.exit(unlink(manifest), add = TRUE)
  paths <- sort(paths)
  writeLines(paste(basename(paths), unname(tools::md5sum(paths))), manifest)
  unname(tools::md5sum(manifest))
}
openalex_batch_files <- list.files(
  openalex_batch_dir,
  pattern = "^batch-[0-9]{4}\\.rds$",
  full.names = TRUE
)
stopifnot(
  identical(metadata$run_mode, "full"),
  identical(unname(tools::md5sum(master_path)), metadata$output$md5),
  length(openalex_batch_files) == metadata$openalex_checkpoints$count,
  identical(
    file_set_md5(openalex_batch_files),
    metadata$openalex_checkpoints$md5
  )
)
master <- read_csv(
  master_path,
  col_types = cols(.default = col_character()),
  na = character(),
  show_col_types = FALSE
)

required_columns <- c(
  "oai_url",
  "repository_name",
  "set_spec",
  "context_name",
  "issn",
  "country",
  "total_record_count",
  "best_doaj_url",
  "identifier_status",
  "openalex_input_issns",
  "openalex_matched_issns",
  "openalex_match_status",
  "openalex_candidate_count",
  "openalex_candidate_ids",
  "crossref_match_status",
  "crossref_candidate_count",
  "crossref_candidate_issn_sets"
)
row_identity <- function(data) {
  paste(data$oai_url, data$repository_name, data$set_spec, sep = "\r")
}
stopifnot(
  nrow(master) == 98273L,
  all(required_columns %in% names(master)),
  !anyDuplicated(row_identity(master)),
  !"admin_email" %in% names(master)
)

source_value <- function(value) {
  if (is.null(value)) return(NA_character_)
  if (is.atomic(value) && length(value) == 1L) return(as.character(value))
  as.character(jsonlite::toJSON(value, auto_unbox = TRUE, null = "null"))
}
openalex_sources <- openalex_batch_files |>
  sort() |>
  map(\(path) readRDS(path)$sources) |>
  unlist(recursive = FALSE)
openalex_fields <- c(
  "type", "display_name", "alternate_titles", "issn", "issn_l", "is_ojs",
  "is_in_doaj", "country_code"
)
openalex_rows <- tibble(
  openalex_candidate_ids = map_chr(openalex_sources, "id")
) |>
  bind_cols(as_tibble(set_names(
    map(openalex_fields, \(field) {
      map_chr(openalex_sources, \(source) source_value(source[[field]]))
    }),
    paste("openalex", openalex_fields, sep = "_")
  ))) |>
  distinct(openalex_candidate_ids, .keep_all = TRUE)
unique_ids <- master$openalex_candidate_ids[
  master$openalex_match_status == "unique"
]
stopifnot(all(unique_ids %in% openalex_rows$openalex_candidate_ids))
master <- left_join(master, openalex_rows, by = "openalex_candidate_ids")
rm(openalex_sources, openalex_rows)
```

## Classify identity and metadata differences

Only rows with one exact-ISSN OpenAlex journal identity enter the metadata
comparisons. Missing or ineligible evidence remains `not_observed`.

```{r}
#| label: classify-disagreements

is_missing_value <- function(value) {
  is.na(value) |
    !nzchar(trimws(value)) |
    toupper(trimws(value)) == "NA"
}

normalize_issn <- function(value) {
  if (length(value) != 1L || is.na(value)) return(NA_character_)
  compact <- gsub("[^0-9X]", "", toupper(trimws(as.character(value))))
  if (!grepl("^[0-9]{7}[0-9X]$", compact)) return(NA_character_)

  digits <- as.integer(strsplit(substr(compact, 1, 7), "")[[1]])
  check_value <- (11 - sum(digits * 8:2) %% 11) %% 11
  expected <- if (check_value == 10) "X" else as.character(check_value)
  if (substr(compact, 8, 8) != expected) return(NA_character_)

  paste0(substr(compact, 1, 4), "-", substr(compact, 5, 8))
}

json_values <- function(value) {
  if (is_missing_value(value)) return(character())
  parsed <- tryCatch(
    jsonlite::fromJSON(value, simplifyVector = TRUE),
    error = function(error) {
      stop("Wide master contains invalid source JSON.", call. = FALSE)
    }
  )
  as.character(unlist(parsed, use.names = FALSE))
}

pipe_values <- function(value) {
  if (is_missing_value(value)) return(character())
  strsplit(value, "|", fixed = TRUE)[[1]]
}

normalize_title <- function(value) {
  if (is_missing_value(value)) return(NA_character_)
  gsub("[[:punct:][:space:]]", "", tolower(enc2utf8(value)))
}

comparison_state <- function(left, right) {
  if (is_missing_value(left) || is_missing_value(right)) {
    return("not_observed")
  }
  if (identical(left, right)) "same" else "different"
}
stopifnot(
  comparison_state("", "US") == "not_observed",
  comparison_state("US", "") == "not_observed",
  comparison_state("US", "US") == "same",
  comparison_state("US", "GB") == "different"
)

category_summary <- function(
  values,
  categories,
  dimension,
  denominator_name
) {
  counts <- table(factor(values, levels = categories))
  denominator <- length(values)
  tibble(
    dimension = dimension,
    category = names(counts),
    denominator_name = denominator_name,
    denominator = denominator,
    count = as.integer(counts),
    percent = round(100 * count / denominator, 3)
  )
}

candidate_count <- as.integer(master$openalex_candidate_count)
input_issn_sets <- map_chr(master$openalex_input_issns, \(value) {
  paste(sort(pipe_values(value)), collapse = "|")
})
matched_issn_sets <- map_chr(master$openalex_matched_issns, \(value) {
  paste(sort(pipe_values(value)), collapse = "|")
})
openalex_source_issn_sets <- map2_chr(
  master$openalex_issn,
  master$openalex_issn_l,
  \(issns, issn_l) {
    values <- c(json_values(issns), issn_l)
    normalized <- vapply(values, normalize_issn, character(1))
    paste(sort(unique(normalized[!is.na(normalized)])), collapse = "|")
  }
)

identity_outcome <- case_when(
  master$identifier_status != "valid" ~ "no_valid_issn",
  candidate_count == 0L ~ "valid_issn_unmatched",
  candidate_count > 1L ~ "conflicting_source_ids",
  is_missing_value(master$openalex_type) |
    tolower(master$openalex_type) != "journal" ~ "non_journal_source",
  input_issn_sets != matched_issn_sets ~ "partial_issn_agreement",
  TRUE ~ "consistent"
)
identity_matched <- identity_outcome %in% c(
  "consistent",
  "partial_issn_agreement"
)

pkp_normalized_titles <- map_chr(master$context_name, normalize_title)
openalex_normalized_titles <- map2(
  master$openalex_display_name,
  master$openalex_alternate_titles,
  \(display_name, alternate_titles) {
    titles <- c(display_name, json_values(alternate_titles))
    normalized <- unique(na.omit(map_chr(titles, normalize_title)))
    normalized[nzchar(normalized)]
  }
)
title_comparison <- map2_chr(
  pkp_normalized_titles,
  openalex_normalized_titles,
  \(pkp_title, openalex_titles) {
    if (!length(openalex_titles) || is.na(pkp_title)) return("not_observed")
    if (pkp_title %in% openalex_titles) "same" else "different"
  }
)
title_comparison[!identity_matched] <- "not_observed"

issn_set_comparison <- map2_chr(
  input_issn_sets,
  openalex_source_issn_sets,
  comparison_state
)
issn_set_comparison[!identity_matched] <- "not_observed"

ojs_evidence_comparison <- case_when(
  !identity_matched | is_missing_value(master$openalex_is_ojs) ~ "not_observed",
  master$openalex_is_ojs == "TRUE" ~ "same",
  TRUE ~ "different"
)

pkp_doaj_evidence <- !is_missing_value(master$best_doaj_url)
openalex_doaj_evidence <- master$openalex_is_in_doaj
doaj_evidence_comparison <- case_when(
  !identity_matched | is_missing_value(openalex_doaj_evidence) ~ "not_observed",
  pkp_doaj_evidence & openalex_doaj_evidence == "TRUE" ~ "both_observed",
  pkp_doaj_evidence ~ "pkp_only",
  openalex_doaj_evidence == "TRUE" ~ "openalex_only",
  TRUE ~ "neither_observed"
)

pkp_country_missing <- is_missing_value(master$country)
pkp_country_code <- countrycode::countrycode(
  master$country,
  origin = "country.name",
  destination = "iso2c",
  warn = FALSE
)
country_mapping_failures <- sum(!pkp_country_missing & is.na(pkp_country_code))
openalex_country_code <- toupper(master$openalex_country_code)
country_evidence_comparison <- map2_chr(
  pkp_country_code,
  openalex_country_code,
  comparison_state
)
country_evidence_comparison[!identity_matched] <- "not_observed"

audit <- master |>
  transmute(
    oai_url,
    repository_name,
    set_spec,
    context_name,
    issn,
    identifier_status,
    openalex_input_issns,
    openalex_matched_issns,
    openalex_match_status,
    openalex_candidate_count,
    openalex_candidate_ids,
    crossref_match_status,
    crossref_candidate_count,
    crossref_candidate_issn_sets,
    identity_outcome,
    openalex_source_type = openalex_type,
    openalex_display_name,
    pkp_normalized_title = pkp_normalized_titles,
    openalex_normalized_titles = map_chr(
      openalex_normalized_titles,
      paste,
      collapse = "|"
    ),
    title_comparison,
    pkp_issn_set = input_issn_sets,
    openalex_issn_set = openalex_source_issn_sets,
    issn_set_comparison,
    ojs_evidence_comparison,
    pkp_doaj_evidence,
    openalex_doaj_evidence,
    doaj_evidence_comparison,
    pkp_country = country,
    pkp_country_code,
    openalex_country_code,
    country_evidence_comparison
  )
```

## Summarize and validate

```{r}
#| label: summarize-and-validate

identity_categories <- c(
  "consistent",
  "partial_issn_agreement",
  "conflicting_source_ids",
  "non_journal_source",
  "valid_issn_unmatched",
  "no_valid_issn"
)
comparison_categories <- c("same", "different", "not_observed")
status_categories <- c(
  "unique",
  "ambiguous",
  "unmatched",
  "not_attempted"
)
matched_denominator <- sum(identity_matched)

source_status_counts <- as_tibble(as.data.frame(table(
  openalex = factor(
    audit$openalex_match_status,
    levels = status_categories
  ),
  crossref = factor(
    audit$crossref_match_status,
    levels = status_categories
  )
))) |>
  transmute(
    dimension = "source_status_crosstab",
    category = paste0("openalex=", openalex, "|crossref=", crossref),
    denominator_name = "all_pkp_v7_rows",
    denominator = nrow(audit),
    count = as.integer(Freq),
    percent = round(100 * count / denominator, 3)
  )

country_mapping_summary <- tibble(
  dimension = "country_mapping",
  category = "unmapped_nonmissing",
  denominator_name = "pkp_country_nonmissing_rows",
  denominator = sum(!pkp_country_missing),
  count = country_mapping_failures,
  percent = round(100 * count / denominator, 3)
)

metadata_disagreement_counts <- c(
  normalized_title = sum(title_comparison == "different"),
  issn_set = sum(issn_set_comparison == "different"),
  ojs_evidence = sum(ojs_evidence_comparison == "different"),
  doaj_evidence = sum(doaj_evidence_comparison %in% c(
    "pkp_only",
    "openalex_only"
  )),
  country_evidence = sum(country_evidence_comparison == "different")
)
metadata_disagreement_summary <- tibble(
  dimension = "metadata_disagreement",
  category = names(metadata_disagreement_counts),
  denominator_name = "identity_matched_openalex_journals",
  denominator = matched_denominator,
  count = as.integer(metadata_disagreement_counts),
  percent = round(100 * count / denominator, 3)
)

audit_summary <- bind_rows(
  category_summary(
    identity_outcome,
    identity_categories,
    "identity_outcome",
    "all_pkp_v7_rows"
  ),
  category_summary(
    title_comparison[identity_matched],
    comparison_categories,
    "normalized_title",
    "identity_matched_openalex_journals"
  ),
  category_summary(
    issn_set_comparison[identity_matched],
    comparison_categories,
    "issn_set",
    "identity_matched_openalex_journals"
  ),
  category_summary(
    ojs_evidence_comparison[identity_matched],
    comparison_categories,
    "ojs_evidence",
    "identity_matched_openalex_journals"
  ),
  category_summary(
    doaj_evidence_comparison[identity_matched],
    c(
      "both_observed",
      "pkp_only",
      "openalex_only",
      "neither_observed",
      "not_observed"
    ),
    "doaj_evidence",
    "identity_matched_openalex_journals"
  ),
  category_summary(
    country_evidence_comparison[identity_matched],
    comparison_categories,
    "country_evidence",
    "identity_matched_openalex_journals"
  ),
  metadata_disagreement_summary,
  country_mapping_summary,
  source_status_counts
)

metadata_dimensions <- c(
  "normalized_title",
  "issn_set",
  "ojs_evidence",
  "doaj_evidence",
  "country_evidence"
)
metadata_reconciliation <- audit_summary |>
  filter(dimension %in% metadata_dimensions) |>
  group_by(dimension) |>
  summarise(
    denominator = unique(denominator),
    count = sum(count),
    .groups = "drop"
  )
stopifnot(
  nrow(audit) == nrow(master),
  identical(audit$identity_outcome, identity_outcome),
  sum(audit_summary$count[
    audit_summary$dimension == "identity_outcome"
  ]) == nrow(master),
  nrow(metadata_reconciliation) == length(metadata_dimensions),
  all(metadata_reconciliation$denominator == matched_denominator),
  all(metadata_reconciliation$count == matched_denominator),
  sum(audit_summary$count[
    audit_summary$dimension == "source_status_crosstab"
  ]) == nrow(master),
  all(is.finite(audit_summary$percent))
)

write_csv_atomic <- function(data, path) {
  temporary_path <- paste0(
    path,
    ".tmp",
    if (endsWith(path, ".gz")) ".gz" else ""
  )
  unlink(temporary_path)
  write_csv(data, temporary_path, na = "")
  observed <- read_csv(
    temporary_path,
    col_types = cols(.default = col_character()),
    na = character(),
    show_col_types = FALSE
  )
  stopifnot(
    nrow(observed) == nrow(data),
    identical(names(observed), names(data))
  )
  if (!file.rename(temporary_path, path)) {
    stop("Could not promote validated analysis output.", call. = FALSE)
  }
}
write_csv_atomic(audit, audit_path)
write_csv_atomic(audit_summary, audit_summary_path)
stopifnot(file.exists(audit_path), file.exists(audit_summary_path))
```

## Identity outcomes

These categories describe exact-ISSN identity evidence. They do not establish
which source is correct and do not explain why a difference exists.

```{r}
#| label: identity-outcome-chart
#| fig-width: 9
#| fig-height: 5

audit_summary |>
  filter(dimension == "identity_outcome") |>
  ggplot(aes(x = reorder(category, count), y = count)) +
  geom_col() +
  geom_text(
    aes(label = sprintf("%s (%.2f%%)", count, percent)),
    hjust = -0.05,
    size = 3
  ) +
  coord_flip(clip = "off") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  labs(x = NULL, y = "PKP V7 rows") +
  theme_minimal()
```

## Metadata comparisons

Title comparison uses the PKP title against the OpenAlex display name and
alternate titles. PKP country is inferred metadata, not authoritative
publisher location. Absent DOAJ evidence is not a negative assertion.

```{r}
#| label: metadata-disagreement-chart
#| fig-width: 9
#| fig-height: 4.5

metadata_disagreement_summary |>
  ggplot(aes(x = reorder(category, count), y = count)) +
  geom_col() +
  geom_text(
    aes(label = sprintf("%s (%.2f%%)", count, percent)),
    hjust = -0.05,
    size = 3
  ) +
  coord_flip(clip = "off") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  labs(x = NULL, y = "Identity-matched OpenAlex journals") +
  theme_minimal()
```

## OpenAlex and Crossref match states

```{r}
#| label: source-status-heatmap
#| fig-width: 9
#| fig-height: 6

audit_summary |>
  filter(dimension == "source_status_crosstab") |>
  mutate(
    openalex = sub("\\|.*$", "", sub("^openalex=", "", category)),
    crossref = sub("^.*\\|crossref=", "", category)
  ) |>
  ggplot(aes(x = openalex, y = crossref, fill = count)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%s\n%.2f%%", count, percent)), size = 3) +
  labs(x = "OpenAlex match state", y = "Crossref match state", fill = "Rows") +
  theme_minimal()
```

## Purposive ten-row review

```{r}
#| label: disagreement-review

sample_spec <- tribble(
  ~context_name, ~oai_url, ~repository_name, ~set_spec,
  "Kuwait Journal of Science",
  "https://journalskuwait.org/kjs/index.php/index/oai",
  "Kuwait Journal of Science",
  "KJS",
  "Cakrawala",
  "https://cakrawalajournal.org/index.php/index/oai",
  "CAKRAWALA",
  "cakrawala",
  "Sintesa: Jurnal Ilmu Pendidikan",
  "https://sintesa.stkip-arrahmaniyah.ac.id/index.php/index/oai",
  "Sintesa: Jurnal Ilmu Pendidikan STKIP Arrahmaniyah Depok",
  "sintesa",
  "CARAKA: Jurnal Teologi Biblika dan Praktika",
  "https://ojs.sttibc.ac.id/index.php/index/oai",
  "CARAKA:Jurnal Teologi",
  "ibc",
  "Jurnal Tata Kelola dan Akuntabilitas Keuangan Negara",
  "https://jurnal.bpk.go.id/index.php/index/oai",
  "Open Journal Systems",
  "TAKEN",
  "JOEEL (Journal of English Education and Literature)",
  "https://journal.stkippamanetalino.ac.id/index.php/index/oai",
  "Rumah Jurnal STKIP Pamane Talino Ngabang",
  "bahasa-inggris",
  "Malaysian Journal of Paediatrics and Child Health",
  "https://mpaeds.my/journals/index.php/index/oai",
  "NA",
  "MJPCH",
  "Jami Scientific Research Quarterly Journal",
  "https://journals.jami.edu.af/index.php/index/oai",
  "Jami University Press",
  "jsrqj",
  "Journal of Polymer & Composites",
  "https://engineeringjournals.stmjournals.in/index.php/index/oai",
  "Engineering Journals",
  "JoPC",
  "JURNAL TEKNIK PERTAMBANGAN",
  "https://e-journal.upr.ac.id/index.php/index/oai",
  "Journal Online Universitas Palangka Raya",
  "JTP"
)
sample_positions <- match(row_identity(sample_spec), row_identity(audit))
stopifnot(!anyNA(sample_positions))
review_rows <- audit[sample_positions, ]

htmltools::div(
  style = "overflow-x: auto;",
  reactable::reactable(
    review_rows,
    searchable = TRUE,
    filterable = TRUE,
    pagination = TRUE,
    defaultPageSize = 10,
    resizable = TRUE,
    compact = TRUE,
    fullWidth = FALSE
  )
)
```

## Why rows were not attempted

`not_attempted` is not an API failure. It means the PKP row had no
checksum-valid ISSN, so the pipeline had no identifier to compare with either
source. The diagnostics below distinguish the raw ISSN state, show the size and
site concentration of affected OJS sets, and identify limited clues for later
manual recovery. Title and `set_spec` similarities are only text-level review
signals; they are not used to assign an ISSN.

```{r}
#| label: not-attempted-profile

master_hosts <- if_else(
  is_missing_value(master$oai_url),
  "Missing URL",
  sub("/.*$", "", sub("^https?://", "", tolower(master$oai_url)))
)
valid_title_rows <- master$identifier_status == "valid" &
  !is.na(pkp_normalized_titles)
valid_title_keys <- unique(pkp_normalized_titles[valid_title_rows])
valid_title_set_text <- paste(
  pkp_normalized_titles[valid_title_rows],
  tolower(trimws(master$set_spec[valid_title_rows])),
  sep = "\r"
)
valid_title_set_text_keys <- unique(valid_title_set_text)
valid_host_title_set_keys <- unique(paste(
  master_hosts[valid_title_rows],
  valid_title_set_text,
  sep = "\r"
))

not_attempted <- master |>
  mutate(
    title_key = pkp_normalized_titles,
    title_set_text_key = paste(
      title_key,
      tolower(trimws(set_spec)),
      sep = "\r"
    ),
    host_title_set_key = paste(master_hosts, title_set_text_key, sep = "\r"),
    raw_issn_state = case_when(
      is.na(issn) ~ "R missing value",
      !nzchar(trimws(issn)) ~ "Blank string",
      toupper(trimws(issn)) == "NA" ~ "Literal NA in PKP",
      TRUE ~ "Present but invalid ISSN"
    ),
    record_count = suppressWarnings(as.numeric(total_record_count)),
    record_band = factor(
      case_when(
        is.na(record_count) ~ "Missing",
        record_count < 10 ~ "Under 10",
        record_count < 25 ~ "10–24",
        record_count < 50 ~ "25–49",
        record_count < 100 ~ "50–99",
        record_count < 1000 ~ "100–999",
        TRUE ~ "1,000+"
      ),
      levels = c(
        "Under 10", "10–24", "25–49", "50–99",
        "100–999", "1,000+", "Missing"
      )
    ),
    host = master_hosts,
    same_title_as_valid = !is.na(title_key) & title_key %in% valid_title_keys,
    same_title_set_text_as_valid = !is.na(title_key) &
      title_set_text_key %in% valid_title_set_text_keys,
    same_host_title_set_as_valid = !is.na(title_key) &
      host_title_set_key %in% valid_host_title_set_keys
  ) |>
  filter(openalex_match_status == "not_attempted")

stopifnot(
  all(not_attempted$identifier_status != "valid"),
  all(not_attempted$crossref_match_status == "not_attempted"),
  all(master$openalex_match_status[master$identifier_status != "valid"] ==
    "not_attempted")
)

not_attempted_summary <- tibble(
  measure = c(
    "All PKP rows not attempted",
    "Raw ISSN is literal NA",
    "Raw ISSN is an R missing value",
    "Raw ISSN is blank",
    "Raw ISSN is present but invalid",
    "Same normalized title has a valid-ISSN row",
    "Same normalized title and set_spec text occurs on a valid-ISSN row",
    "Same host, normalized title, and set_spec has a valid-ISSN row",
    "PKP DOAJ URL is present",
    "At least 100 OAI records"
  ),
  denominator = c(
    nrow(master),
    rep(nrow(not_attempted), 9)
  ),
  rows = c(
    nrow(not_attempted),
    sum(not_attempted$raw_issn_state == "Literal NA in PKP"),
    sum(not_attempted$raw_issn_state == "R missing value"),
    sum(not_attempted$raw_issn_state == "Blank string"),
    sum(not_attempted$raw_issn_state == "Present but invalid ISSN"),
    sum(not_attempted$same_title_as_valid),
    sum(not_attempted$same_title_set_text_as_valid),
    sum(not_attempted$same_host_title_set_as_valid),
    sum(!is_missing_value(not_attempted$best_doaj_url)),
    sum(not_attempted$record_count >= 100, na.rm = TRUE)
  )
) |>
  mutate(percent = round(100 * rows / denominator, 2))

knitr::kable(not_attempted_summary)
```

In this snapshot, `r format(nrow(not_attempted), big.mark = ",")` rows
(`r sprintf("%.2f%%", 100 * nrow(not_attempted) / nrow(master))` of PKP rows)
were not attempted. All of their raw ISSN cells are the literal PKP value
`NA`, rather than malformed identifiers rejected by the checksum rule.

### Size of affected OJS sets

```{r}
#| label: not-attempted-record-distribution
#| fig-width: 9
#| fig-height: 5

not_attempted_record_distribution <- not_attempted |>
  count(record_band, .drop = FALSE, name = "rows") |>
  mutate(percent = 100 * rows / sum(rows))

not_attempted_record_distribution |>
  ggplot(aes(x = record_band, y = rows)) +
  geom_col(fill = "#3B6E8F", color = "#24485F", linewidth = 0.2) +
  geom_text(
    aes(label = sprintf("%s\n%.1f%%", format(rows, big.mark = ","), percent)),
    vjust = -0.15,
    size = 3
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.14))) +
  labs(
    title = "OAI record counts for rows without a usable ISSN",
    subtitle = "PKP V7 rows with match status not_attempted",
    x = "Records reported for the OJS set",
    y = "PKP rows"
  ) +
  theme_minimal()
```

`r sprintf("%.1f%%", 100 * mean(not_attempted$record_count < 25, na.rm = TRUE))`
of the affected OJS sets contain fewer than 25 records, so much of the missing
ISSN volume comes from small collections. However,
`r format(sum(not_attempted$record_count >= 100, na.rm = TRUE), big.mark = ",")`
rows contain at least 100 records and should not be dismissed as empty or test
sets.

### Concentration by OJS host

```{r}
#| label: not-attempted-host-concentration
#| fig-width: 9
#| fig-height: 5.5

not_attempted_hosts <- not_attempted |>
  count(host, sort = TRUE, name = "rows") |>
  mutate(percent = 100 * rows / nrow(not_attempted)) |>
  slice_head(n = 10)

not_attempted_hosts |>
  ggplot(aes(x = reorder(host, rows), y = rows)) +
  geom_col(fill = "#3B6E8F", color = "#24485F", linewidth = 0.2) +
  geom_text(
    aes(label = sprintf("%s (%.1f%%)", format(rows, big.mark = ","), percent)),
    hjust = -0.05,
    size = 3
  ) +
  coord_flip(clip = "off") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  labs(
    title = "OJS hosts with the most rows lacking a usable ISSN",
    subtitle = "Top ten hosts; percentages use all not-attempted rows",
    x = NULL,
    y = "PKP rows"
  ) +
  theme_minimal()
```

The two largest hosts account for
`r sprintf("%.1f%%", 100 * sum(head(not_attempted_hosts$rows, 2)) / nrow(not_attempted))`
of all not-attempted rows. Concentration explains part of the total, but most
missing identifiers remain distributed across the long tail of OJS hosts.

### Fixed random review sample

The sample is reproducible rather than manually selected. It helps distinguish
journal-like titles from books, proceedings, test installations, and other OJS
sets, but those interpretations remain manual judgments.

```{r}
#| label: not-attempted-random-review

set.seed(20260815)
not_attempted_sample <- not_attempted |>
  slice_sample(n = min(12L, nrow(not_attempted))) |>
  transmute(
    title = context_name,
    host,
    records = record_count,
    review_signal = case_when(
      same_host_title_set_as_valid ~ "Same host, title, and set_spec",
      same_title_set_text_as_valid ~ "Same title/set text elsewhere",
      same_title_as_valid ~ "Same title elsewhere",
      TRUE ~ "None"
    )
  )

htmltools::div(
  style = "overflow-x: auto;",
  htmltools::div(
    style = "min-width: 850px;",
    reactable::reactable(
      not_attempted_sample,
      searchable = TRUE,
      filterable = TRUE,
      pagination = FALSE,
      resizable = TRUE,
      compact = TRUE,
      fullWidth = TRUE,
      columns = list(
        title = reactable::colDef(minWidth = 350),
        host = reactable::colDef(minWidth = 250),
        records = reactable::colDef(width = 80),
        review_signal = reactable::colDef(width = 170)
      )
    )
  )
)
```
