---
title: "PKP/OJS multi-source journal enrichment"
format: html
execute:
  echo: true
  warning: false
  message: false
---

This notebook left-joins selected OpenAlex and Crossref journal metadata onto a **purposive validation sample** of ten source-authentic PKP OJS rows by exact ISSN. The sample exercises missing, unmatched, consistent, and inconsistent match branches; it is not representative and must not be used to estimate full-cohort coverage.

Set `OPENALEX_API_KEY` in the environment or the repository-root `.env` file before rendering. The active code is intentionally limited to the ten-row sample. Before a formal full-cohort run, replace the sections marked `FULL-COHORT REPLACEMENT` with reviewed batching, caching, checkpointing, and error-provenance code.

## Packages and paths

This setup chunk loads the required R packages and fixes the input and output paths to the ten-row sample. It uses an already exported `OPENALEX_API_KEY` when available and otherwise reads the repository-root `.env`.

The commented lines mark the path replacement required for a formal full-cohort run. They are deliberately inactive in this sample notebook.

```{r}
#| label: packages-and-paths

library(dplyr)
library(httr2)
library(purrr)
library(readr)
library(tibble)

project_env_path <- file.path("..", "..", "..", ".env")
if (!nzchar(Sys.getenv("OPENALEX_API_KEY")) && file.exists(project_env_path)) {
  readRenviron(project_env_path)
}
if (!nzchar(Sys.getenv("OPENALEX_API_KEY"))) {
  stop("Set OPENALEX_API_KEY before running this notebook.", call. = FALSE)
}

input_path <- file.path(
  "..",
  "data",
  "ojs-journal-enrichment-sample.csv"
)
artifact_dir <- file.path("..", "artifacts", "ojs_journal_enrichment")
run_dir <- file.path(artifact_dir, "sample")
dir.create(run_dir, recursive = TRUE, showWarnings = FALSE)

output_path <- file.path(run_dir, "pkp-ojs-multisource-enriched.csv")

# FULL-COHORT REPLACEMENT (inactive):
# data_root <- Sys.getenv("DATA_ROOT")
# input_path <- file.path(data_root, "raw", "pkp-beacon-v6", "beacon.csv")
# run_dir <- file.path(artifact_dir, "full")
```

## ISSN preparation

Exact matching depends on checksum-valid ISSNs rather than raw identifier
strings. This chunk removes display punctuation, validates the ISSN check
digit, parses rows that contain multiple identifiers, and applies the same
normalization to identifiers returned by external sources. Missing and invalid
identifiers therefore remain distinguishable from valid but unmatched ISSNs.

```{r}
#| label: issn-preparation

normalize_issn <- function(value) {
  compact <- gsub("[^0-9X]", "", toupper(trimws(as.character(value))))
  if (!grepl("^[0-9]{7}[0-9X]$", compact)) return(NA_character_)

  digits <- as.integer(strsplit(substr(compact, 1, 7), "")[[1]])
  check_value <- (11 - sum(digits * 8:2) %% 11) %% 11
  expected <- if (check_value == 10) "X" else as.character(check_value)
  if (substr(compact, 8, 8) != expected) return(NA_character_)

  paste0(substr(compact, 1, 4), "-", substr(compact, 5, 8))
}

parse_issns <- function(value) {
  if (is.na(value) || !nzchar(value)) return(character())

  tokens <- unlist(strsplit(value, "[[:space:],;|]+"))
  normalized <- vapply(tokens[nzchar(tokens)], normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}

source_issns <- function(source) {
  values <- c(unlist(source$issn, use.names = FALSE), source$issn_l)
  normalized <- vapply(values, normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}
```

## Source retrieval

The ten-row sample requires one OpenAlex request and one rate-limited Crossref request per distinct valid ISSN. A Crossref `404` is treated as an unmatched journal. The requests retain timeouts because both APIs are external trust boundaries.

For a formal full-cohort run, replace this chunk with the commented production requirements: OpenAlex batching and cursor pagination, Crossref polite-pool identification, bounded retries, cache checkpoints, and per-ISSN error provenance.

```{r}
#| label: source-retrieval

fetch_openalex_sources <- function(
  issns,
  api_key = Sys.getenv("OPENALEX_API_KEY")
) {
  stopifnot(length(issns) <= 100L, nzchar(api_key))

  request("https://api.openalex.org/sources") |>
    req_url_query(
      filter = paste0("issn:", paste(issns, collapse = "|")),
      per_page = 100,
      api_key = api_key
    ) |>
    req_timeout(seconds = 30) |>
    req_perform() |>
    resp_body_json(simplifyVector = FALSE) |>
    pluck("results")
}

fetch_crossref_journal <- function(issn) {
  response <- request(paste0("https://api.crossref.org/journals/", issn)) |>
    req_timeout(seconds = 30) |>
    req_error(
      is_error = \(response) {
        resp_status(response) != 404L && resp_is_error(response)
      }
    ) |>
    req_perform()
  if (resp_status(response) == 404L) return(NULL)
  resp_body_json(response, simplifyVector = FALSE)$message
}

# FULL-COHORT REPLACEMENT (inactive):
# - split OpenAlex ISSNs into batches of at most 100 and follow next_cursor;
# - identify Crossref requests with CROSSREF_MAILTO and its polite-pool rate;
# - add bounded retries, persistent caches, checkpoints, and error provenance.
```

## Load the PKP/OJS sample

The fixture was extracted from PKP Beacon V6 file `13173372` on 2026-07-17. Its MD5 is `b5fb99c10f38c2a5c3e7e6e0d4aa2212`. This chunk fixes the analysis boundary at ten OJS rows and checks their composite PKP identities before any API request.

For a formal full-cohort run, replace the commented input path in the setup chunk and replace these sample assertions with the pinned full-file checksum and cohort counts.

```{r}
#| label: load-sample

stopifnot(
  file.exists(input_path),
  unname(tools::md5sum(input_path)) == "b5fb99c10f38c2a5c3e7e6e0d4aa2212"
)

raw_input <- read_csv(
  input_path,
  col_types = cols(.default = col_character()),
  na = character(),
  show_col_types = FALSE
)
ojs_rows <- raw_input |>
  filter(tolower(application) == "ojs") |>
  select(-any_of("admin_email"))
row_identity <- paste(
  ojs_rows$oai_url,
  ojs_rows$repository_name,
  ojs_rows$set_spec,
  sep = "\r"
)
stopifnot(nrow(ojs_rows) == 10L, !anyDuplicated(row_identity))

# FULL-COHORT REPLACEMENT (inactive):
# stopifnot(unname(tools::md5sum(input_path)) ==
#   "9f43fa051c7ed1cc45d8592593542011")
# stopifnot(nrow(raw_input) == 87170L, nrow(ojs_rows) == 86282L)
```

## Multi-source enrichment

The sample keeps only the matching rules needed to produce readable scalar CSV columns. OpenAlex candidates are deduplicated by Source ID; Crossref candidates are deduplicated by normalized ISSN set. API failures stop the render, so rerunning repeats the small set of requests.

### Sample matching helpers

These helpers assign the four sample match outcomes, construct Crossref candidate identities, and return source fields only when a row has one consistent candidate.

```{r}
#| label: sample-matching-helpers

classify_status <- function(input_issns, candidate_ids) {
  case_when(
    length(input_issns) == 0L ~ "not_attempted",
    length(candidate_ids) == 0L ~ "unmatched",
    length(candidate_ids) == 1L ~ "consistent",
    TRUE ~ "inconsistent"
  )
}

crossref_key <- function(journal) {
  issns <- vapply(journal$ISSN, normalize_issn, character(1))
  paste(sort(unique(issns[!is.na(issns)])), collapse = "|")
}

consistent_field <- function(match, field) {
  if (match$status != "consistent") return(NA_character_)
  value <- pluck(match$candidates, 1, field, .default = NA_character_)
  if (!length(value)) NA_character_ else as.character(value[[1]])
}
```

### Sample enrichment orchestration

This chunk retrieves the distinct sample ISSNs, applies source-specific candidate identities, selects a fixed review schema, and writes one CSV row per PKP row. It intentionally has no cache, checkpoint, retry loop, dynamic schema, nested candidate JSON, or run metadata.

For a formal full-cohort run, replace this entire chunk with a reviewed pipeline that restores the inactive requirements listed at its end.

```{r}
#| label: sample-enrichment

input_issns <- map(ojs_rows$issn, parse_issns)
tokens <- sort(unique(unlist(input_issns)))

openalex_sources <- fetch_openalex_sources(tokens)
crossref_journals <- set_names(
  map(tokens, \(token) {
    journal <- fetch_crossref_journal(token)
    Sys.sleep(0.25)
    journal
  }),
  tokens
)

openalex_matches <- map(input_issns, \(issns) {
  candidates <- keep(openalex_sources, \(source) {
    any(issns %in% source_issns(source))
  })
  candidate_ids <- map_chr(candidates, "id")
  candidates <- candidates[!duplicated(candidate_ids)]
  candidate_ids <- unique(candidate_ids)

  list(
    status = classify_status(issns, candidate_ids),
    matched_issns = intersect(
      issns,
      unique(unlist(map(candidates, source_issns)))
    ),
    candidate_ids = candidate_ids,
    candidates = candidates
  )
})

crossref_matches <- map(input_issns, \(issns) {
  matched_issns <- keep(issns, \(issn) {
    !is.null(crossref_journals[[issn]])
  })
  candidates <- unname(crossref_journals[matched_issns])
  candidate_ids <- map_chr(candidates, crossref_key)
  candidates <- candidates[!duplicated(candidate_ids)]
  candidate_ids <- unique(candidate_ids)

  list(
    status = classify_status(issns, candidate_ids),
    matched_issns = matched_issns,
    candidate_ids = candidate_ids,
    candidates = candidates
  )
})

result <- ojs_rows |>
  mutate(
    identifier_status = case_when(
      is.na(issn) | !nzchar(issn) ~ "missing",
      lengths(input_issns) == 0L ~ "invalid",
      TRUE ~ "valid"
    ),
    openalex_input_issns = map_chr(input_issns, paste, collapse = "|"),
    openalex_matched_issns = map_chr(
      openalex_matches,
      \(match) paste(match$matched_issns, collapse = "|")
    ),
    openalex_match_status = map_chr(openalex_matches, "status"),
    openalex_candidate_ids = map_chr(
      openalex_matches,
      \(match) paste(match$candidate_ids, collapse = "|")
    ),
    openalex_display_name = map_chr(
      openalex_matches,
      consistent_field,
      field = "display_name"
    ),
    openalex_country_code = map_chr(
      openalex_matches,
      consistent_field,
      field = "country_code"
    ),
    openalex_works_count = map_chr(
      openalex_matches,
      consistent_field,
      field = "works_count"
    ),
    crossref_input_issns = map_chr(input_issns, paste, collapse = "|"),
    crossref_matched_issns = map_chr(
      crossref_matches,
      \(match) paste(match$matched_issns, collapse = "|")
    ),
    crossref_match_status = map_chr(crossref_matches, "status"),
    crossref_candidate_issn_sets = map_chr(
      crossref_matches,
      \(match) paste(match$candidate_ids, collapse = ";")
    ),
    crossref_title = map_chr(
      crossref_matches,
      consistent_field,
      field = "title"
    ),
    crossref_publisher = map_chr(
      crossref_matches,
      consistent_field,
      field = "publisher"
    )
  )

write_csv(result, output_path, na = "")

# FULL-COHORT REPLACEMENT (inactive):
# - batch OpenAlex requests and paginate each batch;
# - cache source responses by normalized ISSN and checkpoint writes;
# - retry transient failures and record per-ISSN API error provenance;
# - restore complete namespaced source fields and candidate records;
# - write run metadata with input, output, and cache checksums.
```

## Checks

These assertions form the executable sample contract. They verify one output row per input row, the fixed review schema, the CSV artifact, and the intended match branches. They do not establish representative full-cohort coverage.

```{r}
#| label: contract-check

stopifnot(
  exists("fetch_openalex_sources", mode = "function"),
  exists("fetch_crossref_journal", mode = "function"),
  nrow(result) == nrow(ojs_rows),
  file.exists(output_path),
  all(c(
    "openalex_match_status",
    "openalex_display_name",
    "crossref_match_status",
    "crossref_title"
  ) %in% names(result)),
  all(c(
    "consistent",
    "unmatched",
    "inconsistent",
    "not_attempted"
  ) %in% result$openalex_match_status),
  "not_attempted" %in% result$crossref_match_status
)
```

## Selected sample review

The table below previews the same fixed scalar fields written to the sample CSV. Complete source records are intentionally outside this exploratory implementation.

```{r}
#| label: selected-review-fields

result |>
  select(any_of(c(
    "context_name",
    "openalex_input_issns",
    "identifier_status",
    "openalex_match_status",
    "openalex_matched_issns",
    "openalex_display_name",
    "openalex_country_code",
    "openalex_works_count",
    "crossref_match_status",
    "crossref_matched_issns",
    "crossref_title",
    "crossref_publisher"
  ))) |>
  knitr::kable()
```

## Complete merged column list

This final audit table enumerates the exact merged schema produced by the run.
It supports inspection of retained source fields and provenance columns but
does not select variables for substantive analysis.

```{r}
#| label: complete-column-list

tibble(column = names(result)) |>
  knitr::kable()
```
