---
title: "OpenAlex Source metadata sample"
format: html
execute:
  echo: true
  warning: false
  message: false
---

This notebook tests the supervisor-requested shape: one OJS journal per row,
with the complete OpenAlex `Source` object retained before later column
selection. It intentionally does not read or run the full OJS cohort.

```{r}
#| label: packages

library(dplyr)
library(httr2)
library(jsonlite)
library(purrr)
library(tibble)
```

## Function

`fetch_openalex_source()` accepts one or more ISSNs belonging to one journal.
It calls the OpenAlex singleton endpoint without `select`, so every top-level
field in the returned `Source` object is retained. Nested objects and arrays
are both stored as lossless JSON and recursively expanded into indexed
`openalex_*` columns such as `openalex_topics_1_display_name` for later
filtering.

```{r}
#| label: fetch-openalex-source

fetch_openalex_source <- function(
  issns,
  api_key = Sys.getenv("OPENALEX_API_KEY")
) {
  raw_tokens <- unlist(
    strsplit(paste(stats::na.omit(issns), collapse = "\n"), "[[:space:],;|]+")
  )
  raw_tokens <- raw_tokens[nzchar(raw_tokens)]
  if (!length(raw_tokens)) {
    stop("Supply at least one ISSN.", call. = FALSE)
  }

  normalize_issn <- function(value) {
    compact <- gsub("[^0-9X]", "", toupper(value))
    if (!grepl("^[0-9]{7}[0-9X]$", compact)) {
      stop(sprintf("Invalid ISSN: %s", value), call. = FALSE)
    }
    digits <- as.integer(strsplit(substr(compact, 1, 7), "")[[1]])
    check_value <- (11 - sum(digits * 8:2) %% 11) %% 11
    expected <- if (check_value == 10) "X" else as.character(check_value)
    if (substr(compact, 8, 8) != expected) {
      stop(sprintf("Invalid ISSN checksum: %s", value), call. = FALSE)
    }
    paste0(substr(compact, 1, 4), "-", substr(compact, 5, 8))
  }

  tokens <- unique(vapply(raw_tokens, normalize_issn, character(1)))
  fetched <- lapply(tokens, function(token) {
    request <- httr2::request(
      paste0("https://api.openalex.org/sources/issn:", token)
    ) |>
      httr2::req_headers(
        Accept = "application/json",
        `User-Agent` = "InvisibleResearch/1.0"
      ) |>
      httr2::req_error(
        is_error = function(response) {
          httr2::resp_status(response) != 404L &&
            httr2::resp_is_error(response)
        }
      ) |>
      httr2::req_retry(max_tries = 3)
    if (nzchar(api_key)) {
      request <- httr2::req_url_query(request, api_key = api_key)
    }

    response <- httr2::req_perform(request)
    source <- if (httr2::resp_status(response) == 404L) {
      NULL
    } else {
      httr2::resp_body_json(response, simplifyVector = FALSE)
    }
    list(issn = token, source = source)
  })

  matched <- fetched[!vapply(fetched, function(item) is.null(item$source), logical(1))]
  candidate_ids <- unique(vapply(
    matched,
    function(item) item$source$id,
    character(1)
  ))
  provenance <- list(
    input_issns = as.character(jsonlite::toJSON(tokens, auto_unbox = FALSE)),
    matched_issns = as.character(jsonlite::toJSON(
      vapply(matched, `[[`, character(1), "issn"),
      auto_unbox = FALSE
    )),
    match_status = if (!length(candidate_ids)) {
      "unmatched"
    } else if (length(candidate_ids) == 1L) {
      "unique"
    } else {
      "ambiguous"
    },
    candidate_source_ids = as.character(jsonlite::toJSON(
      candidate_ids,
      auto_unbox = FALSE
    )),
    openalex_retrieved_at = format(
      Sys.time(),
      tz = "UTC",
      usetz = TRUE
    )
  )

  if (length(candidate_ids) != 1L) {
    return(tibble::as_tibble(provenance))
  }

  source <- matched[[1]]$source
  flatten_value <- function(value, path) {
    if (is.null(value) || !length(value)) {
      return(stats::setNames(list(NA), path))
    }
    if (is.atomic(value) && length(value) == 1L) {
      return(stats::setNames(list(value), path))
    }
    labels <- names(value)
    if (is.null(labels) || any(!nzchar(labels))) {
      labels <- seq_along(value)
    }
    children <- Map(
      function(child, label) {
        flatten_value(child, paste(path, label, sep = "_"))
      },
      value,
      labels
    )
    do.call(c, unname(children))
  }

  source_cells <- list()
  for (field in names(source)) {
    value <- source[[field]]
    path <- paste0("openalex_", field)
    source_cells[[paste0(path, "__json")]] <- as.character(
      jsonlite::toJSON(
        value,
        auto_unbox = TRUE,
        null = "null",
        na = "null",
        digits = NA
      )
    )
    source_cells <- c(source_cells, flatten_value(value, path))
  }
  source_cells$openalex_source__json <- as.character(jsonlite::toJSON(
    source,
    auto_unbox = TRUE,
    null = "null",
    na = "null",
    digits = NA
  ))
  if (anyDuplicated(names(source_cells))) {
    stop("Flattening the OpenAlex Source object produced duplicate columns.")
  }

  output <- tibble::as_tibble(c(provenance, source_cells))
  retained_fields <- vapply(
    names(source),
    function(field) {
      path <- paste0("openalex_", field)
      paste0(path, "__json") %in% names(output)
    },
    logical(1)
  )
  if (!all(retained_fields) || !"openalex_source__json" %in% names(output)) {
    stop("The OpenAlex Source object lost one or more top-level fields.")
  }
  output
}
```

## Ten-journal sample

The sample is embedded here so the notebook makes no full-cohort read. It
includes consistent multi-ISSN matches, one unmatched ISSN, and one journal
whose ISSNs resolve to conflicting OpenAlex Source IDs.

```{r}
#| label: sample-input

sample_journals <- tibble::tibble(
  sample_journal = c(
    "Kuwait Journal of Science",
    "CAKRAWALA",
    "Sintesa: Jurnal Ilmu Pendidikan",
    "CARAKA: Jurnal Teologi",
    "Open Journal Systems",
    "Rumah Jurnal STKIP Pamane Talino Ngabang 1",
    "Journal with ISSNs 1511-4511 and 3093-8112",
    "Rumah Jurnal STKIP Pamane Talino Ngabang 2",
    "Journal with ISSNs 2088-4842 and 2442-8795",
    "Journal Online Universitas Palangka Raya"
  ),
  issns = list(
    c("2307-4108", "2307-4116"),
    c("1978-0354", "2622-013x"),
    "2085-7063",
    c("2722-1393", "2722-1407"),
    c("2460-3937", "2549-452X"),
    c("2723-309X", "2723-3103"),
    c("1511-4511", "3093-8112"),
    c("2722-4376", "2722-4384"),
    c("2088-4842", "2442-8795"),
    c("2087-1058", "2988-7712")
  )
)

sample_results <- purrr::map2_dfr(
  sample_journals$issns,
  sample_journals$sample_journal,
  function(issns, journal) {
    fetch_openalex_source(issns) |>
      dplyr::mutate(sample_journal = journal, .before = 1)
  }
)
```

## Checks

```{r}
#| label: contract-check

source_columns <- names(sample_results)[
  startsWith(names(sample_results), "openalex_") &
    names(sample_results) != "openalex_retrieved_at"
]
status_for <- function(journal) {
  sample_results$match_status[sample_results$sample_journal == journal]
}

stopifnot(
  exists("fetch_openalex_source", mode = "function"),
  nrow(sample_results) == 10L,
  length(source_columns) >= 30L,
  all(c(
    "openalex_id",
    "openalex_topics__json",
    "openalex_topics_1_display_name",
    "openalex_counts_by_year_1_year",
    "openalex_summary_stats_2yr_mean_citedness",
    "openalex_source__json"
  ) %in% source_columns),
  status_for("Kuwait Journal of Science") == "unique",
  status_for("Sintesa: Jurnal Ilmu Pendidikan") == "unmatched",
  status_for("Journal Online Universitas Palangka Raya") == "ambiguous"
)

cat(
  "Checks passed:",
  nrow(sample_results),
  "journals and",
  length(source_columns),
  "OpenAlex Source columns retained.\n"
)
```

## Sample outcomes

```{r}
#| label: sample-outcomes

sample_results |>
  dplyr::select(
    sample_journal,
    input_issns,
    matched_issns,
    match_status,
    candidate_source_ids
  ) |>
  knitr::kable()
```

## Complete OpenAlex columns

```{r}
#| label: complete-column-list

tibble::tibble(column = source_columns) |>
  knitr::kable()
```

The complete values remain in `sample_results`. The
`openalex_source__json` column also preserves each selected Source object
losslessly if a later filtering decision needs a different flattening shape.
