---
title: "OpenAlex Source metadata sample"
format: html
execute:
  echo: true
  warning: false
  message: false
---

This notebook tests the supervisor-requested workflow on ten ISSNs rather than
the full OJS dataset. Each ISSN is matched through the OpenAlex singleton
endpoint, and every top-level field in the returned `Source` object is kept
before any analytical selection.

```{r}
#| label: packages

library(dplyr)
library(httr2)
library(purrr)
library(tibble)
```

## Retrieve one OpenAlex Source

```{r}
#| label: fetch-openalex-source

fetch_openalex_source <- function(
  issn,
  api_key = Sys.getenv("OPENALEX_API_KEY")
) {
  stopifnot(length(issn) == 1L, !is.na(issn))

  source <- request(
    paste0("https://api.openalex.org/sources/issn:", toupper(trimws(issn)))
  ) |>
    req_url_query(api_key = api_key) |>
    req_perform() |>
    resp_body_json(simplifyVector = FALSE)

  as_tibble(lapply(source, list))
}
```

The function does not use the OpenAlex `select` parameter. All returned fields
remain list-columns, so `null` and nested values can coexist across journals
without discarding fields or imposing a project-specific shape during retrieval.

## Ten-ISSN sample

```{r}
#| label: sample-input

sample_issns <- tibble(
  input_issn = c(
    "2307-4108",
    "1978-0354",
    "2722-1393",
    "2460-3937",
    "2723-309X",
    "2722-4376",
    "2088-4842",
    "2087-1058",
    "1511-4511",
    "3093-8112"
  )
)

openalex_sources <- sample_issns$input_issn |>
  map(fetch_openalex_source) |>
  bind_rows() |>
  mutate(input_issn = sample_issns$input_issn, .before = 1)
```

```{r}
#| label: check-sample

stopifnot(
  nrow(openalex_sources) == nrow(sample_issns),
  all(c("id", "display_name", "issn", "topics") %in% names(openalex_sources))
)
```

## Inspect all retrieved columns

```{r}
#| label: all-column-names

tibble(column = names(openalex_sources)) |>
  knitr::kable()
```

## Select analytical columns afterwards

```{r}
#| label: selected-columns

openalex_sources |>
  transmute(
    input_issn,
    id = map_chr(id, identity),
    display_name = map_chr(display_name, identity),
    is_oa = map_lgl(is_oa, identity),
    is_in_doaj = map_lgl(is_in_doaj, identity),
    is_ojs = map_lgl(is_ojs, identity)
  ) |>
  knitr::kable()
```
