---
title: "OpenAlex Source metadata sample"
format: html
execute:
  echo: true
  warning: false
  message: false
---

This notebook tests the supervisor-requested workflow on ten journals rather
than the full OJS dataset. Each ISSN is matched through the OpenAlex singleton
endpoint, and every top-level field in the returned `Source` object is kept
before any analytical selection. Journals with multiple ISSNs are represented
by one row only when their matched ISSNs resolve to the same OpenAlex Source.

```{r}
#| label: packages

library(dplyr)
library(httr2)
library(purrr)
library(tibble)
```

## Retrieve one OpenAlex Source

```{r}
#| label: fetch-openalex-source

fetch_openalex_source <- function(
  issn,
  api_key = Sys.getenv("OPENALEX_API_KEY")
) {
  stopifnot(length(issn) == 1L, !is.na(issn))

  response <- request(
    paste0("https://api.openalex.org/sources/issn:", toupper(trimws(issn)))
  ) |>
    req_url_query(api_key = api_key) |>
    req_error(
      is_error = \(response) {
        resp_status(response) != 404L && resp_is_error(response)
      }
    ) |>
    req_perform()

  if (resp_status(response) == 404L) return(NULL)
  source <- resp_body_json(response, simplifyVector = FALSE)
  as_tibble(lapply(source, list))
}
```

The function does not use the OpenAlex `select` parameter. All returned fields
remain list-columns, so `null` and nested values can coexist across journals
without discarding fields or imposing a project-specific shape during retrieval.

## Ten-journal sample

```{r}
#| label: sample-input

sample_journals <- tibble(
  sample_journal = c(
    "Kuwait Journal of Science",
    "CAKRAWALA",
    "Sintesa: Jurnal Ilmu Pendidikan",
    "CARAKA: Jurnal Teologi",
    "Open Journal Systems",
    "Rumah Jurnal STKIP Pamane Talino Ngabang 1",
    "Journal with ISSNs 1511-4511 and 3093-8112",
    "Rumah Jurnal STKIP Pamane Talino Ngabang 2",
    "Journal with ISSNs 2088-4842 and 2442-8795",
    "Journal Online Universitas Palangka Raya"
  ),
  issns = list(
    c("2307-4108", "2307-4116"),
    c("1978-0354", "2622-013X"),
    "2085-7063",
    c("2722-1393", "2722-1407"),
    c("2460-3937", "2549-452X"),
    c("2723-309X", "2723-3103"),
    c("1511-4511", "3093-8112"),
    c("2722-4376", "2722-4384"),
    c("2088-4842", "2442-8795"),
    c("2087-1058", "2988-7712")
  )
)

openalex_sources <- pmap(
  sample_journals,
  \(sample_journal, issns) {
    matches <- map(issns, fetch_openalex_source)
    matched <- !map_lgl(matches, is.null)
    matched_results <- matches[matched]
    matched_issns <- issns[matched]
    source_ids <- unique(map_chr(
      matched_results,
      \(result) result$id[[1]]
    ))

    match_status <- case_when(
      length(source_ids) == 0L ~ "unmatched",
      length(source_ids) == 1L ~ "consistent",
      TRUE ~ "inconsistent"
    )
    match_info <- tibble(
      sample_journal,
      input_issns = list(issns),
      matched_issns = list(matched_issns),
      match_status,
      candidate_source_ids = list(source_ids)
    )

    if (match_status != "consistent") return(match_info)
    bind_cols(match_info, matched_results[[1]])
  }
) |>
  bind_rows()
```

`consistent` means that every successful lookup for the journal returned the
same Source ID. `input_issns` and `matched_issns` remain separate so partial
matches are visible.

```{r}
#| label: check-sample

stopifnot(
  nrow(openalex_sources) == nrow(sample_journals),
  all(c(
    "display_name",
    "host_organization_name",
    "topics",
    "works_count",
    "cited_by_count",
    "summary_stats"
  ) %in% names(openalex_sources)),
  "inconsistent" %in% openalex_sources$match_status,
  all(map_int(
    openalex_sources$candidate_source_ids[
      openalex_sources$match_status == "consistent"
    ],
    length
  ) == 1L)
)
```

## Inspect all retrieved columns

```{r}
#| label: all-column-names

tibble(column = names(openalex_sources)) |>
  knitr::kable()
```

## Select analytical columns afterwards

```{r}
#| label: selected-columns

openalex_sources |>
  mutate(
    input_issns = map_chr(input_issns, paste, collapse = " | "),
    matched_issns = map_chr(matched_issns, paste, collapse = " | "),
    candidate_source_ids = map_chr(
      candidate_source_ids,
      paste,
      collapse = " | "
    ),
    openalex_title = map_chr(
      display_name,
      \(value) if (is.null(value)) NA_character_ else value
    ),
    publisher = map_chr(
      host_organization_name,
      \(value) if (is.null(value)) NA_character_ else value
    ),
    topics = map_chr(topics, \(value) {
      if (is.null(value) || !length(value)) return(NA_character_)
      paste(map_chr(head(value, 2), "display_name"), collapse = " | ")
    }),
    works_count = map_int(
      works_count,
      \(value) if (is.null(value)) NA_integer_ else value
    ),
    cited_by_count = map_int(
      cited_by_count,
      \(value) if (is.null(value)) NA_integer_ else value
    ),
    mean_citedness_2yr = map_dbl(
      summary_stats,
      \(value) {
        if (is.null(value)) NA_real_ else value[["2yr_mean_citedness"]]
      }
    )
  ) |>
  select(
    sample_journal,
    input_issns,
    matched_issns,
    match_status,
    candidate_source_ids,
    openalex_title,
    publisher,
    topics,
    works_count,
    cited_by_count,
    mean_citedness_2yr
  ) |>
  knitr::kable()
```
