---
title: "Invisible Communication Science"
author: "Saurabh Khanna, Jinyi Yang"
date: "Updated: `r format(Sys.Date(), format='%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 6
    number_sections: yes
    df_print: paged
    code_folding: show
---

<br/>

```{r setup, include=FALSE}
## knitr options
knitr::opts_chunk$set(echo = T, message = FALSE, warning = FALSE, fig.retina = 4)
```


```{r message=FALSE, warning=FALSE}
## Libraries
pacman::p_load(tidyverse, janitor, skimr, estimatr, texreg, countrycode, WDI, cld3, progressr, ggrepel, jsonlite, sf)
```


```{r}
## Define a safe JSON parsing function
safe_fromJSON <- possibly(function(x) {
  x <- str_replace_all(x, "'", "\"")
  fromJSON(x)
}, otherwise = NULL)
```


# The data

We consier peer-reviewed publications with ANZSRC field marked as "Communication and Media Studies". We restrict data to year 2000 and later in order to prevent biases from poorer data quality in older publications. We consider articles, book chapters, and books only in terms of the type of literature, and exclude conferences, preprints, etc.

```{r}
df <- 
  read_csv(here::here("data/cs/dimensions_april2025/consolidated.csv")) %>% 
  clean_names() %>%
  mutate_at(c("year", "times_cited", "recent_citations", "field_citation_ratio", "relative_citation_ratio", "authors_count"), ~ as.numeric(.)) %>% 
  filter(type %in% c("article", "book", "chapter")) %>% 
  mutate(
    invisible = if_else(times_cited == 0, 1L, 0L) %>% replace_na(1L),
    type = str_to_title(type),
  ) %>%
  remove_empty() %>% 
  remove_constant() %>% 
  relocate(id, title, abstract, journal_title, authors_count, authors, type, category_for, times_cited, altmetric, concepts, concepts_scores, date, document_type, open_access, publisher, research_orgs, reference_ids)

glimpse(df %>% arrange(year, -times_cited))
```

```{r}
df %>% tabyl(type) %>% as_tibble() %>% arrange(-n) %>% mutate(percent = round(percent, 4))
```

```{r}
df %>% count(invisible)
```


# Analysis

## Figure 1

Exponential growth in research published per year:

```{r}
df %>% 
  count(type, year) %>% 
  filter(year < 2025) %>% 
  ggplot(aes(year, n, color = type)) +
  geom_line(size=0.8) +
  scale_x_continuous(breaks = scales::breaks_width(10), minor_breaks = NULL) +
  scale_y_continuous(breaks = scales::breaks_width(5000), minor_breaks = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = "Year", y = "Articles published per year", color="")
```


And now we show what % of it stays invisible (no citations even 3 years after publication):

```{r}
# plot year on x axis, and total articles and total invisible articles on y axis,

df %>% 
  filter(year >= 2000 & year <= 2022) %>%
  group_by(year) %>%
  summarise(
    all_articles = n(),
    Invisible = sum(invisible, na.rm = T)
  ) %>% 
  ungroup() %>% 
  mutate(Visible = all_articles - Invisible) %>%
  select(-all_articles) %>% 
  pivot_longer(cols = c(Visible, Invisible), names_to = "type", values_to = "n") %>% 
  mutate(type = factor(type, levels = c("Visible", "Invisible"))) %>%
  ggplot(aes(x = year, y = n, fill = type, color = type, linetype = type)) +
  geom_area(position = "stack") +
  scale_fill_manual(values = c("Visible" = "gray80", "Invisible" = "#FBB4AE")) +
  scale_color_manual(values = c("Visible" = "gray50", "Invisible" = "#E31A1C")) +
  scale_linetype_manual(values = c("Visible" = "solid", "Invisible" = "dashed")) +
  scale_y_continuous(breaks = scales::breaks_width(4000), minor_breaks = NULL) +
  scale_x_continuous(breaks = scales::breaks_width(2), minor_breaks = NULL) +
  annotate("text", x = 2017.5, y = 2250, label = "Research that stays invisible", size = 4, color = "#E31A1C") +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(x = "Year", y = "Peer-reviewed publications per year")

ggsave(here::here("figures/fig1.png"), dpi = "retina")
```


## Figure 2

Check invisibility by publication type:

```{r}
df %>% 
  filter(year >= 2000 & year <= 2022) %>%
  group_by(year, type) %>% 
  summarise(perc_invisible = mean(invisible, na.rm = T)) %>% 
  ungroup() %>% 
  ggplot(aes(x = year, y = perc_invisible)) +
  geom_smooth(method = "loess", aes(color = type), span=1.5) +
  geom_smooth(method = "loess", color = "black", se = F, span = 1.5, linetype = "dotted") +
  scale_color_brewer(palette = "Set1", direction = -1) +
  scale_x_continuous(breaks = scales::breaks_width(3), minor_breaks = NULL) +
  scale_y_continuous(breaks = scales::breaks_width(0.1), minor_breaks = NULL, labels = scales::percent_format(accuracy = 1)) +
  theme_minimal() +
  labs(x = "Year", y = "Proportion of invisible publications", color = "")
```

## Figure 3

Checks by country

```{r}
temp <-
  df %>%
  filter(year >= 2000 & year <= 2022) %>% 
  drop_na(research_org_countries) %>% 
  transmute(
    year, invisible,
    countries = str_extract_all(research_org_countries, "'id':\\s*'([A-Z]{2})'"),
    countries = map(countries, ~str_extract_all(.x, "[A-Z]{2}") %>% unlist())
  ) %>%
  unnest(countries) %>%
  group_by(countries) %>% 
  summarise(
    perc_invisible = mean(invisible, na.rm = T),
    n = n()
  ) %>%
  ungroup() %>%
  left_join(
    WDI(indicator = "NY.GDP.PCAP.CD", start = 2022, end = 2022, extra = TRUE) %>%
      select(countries = iso2c, gdp_per_capita = NY.GDP.PCAP.CD, country)
  ) %>% 
  arrange(-n) %>%
  filter(n >= 50) 

temp %>%
  ggplot(aes(x = gdp_per_capita, y = perc_invisible)) +
  geom_point(aes(size = n), alpha = 0.7, show.legend = F) +
  geom_smooth(method = "lm") +
  geom_label_repel(aes(label = country), data = temp %>% filter(perc_invisible > 0.4, gdp_per_capita < 40000), max.overlaps = 6, nudge_x = 10000, nudge_y = 0.1) +
  geom_label_repel(aes(label = country), data = temp %>% filter(perc_invisible < 0.4, gdp_per_capita > 40000, n > 5000), nudge_y = 0.1, nudge_x = 20000) +
  scale_x_continuous(labels = scales::dollar_format(prefix = "$", big.mark = ",", accuracy = 1)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_size_continuous(range = c(1, 20)) +
  theme_minimal() +
  labs(x = "National wealth (GDP per capita)", y = "Proportion of invisible publications")
```


```{r}
# choropleth map
df_world <-
  df %>%
  filter(year >= 2000 & year <= 2022) %>% 
  drop_na(research_org_countries) %>% 
  transmute(
    year, invisible,
    countries = str_extract_all(research_org_countries, "'id':\\s*'([A-Z]{2})'"),
    countries = map(countries, ~str_extract_all(.x, "[A-Z]{2}") %>% unlist())
  ) %>%
  unnest(countries) %>%
  group_by(countries) %>% 
  summarise(
    perc_invisible = mean(invisible, na.rm = T),
    n = n()
  ) %>%
  ungroup() %>%
  transmute(country = countries, total = n)

# world map shapefile
shapefile <-
  read_sf(here::here("data/TM_WORLD_BORDERS_SIMPL-0.3.shp")) %>%
  clean_names() %>%
  filter(name %in% c("Western Sahara", "Morocco")) %>% # handling boundary issues for morocco and western sahara
  mutate(name = if_else(name == "Western Sahara", "Morocco", name)) %>%
  arrange(name) %>%
  group_by(name) %>%
  summarize(
    geometry = st_union(geometry),
    lat = first(lat),
    lon = first(lon),
    region = first(region),
    subregion = first(subregion)
  ) %>%
  st_cast("MULTIPOLYGON") %>% 
  bind_rows(
    read_sf(here::here("data/TM_WORLD_BORDERS_SIMPL-0.3.shp")) %>%
      clean_names() %>%
      filter(!(name %in% c("Western Sahara", "Morocco")))
  )

labels <- function(x) {
  if_else(x < 10000, as.character(x), "10000+")
}

shapefile %>% 
  clean_names() %>%
  rename(country = iso2) %>%
  left_join(df_world, by = "country") %>%
  filter(total > 0 | area > 1000) %>% 
  filter(country != "Antarctica") %>% 
  mutate(
    total = replace_na(total, 0),
    total = pmin(total, 10000)
  ) %>%
  ggplot() +
  geom_sf(aes(fill = total), lwd = 0, color = "gray", show.legend = T) +
  scale_fill_gradientn(
    #breaks = seq(0, 500, 100),
    labels = labels,
    colors = RColorBrewer::brewer.pal(n = 9, name = "Blues")
  ) +
  guides(
    fill =
      guide_colorbar(
        barheight = 0.5,
        barwidth = 15,
        title = "CS Publications",
        title.vjust = 1,
      )
  ) +
  theme_void() +
  theme(legend.position = "bottom")
```


## Figure 4

By language (english vs non-english)

```{r}
df %>%
  filter(year >= 2000 & year <= 2022) %>% 
  drop_na(language) %>% 
  mutate(non_english = if_else(language == "en", "English", "Non-English")) %>%
  group_by(year, non_english) %>%
  summarise(perc_invisible = mean(invisible, na.rm = T)) %>% 
  ungroup() %>% 
  ggplot(aes(x = year, y = perc_invisible, color = non_english)) +
  geom_smooth(method = "loess", span = 1.5) +
  scale_color_brewer(palette = "Set1", direction = 1) +
  scale_x_continuous(breaks = scales::breaks_width(2), minor_breaks = NULL) +
  scale_y_continuous(breaks = scales::breaks_width(0.05), minor_breaks = NULL, labels = scales::percent_format(accuracy = 1)) +
  theme_minimal() +
  labs(x = "Year", y = "Proportion of invisible publications", color = "")
```

Reducing gaps in visbility rate between English and Non-English publications over time:

```{r}
df %>%
  filter(year >= 2000 & year <= 2022) %>% 
  drop_na(language) %>% 
  mutate(non_english = if_else(language == "en", "English", "Non-English")) %>%
  group_by(year, non_english) %>%
  summarise(perc_invisible = mean(invisible)) %>%
  ungroup() %>%
  pivot_wider(names_from = non_english, values_from = perc_invisible) %>%
  mutate(gap = English - `Non-English`) %>%
  ggplot(aes(x = year, y = gap)) +
  geom_smooth(method = "loess", span = 1) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  scale_x_continuous(breaks = scales::breaks_width(2), minor_breaks = NULL) +
  scale_y_continuous(breaks = scales::breaks_width(0.05), minor_breaks = NULL, labels = scales::percent_format(accuracy = 1)) +
  theme_minimal() +
  labs(x = "Year", y = "Gap in invisibility (English - Non-English)")
```

Reduction might be because of higher volume of English publications, whereas similar volume of non-English publications:

```{r}
df %>% 
  filter(year >= 2000 & year <= 2022) %>%
  drop_na(language) %>% 
  mutate(non_english = if_else(language == "en", "English", "Non-English")) %>%
  count(year, non_english) %>% 
  ggplot(aes(x = year, y = n, color = non_english)) +
  geom_line() +
  scale_color_brewer(palette = "Set1", direction = 1) +
  scale_x_continuous(breaks = scales::breaks_width(2), minor_breaks = NULL) +
  scale_y_continuous(breaks = scales::breaks_width(4000), minor_breaks = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = "Year", y = "Number of publications", color = "")
```


## Figure 5

By mainstream vs non-mainstream topic

```{r}
# mainstream_concepts as top 20%

mainstream_concepts <- 
  df %>% 
  select(concepts, id) %>%
  mutate(concepts = str_remove_all(concepts, "[\\[\\]']")) %>%
  separate_rows(concepts, sep = ", ") %>%
  drop_na(concepts) %>%
  group_by(id) %>% 
  slice_head(n = 1) %>% 
  ungroup() %>%
  count(concepts, sort = TRUE) %>%
  slice_head(prop = 0.2) %>%
  pull(concepts)

df %>%
  filter(year >= 2000 & year <= 2022) %>%
  select(concepts, id, year, invisible) %>%
  mutate(concepts = str_remove_all(concepts, "[\\[\\]']")) %>%
  separate_rows(concepts, sep = ", ") %>%
  drop_na(concepts) %>%
  group_by(id, year, invisible) %>% 
  slice_head(n = 1) %>% 
  ungroup() %>% 
  mutate(mainstream = if_else(concepts %in% mainstream_concepts, "Mainstream", "Non-Mainstream")) %>%
  group_by(year, mainstream) %>%
  summarise(perc_invisible = mean(invisible, na.rm = T)) %>% 
  ungroup() %>% 
  ggplot(aes(x = year, y = perc_invisible, color = mainstream)) +
  geom_smooth(method = "loess", span = 1.5) +
  scale_color_brewer(palette = "Set1", direction = 1) +
  scale_x_continuous(breaks = scales::breaks_width(2), minor_breaks = NULL) +
  scale_y_continuous(breaks = scales::breaks_width(0.05), minor_breaks = NULL, labels = scales::percent_format(accuracy = 1)) +
  theme_minimal() +
  labs(x = "Year", y = "Proportion of invisible publications", color = "")
```

```{r}
temp1 <-
  df %>%
  filter(year >= 2000 & year <= 2022, invisible == 0) %>%
  select(concepts, id) %>%
  mutate(concepts = str_remove_all(concepts, "[\\[\\]']")) %>%
  separate_rows(concepts, sep = ", ") %>%
  drop_na(concepts) %>%
  group_by(id) %>% 
  slice_head(n = 1) %>% 
  ungroup() %>%
  count(concepts, sort = TRUE, name = "visible_n")

temp2 <-
  df %>%
  filter(year >= 2000 & year <= 2022, invisible == 1) %>%
  select(concepts, id) %>%
  mutate(concepts = str_remove_all(concepts, "[\\[\\]']")) %>%
  separate_rows(concepts, sep = ", ") %>%
  drop_na(concepts) %>%
  group_by(id) %>% 
  slice_head(n = 1) %>% 
  ungroup() %>%
  count(concepts, sort = TRUE, name = "invisible_n")

anti_join(temp2, temp1)
```

```{r}
anti_join(temp1, temp2)
```

```{r}
df %>%
  filter(year >= 2000 & year <= 2022) %>%
  sample_n(50) %>% 
  select(contains("category_for_2020")) %>% 
  view()
```


```{r}
temp <-
  df %>% 
  filter(year >= 2000 & year <= 2022, invisible == 1) %>%
  transmute(id, category_for = str_replace_all(category_for, "'", "\"")) %>% 
  mutate(category_for = map(category_for, ~ fromJSON(.) %>% as_tibble())) %>%
  unnest(id, category_for) %>% 
  mutate(
    category_code = str_extract(name, "^\\d+") %>% as.integer(),
    category_name = str_remove(name, "^\\d+\\s+")
  ) %>% 
  filter(category_code != 4701, category_code >= 100) %>% 
  count(category_name, sort = TRUE) %>%
  transmute(category_name, perc_invisible = n/sum(n) %>% replace_na(0)) %>%
  full_join(
    df %>% 
      filter(year >= 2000 & year <= 2022, invisible == 0) %>%
      transmute(id, category_for = str_replace_all(category_for, "'", "\"")) %>% 
      mutate(category_for = map(category_for, ~ fromJSON(.) %>% as_tibble())) %>%
      unnest(id, category_for) %>% 
      mutate(
        category_code = str_extract(name, "^\\d+") %>% as.integer(),
        category_name = str_remove(name, "^\\d+\\s+")
      ) %>% 
      filter(category_code != 4701, category_code >= 100) %>% 
      count(category_name, sort = TRUE) %>%
      transmute(category_name, perc_visible = n/sum(n) %>% replace_na(0))
  )

temp %>%
  filter(!(category_name %in% c("Screen and Digital Media", "Cultural Studies"))) %>%
  mutate_if(is.numeric, ~ round(. * 100, 2)) %>%
  mutate(diff = perc_invisible - perc_visible) %>% 
  arrange(diff)
```

```{r}
df %>%
  filter(year >= 2000 & year <= 2022) %>%
  transmute(invisible, category_for = str_replace_all(category_for, "'", "\"")) %>% 
  mutate(category_for = map(category_for, ~ fromJSON(.) %>% as_tibble())) %>%
  unnest(invisible, category_for) %>% 
  mutate(
    category_code = str_extract(name, "^\\d+") %>% as.integer(),
    category_name = str_remove(name, "^\\d+\\s+")
  ) %>% 
  filter(category_code != 4701, category_code >= 100) %>% 
  transmute(invisible, category_name) %>%
  mutate(category_name = factor(category_name))
```


## Figure 6

By institutional prestige


