---
title: "PKP/OJS multi-source journal enrichment — simple version"
format:
  html:
    toc: true
execute:
  echo: true
  warning: true
  message: false
---

This one-off notebook independently rebuilds the complete PKP/OJS enrichment.
It keeps the original exact-ISSN rule, retrieves OpenAlex and Crossref data when
the reusable caches are missing, expands every observed top-level source field,
and writes the 98,273-row result plus a fixed ten-row review sample as Parquet.
Nested values are retained as JSON strings. Rows with multiple candidates keep
all complete candidate objects in a JSON column instead of selecting one.

The original enrichment caches are optional read-only fallbacks. Any downloaded
input or newly retrieved source data are written under this notebook's own
artifact directory. To retrieve a fresh snapshot, set `reuse_shared_cache` to
`FALSE` and delete this notebook's `source-cache`. Missing OpenAlex batches
require `OPENALEX_API_KEY`; missing Crossref pages require `CROSSREF_MAILTO`.

## Packages and paths

```{r}
#| label: packages-and-paths

library(dplyr)
library(httr2)
library(purrr)
library(readr)
library(tibble)

stopifnot(requireNamespace("jsonlite", quietly = TRUE))
stopifnot(requireNamespace("knitr", quietly = TRUE))
stopifnot(requireNamespace("nanoparquet", quietly = TRUE))

document_path <- knitr::current_input(dir = TRUE)
document_dir <- if (nzchar(document_path)) {
  dirname(normalizePath(document_path))
} else {
  normalizePath(getwd())
}
domain_dir <- normalizePath(file.path(document_dir, "..", ".."))
project_env_path <- file.path(domain_dir, "..", "..", ".env")
if (file.exists(project_env_path)) readRenviron(project_env_path)

artifact_dir <- file.path(
  domain_dir,
  "artifacts",
  "ojs_journal_enrichment_simple_ver"
)
shared_artifact_dir <- file.path(
  domain_dir,
  "artifacts",
  "ojs_journal_enrichment"
)
reuse_shared_cache <- TRUE
cache_dir <- file.path(artifact_dir, "source-cache")
shared_cache_dir <- file.path(shared_artifact_dir, "full-v7")
input_dir <- file.path(artifact_dir, "input")
input_path <- file.path(input_dir, "beacon.csv")
shared_input_path <- file.path(shared_artifact_dir, "pkp-v7", "beacon.csv")
openalex_batch_dir <- file.path(cache_dir, "openalex-batches")
shared_openalex_batch_dir <- file.path(shared_cache_dir, "openalex-batches")
crossref_cache_path <- file.path(cache_dir, "crossref-journals.rds")
shared_crossref_cache_path <- file.path(
  shared_cache_dir,
  "crossref-journals.rds"
)
crossref_page_dir <- file.path(cache_dir, "crossref-pages")
shared_crossref_page_dir <- file.path(shared_cache_dir, "crossref-pages")
output_path <- file.path(
  artifact_dir,
  "pkp-ojs-multisource-enriched-simple_ver.parquet"
)
sample_output_path <- file.path(
  artifact_dir,
  "pkp-ojs-multisource-enriched-simple_ver-sample.parquet"
)
openalex_expanded_path <- file.path(artifact_dir, "openalex-expanded.rds")
crossref_expanded_path <- file.path(artifact_dir, "crossref-expanded.rds")

dir.create(input_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(artifact_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(openalex_batch_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(crossref_page_dir, recursive = TRUE, showWarnings = FALSE)

pkp_url <- paste0(
  "https://dataverse.harvard.edu/api/access/datafile/",
  "14084919?format=original"
)
pkp_md5 <- "3a4ad8ae1ebfcc2b991aaf55b2d82c92"

if (!file.exists(input_path)) {
  if (
    file.exists(shared_input_path) &&
      unname(tools::md5sum(shared_input_path)) == pkp_md5
  ) {
    stopifnot(file.copy(shared_input_path, input_path))
  } else {
    temporary_input <- paste0(input_path, ".download")
    download.file(pkp_url, temporary_input, mode = "wb")
    stopifnot(unname(tools::md5sum(temporary_input)) == pkp_md5)
    stopifnot(file.rename(temporary_input, input_path))
  }
}
stopifnot(unname(tools::md5sum(input_path)) == pkp_md5)
```

## ISSN preparation

```{r}
#| label: issn-preparation

normalize_issn <- function(value) {
  if (length(value) != 1L || is.na(value)) return(NA_character_)
  compact <- gsub("[^0-9X]", "", toupper(trimws(as.character(value))))
  if (!grepl("^[0-9]{7}[0-9X]$", compact)) return(NA_character_)

  digits <- as.integer(strsplit(substr(compact, 1, 7), "")[[1]])
  check_value <- (11 - sum(digits * 8:2) %% 11) %% 11
  expected <- if (check_value == 10) "X" else as.character(check_value)
  if (substr(compact, 8, 8) != expected) return(NA_character_)

  paste0(substr(compact, 1, 4), "-", substr(compact, 5, 8))
}

normalize_issn_set <- function(values) {
  normalized <- vapply(values, normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}

is_missing_value <- function(value) {
  is.na(value) |
    !nzchar(trimws(value)) |
    toupper(trimws(value)) == "NA"
}

parse_issns <- function(value) {
  if (is_missing_value(value)) return(character())
  tokens <- unlist(strsplit(value, "[[:space:],;|]+"))
  normalize_issn_set(tokens[nzchar(tokens)])
}

openalex_issns <- function(source) {
  normalize_issn_set(unlist(source$issn, use.names = FALSE))
}

crossref_issns <- function(journal) {
  normalize_issn_set(journal$ISSN)
}
```

## Source retrieval and caching

The cache code is intentionally small: completed OpenAlex batches and Crossref
pages are reused; missing data are fetched with bounded retries and timeouts.

```{r}
#| label: retrieval-helpers

retryable_response <- function(response) {
  resp_status(response) %in% c(403L, 408L, 429L, 500L, 502L, 503L, 504L)
}

save_cache <- function(value, path) {
  temporary <- paste0(path, ".tmp")
  saveRDS(value, temporary)
  stopifnot(file.rename(temporary, path))
}

read_cache <- function(path) {
  if (!file.exists(path)) return(NULL)
  tryCatch(readRDS(path), error = \(error) NULL)
}

fetch_openalex_page <- function(issns, cursor) {
  api_key <- Sys.getenv("OPENALEX_API_KEY")
  if (!nzchar(api_key)) {
    stop("Set OPENALEX_API_KEY to retrieve missing OpenAlex batches.")
  }

  body <- request("https://api.openalex.org/sources") |>
    req_url_query(
      filter = paste0("issn:", paste(issns, collapse = "|")),
      per_page = 200,
      cursor = cursor,
      api_key = api_key
    ) |>
    req_timeout(seconds = 30) |>
    req_retry(
      max_tries = 4L,
      retry_on_failure = TRUE,
      is_transient = retryable_response
    ) |>
    req_perform() |>
    resp_body_json(simplifyVector = FALSE)

  stopifnot(
    is.list(body$results),
    is.list(body$meta),
    length(body$meta$count) == 1L,
    is.null(body$meta$next_cursor) ||
      length(body$meta$next_cursor) == 1L
  )
  body
}

fetch_openalex_batch <- function(issns) {
  stopifnot(length(issns) <= 100L)
  cursor <- "*"
  seen_cursors <- character()
  sources <- list()
  expected_count <- NULL

  repeat {
    if (cursor %in% seen_cursors) stop("OpenAlex repeated a cursor.")
    seen_cursors <- c(seen_cursors, cursor)
    page <- fetch_openalex_page(issns, cursor)
    if (is.null(expected_count)) expected_count <- as.integer(page$meta$count)
    sources <- c(sources, page$results)
    cursor <- page$meta$next_cursor
    if (is.null(cursor) || !nzchar(cursor)) break
  }

  stopifnot(length(sources) == expected_count)
  sources
}

fetch_crossref_page <- function(cursor) {
  mailto <- Sys.getenv("CROSSREF_MAILTO")
  if (!grepl("^[^@[:space:]]+@[^@[:space:]]+$", mailto)) {
    stop("Set CROSSREF_MAILTO to retrieve missing Crossref pages.")
  }

  body <- request("https://api.crossref.org/journals") |>
    req_url_query(rows = 1000, cursor = cursor, mailto = mailto) |>
    req_user_agent(
      paste(
        "InvisibleResearch journal enrichment;",
        "https://github.com/YannJY02/InvisibleResearch"
      )
    ) |>
    req_timeout(seconds = 60) |>
    req_retry(
      max_tries = 4L,
      retry_on_failure = TRUE,
      is_transient = retryable_response
    ) |>
    req_perform() |>
    resp_body_json(simplifyVector = FALSE) |>
    pluck("message")

  stopifnot(
    is.list(body$items),
    length(body[["total-results"]]) == 1L
  )
  body
}

fetch_crossref_directory <- function() {
  cursor <- "*"
  journals <- list()
  expected_count <- NULL
  previous_items <- NULL
  page_number <- 1L

  repeat {
    page_path <- file.path(
      crossref_page_dir,
      sprintf("page-%04d.rds", page_number)
    )
    page <- read_cache(page_path)
    reusable <- is.list(page) &&
      identical(page$cursor, cursor) &&
      is.list(page$items) &&
      length(page$total_results) == 1L

    if (!reusable) {
      shared_page <- if (reuse_shared_cache) {
        read_cache(file.path(shared_crossref_page_dir, basename(page_path)))
      } else {
        NULL
      }
      reusable <- is.list(shared_page) &&
        identical(shared_page$cursor, cursor) &&
        is.list(shared_page$items) &&
        length(shared_page$total_results) == 1L
      if (reusable) {
        page <- shared_page
      } else {
        response <- fetch_crossref_page(cursor)
        page <- list(
          cursor = cursor,
          total_results = as.integer(response[["total-results"]]),
          next_cursor = response[["next-cursor"]],
          items = response$items
        )
        save_cache(page, page_path)
        Sys.sleep(0.2)
      }
    }

    if (!is.null(previous_items) && identical(page$items, previous_items)) {
      stop("Crossref repeated a page.")
    }
    if (is.null(expected_count)) {
      expected_count <- page$total_results
    } else {
      stopifnot(identical(expected_count, page$total_results))
    }

    journals <- c(journals, page$items)
    stopifnot(length(journals) <= expected_count)
    if (length(page$items) < 1000L) break

    cursor <- page$next_cursor
    if (is.null(cursor) || !nzchar(cursor)) {
      stop("Crossref returned a full page without a next cursor.")
    }
    previous_items <- page$items
    page_number <- page_number + 1L
  }

  stopifnot(length(journals) == expected_count)
  list(journals = journals, total_results = expected_count)
}
```

## Load the complete PKP V7 cohort

```{r}
#| label: load-pkp

pkp_rows <- read_csv(
  input_path,
  col_types = cols(.default = col_character()),
  na = character(),
  show_col_types = FALSE
)

expected_schema <- c(
  "context_name", "issn", "oai_url", "application", "country", "region",
  "total_record_count", "record_count_2025", "earliest_datestamp",
  "repository_name", "set_spec", "stats_id", "best_doaj_url",
  "last_completed_update", "first_beacon", "last_beacon",
  "last_oai_response", "unresponsive_endpoint", "unresponsive_context"
)
row_identity <- function(data) {
  paste(data$oai_url, data$repository_name, data$set_spec, sep = "\r")
}

stopifnot(
  nrow(pkp_rows) == 98273L,
  identical(names(pkp_rows), expected_schema),
  all(tolower(pkp_rows$application) == "ojs"),
  !anyDuplicated(row_identity(pkp_rows))
)

input_issns <- map(pkp_rows$issn, parse_issns)
tokens <- sort(unique(unlist(input_issns)))
stopifnot(
  sum(is_missing_value(pkp_rows$issn)) == 26189L,
  length(tokens) == 103017L
)
```

## Matching and field expansion helpers

```{r}
#| label: matching-helpers

deduplicate_candidates <- function(candidates, identity) {
  ids <- vapply(candidates, identity, character(1))
  keep <- !duplicated(ids)
  list(candidates = candidates[keep], ids = ids[keep])
}

index_candidates <- function(candidate_issns) {
  index <- new.env(hash = TRUE, parent = emptyenv())
  for (position in seq_along(candidate_issns)) {
    for (issn in candidate_issns[[position]]) {
      existing <- get0(
        issn,
        index,
        inherits = FALSE,
        ifnotfound = integer()
      )
      assign(issn, c(existing, position), index)
    }
  }
  index
}

match_candidates <- function(issns, index, candidate_issns) {
  positions <- unique(unlist(map(issns, \(issn) {
    get0(issn, index, inherits = FALSE, ifnotfound = integer())
  })))
  matched_issns <- base::intersect(
    issns,
    unique(unlist(candidate_issns[positions]))
  )
  status <- case_when(
    length(issns) == 0L ~ "not_attempted",
    length(positions) == 0L ~ "unmatched",
    length(positions) == 1L ~ "unique",
    TRUE ~ "ambiguous"
  )
  list(
    status = status,
    matched_issns = matched_issns,
    positions = positions
  )
}

serialize_json <- function(value) {
  as.character(jsonlite::toJSON(
    value,
    auto_unbox = TRUE,
    null = "null",
    na = "null",
    digits = NA
  ))
}

dataframe_value <- function(value) {
  if (is.null(value)) return(NA_character_)
  if (is.atomic(value) && length(value) == 1L) {
    return(as.character(value))
  }
  serialize_json(value)
}

scalar_column <- function(values) {
  present <- keep(values, \(value) !is.null(value) && length(value) == 1L)
  types <- unique(map_chr(present, typeof))
  missing <- \(value) is.null(value) || length(value) == 0L

  if (length(types) > 0L && all(types == "logical")) {
    return(map_lgl(values, \(value) if (missing(value)) NA else value))
  }
  if (length(types) > 0L && all(types == "integer")) {
    return(map_int(values, \(value) if (missing(value)) NA_integer_ else value))
  }
  if (length(types) > 0L && all(types %in% c("integer", "double"))) {
    return(map_dbl(values, \(value) if (missing(value)) NA_real_ else value))
  }
  map_chr(values, \(value) {
    if (missing(value)) NA_character_ else as.character(value)
  })
}

expand_source_fields <- function(matches, candidates, prefix) {
  fields <- sort(unique(unlist(map(candidates, names))))
  single_positions <- map_int(matches, \(match) {
    if (length(match$positions) == 1L) match$positions else NA_integer_
  })

  # ponytail: JSON conversion is single-process for this one-off; use Arrow
  # nested columns only if repeated full rebuilds make serialization a problem.
  columns <- set_names(
    map(fields, \(field) {
      values <- map(candidates, \(candidate) candidate[[field]])
      scalar_field <- all(map_lgl(values, \(value) {
        is.null(value) || (is.atomic(value) && length(value) <= 1L)
      }))
      candidate_values <- if (scalar_field) {
        scalar_column(values)
      } else {
        map_chr(values, dataframe_value)
      }
      candidate_values[single_positions]
    }),
    paste(prefix, fields, sep = "_")
  )
  columns[[paste0(prefix, "_candidates__json")]] <- map_chr(
    matches,
    \(match) {
      if (length(match$positions) > 1L) {
        serialize_json(candidates[match$positions])
      } else {
        NA_character_
      }
    }
  )
  as_tibble(columns, .name_repair = "check_unique")
}
```

## OpenAlex enrichment

```{r}
#| label: openalex-enrichment

token_batches <- split(tokens, ceiling(seq_along(tokens) / 100L))
openalex_batch_results <- imap(token_batches, \(batch, batch_number) {
  checkpoint_path <- file.path(
    openalex_batch_dir,
    sprintf("batch-%04d.rds", as.integer(batch_number))
  )
  checkpoint <- read_cache(checkpoint_path)
  reusable <- is.list(checkpoint) &&
    identical(checkpoint$issns, batch) &&
    is.list(checkpoint$sources)

  if (!reusable && reuse_shared_cache) {
    shared_checkpoint_path <- file.path(
      shared_openalex_batch_dir,
      basename(checkpoint_path)
    )
    checkpoint <- read_cache(shared_checkpoint_path)
    reusable <- is.list(checkpoint) &&
      identical(checkpoint$issns, batch) &&
      is.list(checkpoint$sources)
    if (reusable) checkpoint_path <- shared_checkpoint_path
  }

  if (!reusable) {
    checkpoint <- list(
      issns = batch,
      sources = fetch_openalex_batch(batch)
    )
    save_cache(checkpoint, checkpoint_path)
  }
  list(sources = checkpoint$sources, path = checkpoint_path)
})

openalex_batches <- map(openalex_batch_results, "sources")
openalex_source_key <- paste(
  unname(tools::md5sum(map_chr(openalex_batch_results, "path"))),
  collapse = ""
)
openalex_sources <- unlist(openalex_batches, recursive = FALSE)
openalex_resolved <- deduplicate_candidates(
  openalex_sources,
  \(source) source$id
)
openalex_sources <- openalex_resolved$candidates
openalex_ids <- openalex_resolved$ids
openalex_issns_by_candidate <- map(openalex_sources, openalex_issns)
relevant <- map_lgl(
  openalex_issns_by_candidate,
  \(issns) any(issns %in% tokens)
)
openalex_sources <- openalex_sources[relevant]
openalex_ids <- openalex_ids[relevant]
openalex_issns_by_candidate <- openalex_issns_by_candidate[relevant]
openalex_index <- index_candidates(openalex_issns_by_candidate)
openalex_matches <- map(
  input_issns,
  match_candidates,
  index = openalex_index,
  candidate_issns = openalex_issns_by_candidate
)

base_result <- pkp_rows |>
  mutate(
    identifier_status = case_when(
      is_missing_value(issn) ~ "missing",
      lengths(input_issns) == 0L ~ "invalid",
      TRUE ~ "valid"
    ),
    openalex_input_issns = map_chr(input_issns, paste, collapse = "|"),
    openalex_matched_issns = map_chr(
      openalex_matches,
      \(match) paste(match$matched_issns, collapse = "|")
    ),
    openalex_match_status = map_chr(openalex_matches, "status"),
    openalex_candidate_count = map_int(
      openalex_matches,
      \(match) length(match$positions)
    ),
    openalex_candidate_ids = map_chr(
      openalex_matches,
      \(match) paste(openalex_ids[match$positions], collapse = "|")
    )
  )

openalex_expected_columns <- c(
  paste("openalex", sort(unique(unlist(map(openalex_sources, names)))), sep = "_"),
  "openalex_candidates__json"
)
openalex_expanded <- read_cache(openalex_expanded_path)
if (
  !is.list(openalex_expanded) ||
    !identical(openalex_expanded$source_key, openalex_source_key) ||
    !is.data.frame(openalex_expanded$data) ||
    nrow(openalex_expanded$data) != nrow(pkp_rows) ||
    !identical(names(openalex_expanded$data), openalex_expected_columns)
) {
  openalex_expanded <- list(
    source_key = openalex_source_key,
    data = expand_source_fields(openalex_matches, openalex_sources, "openalex")
  )
  save_cache(openalex_expanded, openalex_expanded_path)
}
openalex_data <- openalex_expanded$data
rm(openalex_batches, openalex_batch_results, openalex_resolved, openalex_index)
gc()
```

## Crossref enrichment

```{r}
#| label: crossref-enrichment

crossref_cache <- read_cache(crossref_cache_path)
cache_is_complete <- is.list(crossref_cache) &&
  is.list(crossref_cache$journals) &&
  identical(
    as.integer(crossref_cache$total_results),
    length(crossref_cache$journals)
  )

crossref_cache_source_path <- crossref_cache_path
if (!cache_is_complete && reuse_shared_cache) {
  crossref_cache <- read_cache(shared_crossref_cache_path)
  cache_is_complete <- is.list(crossref_cache) &&
    is.list(crossref_cache$journals) &&
    identical(
      as.integer(crossref_cache$total_results),
      length(crossref_cache$journals)
    )
  if (cache_is_complete) {
    crossref_cache_source_path <- shared_crossref_cache_path
  }
}
if (!cache_is_complete) {
  crossref_cache <- fetch_crossref_directory()
  save_cache(crossref_cache, crossref_cache_path)
  crossref_cache_source_path <- crossref_cache_path
}
crossref_source_key <- unname(tools::md5sum(crossref_cache_source_path))
crossref_journals <- crossref_cache$journals
crossref_issns_by_candidate <- map(crossref_journals, crossref_issns)
relevant <- map_lgl(
  crossref_issns_by_candidate,
  \(issns) any(issns %in% tokens)
)
crossref_journals <- crossref_journals[relevant]
crossref_ids <- map_chr(
  crossref_journals,
  \(journal) paste(sort(crossref_issns(journal)), collapse = "|")
)
keep <- nzchar(crossref_ids)
crossref_journals <- crossref_journals[keep]
crossref_ids <- crossref_ids[keep]
crossref_issns_by_candidate <- map(crossref_journals, crossref_issns)
crossref_index <- index_candidates(crossref_issns_by_candidate)
crossref_matches <- map(
  input_issns,
  match_candidates,
  index = crossref_index,
  candidate_issns = crossref_issns_by_candidate
)

crossref_expected_columns <- c(
  paste("crossref", sort(unique(unlist(map(crossref_journals, names)))), sep = "_"),
  "crossref_candidates__json"
)
crossref_expanded <- read_cache(crossref_expanded_path)
if (
  !is.list(crossref_expanded) ||
    !identical(crossref_expanded$source_key, crossref_source_key) ||
    !is.data.frame(crossref_expanded$data) ||
    nrow(crossref_expanded$data) != nrow(pkp_rows) ||
    !identical(names(crossref_expanded$data), crossref_expected_columns)
) {
  crossref_expanded <- list(
    source_key = crossref_source_key,
    data = expand_source_fields(crossref_matches, crossref_journals, "crossref")
  )
  save_cache(crossref_expanded, crossref_expanded_path)
}
crossref_data <- crossref_expanded$data

result <- base_result |>
  mutate(
    crossref_input_issns = map_chr(input_issns, paste, collapse = "|"),
    crossref_matched_issns = map_chr(
      crossref_matches,
      \(match) paste(match$matched_issns, collapse = "|")
    ),
    crossref_match_status = map_chr(crossref_matches, "status"),
    crossref_candidate_count = map_int(
      crossref_matches,
      \(match) length(match$positions)
    ),
    crossref_candidate_issn_sets = map_chr(
      crossref_matches,
      \(match) paste(crossref_ids[match$positions], collapse = ";")
    )
  ) |>
  bind_cols(openalex_data, crossref_data)
```

## Parquet outputs and lightweight validation

```{r}
#| label: parquet-output

sample_spec <- tribble(
  ~context_name, ~oai_url, ~repository_name, ~set_spec,
  "Kuwait Journal of Science",
  "https://journalskuwait.org/kjs/index.php/index/oai",
  "Kuwait Journal of Science", "KJS",
  "Cakrawala",
  "https://cakrawalajournal.org/index.php/index/oai",
  "CAKRAWALA", "cakrawala",
  "Sintesa: Jurnal Ilmu Pendidikan",
  "https://sintesa.stkip-arrahmaniyah.ac.id/index.php/index/oai",
  "Sintesa: Jurnal Ilmu Pendidikan STKIP Arrahmaniyah Depok", "sintesa",
  "CARAKA: Jurnal Teologi Biblika dan Praktika",
  "https://ojs.sttibc.ac.id/index.php/index/oai",
  "CARAKA:Jurnal Teologi", "ibc",
  "Jurnal Tata Kelola dan Akuntabilitas Keuangan Negara",
  "https://jurnal.bpk.go.id/index.php/index/oai",
  "Open Journal Systems", "TAKEN",
  "JOEEL (Journal of English Education and Literature)",
  "https://journal.stkippamanetalino.ac.id/index.php/index/oai",
  "Rumah Jurnal STKIP Pamane Talino Ngabang", "bahasa-inggris",
  "Malaysian Journal of Paediatrics and Child Health",
  "https://mpaeds.my/journals/index.php/index/oai",
  "NA", "MJPCH",
  "Jami Scientific Research Quarterly Journal",
  "https://journals.jami.edu.af/index.php/index/oai",
  "Jami University Press", "jsrqj",
  "Journal of Polymer & Composites",
  "https://engineeringjournals.stmjournals.in/index.php/index/oai",
  "Engineering Journals", "JoPC",
  "JURNAL TEKNIK PERTAMBANGAN",
  "https://e-journal.upr.ac.id/index.php/index/oai",
  "Journal Online Universitas Palangka Raya", "JTP"
)
sample_positions <- match(row_identity(sample_spec), row_identity(result))
stopifnot(!anyNA(sample_positions))
sample_result <- result[sample_positions, ]

candidate_json_is_complete <- function(matches, values) {
  map2_lgl(matches, values, \(match, value) {
    candidate_count <- length(match$positions)
    if (candidate_count <= 1L) return(is.na(value))
    parsed <- tryCatch(
      jsonlite::fromJSON(value, simplifyVector = FALSE),
      error = \(error) NULL
    )
    is.list(parsed) && length(parsed) == candidate_count
  })
}

stopifnot(
  nrow(result) == 98273L,
  !"admin_email" %in% names(result),
  identical(row_identity(result), row_identity(pkp_rows)),
  !anyDuplicated(row_identity(result)),
  nrow(sample_result) == 10L,
  all(c(
    "openalex_candidates__json",
    "crossref_candidates__json"
  ) %in% names(result)),
  all(candidate_json_is_complete(
    openalex_matches,
    result$openalex_candidates__json
  )),
  all(candidate_json_is_complete(
    crossref_matches,
    result$crossref_candidates__json
  ))
)

python_candidates <- unname(Sys.which(c("python", "python3")))
python_candidates <- python_candidates[nzchar(python_candidates)]
has_pyarrow <- map_lgl(python_candidates, \(candidate) {
  suppressWarnings(system2(
    candidate,
    c("-c", shQuote("import pyarrow")),
    stdout = FALSE,
    stderr = FALSE
  )) == 0L
})
python <- python_candidates[has_pyarrow][1]
if (is.na(python)) stop("Python with PyArrow is required to write Parquet.")
converter_path <- file.path(artifact_dir, "csv_to_parquet.py")
writeLines(c(
  "import csv",
  "import json",
  "import sys",
  "from pathlib import Path",
  "import pyarrow as pa",
  "import pyarrow.csv as pacsv",
  "import pyarrow.parquet as pq",
  "input_path, schema_path, output_path = map(Path, sys.argv[1:4])",
  "temporary_path = output_path.with_suffix(output_path.suffix + '.tmp')",
  "with open(input_path, newline='', encoding='utf-8') as handle:",
  "    columns = next(csv.reader(handle))",
  "r_types = json.loads(schema_path.read_text(encoding='utf-8'))",
  paste0(
    "arrow_types = {'character': pa.string(), 'logical': pa.bool_(), ",
    "'integer': pa.int32(), 'double': pa.float64()}"
  ),
  "types = {column: arrow_types[r_types[column]] for column in columns}",
  "reader = pacsv.open_csv(",
  "    input_path,",
  "    read_options=pacsv.ReadOptions(block_size=64 * 1024 * 1024),",
  "    convert_options=pacsv.ConvertOptions(",
  paste0(
    "        column_types=types, strings_can_be_null=True, ",
    "null_values=['__OJS_SIMPLE_NULL__']"
  ),
  "    ),",
  ")",
  "temporary_path.unlink(missing_ok=True)",
  "writer = pq.ParquetWriter(temporary_path, reader.schema, compression='zstd')",
  "try:",
  "    for batch in reader:",
  "        writer.write_batch(batch)",
  "finally:",
  "    writer.close()",
  "temporary_path.replace(output_path)"
), converter_path)

write_parquet <- function(data, path) {
  staging_path <- paste0(path, ".csv")
  schema_path <- paste0(path, ".schema.json")
  on.exit(unlink(c(staging_path, schema_path)), add = TRUE)
  write_csv(data, staging_path, na = "__OJS_SIMPLE_NULL__")
  column_types <- map_chr(data, typeof)
  stopifnot(all(column_types %in% c(
    "character", "logical", "integer", "double"
  )))
  jsonlite::write_json(as.list(column_types), schema_path, auto_unbox = TRUE)
  output <- system2(
    python,
    c(converter_path, staging_path, schema_path, path),
    stdout = TRUE,
    stderr = TRUE
  )
  if (!is.null(attr(output, "status")) && attr(output, "status") != 0L) {
    stop(paste(output, collapse = "\n"))
  }
}

pending_output_path <- paste0(output_path, ".pending")
pending_sample_output_path <- paste0(sample_output_path, ".pending")
write_parquet(result, pending_output_path)
write_parquet(sample_result, pending_sample_output_path)
unlink(converter_path)

parquet_info <- nanoparquet::read_parquet_info(pending_output_path)
parquet_schema <- nanoparquet::read_parquet_schema(pending_output_path)
observed <- nanoparquet::read_parquet(pending_output_path)
observed_sample <- nanoparquet::read_parquet(pending_sample_output_path)

stopifnot(
  parquet_info$num_rows == nrow(result),
  parquet_info$num_cols == ncol(result),
  identical(
    parquet_schema$name[!is.na(parquet_schema$r_col)],
    names(result)
  ),
  isTRUE(all.equal(
    as.data.frame(observed),
    as.data.frame(result),
    check.attributes = FALSE
  )),
  isTRUE(all.equal(
    as.data.frame(observed_sample),
    as.data.frame(sample_result),
    check.attributes = FALSE
  ))
)

stopifnot(
  file.rename(pending_output_path, output_path),
  file.rename(pending_sample_output_path, sample_output_path)
)
```

## Ten-row visual review

```{r}
#| label: visual-review

sample_result |>
  mutate(issn = gsub("[\\r\\n]+", " / ", issn, perl = TRUE)) |>
  select(
    context_name,
    issn,
    identifier_status,
    openalex_match_status,
    openalex_candidate_count,
    crossref_match_status,
    crossref_candidate_count,
    any_of(c(
      "openalex_display_name",
      "openalex_works_count",
      "crossref_title",
      "crossref_publisher"
    ))
  ) |>
  knitr::kable()

bind_rows(
  count(result, openalex_match_status, name = "rows") |>
    rename(status = openalex_match_status) |>
    mutate(source = "OpenAlex"),
  count(result, crossref_match_status, name = "rows") |>
    rename(status = crossref_match_status) |>
    mutate(source = "Crossref")
) |>
  select(source, status, rows) |>
  knitr::kable()

tibble(column = names(result)) |>
  knitr::kable()
```
