---
title: "PKP/OJS multi-source journal enrichment — simple version"
format:
  html:
    toc: true
execute:
  echo: true
  warning: true
  message: false
---

This one-off notebook independently rebuilds the complete PKP/OJS enrichment.
It keeps the original exact-ISSN rule, retrieves OpenAlex and Crossref data when
the reusable caches are missing, expands every observed top-level source field,
and writes the 98,273-row result plus a fixed ten-row review sample as Parquet.
Nested values are retained as JSON strings. Rows with multiple candidates keep
all complete candidate objects in a JSON column instead of selecting one.

The original enrichment caches are optional read-only fallbacks. Any downloaded
input or newly retrieved source data are written under this notebook's own
artifact directory. To retrieve a fresh snapshot, set `reuse_shared_cache` to
`FALSE` and delete this notebook's `source-cache`; the expanded-data caches are
then bypassed automatically. Missing OpenAlex batches require
`OPENALEX_API_KEY`; a missing Crossref snapshot requires
`CROSSREF_MAILTO`. Delete `openalex-expanded.rds` and `crossref-expanded.rds`
after changing the matching or field-expansion logic.

## Packages and paths

```{r}
#| label: packages-and-paths

library(dplyr)
library(httr2)
library(purrr)
library(readr)
library(tibble)

document_path <- knitr::current_input(dir = TRUE)
document_dir <- if (nzchar(document_path)) {
  dirname(normalizePath(document_path))
} else {
  normalizePath(getwd())
}
domain_dir <- normalizePath(file.path(document_dir, "..", ".."))
project_env_path <- file.path(domain_dir, "..", "..", ".env")
if (file.exists(project_env_path)) readRenviron(project_env_path)

artifact_dir <- file.path(
  domain_dir,
  "artifacts",
  "ojs_journal_enrichment_simple_ver"
)
shared_artifact_dir <- file.path(
  domain_dir,
  "artifacts",
  "ojs_journal_enrichment"
)
reuse_shared_cache <- TRUE
cache_dir <- file.path(artifact_dir, "source-cache")
shared_cache_dir <- file.path(shared_artifact_dir, "full-v7")
input_dir <- file.path(artifact_dir, "input")
input_path <- file.path(input_dir, "beacon.csv")
shared_input_path <- file.path(shared_artifact_dir, "pkp-v7", "beacon.csv")
openalex_batch_dir <- file.path(cache_dir, "openalex-batches")
shared_openalex_batch_dir <- file.path(shared_cache_dir, "openalex-batches")
crossref_cache_path <- file.path(cache_dir, "crossref-journals.rds")
shared_crossref_cache_path <- file.path(
  shared_cache_dir,
  "crossref-journals.rds"
)
output_path <- file.path(
  artifact_dir,
  "pkp-ojs-multisource-enriched-simple_ver.parquet"
)
sample_output_path <- file.path(
  artifact_dir,
  "pkp-ojs-multisource-enriched-simple_ver-sample.parquet"
)
openalex_expanded_path <- file.path(artifact_dir, "openalex-expanded.rds")
crossref_expanded_path <- file.path(artifact_dir, "crossref-expanded.rds")

dir.create(input_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(openalex_batch_dir, recursive = TRUE, showWarnings = FALSE)

pkp_url <- paste0(
  "https://dataverse.harvard.edu/api/access/datafile/",
  "14084919?format=original"
)
pkp_md5 <- "3a4ad8ae1ebfcc2b991aaf55b2d82c92"

valid_pkp_file <- function(path) {
  file.exists(path) && unname(tools::md5sum(path)) == pkp_md5
}
if (!valid_pkp_file(input_path)) {
  if (valid_pkp_file(shared_input_path)) {
    invisible(file.copy(shared_input_path, input_path, overwrite = TRUE))
  } else {
    download.file(pkp_url, input_path, mode = "wb")
  }
  stopifnot(valid_pkp_file(input_path))
}
```

## ISSN preparation

```{r}
#| label: issn-preparation

normalize_issn <- function(value) {
  if (length(value) != 1L || is.na(value)) return(NA_character_)
  compact <- gsub("[^0-9X]", "", toupper(trimws(as.character(value))))
  if (!grepl("^[0-9]{7}[0-9X]$", compact)) return(NA_character_)

  digits <- as.integer(strsplit(substr(compact, 1, 7), "")[[1]])
  check_value <- (11 - sum(digits * 8:2) %% 11) %% 11
  expected <- if (check_value == 10) "X" else as.character(check_value)
  if (substr(compact, 8, 8) != expected) return(NA_character_)

  paste0(substr(compact, 1, 4), "-", substr(compact, 5, 8))
}

normalize_issn_set <- function(values) {
  normalized <- vapply(values, normalize_issn, character(1))
  unique(normalized[!is.na(normalized)])
}

is_missing_value <- function(value) {
  is.na(value) |
    !nzchar(trimws(value)) |
    toupper(trimws(value)) == "NA"
}

parse_issns <- function(value) {
  if (is_missing_value(value)) return(character())
  tokens <- unlist(strsplit(value, "[[:space:],;|]+"))
  normalize_issn_set(tokens[nzchar(tokens)])
}

openalex_issns <- function(source) {
  normalize_issn_set(unlist(source$issn, use.names = FALSE))
}

crossref_issns <- function(journal) {
  normalize_issn_set(journal$ISSN)
}
```

## Source retrieval and caching

The cache code is intentionally small: completed OpenAlex batches and the full
Crossref directory are reused; missing data are fetched with bounded retries
and timeouts.

```{r}
#| label: retrieval-helpers

read_cache <- function(path) {
  if (!file.exists(path)) return(NULL)
  tryCatch(readRDS(path), error = \(error) NULL)
}

fetch_openalex_page <- function(issns, cursor) {
  api_key <- Sys.getenv("OPENALEX_API_KEY")
  if (!nzchar(api_key)) {
    stop("Set OPENALEX_API_KEY to retrieve missing OpenAlex batches.")
  }

  body <- request("https://api.openalex.org/sources") |>
    req_url_query(
      filter = paste0("issn:", paste(issns, collapse = "|")),
      per_page = 100,
      cursor = cursor,
      api_key = api_key
    ) |>
    req_timeout(seconds = 30) |>
    req_retry(
      max_tries = 4L,
      retry_on_failure = TRUE
    ) |>
    req_perform() |>
    resp_body_json(simplifyVector = FALSE)

  stopifnot(
    is.list(body$results),
    is.list(body$meta),
    length(body$meta$count) == 1L,
    is.null(body$meta$next_cursor) ||
      length(body$meta$next_cursor) == 1L
  )
  body
}

fetch_openalex_batch <- function(issns) {
  stopifnot(length(issns) <= 100L)
  cursor <- "*"
  sources <- list()
  expected_count <- NULL

  repeat {
    page <- fetch_openalex_page(issns, cursor)
    if (is.null(expected_count)) expected_count <- as.integer(page$meta$count)
    sources <- c(sources, page$results)
    next_cursor <- page$meta$next_cursor
    if (identical(next_cursor, cursor)) stop("OpenAlex cursor did not advance.")
    cursor <- next_cursor
    if (is.null(cursor) || !nzchar(cursor)) break
  }

  stopifnot(length(sources) == expected_count)
  sources
}

fetch_crossref_page <- function(cursor) {
  mailto <- Sys.getenv("CROSSREF_MAILTO")
  if (!grepl("^[^@[:space:]]+@[^@[:space:]]+$", mailto)) {
    stop("Set CROSSREF_MAILTO to retrieve the missing Crossref directory.")
  }

  body <- request("https://api.crossref.org/journals") |>
    req_url_query(rows = 1000, cursor = cursor, mailto = mailto) |>
    req_user_agent(
      paste(
        "InvisibleResearch journal enrichment;",
        "https://github.com/YannJY02/InvisibleResearch"
      )
    ) |>
    req_timeout(seconds = 60) |>
    req_retry(
      max_tries = 4L,
      retry_on_failure = TRUE
    ) |>
    req_perform() |>
    resp_body_json(simplifyVector = FALSE) |>
    pluck("message")

  stopifnot(is.list(body$items))
  body
}

fetch_crossref_directory <- function() {
  cursor <- "*"
  journals <- list()

  repeat {
    page <- fetch_crossref_page(cursor)
    journals <- c(journals, page$items)
    if (length(page$items) < 1000L) break

    next_cursor <- page[["next-cursor"]]
    if (is.null(next_cursor) || !nzchar(next_cursor)) {
      stop("Crossref returned a full page without a next cursor.")
    }
    if (identical(next_cursor, cursor)) stop("Crossref cursor did not advance.")
    cursor <- next_cursor
    Sys.sleep(0.2)
  }

  journals
}
```

## Load the complete PKP V7 cohort

```{r}
#| label: load-pkp

pkp_rows <- read_csv(
  input_path,
  col_types = cols(.default = col_character()),
  na = character(),
  show_col_types = FALSE
)

row_identity <- function(data) {
  paste(data$oai_url, data$repository_name, data$set_spec, sep = "\r")
}

stopifnot(
  all(c("issn", "oai_url", "repository_name", "set_spec") %in% names(pkp_rows)),
  all(tolower(pkp_rows$application) == "ojs")
)

input_issns <- map(pkp_rows$issn, parse_issns)
tokens <- sort(unique(unlist(input_issns)))
```

## Matching and field expansion helpers

```{r}
#| label: matching-helpers

index_candidates <- function(candidate_issns) {
  index <- new.env(hash = TRUE, parent = emptyenv())
  for (position in seq_along(candidate_issns)) {
    for (issn in candidate_issns[[position]]) {
      existing <- get0(
        issn,
        index,
        inherits = FALSE,
        ifnotfound = integer()
      )
      assign(issn, c(existing, position), index)
    }
  }
  index
}

match_candidates <- function(issns, index, candidate_issns) {
  positions <- unique(unlist(map(issns, \(issn) {
    get0(issn, index, inherits = FALSE, ifnotfound = integer())
  })))
  matched_issns <- base::intersect(
    issns,
    unique(unlist(candidate_issns[positions]))
  )
  status <- case_when(
    length(issns) == 0L ~ "not_attempted",
    length(positions) == 0L ~ "unmatched",
    length(positions) == 1L ~ "unique",
    TRUE ~ "ambiguous"
  )
  list(
    status = status,
    matched_issns = matched_issns,
    positions = positions
  )
}

serialize_json <- function(value) {
  as.character(jsonlite::toJSON(
    value,
    auto_unbox = TRUE,
    null = "null",
    na = "null",
    digits = NA
  ))
}

dataframe_value <- function(value) {
  if (is.null(value)) return(NA_character_)
  if (is.atomic(value) && length(value) == 1L) {
    return(as.character(value))
  }
  serialize_json(value)
}

scalar_column <- function(values) {
  present <- keep(values, \(value) !is.null(value) && length(value) == 1L)
  types <- unique(map_chr(present, typeof))
  missing <- \(value) is.null(value) || length(value) == 0L

  if (length(types) > 0L && all(types == "logical")) {
    return(map_lgl(values, \(value) if (missing(value)) NA else value))
  }
  if (length(types) > 0L && all(types == "integer")) {
    return(map_int(values, \(value) if (missing(value)) NA_integer_ else value))
  }
  if (length(types) > 0L && all(types %in% c("integer", "double"))) {
    return(map_dbl(values, \(value) if (missing(value)) NA_real_ else value))
  }
  map_chr(values, \(value) {
    if (missing(value)) NA_character_ else as.character(value)
  })
}

expand_source_fields <- function(matches, candidates, prefix) {
  fields <- sort(unique(unlist(map(candidates, names))))
  single_positions <- map_int(matches, \(match) {
    if (length(match$positions) == 1L) match$positions else NA_integer_
  })

  # ponytail: JSON conversion is single-process for this one-off; use Arrow
  # nested columns only if repeated full rebuilds make serialization a problem.
  columns <- set_names(
    map(fields, \(field) {
      values <- map(candidates, \(candidate) candidate[[field]])
      scalar_field <- all(map_lgl(values, \(value) {
        is.null(value) || (is.atomic(value) && length(value) <= 1L)
      }))
      candidate_values <- if (scalar_field) {
        scalar_column(values)
      } else {
        map_chr(values, dataframe_value)
      }
      candidate_values[single_positions]
    }),
    paste(prefix, fields, sep = "_")
  )
  columns[[paste0(prefix, "_candidates__json")]] <- map_chr(
    matches,
    \(match) {
      if (length(match$positions) > 1L) {
        serialize_json(candidates[match$positions])
      } else {
        NA_character_
      }
    }
  )
  as_tibble(columns, .name_repair = "check_unique")
}
```

## OpenAlex enrichment

```{r}
#| label: openalex-enrichment

token_batches <- split(tokens, ceiling(seq_along(tokens) / 100L))
openalex_retrieved <- FALSE
openalex_batches <- imap(token_batches, \(batch, batch_number) {
  checkpoint_path <- file.path(
    openalex_batch_dir,
    sprintf("batch-%04d.rds", as.integer(batch_number))
  )
  checkpoint <- read_cache(checkpoint_path)
  reusable <- is.list(checkpoint) &&
    identical(checkpoint$issns, batch) &&
    is.list(checkpoint$sources)

  if (!reusable && reuse_shared_cache) {
    shared_checkpoint_path <- file.path(
      shared_openalex_batch_dir,
      basename(checkpoint_path)
    )
    checkpoint <- read_cache(shared_checkpoint_path)
    reusable <- is.list(checkpoint) &&
      identical(checkpoint$issns, batch) &&
      is.list(checkpoint$sources)
    if (reusable) checkpoint_path <- shared_checkpoint_path
  }

  if (!reusable) {
    openalex_retrieved <<- TRUE
    checkpoint <- list(
      issns = batch,
      sources = fetch_openalex_batch(batch)
    )
    saveRDS(checkpoint, checkpoint_path)
  }
  checkpoint$sources
})

openalex_sources <- unlist(openalex_batches, recursive = FALSE)
openalex_ids <- map_chr(openalex_sources, "id")
keep <- !duplicated(openalex_ids)
openalex_sources <- openalex_sources[keep]
openalex_ids <- openalex_ids[keep]
openalex_issns_by_candidate <- map(openalex_sources, openalex_issns)
relevant <- map_lgl(
  openalex_issns_by_candidate,
  \(issns) any(issns %in% tokens)
)
openalex_sources <- openalex_sources[relevant]
openalex_ids <- openalex_ids[relevant]
openalex_issns_by_candidate <- openalex_issns_by_candidate[relevant]
openalex_index <- index_candidates(openalex_issns_by_candidate)
openalex_matches <- map(
  input_issns,
  match_candidates,
  index = openalex_index,
  candidate_issns = openalex_issns_by_candidate
)

base_result <- pkp_rows |>
  mutate(
    identifier_status = case_when(
      is_missing_value(issn) ~ "missing",
      lengths(input_issns) == 0L ~ "invalid",
      TRUE ~ "valid"
    ),
    openalex_input_issns = map_chr(input_issns, paste, collapse = "|"),
    openalex_matched_issns = map_chr(
      openalex_matches,
      \(match) paste(match$matched_issns, collapse = "|")
    ),
    openalex_match_status = map_chr(openalex_matches, "status"),
    openalex_candidate_count = map_int(
      openalex_matches,
      \(match) length(match$positions)
    ),
    openalex_candidate_ids = map_chr(
      openalex_matches,
      \(match) paste(openalex_ids[match$positions], collapse = "|")
    )
  )

load_expanded <- function(path, source_retrieved) {
  # ponytail: this one-off cache has no version metadata; delete it after
  # changing matching or field-expansion logic.
  if (!reuse_shared_cache || source_retrieved) return(NULL)
  cached <- read_cache(path)
  if (is.list(cached) && is.data.frame(cached$data)) cached <- cached$data
  if (is.data.frame(cached) && nrow(cached) == nrow(pkp_rows)) cached else NULL
}

openalex_data <- load_expanded(openalex_expanded_path, openalex_retrieved)
if (is.null(openalex_data)) {
  openalex_data <- expand_source_fields(
    openalex_matches,
    openalex_sources,
    "openalex"
  )
  saveRDS(openalex_data, openalex_expanded_path)
}
rm(openalex_batches, openalex_index)
gc()
```

## Crossref enrichment

```{r}
#| label: crossref-enrichment

crossref_retrieved <- FALSE
crossref_journals <- read_cache(crossref_cache_path)
if (is.list(crossref_journals$journals)) {
  crossref_journals <- crossref_journals$journals
}
if (!is.list(crossref_journals) && reuse_shared_cache) {
  crossref_journals <- read_cache(shared_crossref_cache_path)
  if (is.list(crossref_journals$journals)) {
    crossref_journals <- crossref_journals$journals
  }
}
if (!is.list(crossref_journals)) {
  crossref_journals <- fetch_crossref_directory()
  saveRDS(crossref_journals, crossref_cache_path)
  crossref_retrieved <- TRUE
}
crossref_issns_by_candidate <- map(crossref_journals, crossref_issns)
relevant <- map_lgl(
  crossref_issns_by_candidate,
  \(issns) any(issns %in% tokens)
)
crossref_journals <- crossref_journals[relevant]
crossref_issns_by_candidate <- crossref_issns_by_candidate[relevant]
crossref_ids <- map_chr(
  crossref_issns_by_candidate,
  \(issns) paste(sort(issns), collapse = "|")
)
keep <- nzchar(crossref_ids)
crossref_journals <- crossref_journals[keep]
crossref_ids <- crossref_ids[keep]
crossref_issns_by_candidate <- crossref_issns_by_candidate[keep]
crossref_index <- index_candidates(crossref_issns_by_candidate)
crossref_matches <- map(
  input_issns,
  match_candidates,
  index = crossref_index,
  candidate_issns = crossref_issns_by_candidate
)

crossref_data <- load_expanded(crossref_expanded_path, crossref_retrieved)
if (is.null(crossref_data)) {
  crossref_data <- expand_source_fields(
    crossref_matches,
    crossref_journals,
    "crossref"
  )
  saveRDS(crossref_data, crossref_expanded_path)
}

result <- base_result |>
  mutate(
    crossref_input_issns = map_chr(input_issns, paste, collapse = "|"),
    crossref_matched_issns = map_chr(
      crossref_matches,
      \(match) paste(match$matched_issns, collapse = "|")
    ),
    crossref_match_status = map_chr(crossref_matches, "status"),
    crossref_candidate_count = map_int(
      crossref_matches,
      \(match) length(match$positions)
    ),
    crossref_candidate_issn_sets = map_chr(
      crossref_matches,
      \(match) paste(crossref_ids[match$positions], collapse = ";")
    )
  ) |>
  bind_cols(openalex_data, crossref_data)
```

## Parquet outputs and lightweight validation

```{r}
#| label: parquet-output

sample_spec <- tribble(
  ~context_name, ~oai_url, ~repository_name, ~set_spec,
  "Kuwait Journal of Science",
  "https://journalskuwait.org/kjs/index.php/index/oai",
  "Kuwait Journal of Science", "KJS",
  "Cakrawala",
  "https://cakrawalajournal.org/index.php/index/oai",
  "CAKRAWALA", "cakrawala",
  "Sintesa: Jurnal Ilmu Pendidikan",
  "https://sintesa.stkip-arrahmaniyah.ac.id/index.php/index/oai",
  "Sintesa: Jurnal Ilmu Pendidikan STKIP Arrahmaniyah Depok", "sintesa",
  "CARAKA: Jurnal Teologi Biblika dan Praktika",
  "https://ojs.sttibc.ac.id/index.php/index/oai",
  "CARAKA:Jurnal Teologi", "ibc",
  "Jurnal Tata Kelola dan Akuntabilitas Keuangan Negara",
  "https://jurnal.bpk.go.id/index.php/index/oai",
  "Open Journal Systems", "TAKEN",
  "JOEEL (Journal of English Education and Literature)",
  "https://journal.stkippamanetalino.ac.id/index.php/index/oai",
  "Rumah Jurnal STKIP Pamane Talino Ngabang", "bahasa-inggris",
  "Malaysian Journal of Paediatrics and Child Health",
  "https://mpaeds.my/journals/index.php/index/oai",
  "NA", "MJPCH",
  "Jami Scientific Research Quarterly Journal",
  "https://journals.jami.edu.af/index.php/index/oai",
  "Jami University Press", "jsrqj",
  "Journal of Polymer & Composites",
  "https://engineeringjournals.stmjournals.in/index.php/index/oai",
  "Engineering Journals", "JoPC",
  "JURNAL TEKNIK PERTAMBANGAN",
  "https://e-journal.upr.ac.id/index.php/index/oai",
  "Journal Online Universitas Palangka Raya", "JTP"
)
sample_positions <- match(row_identity(sample_spec), row_identity(result))
stopifnot(!anyNA(sample_positions))
sample_result <- result[sample_positions, ]

candidate_json_is_complete <- function(matches, values) {
  map2_lgl(matches, values, \(match, value) {
    candidate_count <- length(match$positions)
    if (candidate_count <= 1L) return(is.na(value))
    parsed <- tryCatch(
      jsonlite::fromJSON(value, simplifyVector = FALSE),
      error = \(error) NULL
    )
    is.list(parsed) && length(parsed) == candidate_count
  })
}

stopifnot(
  !"admin_email" %in% names(result),
  identical(row_identity(result), row_identity(pkp_rows)),
  all(candidate_json_is_complete(
    openalex_matches,
    result$openalex_candidates__json
  )),
  all(candidate_json_is_complete(
    crossref_matches,
    result$crossref_candidates__json
  ))
)

pending_output_path <- paste0(output_path, ".pending")
pending_sample_output_path <- paste0(sample_output_path, ".pending")
arrow::write_parquet(
  as.data.frame(result),
  pending_output_path,
  compression = "zstd"
)
arrow::write_parquet(
  as.data.frame(sample_result),
  pending_sample_output_path,
  compression = "zstd"
)

parquet_info <- nanoparquet::read_parquet_info(pending_output_path)
parquet_schema <- nanoparquet::read_parquet_schema(pending_output_path)
observed_sample <- nanoparquet::read_parquet(pending_sample_output_path)

stopifnot(
  parquet_info$num_rows == nrow(result),
  parquet_info$num_cols == ncol(result),
  identical(
    parquet_schema$name[!is.na(parquet_schema$r_col)],
    names(result)
  ),
  isTRUE(all.equal(
    as.data.frame(observed_sample),
    as.data.frame(sample_result),
    check.attributes = FALSE
  ))
)

stopifnot(
  file.rename(pending_output_path, output_path),
  file.rename(pending_sample_output_path, sample_output_path)
)
```

## Ten-row visual review

```{r}
#| label: visual-review

sample_result |>
  mutate(issn = gsub("[\\r\\n]+", " / ", issn, perl = TRUE)) |>
  DT::datatable(
    rownames = FALSE,
    filter = "top",
    options = list(
      scrollX = TRUE,
      pageLength = 10L,
      lengthChange = FALSE
    )
  )

bind_rows(
  count(result, openalex_match_status, name = "rows") |>
    rename(status = openalex_match_status) |>
    mutate(source = "OpenAlex"),
  count(result, crossref_match_status, name = "rows") |>
    rename(status = crossref_match_status) |>
    mutate(source = "Crossref")
) |>
  select(source, status, rows) |>
  knitr::kable()

tibble(column = names(result)) |>
  knitr::kable()
```
